### 2019-09-24 source/cadabra/checks/check-calzetta.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

% ============================================================================================
\section*{Checking the 2nd and 3rd order terms of Calzetta etal.}

The following calculations show that my results for the RNC connection agree with those of Calzetta etal. to third order terms.

Note that I take $\nabla_{ab}$ to be $\nabla_a\left(\nabla_b\right)$.

Note also that $(LCB)\>R_{abcd} = - (Calzetta)\> R_{abcd}$. Consequently, I replace $R_{abcd}$ with
$-R_{abcd}$ in the Calzetta expressions (done as a Cadabra substitution rule).

This is relatively straightforward. We just apply a few carefully chosen applications of the first and second Bianchi identities.

Note that in this example, 2nd and 3rd order refer to the powers of $x$ in the expression. This differs from
the usage elsewhere in these examples (3rd and 4th order with respect to $R$ and its derivatives).

\clearpage

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,u,v,w#}::Indices("latin",position=independent).
   {\mu,\nu,\rho,\sigma,\tau,\lambda,\xi#}::Indices("greek",position=independent).

   \nabla{#}::Derivative.

   g_{a b}::Metric.
   g^{a b}::InverseMetric.
   g^{a b}::Weight(label=gnum,value=1).

   \delta{#}::KroneckerDelta.

   R_{a b c d}::RiemannTensor.
   R_{a b c d}::Depends(\nabla{#}).

   x^{a}::Weight(label=xnum,value=1).

   def add_tags (obj,tag):

      n = 0
      ans = Ex('0')

      for i in obj.top().terms():
         foo = obj[i]
         bah = Ex(tag+'_{'+str(n)+'}')
         ans := @(ans) + @(bah) @(foo).
         n = n + 1

      return ans

   def clear_tags (obj,tag):

      ans := @(obj).
      foo  = Ex(tag+'_{a?} -> 1')
      substitute (ans,foo)

      return ans

   def get_xterm (obj,n):

       foo := @(obj).
       bah  = Ex("xnum = " + str(n))
       distribute  (foo)
       keep_weight (foo, bah)

       return foo

   def get_gterm (obj,n):

       foo := @(obj).
       bah  = Ex("gnum = " + str(n))
       distribute  (foo)
       keep_weight (foo, bah)

       return foo

   # note: keeping numbering as is (out of order) to ensure R appears before \nabla R etc.
   def product_sort (obj):
       substitute (obj,$ g^{a b}                   -> A001^{a b}                $)
       substitute (obj,$ x^{a}                     -> A002^{a}                  $)
       substitute (obj,$ z^{a}                     -> A003^{a}                  $)
       substitute (obj,$ \nabla_{e f}{R_{a b c d}} -> A006_{a b c d e f}        $)
       substitute (obj,$ \nabla_{e}{R_{a b c d}}   -> A005_{a b c d e}          $)
       substitute (obj,$ R_{a b c d}               -> A004_{a b c d}            $)
       sort_sum       (obj)
       sort_product   (obj)
       rename_dummies (obj)
       substitute (obj,$ A001^{a b}                -> g^{a b}                   $)
       substitute (obj,$ A002^{a}                  -> x^{a}                     $)
       substitute (obj,$ A003^{a}                  -> z^{a}                     $)
       substitute (obj,$ A004_{a b c d}            -> R_{a b c d}               $)
       substitute (obj,$ A005_{a b c d e}          -> \nabla_{e}{R_{a b c d}}   $)
       substitute (obj,$ A006_{a b c d e f}        -> \nabla_{e f}{R_{a b c d}} $)

       return obj

   def reformat (obj,scaleA,scaleB):

      foo  = Ex(str(scaleA))
      moo  = Ex(str(scaleB))
      bah := @(foo) @(obj) / @(moo).

      distribute     (bah)
      bah = product_sort (bah)
      rename_dummies (bah)
      canonicalise   (bah)
      factor_out     (bah,$g^{c? d?}$)
      factor_out     (bah,$x^{a?},z^{b?}$)
      ans := @(moo) @(bah) / @(foo).

      return ans

   # ==========================================================================
   # LCB

   import cdblib
   Gamma  = cdblib.get ('Gamma','../connection.json')             # cdb(ex-12.100,Gamma)

   # note that the next two lines require careful inspection of the free indices on Gamma
   # expecting Gamma = \Gamma^{d}_{ab}
   Gamma := z^{a} z^{b} @(Gamma).

   # lower index ^{d} to _{v}

   Gamma := g_{v d} @(Gamma).

   distribute (Gamma)
   substitute (Gamma, $g_{a d} g^{d b} -> \delta_{a}^{b}$)
   eliminate_kronecker (Gamma)                                    # cdb(ex-12.101,Gamma)

   # change free index _{v} to _{a}

   foo := tmp_{v} -> @(Gamma).                                    # cdb(ex-12.191,foo)
   bah := tmp_{a}.                                                # cdb(ex-12.192,bah)
   substitute (bah, foo)                                          # cdb(ex-12.193,bah)

   Gamma := @(bah).                                               # cdb(ex-12.102,Gamma)

   Gamma = product_sort (Gamma)                                   # cdb(ex-12.103,Gamma)

   gam1  = get_xterm (Gamma,1)                                    # cdb(ex-12.200,gam1)
   gam2  = get_xterm (Gamma,2)                                    # cdb(ex-12.201,gam2)
   gam3  = get_xterm (Gamma,3)                                    # cdb(ex-12.202,gam3)

   gam30 = get_gterm (gam3,0)                                     # cdb(ex-12.203,gam30)
   gam31 = get_gterm (gam3,1)                                     # cdb(ex-12.204,gam31)

   gam1  = reformat (gam1, 3,1)                                   # cdb(ex-12.300,gam1)
   gam2  = reformat (gam2,12,1)                                   # cdb(ex-12.301,gam2)

   gam30 = reformat (gam30,40,1)                                  # cdb(ex-12.302,gam30)
   gam31 = reformat (gam31,45,2)                                  # cdb(ex-12.303,gam31)

   gam3 := @(gam30) + @(gam31).                                   # cdb(ex-12.304,gam3)

   Gamma := @(gam1) + @(gam2) + @(gam3).                          # cdb(ex-12.305,Gamma)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.100}}
   \Dmath*{\cdb*{ex-12.191}}
   \Dmath*{\cdb*{ex-12.192}}
   \Dmath*{\cdb*{ex-12.193}}
   \Dmath*{\cdb*{ex-12.101}}
   \Dmath*{\cdb*{ex-12.102}}
   \Dmath*{\cdb*{ex-12.103}}
\end{dgroup*}

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.200}}
   \Dmath*{\cdb*{ex-12.201}}
   \Dmath*{\cdb*{ex-12.202}}
   \Dmath*{\cdb*{ex-12.203}}
   \Dmath*{\cdb*{ex-12.204}}
\end{dgroup*}

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.300}}
   \Dmath*{\cdb*{ex-12.301}}
   \Dmath*{\cdb*{ex-12.302}}
   \Dmath*{\cdb*{ex-12.303}}
   \Dmath*{\cdb*{ex-12.304}}
   \Dmath*{\cdb*{ex-12.305}}
\end{dgroup*}

\clearpage

\begin{cadabra}
   # ==========================================================================
   # Calzetta
   # note: \nabla_{a b} defined as \nabla_{a}\nabla_{b}

   GammaBar := z^{\nu} z^{\rho} (
                 (2/3) R^{\mu}_{\nu\rho\sigma} x^{\sigma}
               + (1/12) (5 \nabla_{\lambda}{R^{\mu}_{\nu\rho\sigma}}
                         + \nabla_{\rho}{R^{\mu}_{\sigma\nu\lambda}}) x^{\sigma} x^{\lambda}
               + (1/6) (  (9/10) \nabla_{\tau\lambda}{R^{\mu}_{\rho\nu\sigma}}
                        + (3/20) (  \nabla_{\tau\rho}{R^{\mu}_{\sigma\nu\lambda}}
                                  + \nabla_{\rho\tau}{R^{\mu}_{\sigma\nu\lambda}} )
                        + (1/60) (  21 R^{\mu}_{\lambda\xi\rho} R^{\xi}_{\sigma\nu\tau}
                                  + 48 R^{\mu}_{\xi\rho\lambda} R^{\xi}_{\sigma\nu\tau}
                                  - 37 R^{\mu}_{\sigma\xi\lambda} R^{\xi}_{\nu\rho\tau} ) ) x^{\sigma} x^{\lambda} x^{\tau} ).
                                                                  # cdb(ex-12.400,GammaBar)

   # convert from Greek to Latin indices

   distribute (GammaBar)
   rename_dummies (GammaBar,"greek","latin")                      # cdb(ex-12.401,GammaBar)

   # lower the \mu index

   GammaBar := \delta_{a \mu} @(GammaBar).                        # cdb(ex-12.402,GammaBar)
   distribute (GammaBar)                                          # cdb(ex-12.403,GammaBar)
   eliminate_kronecker (GammaBar)                                 # cdb(ex-12.404,GammaBar)

   # sort products

   GammaBar = product_sort (GammaBar)                             # cdb(ex-12.405,GammaBar)

   # Replace R with - R (Calzetta uses the non-MTW convention for Riemann)

   substitute (GammaBar, $R_{a b c d} -> - R_{a b c d}$)          # cdb(ex-12.406,GammaBar)
   substitute (GammaBar, $R^{a}_{b c d} -> - R^{a}_{b c d}$)      # cdb(ex-12.407,GammaBar)

   substitute (GammaBar, $R^{a}_{b c d} -> g^{a e} R_{e b c d}$)  # cdb(ex-12.408,GammaBar)

   cal1 = get_xterm (GammaBar,1)                                  # cdb(ex-12.500,cal1)
   cal2 = get_xterm (GammaBar,2)                                  # cdb(ex-12.501,cal2)
   cal3 = get_xterm (GammaBar,3)                                  # cdb(ex-12.502,cal3)

   cal1 = reformat (cal1,3,1)                                     # cdb(ex-12.600,cal1)
   cal2 = reformat (cal2,12,1)                                    # cdb(ex-12.601,cal2)
   # cal3 = reformat (cal3,360,1)                                   # cdb(ex-12.602,cal3)

   cal30 = get_gterm (cal3,0)                                     # cdb(ex-12.602,cal30)
   cal31 = get_gterm (cal3,1)                                     # cdb(ex-12.603,cal31)

   cal1  = reformat (cal1, 3,1)                                   # cdb(ex-12.604,cal1)
   cal2  = reformat (cal2,12,1)                                   # cdb(ex-12.605,cal2)

   cal30 = reformat (cal30,40,1)                                  # cdb(ex-12.606,cal30)
   cal31 = reformat (cal31,360,1)                                 # cdb(ex-12.607,cal31)

   cal3 := @(cal30) + @(cal31).                                   # cdb(ex-12.608,cal3)

   GammaBar := @(cal1) + @(cal2) + @(cal3).                       # cdb(ex-12.409,GammaBar)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.400}}
   \Dmath*{\cdb*{ex-12.401}}
   \Dmath*{\cdb*{ex-12.402}}
   \Dmath*{\cdb*{ex-12.403}}
   \Dmath*{\cdb*{ex-12.404}}
   \Dmath*{\cdb*{ex-12.405}}
   \Dmath*{\cdb*{ex-12.406}}
   \Dmath*{\cdb*{ex-12.407}}
   \Dmath*{\cdb*{ex-12.408}}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.500}}
   \Dmath*{\cdb*{ex-12.501}}
   \Dmath*{\cdb*{ex-12.502}}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.600}}
   \Dmath*{\cdb*{ex-12.601}}
   \Dmath*{\cdb*{ex-12.602}}
   \Dmath*{\cdb*{ex-12.603}}
   \Dmath*{\cdb*{ex-12.604}}
   \Dmath*{\cdb*{ex-12.605}}
   \Dmath*{\cdb*{ex-12.606}}
   \Dmath*{\cdb*{ex-12.607}}
   \Dmath*{\cdb*{ex-12.608}}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.409}}
\end{dgroup*}

\clearpage

\def\GammaBar{{\bar{\Gamma}}}

% -----------------------------------------------------------------------------
\subsection*{The fun begins $\Gamma - \GammaBar$}

It's now time to compute the difference $\Gamma-\GammaBar$. Here it is.

\begin{cadabra}
   def reformat_diff (obj):

       distribute (obj)

       obj1  = get_xterm (obj,1)
       obj2  = get_xterm (obj,2)
       obj3  = get_xterm (obj,3)

       obj30 = get_gterm (obj3,0)
       obj31 = get_gterm (obj3,1)

       obj1  = reformat (obj1, 3,1)
       obj2  = reformat (obj2,12,1)

       obj30 = reformat (obj30,40,1)
       obj31 = reformat (obj31,360,1)

       obj3 := @(obj30) + @(obj31).

       ans := @(obj1) + @(obj2) + @(obj3).

       return ans

   # We could use reformat_diff here but instead we'll do it one step at a time so that
   # we can see exactly what's going on. Later on we will use reformat_diff to do the job.

   diff := @(Gamma) - @(GammaBar).                                # cdb(ex-12.diff.100,diff)
   distribute (diff)

   diff1  = get_xterm (diff,1)                                    # cdb(ex-12.diff.200,diff1)
   diff2  = get_xterm (diff,2)                                    # cdb(ex-12.diff.201,diff2)
   diff3  = get_xterm (diff,3)                                    # cdb(ex-12.diff.202,diff3)

   diff30 = get_gterm (diff3,0)                                   # cdb(ex-12.diff.203,diff30)
   diff31 = get_gterm (diff3,1)                                   # cdb(ex-12.diff.204,diff31)

   diff1  = reformat (diff1, 3,1)                                 # cdb(ex-12.diff.300,diff1)
   diff2  = reformat (diff2,12,1)                                 # cdb(ex-12.diff.301,diff2)

   diff30 = reformat (diff30,40,1)                                # cdb(ex-12.diff.302,diff30)
   diff31 = reformat (diff31,360,1)                               # cdb(ex-12.diff.303,diff31)

   diff3 := @(diff30) + @(diff31).                                # cdb(ex-12.diff.304,diff3)

   diff := @(diff1) + @(diff2) + @(diff3).                        # cdb(ex-12.diff.305,diff)

\end{cadabra}

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.diff.100}}
\end{dgroup*}

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.diff.200}}
   \Dmath*{\cdb*{ex-12.diff.201}}
   \Dmath*{\cdb*{ex-12.diff.202}}
   \Dmath*{\cdb*{ex-12.diff.203}}
   \Dmath*{\cdb*{ex-12.diff.204}}
\end{dgroup*}

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.diff.300}}
   \Dmath*{\cdb*{ex-12.diff.301}}
   \Dmath*{\cdb*{ex-12.diff.302}}
   \Dmath*{\cdb*{ex-12.diff.303}}
   \Dmath*{\cdb*{ex-12.diff.304}}
   \Dmath*{\cdb*{ex-12.diff.305}}
\end{dgroup*}

\clearpage

% -----------------------------------------------------------------------------
\subsection*{Second order terms}

\begin{cadabra}
   diff2  = get_xterm (diff,2)
   diff2 := 12 @(diff2).                                                        # cdb (ex-12.701,diff2)
   distribute (diff2)                                                           # cdb (ex-12.702,diff2)

   diff2 = add_tags (diff2,'\\mu')                                              # cdb (ex-12.711,diff2)

   # swap indices on middle term, then apply 2nd Bianchi identity

   zoom       (diff2, $\mu_{1} Q??$)                                            # cdb (ex-12.712,diff2)
   substitute (diff2, $\nabla_{b}{R_{a d c e}} -> - \nabla_{b}{R_{d a c e}}$)   # cdb (ex-12.713,diff2)
   unzoom     (diff2)

   substitute (diff2, $\mu_{1} -> \mu_{0}, \mu_{2} -> \mu_{0}$)                 # cdb (ex-12.714,diff2)
   substitute (diff2, $\mu_{0} -> 0$)                                           # cdb (ex-12.715,diff2)

   diff2 = clear_tags (diff2,'\\mu')                                            # cdb (ex-12.716,diff2)

   diff2 := @(diff2) / 12 .

   diff := @(diff1) + @(diff2) + @(diff3).

   diff  = reformat_diff (diff)                                                 # cdb(ex-12.diff.306,diff)
\end{cadabra}

\clearpage

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.701}}
   \Dmath*{\cdb*{ex-12.702}}
\end{dgroup*}

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.711}}
   \Dmath*{\cdb*{ex-12.712}}
   \Dmath*{\cdb*{ex-12.713}}
   \Dmath*{\cdb*{ex-12.714}}
   \Dmath*{\cdb*{ex-12.715}}
   \Dmath*{\cdb*{ex-12.716}}
\end{dgroup*}

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.diff.306}}
\end{dgroup*}

\clearpage

% -----------------------------------------------------------------------------
\subsection*{Third order terms, commute $\nabla\nabla R$ terms}

\begin{cadabra}
   diff3  = get_xterm (diff,3)
   diff3 := 360 @(diff3).                              # cdb (ex-12.801,diff3)
   distribute (diff3)                                  # cdb (ex-12.802,diff3)

   # commutation rule for covariant derivs on Rabcd, see exrecise 3.6
   # note: \nabla_{a b} defined as \nabla_{a}\nabla_{b}
   CommuteNablaRiemann := \nabla_{f e}(R_{a b c d}) -> \nabla_{e f}(R_{a b c d})
                                                     + g^{u v} R_{u a e f} R_{v b c d}
                                                     + g^{u v} R_{u b e f} R_{a v c d}
                                                     + g^{u v} R_{u c e f} R_{a b v d}
                                                     + g^{u v} R_{u d e f} R_{a b c v}.

   diff3 = add_tags (diff3,'\\mu')                     # cdb (ex-12.901,diff3)

   # commute derivs on Rabcd so that each double deriv is of the form \nabla_{b*}

   substitute (diff3, $\mu_{3} -> \mu_{1}$)            # cdb (ex-12.902,diff3)

   zoom       (diff3, $\mu_{1} Q??$)                   # cdb (ex-12.903,diff3)
   substitute (diff3, CommuteNablaRiemann)             # cdb (ex-12.904,diff3)
   unzoom     (diff3)

   diff3 = clear_tags (diff3,'\\mu')
   diff3 := @(diff3) / 360 .

   distribute   (diff3)
   canonicalise (diff3)                                # cdb (ex-12.905,diff3)

   diff := @(diff1) + @(diff2) + @(diff3).

   diff  = reformat_diff (diff)                        # cdb(ex-12.diff.307,diff)
\end{cadabra}

\clearpage

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.801}}
   \Dmath*{\cdb*{ex-12.802}}
\end{dgroup*}

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.901}}
   \Dmath*{\cdb*{ex-12.902}}
   \Dmath*{\cdb*{ex-12.903}}
   \Dmath*{\cdb*{ex-12.904}}
   \Dmath*{\cdb*{ex-12.905}}
\end{dgroup*}

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.diff.307}}
\end{dgroup*}

\clearpage

% -----------------------------------------------------------------------------
\subsection*{Third order terms, use 2nd Bianchi identity on $\nabla\nabla R$ terms}

\begin{cadabra}
   diff3  = get_xterm (diff,3)
   diff3 := 360 @(diff3).                                                           # cdb (ex-12.910,diff3)
   distribute (diff3)                                                               # cdb (ex-12.911,diff3)

   diff3 = add_tags (diff3,'\\mu')                                                  # cdb (ex-12.912,diff3)

   # swap indices on middle second deriv term, then apply 2nd Bianchi identity

   zoom       (diff3, $\mu_{1} Q??$)                                                # cdb (ex-12.913,diff3)
   substitute (diff3, $\nabla_{b c}{R_{a e d f}} -> - \nabla_{b c}{R_{e a d f}}$)   # cdb (ex-12.914,diff3)
   unzoom     (diff3)

   substitute (diff3, $\mu_{1} -> \mu_{0}, \mu_{2} -> \mu_{0}$)                     # cdb (ex-12.915,diff3)
   substitute (diff3, $\mu_{0} -> 0$)                                               # cdb (ex-12.916,diff3)

   diff3 = clear_tags (diff3,'\\mu')
   diff3 := @(diff3) / 360 .

   distribute   (diff3)
   canonicalise (diff3)                                                             # cdb (ex-12.917,diff3)

   diff := @(diff1) + @(diff2) + @(diff3).

   diff  = reformat_diff (diff)                                                     # cdb(ex-12.diff.308,diff)
\end{cadabra}

\clearpage

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.910}}
   \Dmath*{\cdb*{ex-12.911}}
   \Dmath*{\cdb*{ex-12.912}}
   \Dmath*{\cdb*{ex-12.913}}
   \Dmath*{\cdb*{ex-12.914}}
   \Dmath*{\cdb*{ex-12.915}}
   \Dmath*{\cdb*{ex-12.916}}
   \Dmath*{\cdb*{ex-12.917}}
\end{dgroup*}

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.diff.308}}
\end{dgroup*}

\clearpage

% -----------------------------------------------------------------------------
\subsection*{Third order terms, use 1st Bianchi identity on $R R$ terms}

\begin{cadabra}
   diff3  = get_xterm (diff,3)
   diff3 := 360 @(diff3).
   distribute (diff3)

   diff3 = add_tags (diff3,'\\mu')                                              # cdb (ex-12.921,diff3)

   # swap indices on middle term, then apply 1st Bianchi identity

   zoom       (diff3, $\mu_{1} Q??$)                                            # cdb (ex-12.922,diff3)
   substitute (diff3, $R_{a d b g} R_{c e f h} -> - R_{a d g b} R_{c e f h}$)   # cdb (ex-12.923,diff3)
   unzoom     (diff3)

   substitute (diff3, $\mu_{1} -> \mu_{0}, \mu_{2} -> \mu_{0}$)                 # cdb (ex-12.924,diff3)
   substitute (diff3, $\mu_{0} -> 0$)                                           # cdb (ex-12.925,diff3)

   diff3 = clear_tags (diff3,'\\mu')                                            # cdb (ex-12.926,diff3)

   diff := @(diff1) + @(diff2) + @(diff3).

   diff  = reformat_diff (diff)                                                 # cdb(ex-12.diff.309,diff)
\end{cadabra}

\clearpage

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.921}}
   \Dmath*{\cdb*{ex-12.922}}
   \Dmath*{\cdb*{ex-12.923}}
   \Dmath*{\cdb*{ex-12.924}}
   \Dmath*{\cdb*{ex-12.925}}
   \Dmath*{\cdb*{ex-12.926}}
\end{dgroup*}

\begin{dgroup*}
   \Dmath*{\cdb*{ex-12.diff.309}}
\end{dgroup*}

\end{document}

### 2019-09-24 source/cadabra/checks/check-genGamma.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

% =================================================================================================
\section*{Compute $\Gamma^{a}{}_{bcd}$ directly}

Here we will compute one of generalised connections directly from the connection. That is, we will
compute
\begin{align}
   \Gamma^{a}{}_{bcd} = \Gamma^{a}{}_{(bc,d)} - 2 \Gamma^{a}{}_{p(c} \Gamma^{p}{}_{bd)}
\end{align}
given an explict expression for the RNC connection $\Gamma^{a}{}_{bc}$.

This code was written as a check for the {\tt genGamma.tex} code. I had a discrepency between my
newly created Cadabra v2.0 codes and my old Cadabra v1.0 codes (which were the basis of my lcb09-03 paper).
I found that my new codes agreed with this code and thus my old codes were wrong. I found the errors
in the old code (see the updated codes in v1.0/rnc-new/gen-gamma.cdbp, see also the file v1.0/rnc-new/NOTES.txt).

The head-on approach shown in this code works for this simple compuation but for the higher
generalised connections it's almost certainly going to be way too slow to be use. So this code is
only a check of the results from {\tt genGamma.tex} code. The good news is -- they agree.

\clearpage

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   D{#}::Derivative.
   \nabla{#}::Derivative.
   \partial{#}::PartialDerivative.

   g_{a b}::Metric.
   g^{a b}::InverseMetric.
   g_{a}^{b}::KroneckerDelta.
   g^{a}_{b}::KroneckerDelta.
   \delta^{a}_{b}::KroneckerDelta.
   \delta_{a}^{b}::KroneckerDelta.

   R_{a b c d}::RiemannTensor.
   R_{a b c d}::Depends(\nabla{#}).

   x^{a}::Depends(\partial{#}).

   \Gamma^{a}_{b c}::Depends(\partial{#}).
   \Gamma^{a}_{b c}::TableauSymmetry(shape={2}, indices={1,2}).

   Q_{a b c d}::Weight(label=numR,value=2).
   Q_{a b c d e}::Weight(label=numR,value=3).
   Q_{a b c d e f}::Weight(label=numR,value=4).
   Q_{a b c d e f g}::Weight(label=numR,value=5).

   # note: keeping numbering as is (out of order) to ensure R appears before \nabla R etc.
   def product_sort (obj):
       substitute (obj,$ A^{a}                            -> A001^{a}               $)
       substitute (obj,$ x^{a}                            -> A002^{a}               $)
       substitute (obj,$ g^{a b}                          -> A003^{a b}             $)
       substitute (obj,$ \nabla_{e f g h}{R_{a b c d}}    -> A008_{a b c d e f g h} $)
       substitute (obj,$ \nabla_{e f g}{R_{a b c d}}      -> A007_{a b c d e f g}   $)
       substitute (obj,$ \nabla_{e f}{R_{a b c d}}        -> A006_{a b c d e f}     $)
       substitute (obj,$ \nabla_{e}{R_{a b c d}}          -> A005_{a b c d e}       $)
       substitute (obj,$ R_{a b c d}                      -> A004_{a b c d}         $)
       sort_product   (obj)
       rename_dummies (obj)
       substitute (obj,$ A001^{a}                  -> A^{a}                         $)
       substitute (obj,$ A002^{a}                  -> x^{a}                         $)
       substitute (obj,$ A003^{a b}                -> g^{a b}                       $)
       substitute (obj,$ A004_{a b c d}            -> R_{a b c d}                   $)
       substitute (obj,$ A005_{a b c d e}          -> \nabla_{e}{R_{a b c d}}       $)
       substitute (obj,$ A006_{a b c d e f}        -> \nabla_{e f}{R_{a b c d}}     $)
       substitute (obj,$ A007_{a b c d e f g}      -> \nabla_{e f g}{R_{a b c d}}   $)
       substitute (obj,$ A008_{a b c d e f g h}    -> \nabla_{e f g h}{R_{a b c d}} $)

       return obj

   def truncate (obj,n):

   # I would like to assign different weights to \nabla_{a}, \nabla_{a b}, \nabla_{a b c} etc. but no matter
   # what I do it appears that Cadabra assigns the same weight to all of these regardless of the number of subscripts.
   # It seems that the weight is assigned to the symbol \nabla alone. So I'm forced to use the following substitution trick.

       tmp := @(obj).

       substitute (tmp, $\nabla_{e f g}{R_{a b c d}} -> Q_{a b c d e f g}$)
       substitute (tmp, $\nabla_{e f}{R_{a b c d}} -> Q_{a b c d e f}$)
       substitute (tmp, $\nabla_{e}{R_{a b c d}} -> Q_{a b c d e}$)
       substitute (tmp, $R_{a b c d} -> Q_{a b c d}$)

       ans = Ex(0)

       for i in range (0,n+1):
          foo := @(tmp).
          bah = Ex("numR = " + str(i))
          keep_weight (foo, bah)
          ans = ans + foo

       substitute (ans, $Q_{a b c d e f g} -> \nabla_{e f g}{R_{a b c d}}$)
       substitute (ans, $Q_{a b c d e f} -> \nabla_{e f}{R_{a b c d}}$)
       substitute (ans, $Q_{a b c d e} -> \nabla_{e}{R_{a b c d}}$)
       substitute (ans, $Q_{a b c d} -> R_{a b c d}$)

       return ans

   def get_term (obj,n):

       bah := @(obj).
       distribute (bah)

       substitute (bah, $\nabla_{e f g}{R_{a b c d}} -> Q_{a b c d e f g}$)
       substitute (bah, $\nabla_{e f}{R_{a b c d}} -> Q_{a b c d e f}$)
       substitute (bah, $\nabla_{e}{R_{a b c d}} -> Q_{a b c d e}$)
       substitute (bah, $R_{a b c d} -> Q_{a b c d}$)

       foo = Ex("numR = " + str(n))
       keep_weight (bah, foo)

       substitute (bah, $Q_{a b c d e f g} -> \nabla_{e f g}{R_{a b c d}}$)
       substitute (bah, $Q_{a b c d e f} -> \nabla_{e f}{R_{a b c d}}$)
       substitute (bah, $Q_{a b c d e} -> \nabla_{e}{R_{a b c d}}$)
       substitute (bah, $Q_{a b c d} -> R_{a b c d}$)

       return bah

   def tidy (obj,number):
      bah  = Ex(str(number))
      tmp := @(bah) @(obj).
      distribute  (tmp)
      factor_out  (tmp,$A^{a?},x^{b?}$)
      ans := @(tmp) / @(bah).
      return ans

   import cdblib

   Gamma = cdblib.get ('Gamma','../connection.json')

   defGamma := \Gamma^{d}_{a b} -> @(Gamma).

   genGam := A^{d} A^{b} A^{c} (\partial_{d}{\Gamma^{a}_{b c}} - 2 \Gamma^{a}_{p c} \Gamma^{p}_{b d}).

   substitute     (genGam,defGamma)   # cdb (genGam.001,genGam)
   distribute     (genGam)
   unwrap         (genGam)            # cdb (genGam.002,genGam)
   distribute     (genGam)            # cdb (genGam.003,genGam)
   product_rule   (genGam)            # cdb (genGam.004,genGam)
   distribute     (genGam)            # cdb (genGam.005,genGam)
   substitute     (genGam,$\partial_{a}{x^{b}}->\delta_{a}^{b}$)   # cdb (genGam.006,genGam)
   eliminate_kronecker (genGam)       # cdb (genGam.007,genGam)
   genGam = truncate     (genGam,5)   # cdb (genGam.008,genGam)
   genGam = product_sort (genGam)     # cdb (genGam.009,genGam)
   rename_dummies (genGam)            # cdb (genGam.010,genGam)
   canonicalise   (genGam)            # cdb (genGam.011,genGam)

   # all done, now for some housekeeping

   term3 = get_term (genGam,3)   # cdb (term3.001,term3)
   term4 = get_term (genGam,4)   # cdb (term4.001,term4)
   term5 = get_term (genGam,5)   # cdb (term5.001,term5)

   term3 = tidy (term3,2)
   term4 = tidy (term4,120)
   term5 = tidy (term5,180)

   genGam := @(term3) + @(term4) + @(term5).

   scaledGamma := 360 @(genGam).  # cdb (scaledGamma.001,scaledGamma)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{term3.001} \end{dmath*}
   \begin{dmath*} \cdb*{term4.001} \end{dmath*}
   \begin{dmath*} \cdb*{term5.001} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Summary}

\begin{dgroup*}
   \begin{dmath*} 360 A^b A^c A^d \Gamma^{a}_{b c d} = \cdb{scaledGamma.001} \end{dmath*}
\end{dgroup*}

\end{document}

### 2019-09-24 source/cadabra/connection.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

% =================================================================================================
% create checkpoint file

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/connection.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{The connection}

Here we use the output from {\tt metric.tex} and {\tt metric-inv.tex} to compute the metric connection
$\Gamma^{d}_{ab}$. We use the standard metric compatible connection
\begin{align}
   \label{eqn:Gamma}
   \Gamma^{d}_{ab} = \frac{1}{2} g^{dc}\left( g_{cb,a} + g_{ac,b} - g_{ab,c} \right)
\end{align}

Since {\tt metric.tex} and {\tt metric-inv.tex} generate truncated expressions for $g_{ab}$ and
$g^{ab}$ a similar truncation must be applied to this computation of $\Gamma^{d}_{ab}$. The naive
choice is to truncate $\Gamma^{d}_{ab}$ \emph{after} it has been fully evaluated on the truncated
expersions for $g_{ab}$ and $g^{ab}$. This will work but it wastes time and memory (big time).

A better approach is to truncate $\Gamma^{d}_{ab}$ during its construction. That is, we take
careful note of how the terms in the finite series for $g_{ab}$ and $g^{ab}$ combine to produce
the terms of a particular order in the expansion of $\Gamma^{d}_{ab}$.

Suppose $g_{ab}$ and $g^{ab}$ are known to say fourth order. We can write each of these as follows
\begin{align}
   g_{ab} &= \ngab{0}_{ab} + \ngab{1}_{ab} + \ngab{2}_{ab} + \ngab{3}_{ab} + \ngab{4}_{ab}\\
   g^{ab} &= \ngab{0}^{ab} + \ngab{1}^{ab} + \ngab{2}^{ab} + \ngab{3}^{ab} + \ngab{4}^{ab}
\end{align}
where $\ngab{n}$ denotes a term of order $\BigO{\eps^n}$. A similar expansion applies for $\Gamma^{d}_{ab}$, that is
\begin{align}
   \Gamma^{d}_{ab} = \nGamma{0}^{d}_{ab}
                   + \nGamma{1}^{d}_{ab}
                   + \nGamma{2}^{d}_{ab}
                   + \nGamma{3}^{d}_{ab}
                   + \nGamma{4}^{d}_{ab}
\end{align}
After substituting these formal expansions into the equation \eqref{eqn:Gamma} and then matching
corresponing terms we obtain
\begin{align}
   \nGamma{n}^{d}_{ab}
   =
   \frac{1}{2} \sum_{i=0}^{i=n} \ngab{i}^{dc}\left( \ngab{n-i}_{cb,a} + \ngab{n-i}_{ac,b} - \ngab{n-i}_{ab,c} \right)
\end{align}

We use this equation to compute the successive terms in $\Gamma^{d}_{ab}$.

\clearpage

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   D{#}::Derivative.
   \nabla{#}::Derivative.
   \partial{#}::PartialDerivative.

   g_{a b}::Metric.
   g^{a b}::InverseMetric.
   g_{a}^{b}::KroneckerDelta.
   g^{a}_{b}::KroneckerDelta.
   \delta^{a}_{b}::KroneckerDelta.
   \delta_{a}^{b}::KroneckerDelta.

   R_{a b c d}::RiemannTensor.
   R^{a}_{b c d}::RiemannTensor.

   x^{a}::Depends(D{#}).

   R_{a b c d}::Depends(\nabla{#}).
   R^{a}_{b c d}::Depends(\nabla{#}).

   import cdblib

   gab = cdblib.get ('g_ab','metric.json')      # cdb(gab.000,gab)
   iab = cdblib.get ('g^ab','metric-inv.json')  # cdb(iab.000,iab)

   defgab := g_{a b} -> @(gab).
   defiab := g^{a b} -> @(iab).

   dgab := D_{a}{g_{c b}} + D_{b}{g_{a c}} - D_{c}{g_{a b}}.  # cdb(dgab.001,dgab)

   substitute   (dgab,defgab)

   distribute   (dgab)              # cdb(dgab.002,dgab)
   unwrap       (dgab)              # cdb(dgab.003,dgab)
   product_rule (dgab)              # cdb(dgab.004,dgab)
   distribute   (dgab)              # cdb(dgab.005,dgab)
   substitute   (dgab,$D_{a}{x^{b}}->\delta^{b}_{a}$,repeat=True)  # cdb(dgab.006,dgab)
   eliminate_kronecker (dgab)       # cdb(dgab.007,dgab)
   sort_product   (dgab)            # cdb(dgab.008,dgab)
   rename_dummies (dgab)            # cdb(dgab.009,dgab)
   canonicalise   (dgab)            # cdb(dgab.010,dgab)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{gab.000} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dgab.001} \end{dmath*}
   \begin{dmath*} \cdb*{dgab.002} \end{dmath*}
   % \begin{dmath*} \cdb*{dgab.003} \end{dmath*} % too big for pdfLaTeX
   % \begin{dmath*} \cdb*{dgab.004} \end{dmath*}
   % \begin{dmath*} \cdb*{dgab.005} \end{dmath*}
   % \begin{dmath*} \cdb*{dgab.006} \end{dmath*}
   % \begin{dmath*} \cdb*{dgab.007} \end{dmath*}
   % \begin{dmath*} \cdb*{dgab.008} \end{dmath*}
   % \begin{dmath*} \cdb*{dgab.009} \end{dmath*}
   \begin{dmath*} \cdb*{dgab.010} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{cadabra}
   # Note:
   # Computing Gamma directly by (1/2) iab dgab and *then* truncating to lower order
   # is not optimal. We only want the leading oder terms (to 4th order in x). But the direct
   # calculation would compute *all* terms before the truncation. This does work but it
   # is slower than the following code.
   #
   # The better approach (as adopted in this code) is to extract all of the terms of iab
   # and dgab then construct the leading order terms of Gamma (to fifth order) term by term.

   def get_Rterm (obj,n):

   # I would like to assign different weights to \nabla_{a}, \nabla_{a b}, \nabla_{a b c} etc. but no matter
   # what I do it appears that Cadabra assigns the same weight to all of these regardless of the number of subscripts.
   # It seems that the weight is assigned to the symbol \nabla alone. So I'm forced to use the following substitution trick.

       Q_{a b c d}::Weight(label=numR,value=2).
       Q_{a b c d e}::Weight(label=numR,value=3).
       Q_{a b c d e f}::Weight(label=numR,value=4).
       Q_{a b c d e f g}::Weight(label=numR,value=5).

       tmp := @(obj).

       distribute (tmp)

       substitute (tmp, $\nabla_{e f g}{R_{a b c d}} -> Q_{a b c d e f g}$)
       substitute (tmp, $\nabla_{e f}{R_{a b c d}} -> Q_{a b c d e f}$)
       substitute (tmp, $\nabla_{e}{R_{a b c d}} -> Q_{a b c d e}$)
       substitute (tmp, $R_{a b c d} -> Q_{a b c d}$)

       foo := @(tmp).
       bah = Ex("numR = " + str(n))
       keep_weight (foo, bah)

       substitute (foo, $Q_{a b c d e f g} -> \nabla_{e f g}{R_{a b c d}}$)
       substitute (foo, $Q_{a b c d e f} -> \nabla_{e f}{R_{a b c d}}$)
       substitute (foo, $Q_{a b c d e} -> \nabla_{e}{R_{a b c d}}$)
       substitute (foo, $Q_{a b c d} -> R_{a b c d}$)

       return foo

   # terms of the curvature expansion of dg_{ab}

   dgab00 = get_Rterm (dgab,0)   # cdb(dgab00.105,dgab00)  # zero
   dgab01 = get_Rterm (dgab,1)   # cdb(dgab01.105,dgab01)  # zero
   dgab02 = get_Rterm (dgab,2)   # cdb(dgab02.105,dgab02)
   dgab03 = get_Rterm (dgab,3)   # cdb(dgab03.105,dgab03)
   dgab04 = get_Rterm (dgab,4)   # cdb(dgab04.105,dgab04)
   dgab05 = get_Rterm (dgab,5)   # cdb(dgab05.105,dgab05)

   # Convert free indices on iab from ^{a b} to ^{d c}
   # This ensures we can later build products like @(iab) @(dgab) knowing that the indices are correctly ordered.
   # Without this step we would be using free indices ^{a b} and _{a b c}. Thus the product @(iab) @(dgab) would
   # have just one free index _{c}. This is clearly wrong.

   tmp := @(iab) \delta_{a}^{d} \delta_{b}^{c}.

   distribute     (tmp)
   eliminate_kronecker (tmp)
   sort_product   (tmp)
   rename_dummies (tmp)
   canonicalise   (tmp)

   idc := @(tmp).

   # terms of the curvature expansion of g^{ab}

   idc00 = get_Rterm (idc,0)   # cdb(idc00.105,idc00)
   idc01 = get_Rterm (idc,1)   # cdb(idc01.105,idc01)  # zero
   idc02 = get_Rterm (idc,2)   # cdb(idc02.105,idc02)
   idc03 = get_Rterm (idc,3)   # cdb(idc03.105,idc03)
   idc04 = get_Rterm (idc,4)   # cdb(idc04.105,idc04)
   idc05 = get_Rterm (idc,5)   # cdb(idc05.105,idc05)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{dgab00.105} \end{dmath*}
   \begin{dmath*} \cdb*{dgab01.105} \end{dmath*}
   \begin{dmath*} \cdb*{dgab02.105} \end{dmath*}
   \begin{dmath*} \cdb*{dgab03.105} \end{dmath*}
   \begin{dmath*} \cdb*{dgab04.105} \end{dmath*}
   \begin{dmath*} \cdb*{dgab05.105} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{idc00.105} \end{dmath*}
   \begin{dmath*} \cdb*{idc01.105} \end{dmath*}
   \begin{dmath*} \cdb*{idc02.105} \end{dmath*}
   \begin{dmath*} \cdb*{idc03.105} \end{dmath*}
   \begin{dmath*} \cdb*{idc04.105} \end{dmath*}
   \begin{dmath*} \cdb*{idc05.105} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{cadabra}
   # idc  = g^{d c}
   # dgab = D_{a}{g_{c b}} + D_{b}{g_{a c}} - D_{c}{g_{a b}}

   # terms of the curvature expansion of \Gamma^{d}_{a b}

   # term0 := (1/2)  @(idc00) @(dgab00).
   # term1 := (1/2) (@(idc01) @(dgab00) + @(idc00) @(dgab01)).
   # term2 := (1/2) (@(idc02) @(dgab00) + @(idc01) @(dgab01) + @(idc00) @(dgab02)).
   # term3 := (1/2) (@(idc03) @(dgab00) + @(idc02) @(dgab01) + @(idc01) @(dgab02) + @(idc00) @(dgab03)).
   # term4 := (1/2) (@(idc04) @(dgab00) + @(idc03) @(dgab01) + @(idc02) @(dgab02) + @(idc01) @(dgab03) + @(idc00) @(dgab04)).
   # term5 := (1/2) (@(idc05) @(dgab00) + @(idc04) @(dgab01) + @(idc03) @(dgab02) + @(idc02) @(dgab03) + @(idc01) @(dgab04) + @(idc00) @(dgab05)).

   # simplidied version of the above after noting dgab00 = dgab01 = 0

   term0 := 0.
   term1 := 0.
   term2 := (1/2) (@(idc00) @(dgab02)).
   term3 := (1/2) (@(idc01) @(dgab02) + @(idc00) @(dgab03)).
   term4 := (1/2) (@(idc02) @(dgab02) + @(idc01) @(dgab03) + @(idc00) @(dgab04)).
   term5 := (1/2) (@(idc03) @(dgab02) + @(idc02) @(dgab03) + @(idc01) @(dgab04) + @(idc00) @(dgab05)).

   def tidy_terms (obj):
       substitute     (obj,$x^{a}->AA^{a}$,repeat=True)  # will force AA to the left of all terms
       distribute     (obj)
       sort_product   (obj)
       rename_dummies (obj)
       canonicalise   (obj)
       substitute     (obj,$AA^{a}->x^{a}$,repeat=True)  # replace AA with x
       factor_out     (obj,$x^{a?}$)

       return obj

   term0 = tidy_terms (term0)   # cdb(term0.201,term0)  # zero
   term1 = tidy_terms (term1)   # cdb(term1.201,term1)  # zero
   term2 = tidy_terms (term2)   # cdb(term2.201,term2)
   term3 = tidy_terms (term3)   # cdb(term3.201,term3)
   term4 = tidy_terms (term4)   # cdb(term4.201,term4)
   term5 = tidy_terms (term5)   # cdb(term5.201,term5)

   Gamma := @(term0) + @(term1) + @(term2) + @(term3) + @(term4) + @(term5). # cdb(Gamma.200,Gamma)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{term0.201} \end{dmath*}
   \begin{dmath*} \cdb*{term1.201} \end{dmath*}
   \begin{dmath*} \cdb*{term2.201} \end{dmath*}
   \begin{dmath*} \cdb*{term3.201} \end{dmath*}
   \begin{dmath*} \cdb*{term4.201} \end{dmath*}
   \begin{dmath*} \cdb*{term5.201} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{Gamma.200} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{cadabra}
   cdblib.create ('connection.json')

   cdblib.put ('Gamma',Gamma,'connection.json')

   cdblib.put ('GammaRterm0',term0,'connection.json')
   cdblib.put ('GammaRterm1',term1,'connection.json')
   cdblib.put ('GammaRterm2',term2,'connection.json')
   cdblib.put ('GammaRterm3',term3,'connection.json')
   cdblib.put ('GammaRterm4',term4,'connection.json')
   cdblib.put ('GammaRterm5',term5,'connection.json')

   checkpoint.append (term0)
   checkpoint.append (term1)
   checkpoint.append (term2)
   checkpoint.append (term3)
   checkpoint.append (term4)
   checkpoint.append (term5)

\end{cadabra}

% =================================================================================================
% the remaining code is just for pretty printing

\clearpage

\begin{cadabra}
   # note: keeping numbering as is (out of order) to ensure R appears before \nabla R etc.
   def product_sort (obj):
       substitute (obj,$ A^{a}                            -> A001^{a}               $)
       substitute (obj,$ x^{a}                            -> A002^{a}               $)
       substitute (obj,$ g^{a b}                          -> A003^{a b}             $)
       substitute (obj,$ \nabla_{e f g h}{R_{a b c d}}    -> A008_{a b c d e f g h} $)
       substitute (obj,$ \nabla_{e f g}{R_{a b c d}}      -> A007_{a b c d e f g}   $)
       substitute (obj,$ \nabla_{e f}{R_{a b c d}}        -> A006_{a b c d e f}     $)
       substitute (obj,$ \nabla_{e}{R_{a b c d}}          -> A005_{a b c d e}       $)
       substitute (obj,$ R_{a b c d}                      -> A004_{a b c d}         $)
       sort_product   (obj)
       rename_dummies (obj)
       substitute (obj,$ A001^{a}                  -> A^{a}                         $)
       substitute (obj,$ A002^{a}                  -> x^{a}                         $)
       substitute (obj,$ A003^{a b}                -> g^{a b}                       $)
       substitute (obj,$ A008_{a b c d e f g h}    -> \nabla_{e f g h}{R_{a b c d}} $)
       substitute (obj,$ A007_{a b c d e f g}      -> \nabla_{e f g}{R_{a b c d}}   $)
       substitute (obj,$ A006_{a b c d e f}        -> \nabla_{e f}{R_{a b c d}}     $)
       substitute (obj,$ A005_{a b c d e}          -> \nabla_{e}{R_{a b c d}}       $)
       substitute (obj,$ A004_{a b c d}            -> R_{a b c d}                   $)

       return obj

   def reformat (obj,scale):
      foo  = Ex(str(scale))
      bah := @(foo) @(obj).
      distribute     (bah)
      bah = product_sort (bah)
      rename_dummies (bah)
      canonicalise   (bah)
      factor_out     (bah,$A^{a?},x^{b?}$)
      ans := @(bah) / @(foo).
      return ans

   def rescale (obj,scale):
      foo  = Ex(str(scale))
      bah := @(foo) @(obj).
      distribute  (bah)
      factor_out  (bah,$A^{a?},x^{b?}$)
      return bah

   Rterm2 := @(term2) A^{a} A^{b}.
   Rterm3 := @(term3) A^{a} A^{b}.
   Rterm4 := @(term4) A^{a} A^{b}.
   Rterm5 := @(term5) A^{a} A^{b}.

   Rterm2 = reformat (Rterm2,  3)    # cdb(Rterm2.301,Rterm2)
   Rterm3 = reformat (Rterm3, 12)    # cdb(Rterm3.301,Rterm3)
   Rterm4 = reformat (Rterm4,360)    # cdb(Rterm4.301,Rterm4)
   Rterm5 = reformat (Rterm5,180)    # cdb(Rterm5.301,Rterm5)

   Gamma  := @(Rterm2) + @(Rterm3) + @(Rterm4) + @(Rterm5).  # cdb (Gamma.301,Gamma)
   Scaled := 360 @(Gamma).                                   # cdb (Scaled.301,Scaled)

   scaled2 = rescale (Rterm2,   3)   # cdb (scaled2.301,scaled2)
   scaled3 = rescale (Rterm3,  12)   # cdb (scaled3.301,scaled3)
   scaled4 = rescale (Rterm4, 360)   # cdb (scaled4.301,scaled4)
   scaled5 = rescale (Rterm5, 180)   # cdb (scaled5.301,scaled5)
\end{cadabra}

\clearpage

% =================================================================================================
\section*{The connection in Riemann normal coordinates}

\begin{dgroup*}
   \begin{dmath*} A^a A^b \Gamma^{d}_{a b} = \cdb{Gamma.301} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} 360 A^a A^b \Gamma^{d}_{a b} = \cdb{Scaled.301} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Curvature expansion of the connection}

\begin{align*}
     A^a A^b \Gamma^{d}_{a b} =
     A^a A^b \nGamma{2}^{d}{}_{a b}
   + A^a A^b \nGamma{3}^{d}{}_{a b}
   + A^a A^b \nGamma{4}^{d}{}_{a b}
   + A^a A^b \nGamma{5}^{d}{}_{a b}+\BigO{\eps^6}
\end{align*}
\begin{dgroup*}
   \begin{dmath*}   3 A^a A^b \nGamma{2}^{d}_{a b} = \cdb{scaled2.301} \end{dmath*}
   \begin{dmath*}  12 A^a A^b \nGamma{3}^{d}_{a b} = \cdb{scaled3.301} \end{dmath*}
   \begin{dmath*} 360 A^a A^b \nGamma{4}^{d}_{a b} = \cdb{scaled4.301} \end{dmath*}
   \begin{dmath*} 180 A^a A^b \nGamma{5}^{d}_{a b} = \cdb{scaled5.301} \end{dmath*}
\end{dgroup*}

% =================================================================================================
% export checkpoints in json format

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   for i in range( len(checkpoint) ):
      cdblib.put ('check{:03d}'.format(i),checkpoint[i],checkpoint_file)
\end{cadabra}
\egroup

\end{document}

### 2019-09-24 source/cadabra/dGamma.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

% =================================================================================================
% create checkpoint file

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/dGamma.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{Symmetrized partial derivatives of the connection}

Here we calculate the recursive sequences
\begin{align*}
(n+3)\Gamma^a{}_{d(b,c\uen)}
   &= (n+1)\left(R^a{}_{(bc\Dot d,\uen)}
    - \left(\Gamma^a{}_{f(c}\Gamma^f{}_{b{\Dot d}}\right){\vphantom{\Gamma}}_{,\uen)}\right)
\end{align*}
for $n=1,2,3,\cdots$. Note that the (extended) index $\uen$ contains $n$ normal indices.

The result will be expressions for the $\Gamma^a{}_{d(b,c\uen)}$ in terms of the Riemann tensor and its
partial derivatives.

% =================================================================================================
\section*{Stage 1: Compute symmetrised derivatives}

In the first stage we simply apply the above recursive equation using a simple trick to impose
the symmetries. Start with the original equation and dot out the symmetric indices with $A^a$ then
factor out the partial derivatives. This leads to
\begin{equation}
(n+3)\Gamma^a{}_{db,c\uen} A^b A^c A^{\cdot\uen}
    = (n+1)\left(R^a{}_{bcd}
    - \Gamma^a{}_{fc}\Gamma^f{}_{bd}\right)_{,\uen} A^b A^c A^{\cdot\uen}
\end{equation}
Thus we also have (for the next iteration)
\begin{equation}
(n+4)\Gamma^a{}_{db,c\uenp} A^b A^c A^{\cdot\uenp}
    = (n+2)\left(R^a{}_{bcd}
    - \Gamma^a{}_{fc}\Gamma^f{}_{bd}\right)_{,\uenp} A^b A^c A^{\cdot\uenp}
\end{equation}
The $A^a$ can be freely chosen so choose $A^a$ to be a constant (i.e., zero derivative). Now
define $P_n$ by
\begin{equation}
   P_n = \Gamma^a{}_{db,c\uen} A^b A^c A^{\cdot\uen}
\end{equation}
then the above pair of equations can be combinded to give
\begin{equation}
   P_{n+1} = \frac{(n+2)(n+3)}{(n+4)(n+1)} A^f\partial_f\left(P_n\right)
\end{equation}
This is a very easy equation to compute as it just requires successive rounds of differentiation.

The first term in the sequence is $P_0$ given by
\begin{equation}
   P_0 = \cdb{dGamma01.101}
\end{equation}

The first few results are

\begin{dgroup*}
   \begin{dmath*} P_0 \hiderel{=} A^{b} A^{c} \Gamma^a{}_{d(b,c)} = \cdb{dGamma01.101} \end{dmath*}
   \begin{dmath*} P_1 \hiderel{=} A^{b} A^{c} A^{e} \Gamma^a{}_{d(b,ce)} = \cdb{dGamma02.105} \end{dmath*}
   \begin{dmath*} P_2 \hiderel{=} A^{b} A^{c} A^{e} A^{f} \Gamma^a{}_{d(b,cef)} = \cdb{dGamma03.105} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 2: Impose Riemann normal coordinates}

Here we impose the RNC conition by setting the $\Gamma^{a}{}_{bc}$ to zero (but not their derivatives).

\begin{dgroup*}
   \begin{dmath*} A^{b} A^{c} \Gamma^a{}_{d(b,c)} = \cdb{dGamma01.201} \end{dmath*}
   \begin{dmath*} A^{b} A^{c} A^{e} \Gamma^a{}_{d(b,ce)} = \cdb{dGamma02.202} \end{dmath*}
   \begin{dmath*} A^{b} A^{c} A^{e} A^{f} \Gamma^a{}_{d(b,cef)} = \cdb{dGamma03.203} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 3: Replace partial derivatives $\Gamma$ with partial derivatives $R$}

The key point to note is that the partial derivatives of $\Gamma$ on the right hand side are all
symmetrized in exactly the same manner as the partial derivatives on the left hand side. Thus
results from the lower order equations can be fed into the later equations to completely eliminate
the partial derivatives of $\Gamma$.

\begin{dgroup*}
   \begin{dmath*} A^{b} A^{c} \Gamma^a{}_{d(b,c)} = \cdb{dGamma01.201} \end{dmath*}
   \begin{dmath*} A^{b} A^{c} A^{e} \Gamma^a{}_{d(b,ce)} = \cdb{dGamma02.202} \end{dmath*}
   \begin{dmath*} A^{b} A^{c} A^{e} A^{f} \Gamma^a{}_{d(b,cef)} = \cdb{dGamma03.402} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 4: Reformatting}

This is just simple reformatting.

\begin{dgroup*}
   \begin{dmath*}   3 A^b A^c\Gamma^a{}_{d(b,c)} = \cdb{scaled1.002} \end{dmath*}
   \begin{dmath*}   6 A^b A^c A^e \Gamma^a{}_{d(b,ce)} = \cdb{scaled2.002} \end{dmath*}
   \begin{dmath*}  15 A^b A^c A^e A^f \Gamma^a{}_{d(b,cef)} = \cdb{scaled3.002} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 1: Compute symmetrised derivatives}

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   \nabla{#}::Derivative.
   \partial{#}::PartialDerivative.

   g_{a b}::Metric.
   g^{a b}::InverseMetric.
   g_{a}^{b}::KroneckerDelta.
   g^{a}_{b}::KroneckerDelta.

   R_{a b c d}::RiemannTensor.
   R^{a}_{b c d}::RiemannTensor.
   R_{a b c}^{d}::RiemannTensor.

   \Gamma^{a}_{b c}::TableauSymmetry(shape={2}, indices={1,2}).

   g_{a b}::Depends(\partial{#}).
   R_{a b c d}::Depends(\partial{#}).
   R^{a}_{b c d}::Depends(\partial{#}).
   \Gamma^{a}_{b c}::Depends(\partial{#}).

   # symmetrized partial derivatives of \Gamma

   dGamma01:= (1/3) A^{b} A^{c} ( R^{a}_{b c d} - \Gamma^{a}_{c e}\Gamma^{e}_{b d} ).
                                                        # cdb (dGamma01.101,dGamma01)

   dGamma02:= (6/4) A^{a}\partial_{a}{ @(dGamma01) }.   # cdb (dGamma02.101,dGamma02)
   distribute   (dGamma02)                              # cdb (dGamma02.102,dGamma02)
   product_rule (dGamma02)                              # cdb (dGamma02.103,dGamma02)
   unwrap       (dGamma02)                              # cdb (dGamma02.104,dGamma02)
   distribute   (dGamma02)                              # cdb (dGamma02.105,dGamma02)

   dGamma03:= (12/10) A^{a}\partial_{a}{ @(dGamma02) }. # cdb (dGamma03.101,dGamma03)
   distribute   (dGamma03)                              # cdb (dGamma03.102,dGamma03)
   product_rule (dGamma03)                              # cdb (dGamma03.103,dGamma03)
   unwrap       (dGamma03)                              # cdb (dGamma03.104,dGamma03)
   distribute   (dGamma03)                              # cdb (dGamma03.105,dGamma03)

   dGamma04:= (20/18) A^{a}\partial_{a}{ @(dGamma03) }. # cdb (dGamma04.101,dGamma04)
   distribute   (dGamma04)                              # cdb (dGamma04.102,dGamma04)
   product_rule (dGamma04)                              # cdb (dGamma04.103,dGamma04)
   unwrap       (dGamma04)                              # cdb (dGamma04.104,dGamma04)
   distribute   (dGamma04)                              # cdb (dGamma04.105,dGamma04)

   dGamma05:= (30/28) A^{a}\partial_{a}{ @(dGamma04) }. # cdb (dGamma05.101,dGamma05)
   distribute   (dGamma05)                              # cdb (dGamma05.102,dGamma05)
   product_rule (dGamma05)                              # cdb (dGamma05.103,dGamma05)
   unwrap       (dGamma05)                              # cdb (dGamma05.104,dGamma05)
   distribute   (dGamma05)                              # cdb (dGamma05.105,dGamma05)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma01.101} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma02.101} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma02.102} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma02.103} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma02.104} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma02.105} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma03.101} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.102} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.103} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.104} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.105} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma04.101} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.102} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.103} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.104} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.105} \end{dmath*}
\end{dgroup*}

% too long for pdflatex

% \begin{dgroup*}
%    \begin{dmath*} \cdb*{dGamma05.101} \end{dmath*}
%    \begin{dmath*} \cdb*{dGamma05.102} \end{dmath*}
%    \begin{dmath*} \cdb*{dGamma05.103} \end{dmath*}
%    \begin{dmath*} \cdb*{dGamma05.104} \end{dmath*}
%    \begin{dmath*} \cdb*{dGamma05.105} \end{dmath*}
% \end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 2: Impose Riemann normal coordinates}

\begin{cadabra}
   def impose_rnc (obj):
       # hide the derivatives of Gamma
       substitute (obj,$\partial_{d}{\Gamma^{a}_{b c}} -> zzz_{d}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e}{\Gamma^{a}_{b c}} -> zzz_{d e}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f}{\Gamma^{a}_{b c}} -> zzz_{d e f}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f g}{\Gamma^{a}_{b c}} -> zzz_{d e f g}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f g h}{\Gamma^{a}_{b c}} -> zzz_{d e f g h}^{a}_{b c}$,repeat=True)
       # set Gamma to zero
       substitute (obj,$\Gamma^{a}_{b c} -> 0$,repeat=True)
       # recover the derivatives Gamma
       substitute (obj,$zzz_{d}^{a}_{b c} -> \partial_{d}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e}^{a}_{b c} -> \partial_{d e}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f}^{a}_{b c} -> \partial_{d e f}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f g}^{a}_{b c} -> \partial_{d e f g}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f g h}^{a}_{b c} -> \partial_{d e f g h}{\Gamma^{a}_{b c}}$,repeat=True)
       return obj

   # switch to RNC

   dGamma01 = impose_rnc (dGamma01)   # cdb (dGamma01.201,dGamma01)
   dGamma02 = impose_rnc (dGamma02)   # cdb (dGamma02.202,dGamma02)
   dGamma03 = impose_rnc (dGamma03)   # cdb (dGamma03.203,dGamma03)
   dGamma04 = impose_rnc (dGamma04)   # cdb (dGamma04.204,dGamma04)
   dGamma05 = impose_rnc (dGamma05)   # cdb (dGamma05.205,dGamma05)

\end{cadabra}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma01.201} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma02.202} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.203} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.204} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.205} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 3: Replace partial derivatives $\Gamma$ with partial derivatives $R$}

\begin{cadabra}
   # use lower equations to eliminate partial derivs of Gamma from rhs

   # this produces experssions for the partial derivs of the Gamma's in terms of the Rabcd and its partial derivs

   substitute (dGamma03,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{b d}} -> @(dGamma01)$,repeat=True)                 # cdb(dGamma03.301,dGamma03)
   substitute (dGamma03,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{d b}} -> @(dGamma01)$,repeat=True)                 # cdb(dGamma03.302,dGamma03)
   distribute (dGamma03)                                                                                       # cdb(dGamma03.303,dGamma03)

   substitute (dGamma04,$A^{c}A^{b}A^{e}\partial_{c e}{\Gamma^{a}_{d b}} -> @(dGamma02)$,repeat=True)          # cdb(dGamma04.301,dGamma04)
   substitute (dGamma04,$A^{c}A^{b}A^{e}\partial_{c e}{\Gamma^{a}_{b d}} -> @(dGamma02)$,repeat=True)          # cdb(dGamma04.302,dGamma04)
   substitute (dGamma04,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{b d}} -> @(dGamma01)$,repeat=True)                 # cdb(dGamma04.303,dGamma04)
   substitute (dGamma04,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{d b}} -> @(dGamma01)$,repeat=True)                 # cdb(dGamma04.304,dGamma04)
   distribute (dGamma04)                                                                                       # cdb(dGamma04.305,dGamma04)

   substitute (dGamma05,$A^{c}A^{b}A^{e}A^{f}\partial_{c e f}{\Gamma^{a}_{d b}} -> @(dGamma03)$,repeat=True)   # cdb(dGamma05.301,dGamma05)
   substitute (dGamma05,$A^{c}A^{b}A^{e}A^{f}\partial_{c e f}{\Gamma^{a}_{b d}} -> @(dGamma03)$,repeat=True)   # cdb(dGamma05.302,dGamma05)
   substitute (dGamma05,$A^{c}A^{b}A^{e}\partial_{c e}{\Gamma^{a}_{d b}} -> @(dGamma02)$,repeat=True)          # cdb(dGamma05.303,dGamma05)
   substitute (dGamma05,$A^{c}A^{b}A^{e}\partial_{c e}{\Gamma^{a}_{b d}} -> @(dGamma02)$,repeat=True)          # cdb(dGamma05.304,dGamma05)
   substitute (dGamma05,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{b d}} -> @(dGamma01)$,repeat=True)                 # cdb(dGamma05.305,dGamma05)
   substitute (dGamma05,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{d b}} -> @(dGamma01)$,repeat=True)                 # cdb(dGamma05.306,dGamma05)
   distribute (dGamma05)                                                                                       # cdb(dGamma05.307,dGamma05)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma03.301} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.302} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.303} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma04.301} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.302} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.303} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.304} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.305} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma05.301} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.302} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.303} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.304} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.305} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.306} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.307} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{cadabra}
   # note:
   # canonicalise must not be used here because it may make changes like
   #    R^{a}_{b c d} -> - R_{b}^{a}_{c d}
   # these changes can not be applied inside a \partial, must defer use
   # of canocialise until we have \nabla acting on curvatures

   sort_product   (dGamma03) # cdb(dGamma03.401,dGamma03)
   rename_dummies (dGamma03) # cdb(dGamma03.402,dGamma03)
   # canonicalise   (dGamma03) # cdb(dGamma03.403,dGamma03)

   sort_product   (dGamma04) # cdb(dGamma04.401,dGamma04)
   rename_dummies (dGamma04) # cdb(dGamma04.402,dGamma04)
   # canonicalise   (dGamma04) # cdb(dGamma04.403,dGamma04)

   sort_product   (dGamma05) # cdb(dGamma05.401,dGamma05)
   rename_dummies (dGamma05) # cdb(dGamma05.402,dGamma05)
   # canonicalise   (dGamma05) # cdb(dGamma05.403,dGamma05)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma03.401} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma03.402} \end{dmath*}
   % \begin{dmath*} \cdb*{dGamma03.403} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma04.401} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma04.402} \end{dmath*}
   % \begin{dmath*} \cdb*{dGamma04.403} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dGamma05.401} \end{dmath*}
   \begin{dmath*} \cdb*{dGamma05.402} \end{dmath*}
   % \begin{dmath*} \cdb*{dGamma05.403} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{cadabra}
   import cdblib

   cdblib.create ('dGamma.json')

   cdblib.put ('dGamma01',dGamma01,'dGamma.json')
   cdblib.put ('dGamma02',dGamma02,'dGamma.json')
   cdblib.put ('dGamma03',dGamma03,'dGamma.json')
   cdblib.put ('dGamma04',dGamma04,'dGamma.json')
   cdblib.put ('dGamma05',dGamma05,'dGamma.json')

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Stage 4: Reformatting}

\begin{cadabra}
   # note: keeping numbering as is (out of order) to ensure R appears before \nabla R etc.
   def product_sort (obj):
       substitute (obj,$ A^{a}                              -> A001^{a}                 $)
       substitute (obj,$ x^{a}                              -> A002^{a}                 $)
       substitute (obj,$ g^{a b}                            -> A003^{a b}               $)
       substitute (obj,$ \partial_{e f g h}{R^{a}_{b c d}}  -> A008^{a}_{b c d e f g h} $)
       substitute (obj,$ \partial_{e f g}{R^{a}_{b c d}}    -> A007^{a}_{b c d e f g}   $)
       substitute (obj,$ \partial_{e f}{R^{a}_{b c d}}      -> A006^{a}_{b c d e f}     $)
       substitute (obj,$ \partial_{e}{R^{a}_{b c d}}        -> A005^{a}_{b c d e}       $)
       substitute (obj,$ R^{a}_{b c d}                      -> A004^{a}_{b c d}         $)
       sort_product   (obj)
       rename_dummies (obj)
       substitute (obj,$ A001^{a}                  -> A^{a}                             $)
       substitute (obj,$ A002^{a}                  -> x^{a}                             $)
       substitute (obj,$ A003^{a b}                -> g^{a b}                           $)
       substitute (obj,$ A004^{a}_{b c d}          -> R^{a}_{b c d}                     $)
       substitute (obj,$ A005^{a}_{b c d e}        -> \partial_{e}{R^{a}_{b c d}}       $)
       substitute (obj,$ A006^{a}_{b c d e f}      -> \partial_{e f}{R^{a}_{b c d}}     $)
       substitute (obj,$ A007^{a}_{b c d e f g}    -> \partial_{e f g}{R^{a}_{b c d}}   $)
       substitute (obj,$ A008^{a}_{b c d e f g h}  -> \partial_{e f g h}{R^{a}_{b c d}} $)

       return obj

   def reformat (obj,scale):
       bah  = Ex(str(scale))
       tmp := @(bah) @(obj).
       distribute     (tmp)
       tmp = product_sort (tmp)
       rename_dummies (tmp)
       factor_out     (tmp,$A^{a?}$)
       return tmp

   def get_term (obj,n):

       A^{a}::Weight(label=numA).

       foo := @(obj).
       bah  = Ex("numA = " + str(n))
       distribute  (foo)
       keep_weight (foo, bah)

       return foo

   Gterm01 := @(dGamma01).
   Gterm02 := @(dGamma02).
   Gterm03 := @(dGamma03).
   Gterm04 := @(dGamma04).
   Gterm05 := @(dGamma05).

   scaled1 = reformat (Gterm01,   3)   # cdb (scaled1.002,scaled1)
   scaled2 = reformat (Gterm02,   6)   # cdb (scaled2.002,scaled2)
   scaled3 = reformat (Gterm03,  15)   # cdb (scaled3.002,scaled3)
   scaled4 = reformat (Gterm04,   9)   # cdb (scaled4.002,scaled4)
   scaled5 = reformat (Gterm05, 252)   # cdb (scaled5.002,scaled5)

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Symmetrised partial derivatives of the connection}

\begin{dgroup*}
   \begin{dmath*}   3 A^b A^c\Gamma^a{}_{d(b,c)} = \cdb{scaled1.002} \end{dmath*}
   \begin{dmath*}   6 A^b A^c A^e \Gamma^a{}_{d(b,ce)} = \cdb{scaled2.002} \end{dmath*}
   \begin{dmath*}  15 A^b A^c A^e A^f \Gamma^a{}_{d(b,cef)} = \cdb{scaled3.002} \end{dmath*}
   \begin{dmath*}   9 A^b A^c A^e A^f A^g \Gamma^a{}_{d(b,cefg)} = \cdb{scaled4.002} \end{dmath*}
   \begin{dmath*} 252 A^b A^c A^e A^f A^g A^h\Gamma^a{}_{d(b,cefgh)} = \cdb{scaled5.002} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
% export selected objects, these will later be imported into a library
% these are the objects that will appear in the paper

\begin{cadabra}
   substitute (scaled1,$A^{a}->1$)
   substitute (scaled2,$A^{a}->1$)
   substitute (scaled3,$A^{a}->1$)
   substitute (scaled4,$A^{a}->1$)
   substitute (scaled5,$A^{a}->1$)

   cdblib.create ('dGamma.export')

   # 6th order dGamma, scaled
   cdblib.put ('dGamma61scaled',scaled1,'dGamma.export')
   cdblib.put ('dGamma62scaled',scaled2,'dGamma.export')
   cdblib.put ('dGamma63scaled',scaled3,'dGamma.export')
   cdblib.put ('dGamma64scaled',scaled4,'dGamma.export')
   cdblib.put ('dGamma65scaled',scaled5,'dGamma.export')

   checkpoint.append (scaled1)
   checkpoint.append (scaled2)
   checkpoint.append (scaled3)
   checkpoint.append (scaled4)
   checkpoint.append (scaled5)

\end{cadabra}

% =================================================================================================
% export checkpoints in json format

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   for i in range( len(checkpoint) ):
      cdblib.put ('check{:03d}'.format(i),checkpoint[i],checkpoint_file)
\end{cadabra}
\egroup

\end{document}

### 2019-09-24 source/cadabra/dRabcd.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

% =================================================================================================
% create checkpoint file

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/dRabcd.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{Symmetrised partial derivatives of the Riemann tensor}

Here we compute the symmetrised partial derivatoves $R^{a}{}_{(b{\Dot c}d,\ue)}$ in terms of
the symmetrised covariant derivatives $R^{a}{}_{(b{\Dot c}d;\ue)}$. Note that the dot over an
index indicates that that index does not take part in the symmetrisation.

We will use the algorithm described in section (10.3) of my lcb09-03 paper. Here we will make
one small change of notation -- the symbol $D^a$ will replaced with $A^a$.

We have lots of space (and no annoying editors to appease with brevity) so I will take the
liberty to expand slightly on what I wrote in the lcb0-03 paper.

Our starting point is the simple identity
\begin{align}
   \label{eqn:submain}
   \left( R^{a}{}_{cdb} B^{b}{}_{a} A^{c} A^{d} \right)_{;e} A^{e}
   =
   \left( R^{a}{}_{cdb} B^{b}{}_{a} A^{c} A^{d} \right)_{,e} A^{e}
\end{align}
This is true in all frames since the quantity inside the brackets is a scalar. We are free to
make any choice we like for $A^{a}$ and $B^{a}{}_{b}$ so let's choose $A^{a}$ to be the tangent
vector to any geodesic through the origin and choose the $B^{a}{}_{b}$ to be constants (i.e,
all partial derivatives are zero). We will also use local Riemann normal coordinates and as a
consequence, the $A^{a}$ will also be constant along the integral curves of $A$ (the geodesics
in an RNC are always of the form $x^{a}(s) = s A^{a}$ for some affine parameter $s$ on the
geodesic). Let $df/ds$ be the directional derivative of the function $f$ along the geodesics
defined by $A^{a}$ and assume that $s$ is the proper length along the geodesic (although any
affine parameter would be sufficient).

Thus at the origin we have, by choice,
\begin{gather*}
   0 = B^{a}{}_{b,c} = B^{a}{}_{b,cd} = B^{a}{}_{b,cde} = \dots\\[5pt]
   0 = dA^{a}/ds = d^2A^{a}/ds^2 = d^3A^{a}/ds^3 = \dots\\[5pt]
   0 = A^{a}{}_{,b} A^{b} = \left(A^{a}{}_{,b} A^{b}\right)_{,c} A^{c} = \left(\left(A^{a}{}_{,b} A^{b}\right)_{,c} A^{c}\right)_{,d} A^{d}\\[5pt]
   0 = A^{a}{}_{;b} A^{b} = \left(A^{a}{}_{;b} A^{b}\right)_{;c} A^{c} = \left(\left(A^{a}{}_{;b} A^{b}\right)_{;c} A^{c}\right)_{;d} A^{d}\\[5pt]
   df/ds = f_{,a} A^{a} = f_{;a} A^{a}\\[5pt]
   d^2f/ds^2 = \left ( f_{,a} A^{a} \right)_{,b} A^{b} = \left ( f_{;a} A^{a} \right)_{;b} A^{b}\\[5pt]
   d^3f/ds^3 = \left(\left ( f_{,a} A^{a} \right)_{,b} A^{b}\right)_{,c} A^{c} = \left(\left ( f_{;a} A^{a} \right)_{;b} A^{b}\right)_{;c} A^{c}
\end{gather*}
I admit I've gone overboard here in writing out more than I need to but it's handy to have all
of these equations laid bare in one convenient place.

Now put $f = R^{p}{}_{abq} B^{q}{}_{p} A^{a} A^{b}$. Then upon taking successive derivatives,
while taking full advantage of the asummptions just noted, we can eaily see that
\begin{align}
   \label{eqn:main}
   \left( R^{a}{}_{cdb} B^{b}{}_{a} \right)_{;\ue} A^{c} A^{d} A^{\ue}
   =
   \left( R^{a}{}_{cdb} \right)_{,\ue} B^{b}{}_{a} A^{c} A^{d} A^{\ue}
\end{align}
This is the equation that will be computed by the following Cadabra code. All of the
compuations will be carried out on the left hand side (in the first version of the paper I
swapped the left and righ hand sides).

We will need the successive covariant derivatives of $B$. The first covariant derivative is
just \begin{align*} B^{a}{}_{b;c} A^c &= \Gamma^{a}{}_{dc}B^{d}{}_{b} A^c -
\Gamma^{d}{}_{bc}B^{a}{}_{d} A^c \end{align*} The quantities on the left hand side are the
components of a tensor so further covariant derivatives of the right hand side can be computed
(despite the presence of the $\Gamma$'s) by application of the usual rule for a covariant
derivative of a mixed tensor.

% =================================================================================================
\section*{Stage 1: Symmetrised partial derivatives of $R$}

The first stage involves the expansion of the left side of (\ref{eqn:main}). This leads to
expressions for the symmetrized partial derivatives of $R_{abcd}$ in terms of the symmetrized
covariant derivatives of $R_{abcd}$ and $B^{a}{}_{b}$.

\begin{dgroup*}
   \begin{dmath*} \left( R^{a}{}_{cdb} \right)_{,e} B^{b}{}_{a} A^{c} A^{d} A^{e}
                  = \cdb{dRabcd01.108} \end{dmath*}
   \begin{dmath*} \left( R^{a}{}_{cdb} \right)_{,ef} B^{b}{}_{a} A^{c} A^{d} A^{e} A^{f}
                  = \cdb{dRabcd02.108} \end{dmath*}
   \begin{dmath*} \left( R^{a}{}_{cdb} \right)_{,efg} B^{b}{}_{a} A^{c} A^{d} A^{e} A^{f} A^{g}
                  = \cdb{dRabcd03.108} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 2: Symmetrised covariant derivatives of $B$}

In this stage the symmetrized covariant derivatives of $B^{a}{}_{b}$ are computed in terms of
its partial derivatives (which by choice are all zero) and the connection and its partial
derivatives (which in general are not zero).

\begin{dgroup*}
   \begin{dmath*} A^{c}\nabla_{c}\left(B^{a}{}_{b}\right)
                  = \cdb{dBab01.209} \end{dmath*}
   \begin{dmath*} A^{d}A^{c}\nabla_{d}\left(\nabla_{c}\left(B^{a}{}_{b}\right)\right)
                  = \cdb{dBab02.209} \end{dmath*}
   \begin{dmath*} A^{e}A^{d}A^{c}\nabla_{e}\left(\nabla_{d}\left(\nabla_{c}\left(B^{a}{}_{b}\right)\right)\right)
                  = \cdb{dBab03.209} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 3: Impose the Riemann normal coordinate condition on covariant derivs of $B$}

Here we impose the RNC condition (that $\Gamma = 0$ while $\partial\Gamma\not=0$).

\begin{dgroup*}
   \begin{dmath*} A^{c}\nabla_{c}\left(B^{a}{}_{b}\right)
                  = \cdb{dBab01.301} \end{dmath*}
   \begin{dmath*} A^{d}A^{c}\nabla_{d}\left(\nabla_{c}\left(B^{a}{}_{b}\right)\right)
                  = \cdb{dBab02.301} \end{dmath*}
   \begin{dmath*} A^{e}A^{d}A^{c}\nabla_{e}\left(\nabla_{d}\left(\nabla_{c}\left(B^{a}{}_{b}\right)\right)\right)
                  = \cdb{dBab03.301} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 4: Replace covariant derivs of $B$ with partial derivs of $\Gamma$}

This stage uses the results from the second stage to eliminate the $\nabla B$ terms from the
results of the first stage. This produces expressions for the symmetrized partial derivatives
of $R_{abcd}$ in terms of the symmetrized covariant derivatives of $R_{abcd}$ and the partial
derivatives of the connection. In this stage we also set the $B^{a}{}_{b}$ to equal 1.

\begin{dgroup*}
   \begin{dmath*} \left( R^{a}{}_{cdb} \right)_{,e} A^{c} A^{d} A^{e}
                  = \cdb{dRabcd01.401} \end{dmath*}
   \begin{dmath*} \left( R^{a}{}_{cdb} \right)_{,ef} A^{c} A^{d} A^{e} A^{e}
                  = \cdb{dRabcd02.401} \end{dmath*}
   \begin{dmath*} \left( R^{a}{}_{cdb} \right)_{,efg} A^{c} A^{d} A^{e} A^{f} A^{g}
                  = \cdb{dRabcd03.401} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 5: Replace partial derivs of $\Gamma$ with partial derivs of $R$}

The fifth stage draws in results from {\tt dGamma.tex} to replace the partial derivatives of
$\Gamma$ with partial derivatives of $R_{abcd}$.

\begin{dgroup*}
   \begin{dmath*} \left( R^{a}{}_{cdb} \right)_{,e} A^{c} A^{d} A^{e}
                  = \cdb{dRabcd01.500} \end{dmath*}
   \begin{dmath*} \left( R^{a}{}_{cdb} \right)_{,ef} A^{c} A^{d} A^{e} A^{f}
                  = \cdb{dRabcd02.505} \end{dmath*}
   \begin{dmath*} \left( R^{a}{}_{cdb} \right)_{,efg} A^{c} A^{d} A^{e} A^{f} A^{g}
                  = \cdb{dRabcd03.507} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 6: Replace partial derivs of $R$ with covariant derivs of $R$}

The final stage is to eliminate the $\partial R$ by using earlier results. For example, in the
equation for $\partial^3 R$ we see terms involving $\partial R$. These first order partial
derivatives can be replaced with the expression previously computed for $\partial R$ in terms
of $\nabla R$.

\begin{dgroup*}
   \begin{dmath*} \left( R^{a}{}_{cdb} \right)_{,e} A^{c} A^{d} A^{e}
                  = \cdb{dRabcd01.702} \end{dmath*}
   \begin{dmath*} \left( R^{a}{}_{cdb} \right)_{,ef} A^{c} A^{d} A^{e} A^{e}
                  = \cdb{dRabcd02.702} \end{dmath*}
   \begin{dmath*} \left( R^{a}{}_{cdb} \right)_{,efg} A^{c} A^{d} A^{e} A^{f} A^{g}
                  = \cdb{dRabcd03.702} \end{dmath*}
\end{dgroup*}

The end result are expressions for the symmetrized partial derivatives of $R_{abcd}$ solely in
terms of the symmetrized covariant derivatives of $R_{abcd}$.

\clearpage

% =================================================================================================
\section*{Shared properties}

\begin{cadabra}
   import time

   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   \nabla{#}::Derivative.
   \partial{#}::PartialDerivative.

   g_{a b}::Metric.
   g^{a b}::InverseMetric.
   g_{a}^{b}::KroneckerDelta.
   g^{a}_{b}::KroneckerDelta.

   R_{a b c d}::RiemannTensor.
   R^{a}_{b c d}::RiemannTensor.

   \Gamma^{a}_{b c}::TableauSymmetry(shape={2}, indices={1,2}).

   g_{a b}::Depends(\partial{#}).
   R_{a b c d}::Depends(\partial{#}).
   R^{a}_{b c d}::Depends(\partial{#}).
   \Gamma^{a}_{b c}::Depends(\partial{#}).

   B^{a}_{b::Depends(\nabla{#}).
   R_{a b c d}::Depends(\nabla{#}).
   R^{a}_{b c d}::Depends(\nabla{#}).

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Stage 1: Symmetrised partial derivatives of $R$}

\begin{cadabra}
   def flatten_Rabcd (obj):
       substitute (obj,$R^{a}_{b c d}   -> g^{a e} R_{e b c d}$)
       substitute (obj,$R_{a}^{b}_{c d} -> g^{b e} R_{a e c d}$)
       substitute (obj,$R_{a b}^{c}_{b} -> g^{c e} R_{a b e d}$)
       substitute (obj,$R_{a b c}^{d}   -> g^{d e} R_{a b c e}$)
       unwrap     (obj)
       sort_product   (obj)
       rename_dummies (obj)
       return obj

   # compute the symmetric covariant derivatives of R^{a}_{bcd} B^{d}_{a}

   beg_stage_1 = time.time()

   dRabcd00:=R^{a}_{b c d} B^{d}_{a} A^{b} A^{c}.        # cdb(dRabcd00.101,dRabcd00)

   dRabcd01:=A^{a}\nabla_{a}{ @(dRabcd00) }.             # cdb(dRabcd01.101,dRabcd01)
   distribute     (dRabcd01)                             # cdb(dRabcd01.102,dRabcd01)
   product_rule   (dRabcd01)                             # cdb(dRabcd01.103,dRabcd01)
   distribute     (dRabcd01)                             # cdb(dRabcd01.104,dRabcd01)
   substitute     (dRabcd01,$\nabla_{a}{A^{b}} -> 0$)    # cdb(dRabcd01.105,dRabcd01)
   substitute     (dRabcd01,$\nabla_{a}{g^{b c}} -> 0$)  # cdb(dRabcd01.106,dRabcd01)

   sort_product   (dRabcd01)
   rename_dummies (dRabcd01)
   canonicalise   (dRabcd01)                             # cdb(dRabcd01.107,dRabcd01)
   dRabcd01 = flatten_Rabcd (dRabcd01)                   # cdb(dRabcd01.108,dRabcd01)

   dRabcd02:=A^{a}\nabla_{a}{ @(dRabcd01) }.             # cdb(dRabcd02.101,dRabcd02)
   distribute     (dRabcd02)                             # cdb(dRabcd02.102,dRabcd02)
   product_rule   (dRabcd02)                             # cdb(dRabcd02.103,dRabcd02)
   distribute     (dRabcd02)                             # cdb(dRabcd02.104,dRabcd02)
   substitute     (dRabcd02,$\nabla_{a}{A^{b}} -> 0$)    # cdb(dRabcd02.105,dRabcd02)
   substitute     (dRabcd02,$\nabla_{a}{g^{b c}} -> 0$)  # cdb(dRabcd02.106,dRabcd02)

   sort_product   (dRabcd02)
   rename_dummies (dRabcd02)
   canonicalise   (dRabcd02)                             # cdb(dRabcd02.107,dRabcd02)
   dRabcd02 = flatten_Rabcd (dRabcd02)                   # cdb(dRabcd02.108,dRabcd02)

   dRabcd03:=A^{a}\nabla_{a}{ @(dRabcd02) }.             # cdb(dRabcd03.101,dRabcd03)
   distribute     (dRabcd03)                             # cdb(dRabcd03.102,dRabcd03)
   product_rule   (dRabcd03)                             # cdb(dRabcd03.103,dRabcd03)
   distribute     (dRabcd03)                             # cdb(dRabcd03.104,dRabcd03)
   substitute     (dRabcd03,$\nabla_{a}{A^{b}} -> 0$)    # cdb(dRabcd03.105,dRabcd03)
   substitute     (dRabcd03,$\nabla_{a}{g^{b c}} -> 0$)  # cdb(dRabcd03.106,dRabcd03)

   sort_product   (dRabcd03)
   rename_dummies (dRabcd03)
   canonicalise   (dRabcd03)                             # cdb(dRabcd03.107,dRabcd03)
   dRabcd03 = flatten_Rabcd (dRabcd03)                   # cdb(dRabcd03.108,dRabcd03)

   dRabcd04:=A^{a}\nabla_{a}{ @(dRabcd03) }.
   distribute     (dRabcd04)
   product_rule   (dRabcd04)
   distribute     (dRabcd04)
   substitute     (dRabcd04,$\nabla_{a}{A^{b}} -> 0$)
   substitute     (dRabcd04,$\nabla_{a}{g^{b c}} -> 0$)

   sort_product   (dRabcd04)
   rename_dummies (dRabcd04)
   canonicalise   (dRabcd04)
   dRabcd04 = flatten_Rabcd (dRabcd04)

   dRabcd05:=A^{a}\nabla_{a}{ @(dRabcd04) }.
   distribute     (dRabcd05)
   product_rule   (dRabcd05)
   distribute     (dRabcd05)
   substitute     (dRabcd05,$\nabla_{a}{A^{b}} -> 0$)
   substitute     (dRabcd05,$\nabla_{a}{g^{b c}} -> 0$)

   sort_product   (dRabcd05)
   rename_dummies (dRabcd05)
   canonicalise   (dRabcd05)
   dRabcd05 = flatten_Rabcd (dRabcd05)

   def combine_nabla (obj):
       substitute (obj,$\nabla_{p}{\nabla_{q}{\nabla_{r}{\nabla_{s}{\nabla_{t}{A??}}}}}->\nabla_{p q r s t}{A??}$,repeat=True)
       substitute (obj,$\nabla_{p}{\nabla_{q}{\nabla_{r}{\nabla_{s}{A??}}}}->\nabla_{p q r s}{A??}$,repeat=True)
       substitute (obj,$\nabla_{p}{\nabla_{q}{\nabla_{r}{A??}}}->\nabla_{p q r}{A??}$,repeat=True)
       substitute (obj,$\nabla_{p}{\nabla_{q}{A??}}->\nabla_{p q}{A??}$,repeat=True)
       return obj

   dRabcd01 = combine_nabla (dRabcd01)
   dRabcd02 = combine_nabla (dRabcd02)
   dRabcd03 = combine_nabla (dRabcd03)
   dRabcd04 = combine_nabla (dRabcd04)
   dRabcd05 = combine_nabla (dRabcd05)

   end_stage_1 = time.time()
\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{dRabcd00.101} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dRabcd01.101} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd01.102} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd01.103} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd01.104} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd01.105} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd01.106} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd01.107} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd01.108} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dRabcd02.101} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.102} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.103} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.104} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.105} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.106} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.107} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.108} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dRabcd03.101} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.102} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.103} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.104} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.105} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.106} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.107} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.108} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 2: Symmetrised covariant derivatives of $B$}

\begin{cadabra}
   # compute the covariant derivatives of B^{a}_{b}, note B^{a}_{b,c} is zero, by choice
   # this method of computing covariant derivatives does not use auxillary fields

   beg_stage_2 = time.time()

   dBab00:=B^{a}_{b}.      # cdb(dBab00.201,dBab00)

   dBab01:=A^{c}\partial_{c}{ @(dBab00) } + \Gamma^{a}_{p q} W^{p}_{b} A^{q}
                                          - \Gamma^{p}_{b q} W^{a}_{p} A^{q}.
                                                         # cdb(dBab01.201,dBab01)
   distribute   (dBab01)                                 # cdb(dBab01.202,dBab01)
   product_rule (dBab01)                                 # cdb(dBab01.203,dBab01)
   distribute   (dBab01)                                 # cdb(dBab01.204,dBab01)
   substitute   (dBab01,$\partial_{a}{A^{b}} -> 0$)      # cdb(dBab01.205,dBab01)
   substitute   (dBab01,$\partial_{a}{B^{b}_{c}} -> 0$)  # cdb(dBab01.206,dBab01)
   substitute   (dBab01,$W^{a}_{b} -> @(dBab00)$)        # cdb(dBab01.207,dBab01)
   distribute   (dBab01)                                 # cdb(dBab01.208,dBab01)
   canonicalise (dBab01)                                 # cdb(dBab01.209,dBab01)

   dBab02:=A^{c}\partial_{c}{ @(dBab01) } + \Gamma^{a}_{p q} W^{p}_{b} A^{q}
                                          - \Gamma^{p}_{b q} W^{a}_{p} A^{q}.
                                                         # cdb(dBab02.201,dBab02)
   distribute   (dBab02)                                 # cdb(dBab02.202,dBab02)
   product_rule (dBab02)                                 # cdb(dBab02.203,dBab02)
   distribute   (dBab02)                                 # cdb(dBab02.204,dBab02)
   substitute   (dBab02,$\partial_{a}{A^{b}} -> 0$)      # cdb(dBab02.205,dBab02)
   substitute   (dBab02,$\partial_{a}{B^{b}_{c}} -> 0$)  # cdb(dBab02.206,dBab02)
   substitute   (dBab02,$W^{a}_{b} -> @(dBab01)$)        # cdb(dBab02.207,dBab02)
   distribute   (dBab02)                                 # cdb(dBab02.208,dBab02)
   canonicalise (dBab02)                                 # cdb(dBab02.209,dBab02)

   dBab03:=A^{c}\partial_{c}{ @(dBab02) } + \Gamma^{a}_{p q} W^{p}_{b} A^{q}
                                          - \Gamma^{p}_{b q} W^{a}_{p} A^{q}.
                                                         # cdb(dBab03.201,dBab03)
   distribute   (dBab03)                                 # cdb(dBab03.202,dBab03)
   product_rule (dBab03)                                 # cdb(dBab03.203,dBab03)
   distribute   (dBab03)                                 # cdb(dBab03.204,dBab03)
   substitute   (dBab03,$\partial_{a}{A^{b}} -> 0$)      # cdb(dBab03.205,dBab03)
   substitute   (dBab03,$\partial_{a}{B^{b}_{c}} -> 0$)  # cdb(dBab03.206,dBab03)
   substitute   (dBab03,$W^{a}_{b} -> @(dBab02)$)        # cdb(dBab03.207,dBab03)
   distribute   (dBab03)                                 # cdb(dBab03.208,dBab03)
   canonicalise (dBab03)                                 # cdb(dBab03.209,dBab03)

   dBab04:=A^{c}\partial_{c}{ @(dBab03) } + \Gamma^{a}_{p q} W^{p}_{b} A^{q}
                                          - \Gamma^{p}_{b q} W^{a}_{p} A^{q}.
   distribute   (dBab04)
   product_rule (dBab04)
   distribute   (dBab04)
   substitute   (dBab04,$\partial_{a}{A^{b}} -> 0$)
   substitute   (dBab04,$\partial_{a}{B^{b}_{c}} -> 0$)
   substitute   (dBab04,$W^{a}_{b} -> @(dBab03)$)
   distribute   (dBab04)
   canonicalise (dBab04)

   dBab05:=A^{c}\partial_{c}{ @(dBab04) } + \Gamma^{a}_{p q} W^{p}_{b} A^{q}
                                          - \Gamma^{p}_{b q} W^{a}_{p} A^{q}.
   distribute   (dBab05)
   product_rule (dBab05)
   distribute   (dBab05)
   substitute   (dBab05,$\partial_{a}{A^{b}} -> 0$)
   substitute   (dBab05,$\partial_{a}{B^{b}_{c}} -> 0$)
   substitute   (dBab05,$W^{a}_{b} -> @(dBab04)$)
   distribute   (dBab05)
   canonicalise (dBab05)

   end_stage_2 = time.time()
\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{dBab00.201} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dBab01.201} \end{dmath*}
   \begin{dmath*} \cdb*{dBab01.202} \end{dmath*}
   \begin{dmath*} \cdb*{dBab01.203} \end{dmath*}
   \begin{dmath*} \cdb*{dBab01.204} \end{dmath*}
   \begin{dmath*} \cdb*{dBab01.205} \end{dmath*}
   \begin{dmath*} \cdb*{dBab01.206} \end{dmath*}
   \begin{dmath*} \cdb*{dBab01.207} \end{dmath*}
   \begin{dmath*} \cdb*{dBab01.208} \end{dmath*}
   \begin{dmath*} \cdb*{dBab01.209} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dBab02.201} \end{dmath*}
   \begin{dmath*} \cdb*{dBab02.202} \end{dmath*}
   \begin{dmath*} \cdb*{dBab02.203} \end{dmath*}
   \begin{dmath*} \cdb*{dBab02.204} \end{dmath*}
   \begin{dmath*} \cdb*{dBab02.205} \end{dmath*}
   \begin{dmath*} \cdb*{dBab02.206} \end{dmath*}
   \begin{dmath*} \cdb*{dBab02.207} \end{dmath*}
   \begin{dmath*} \cdb*{dBab02.208} \end{dmath*}
   \begin{dmath*} \cdb*{dBab02.209} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dBab03.201} \end{dmath*}
   \begin{dmath*} \cdb*{dBab03.202} \end{dmath*}
   \begin{dmath*} \cdb*{dBab03.203} \end{dmath*}
   \begin{dmath*} \cdb*{dBab03.204} \end{dmath*}
   \begin{dmath*} \cdb*{dBab03.205} \end{dmath*}
   \begin{dmath*} \cdb*{dBab03.206} \end{dmath*}
   \begin{dmath*} \cdb*{dBab03.207} \end{dmath*}
   \begin{dmath*} \cdb*{dBab03.208} \end{dmath*}
   \begin{dmath*} \cdb*{dBab03.209} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 3: Impose the Riemann normal coordinate condition on covariant derivs of $B$}

\begin{cadabra}
   def impose_rnc (obj):
       # hide the derivatives of Gamma
       substitute (obj,$\partial_{d}{\Gamma^{a}_{b c}} -> zzz_{d}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e}{\Gamma^{a}_{b c}} -> zzz_{d e}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f}{\Gamma^{a}_{b c}} -> zzz_{d e f}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f g}{\Gamma^{a}_{b c}} -> zzz_{d e f g}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f g h}{\Gamma^{a}_{b c}} -> zzz_{d e f g h}^{a}_{b c}$,repeat=True)
       # set Gamma to zero
       substitute (obj,$\Gamma^{a}_{b c} -> 0$,repeat=True)
       # recover the derivatives Gamma
       substitute (obj,$zzz_{d}^{a}_{b c} -> \partial_{d}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e}^{a}_{b c} -> \partial_{d e}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f}^{a}_{b c} -> \partial_{d e f}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f g}^{a}_{b c} -> \partial_{d e f g}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f g h}^{a}_{b c} -> \partial_{d e f g h}{\Gamma^{a}_{b c}}$,repeat=True)
       return obj

   # switch to RNC

   beg_stage_3 = time.time()

   dBab01 = impose_rnc (dBab01)   # cdb (dBab01.301,dBab01)
   dBab02 = impose_rnc (dBab02)   # cdb (dBab02.301,dBab02)
   dBab03 = impose_rnc (dBab03)   # cdb (dBab03.301,dBab03)
   dBab04 = impose_rnc (dBab04)   # cdb (dBab04.301,dBab04)
   dBab05 = impose_rnc (dBab05)   # cdb (dBab05.301,dBab05)

   end_stage_3 = time.time()
\end{cadabra}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dBab01.301} \end{dmath*}
   \begin{dmath*} \cdb*{dBab02.301} \end{dmath*}
   \begin{dmath*} \cdb*{dBab03.301} \end{dmath*}
   \begin{dmath*} \cdb*{dBab04.301} \end{dmath*}
   \begin{dmath*} \cdb*{dBab05.301} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 4: Replace covariant derivs of $B$ with partial derivs of $\Gamma$}

\begin{cadabra}
   # substitute covariant derivs of B^{a}_{b} into covariant derivs of R^{a}_{bcd}B^{d}_{a}
   # this produces expressions for the partial derivs of Rabcd its covariant derivs and partial derivs of Gamma
   # the partial derivs of Gamma will be eliminted later by using results imported from dGamma.json

   beg_stage_4 = time.time()

   substitute (dRabcd01,$A^{c}\nabla_{c}{B^{a}_{b}} -> @(dBab01)$,repeat=True);   distribute (dRabcd01)
   substitute (dRabcd02,$A^{c}\nabla_{c}{B^{a}_{b}} -> @(dBab01)$,repeat=True);   distribute (dRabcd02)
   substitute (dRabcd03,$A^{c}\nabla_{c}{B^{a}_{b}} -> @(dBab01)$,repeat=True);   distribute (dRabcd03)
   substitute (dRabcd04,$A^{c}\nabla_{c}{B^{a}_{b}} -> @(dBab01)$,repeat=True);   distribute (dRabcd04)
   substitute (dRabcd05,$A^{c}\nabla_{c}{B^{a}_{b}} -> @(dBab01)$,repeat=True);   distribute (dRabcd05)

   substitute (dRabcd02,$A^{c}A^{d}\nabla_{c d}{B^{a}_{b}} -> @(dBab02)$,repeat=True);   distribute (dRabcd02)
   substitute (dRabcd03,$A^{c}A^{d}\nabla_{c d}{B^{a}_{b}} -> @(dBab02)$,repeat=True);   distribute (dRabcd03)
   substitute (dRabcd04,$A^{c}A^{d}\nabla_{c d}{B^{a}_{b}} -> @(dBab02)$,repeat=True);   distribute (dRabcd04)
   substitute (dRabcd05,$A^{c}A^{d}\nabla_{c d}{B^{a}_{b}} -> @(dBab02)$,repeat=True);   distribute (dRabcd05)

   substitute (dRabcd03,$A^{c}A^{d}A^{e}\nabla_{c d e}{B^{a}_{b}} -> @(dBab03)$,repeat=True);   distribute (dRabcd03)
   substitute (dRabcd04,$A^{c}A^{d}A^{e}\nabla_{c d e}{B^{a}_{b}} -> @(dBab03)$,repeat=True);   distribute (dRabcd04)
   substitute (dRabcd05,$A^{c}A^{d}A^{e}\nabla_{c d e}{B^{a}_{b}} -> @(dBab03)$,repeat=True);   distribute (dRabcd05)

   substitute (dRabcd04,$A^{c}A^{d}A^{e}A^{f}\nabla_{c d e f}{B^{a}_{b}} -> @(dBab04)$,repeat=True); distribute (dRabcd04)
   substitute (dRabcd05,$A^{c}A^{d}A^{e}A^{f}\nabla_{c d e f}{B^{a}_{b}} -> @(dBab04)$,repeat=True); distribute (dRabcd05)

   substitute (dRabcd05,$A^{c}A^{d}A^{e}A^{f}A^{g}\nabla_{c d e f g}{B^{a}_{b}} -> @(dBab05)$,repeat=True); distribute (dRabcd05)

   # no longer need B, so let's get rid of it

   # two subtle tricks are used here
   # 1) rename A and B as A002 and A001 before sort_product,
   #    this ensures B will be to left of A after the sort
   # 2) indices on B changed from B^{a}_{b} to B_{b}^{a},
   #    this ensures that after factor_out B will have dummy indices B_{a}^{b}

   def remove_Bab (obj):
       foo := @(obj).
       substitute     (foo,$A^{a}->A002^{a},B^{a}_{b}->A001_{b}^{a}$)  # need this to sort B to the left of A
       sort_product   (foo)
       rename_dummies (foo)
       factor_out     (foo,$A001^{a?}_{b?},A002^{c?}$)
       substitute     (foo,$A001_{a}^{b}->1,A002^{a}->A^{a}$)  # recover A and set B = 1, free indices now ^{a}_{b}
       return foo

   dRabcd01 = remove_Bab (dRabcd01)   # cdb(dRabcd01.401,dRabcd01)
   dRabcd02 = remove_Bab (dRabcd02)   # cdb(dRabcd02.401,dRabcd02)
   dRabcd03 = remove_Bab (dRabcd03)   # cdb(dRabcd03.401,dRabcd03)
   dRabcd04 = remove_Bab (dRabcd04)   # cdb(dRabcd04.401,dRabcd04)
   dRabcd05 = remove_Bab (dRabcd05)   # cdb(dRabcd05.401,dRabcd05)

   end_stage_4 = time.time()
\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{dRabcd01.401} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.401} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.401} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd04.401} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd05.401} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 5: Replace partial derivs of $\Gamma$ with partial derivs of $R$}

\begin{cadabra}
   import cdblib

   beg_stage_5 = time.time()

   dGamma01 = cdblib.get ('dGamma01','dGamma.json')  # cdb(dGamma01.500,dGamma01)
   dGamma02 = cdblib.get ('dGamma02','dGamma.json')  # cdb(dGamma02.500,dGamma02)
   dGamma03 = cdblib.get ('dGamma03','dGamma.json')  # cdb(dGamma03.500,dGamma03)
   dGamma04 = cdblib.get ('dGamma04','dGamma.json')  # cdb(dGamma04.500,dGamma04)
   dGamma05 = cdblib.get ('dGamma05','dGamma.json')  # cdb(dGamma05.500,dGamma05)

   distribute (dRabcd01)   # cdb(dRabcd01.500,dRabcd01)
   distribute (dRabcd02)   # cdb(dRabcd02.500,dRabcd02)
   distribute (dRabcd03)   # cdb(dRabcd03.500,dRabcd03)
   distribute (dRabcd04)   # cdb(dRabcd04.500,dRabcd04)
   distribute (dRabcd05)   # cdb(dRabcd05.500,dRabcd05)

   # use dGamma to eliminate the partial derivs of Gamma
   # this will introduces some lower order partial dervis of Rabcd on the rhs
   # these extra partial derivs of Rabcd will be eliminated (later) by substiting lower order dRabcd into the higher order dRabcd

   substitute (dRabcd02,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{d b}} -> @(dGamma01)$,repeat=True)                # cdb(dRabcd02.501,dRabcd02)
   substitute (dRabcd02,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{b d}} -> @(dGamma01)$,repeat=True)                # cdb(dRabcd02.502,dRabcd02)
   distribute (dRabcd02)                                                                                      # cdb(dRabcd02.503,dRabcd02)
   sort_product   (dRabcd02)                                                                                  # cdb(dRabcd02.504,dRabcd02)
   rename_dummies (dRabcd02)                                                                                  # cdb(dRabcd02.505,dRabcd02)

   substitute (dRabcd03,$A^{c}A^{b}A^{e}\partial_{c e}{\Gamma^{a}_{d b}} -> @(dGamma02)$,repeat=True)         # cdb(dRabcd03.501,dRabcd03)
   substitute (dRabcd03,$A^{c}A^{b}A^{e}\partial_{c e}{\Gamma^{a}_{b d}} -> @(dGamma02)$,repeat=True)         # cdb(dRabcd03.502,dRabcd03)
   substitute (dRabcd03,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{d b}} -> @(dGamma01)$,repeat=True)                # cdb(dRabcd03.503,dRabcd03)
   substitute (dRabcd03,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{b d}} -> @(dGamma01)$,repeat=True)                # cdb(dRabcd03.504,dRabcd03)
   distribute (dRabcd03)                                                                                      # cdb(dRabcd03.505,dRabcd03)
   sort_product   (dRabcd03)                                                                                  # cdb(dRabcd03.506,dRabcd03)
   rename_dummies (dRabcd03)                                                                                  # cdb(dRabcd03.507,dRabcd03)

   substitute (dRabcd04,$A^{c}A^{b}A^{e}A^{f}\partial_{c e f}{\Gamma^{a}_{d b}} -> @(dGamma03)$,repeat=True)  # cdb(dRabcd04.501,dRabcd04)
   substitute (dRabcd04,$A^{c}A^{b}A^{e}A^{f}\partial_{c e f}{\Gamma^{a}_{b d}} -> @(dGamma03)$,repeat=True)  # cdb(dRabcd04.502,dRabcd04)
   substitute (dRabcd04,$A^{c}A^{b}A^{e}\partial_{c e}{\Gamma^{a}_{d b}} -> @(dGamma02)$,repeat=True)         # cdb(dRabcd04.503,dRabcd04)
   substitute (dRabcd04,$A^{c}A^{b}A^{e}\partial_{c e}{\Gamma^{a}_{b d}} -> @(dGamma02)$,repeat=True)         # cdb(dRabcd04.504,dRabcd04)
   substitute (dRabcd04,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{d b}} -> @(dGamma01)$,repeat=True)                # cdb(dRabcd04.505,dRabcd04)
   substitute (dRabcd04,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{b d}} -> @(dGamma01)$,repeat=True)                # cdb(dRabcd04.506,dRabcd04)
   distribute (dRabcd04)                                                                                      # cdb(dRabcd04.507,dRabcd04)
   sort_product   (dRabcd04)                                                                                  # cdb(dRabcd04.508,dRabcd04)
   rename_dummies (dRabcd04)                                                                                  # cdb(dRabcd04.509,dRabcd04)

   substitute (dRabcd05,$A^{c}A^{b}A^{e}A^{f}A^{g}\partial_{c e f g}{\Gamma^{a}_{d b}} -> @(dGamma04)$,repeat=True)
   substitute (dRabcd05,$A^{c}A^{b}A^{e}A^{f}A^{g}\partial_{c e f g}{\Gamma^{a}_{b d}} -> @(dGamma04)$,repeat=True)
   substitute (dRabcd05,$A^{c}A^{b}A^{e}A^{f}\partial_{c e f}{\Gamma^{a}_{d b}} -> @(dGamma03)$,repeat=True)
   substitute (dRabcd05,$A^{c}A^{b}A^{e}A^{f}\partial_{c e f}{\Gamma^{a}_{b d}} -> @(dGamma03)$,repeat=True)
   substitute (dRabcd05,$A^{c}A^{b}A^{e}\partial_{c e}{\Gamma^{a}_{d b}} -> @(dGamma02)$,repeat=True)
   substitute (dRabcd05,$A^{c}A^{b}A^{e}\partial_{c e}{\Gamma^{a}_{b d}} -> @(dGamma02)$,repeat=True)
   substitute (dRabcd05,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{d b}} -> @(dGamma01)$,repeat=True)
   substitute (dRabcd05,$A^{c}A^{b}\partial_{c}{\Gamma^{a}_{b d}} -> @(dGamma01)$,repeat=True)
   distribute (dRabcd05)
   sort_product   (dRabcd05)
   rename_dummies (dRabcd05)

   end_stage_5 = time.time()
\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{dRabcd01.500} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dRabcd02.500} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.501} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.502} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.503} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.504} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.505} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dRabcd03.500} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.501} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.502} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.503} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.504} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.505} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.506} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.507} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dRabcd04.500} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd04.501} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd04.502} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd04.503} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd04.504} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd04.505} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd04.506} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd04.507} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd04.508} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd04.509} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 6: Replace partial derivs of $R$ with covariant derivs of $R$}

\begin{cadabra}
   # now eliminate remaining partial derivs of Rabcd by substitution from the lower order dRabcd

   # note that
   #   dRabcd01 = R^a_{cdb,e} A^c A^d A^e
   #   dRabcd02 = R^a_{cdb,ef} A^c A^d A^e A^f
   #   dRabcd03 = R^a_{cdb,efg} A^c A^d A^e A^f A^g

   # thus we can use
   #   dRabcd01 to eliminate 1st partial derivs of R in dRabcd03, dRabcd04, etc.
   #   dRabcd02 to eliminate 2nd partial derivs of R in dRabcd04, dRabcd05, etc.
   #   dRabcd03 to eliminate 3rd partial derivs of R in dRabcd05, dRabcd06, etc.

   beg_stage_6 = time.time()

   substitute (dRabcd03,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c d b}} -> @(dRabcd01)$,repeat=True)         # cdb(dRabcd03.601,dRabcd03)
   distribute (dRabcd03)                                                                                 # cdb(dRabcd03.602,dRabcd03)

   # note: dRabcd04 and dRabcd05 unused in this code (or any other code)

   substitute (dRabcd04,$A^{c}A^{d}A^{e}A^{f}\partial_{e f}{R^{a}_{c d b}} -> @(dRabcd02)$,repeat=True)  # cdb(dRabcd04.601,dRabcd04)
   substitute (dRabcd04,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c d b}} -> @(dRabcd01)$,repeat=True)         # cdb(dRabcd04.602,dRabcd04)
   distribute (dRabcd04)                                                                                 # cdb(dRabcd04.603,dRabcd04)

   substitute (dRabcd05,$A^{c}A^{d}A^{e}A^{f}A^{g}\partial_{e f g}{R^{a}_{c d b}} -> @(dRabcd03)$,repeat=True)
   substitute (dRabcd05,$A^{c}A^{d}A^{e}A^{f}\partial_{e f}{R^{a}_{c d b}} -> @(dRabcd02)$,repeat=True)
   substitute (dRabcd05,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c d b}} -> @(dRabcd01)$,repeat=True)
   distribute (dRabcd05)

   end_stage_6 = time.time()
\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{dRabcd03.601} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.602} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dRabcd04.601} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd04.602} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd04.603} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 7: Reformatting}

\begin{cadabra}
   beg_stage_7 = time.time()

   dRabcd01 = flatten_Rabcd (dRabcd01)  # cdb(dRabcd01.701,dRabcd01)
   dRabcd02 = flatten_Rabcd (dRabcd02)  # cdb(dRabcd02.701,dRabcd02)
   dRabcd03 = flatten_Rabcd (dRabcd03)  # cdb(dRabcd03.701,dRabcd03)
   dRabcd04 = flatten_Rabcd (dRabcd04)  # cdb(dRabcd04.701,dRabcd04)
   dRabcd05 = flatten_Rabcd (dRabcd05)  # cdb(dRabcd05.701,dRabcd05)

   canonicalise (dRabcd01)   # cdb(dRabcd01.702,dRabcd01)
   canonicalise (dRabcd02)   # cdb(dRabcd02.702,dRabcd02)
   canonicalise (dRabcd03)   # cdb(dRabcd03.702,dRabcd03)
   canonicalise (dRabcd04)   # cdb(dRabcd04.702,dRabcd04)
   canonicalise (dRabcd05)   # cdb(dRabcd05.702,dRabcd05)

   end_stage_7 = time.time()

   # cdbBeg (timing)
   print ("Stage 1: {:7.1f} secs\\hfill\\break".format(end_stage_1-beg_stage_1))
   print ("Stage 2: {:7.1f} secs\\hfill\\break".format(end_stage_2-beg_stage_2))
   print ("Stage 3: {:7.1f} secs\\hfill\\break".format(end_stage_3-beg_stage_3))
   print ("Stage 4: {:7.1f} secs\\hfill\\break".format(end_stage_4-beg_stage_4))
   print ("Stage 5: {:7.1f} secs\\hfill\\break".format(end_stage_5-beg_stage_5))
   print ("Stage 6: {:7.1f} secs\\hfill\\break".format(end_stage_6-beg_stage_6))
   print ("Stage 7: {:7.1f} secs".format(end_stage_7-beg_stage_7))
   # cdbEnd (timing)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{dRabcd01.701} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.701} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.701} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd04.701} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd05.701} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{dRabcd01.702} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd02.702} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd03.702} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd04.702} \end{dmath*}
   \begin{dmath*} \cdb*{dRabcd05.702} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{cadabra}
   cdblib.create ('dRabcd.json')

   cdblib.put ('dRabcd01',dRabcd01,'dRabcd.json')
   cdblib.put ('dRabcd02',dRabcd02,'dRabcd.json')
   cdblib.put ('dRabcd03',dRabcd03,'dRabcd.json')
   cdblib.put ('dRabcd04',dRabcd04,'dRabcd.json')
   cdblib.put ('dRabcd05',dRabcd05,'dRabcd.json')

\end{cadabra}

% =================================================================================================
% the remaining code is just for pretty printing

\clearpage

\begin{cadabra}
   # note: keeping numbering as is (out of order) to ensure R appears before \nabla R etc.
   def product_sort (obj):
       substitute (obj,$ A^{a}                            -> A001^{a}               $)
       substitute (obj,$ x^{a}                            -> A002^{a}               $)
       substitute (obj,$ g^{a b}                          -> A003^{a b}             $)
       substitute (obj,$ \nabla_{e f g h}{R_{a b c d}}    -> A008_{a b c d e f g h} $)
       substitute (obj,$ \nabla_{e f g}{R_{a b c d}}      -> A007_{a b c d e f g}   $)
       substitute (obj,$ \nabla_{e f}{R_{a b c d}}        -> A006_{a b c d e f}     $)
       substitute (obj,$ \nabla_{e}{R_{a b c d}}          -> A005_{a b c d e}       $)
       substitute (obj,$ R_{a b c d}                      -> A004_{a b c d}         $)
       sort_product   (obj)
       rename_dummies (obj)
       substitute (obj,$ A001^{a}                  -> A^{a}                         $)
       substitute (obj,$ A002^{a}                  -> x^{a}                         $)
       substitute (obj,$ A003^{a b}                -> g^{a b}                       $)
       substitute (obj,$ A004_{a b c d}            -> R_{a b c d}                   $)
       substitute (obj,$ A005_{a b c d e}          -> \nabla_{e}{R_{a b c d}}       $)
       substitute (obj,$ A006_{a b c d e f}        -> \nabla_{e f}{R_{a b c d}}     $)
       substitute (obj,$ A007_{a b c d e f g}      -> \nabla_{e f g}{R_{a b c d}}   $)
       substitute (obj,$ A008_{a b c d e f g h}    -> \nabla_{e f g h}{R_{a b c d}} $)

       return obj

   def reformat (obj,scale):
       foo  = Ex(str(scale))
       bah := @(foo) @(obj).
       distribute     (bah)
       bah = product_sort (bah)
       rename_dummies (bah)
       canonicalise   (bah)
       factor_out     (bah,$A^{a?}$)
       ans := @(bah).
       return ans

   scaled1 = reformat (dRabcd01, 1)    # cdb(scaled1.601,scaled1)
   scaled2 = reformat (dRabcd02, 1)    # cdb(scaled2.601,scaled2)
   scaled3 = reformat (dRabcd03,-2)    # cdb(scaled3.601,scaled3)
   scaled4 = reformat (dRabcd04,-5)    # cdb(scaled4.601,scaled4)
   scaled5 = reformat (dRabcd05,-3)    # cdb(scaled5.601,scaled5)

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Symmetrised partial derivatives of $R^a{}_{bcd}$}
\begin{dgroup*}
   \begin{dmath*}    A^c A^d A^e R^a{}_{cdb,e} = \cdb{scaled1.601} \end{dmath*}
   \begin{dmath*}    A^c A^d A^e A^{f} R^a{}_{cdb,ef} = \cdb{scaled2.601} \end{dmath*}
   \begin{dmath*} -2 A^c A^d A^e A^{f} A^{g} R^a{}_{cdb,efg} = \cdb{scaled3.601} \end{dmath*}
   \begin{dmath*} -5 A^c A^d A^e A^{f} A^{g} A^{h} R^a{}_{cdb,efgh} = \cdb{scaled4.601} \end{dmath*}
   \begin{dmath*} -3 A^c A^d A^e A^{f} A^{g} A^{h} A^{i}R^a{}_{cdb,efghi} = \cdb{scaled5.601} \end{dmath*}
\end{dgroup*}

% LCB: only need dRabcd01,02,03. So I could save time by not
%      computing dRabcd04,05

% computing just dRabcd0n to n=4 takes about 1 min 40 sec.
% but going to n=5 takes about 7 min

\clearpage

% =================================================================================================
% export selected objects, these will later be imported into a library
% these are the objects that will appear in the paper

\begin{cadabra}
   substitute (scaled1,$A^{a}->1$)
   substitute (scaled2,$A^{a}->1$)
   substitute (scaled3,$A^{a}->1$)
   substitute (scaled4,$A^{a}->1$)
   substitute (scaled5,$A^{a}->1$)

   cdblib.create ('dRabcd.export')

   # 6th order dRabcd, scaled
   cdblib.put ('dRabcd61scaled',scaled1,'dRabcd.export')
   cdblib.put ('dRabcd62scaled',scaled2,'dRabcd.export')
   cdblib.put ('dRabcd63scaled',scaled3,'dRabcd.export')
   cdblib.put ('dRabcd64scaled',scaled4,'dRabcd.export')
   cdblib.put ('dRabcd65scaled',scaled5,'dRabcd.export')

   checkpoint.append (scaled1)
   checkpoint.append (scaled2)
   checkpoint.append (scaled3)
   checkpoint.append (scaled4)
   checkpoint.append (scaled5)

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Timing}

\cdb{timing}

% =================================================================================================
% export checkpoints in json format

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   for i in range( len(checkpoint) ):
      cdblib.put ('check{:03d}'.format(i),checkpoint[i],checkpoint_file)
\end{cadabra}
\egroup

\end{document}

### 2019-09-24 source/cadabra/detg2.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

% =================================================================================================
% create checkpoint file

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/detg2.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{The determinant of the metric}

Our game here is to compute (the leading terms) in $\det g$ of the metric in RNC form
\begin{dgroup*}
   \begin{dmath*} g_{a b}(x) = \cdb{gab.001}+\BigO{\eps^5} \end{dmath*}
\end{dgroup*}
For the sake of simplicity let's assume that we are working in 3-dimensions. The following
analysis is easily generalsied to other dimensions (and the final answers for $\det g$ and
friends are unchanged).

Define $\eps^{abc}_{ijk}$ by
\begin{align}
   \eps^{abc}_{ijk} =
        \delta^a_i \delta^b_j \delta^c_k - \delta^b_i \delta^a_j \delta^c_k
      + \delta^c_i \delta^a_j \delta^b_k - \delta^c_i \delta^b_j \delta^a_k
      + \delta^b_i \delta^c_j \delta^a_k - \delta^a_i \delta^c_j \delta^b_k
\end{align}
It is easy to see that $\eps^{abc}_{ijk}$ is anti-symmetric in both its upper and lower
indices. A trivial computation shows that for any $3{}\times{}3$ square matrix $M_{ab}$,
\begin{align}
   \eps^{abc}_{123} M_{1a} M_{2b} M_{3c}
   = \left(
          \delta^a_1 \delta^b_2 \delta^c_3 - \delta^b_1 \delta^a_2 \delta^c_3
        + \delta^c_1 \delta^a_2 \delta^b_3 - \delta^c_1 \delta^b_2 \delta^a_3
        + \delta^b_1 \delta^c_2 \delta^a_3 - \delta^a_1 \delta^c_2 \delta^b_3
     \right)M_{1a} M_{2b} M_{3c}
   = \det M
\end{align}
This can be easily generalised to
\begin{align}
   \eps^{abc}_{ijk} M_{pa} M_{qb} M_{rc}
   =
   \begin{cases}
      \pm \det M &\text{when $(ijk)$ and $(pqr)$ are permutations of $(123)$}\\
      0 & \text{otherwise}
   \end{cases}
\end{align}
The $\pm$ sign in the above depends on the particular permutations of $(ijk)$ and $(pqr)$. If
both permutations are even or both odd then the sign is $+1$ otherwise the sign is $-1$.
The same arguments can also be applied to a matrix inverse $N^{-1}$ leading to
\begin{align}
   \eps^{ijk}_{uvw} N^{pu} N^{qv} M^{rw}
   =
   \begin{cases}
      \pm \det {N^{-1}} &\text{when $(ijk)$ and $(pqr)$ are permutations of $(123)$}\\
      0 & \text{otherwise}
   \end{cases}
\end{align}
Note that the $\pm$ in this case will match exactly that for the case of $\det M$. Thus,
multiplying both expressions and summing over all choices for $(ijk)$ and $(pqr)$ leads
to
\begin{align}
   \sum_{\substack{(ijk)\\(pqr)}}\left(\det N^{-1}\right) \det M
   = \eps^{ijk}_{uvw} N^{pu} N^{qv} M^{rw} \eps^{abc}_{ijk} M_{pa} M_{qb} M_{rc}
\end{align}
where the sum on the left hand side includes just those $(ijk)$ and $(prq)$ that are
permutations of $(123)$. There are $3!$ choices for $(ijk)$ and $3!$ choices for
$(pqr)$ and thus the left hand side is easily reduced to $(3!)^2 \det M/\det N$ where
$\det N = 1/\det N^{-1}$. For the right hand side notice that
\begin{align}
   \eps^{ijk}_{uvw} \eps^{abc}_{ijk} = 3! \eps^{abc}_{uvw}
\end{align}
which leads to
\begin{align}
   \det M = \frac{1}{3!} \det N \eps^{abc}_{uvw} M_{pa} M_{qb} M_{rc} N^{pu} N^{qv} N^{rw}
\end{align}

For our RNC metric we will set $N^{ab} = g^{ab}$ and $M_{ij} = g_{ij}(x)$. Since $g^{ab}$ is
of the form ${\rm diag}(-1,1,1,1)$ we have $\det g = -1$ and thus
\begin{align}
   \det g(x) = - \frac{1}{3!} \eps^{abc}_{ijk}\, g_{pa}(x)\, g_{qb}(x)\, g_{rc}(x)\, g^{ip} g^{jq} g^{kr}
\end{align}

The $\eps^{abc}_{ijk}$ can be constructed in Cadabra by applying the \verb|asym| algorithm
to the upper indices of $\delta^a_i \delta^b_j \delta^c_k$. Note that \verb|asym| will
include the $1/3!$ coeffcient as part of its output.

The following code computes $-\det g$ rather than $\det g$.

{\bf Note that Calzetta uses an opposite sign for $R_{abcd}$ so when comparing my results
against Calzetta do take note of this flipped sign in $R_{abcd}$.}

\clearpage

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Integer(1..2).

   \nabla{#}::Derivative.

   d{#}::KroneckerDelta.

   g^{a b}::Symmetric.
   g_{a b}::Symmetric.

   R_{a b c d}::RiemannTensor.

   x^{a}::Weight(label=numx,value=1).

   def truncate (obj,n):

       ans = Ex(0)

       for i in range (0,n+1):
          foo := @(obj).
          bah = Ex("numx = " + str(i))
          keep_weight (foo, bah)
          ans = ans + foo

       return ans

   import cdblib

   g0ab = cdblib.get('g_ab_0','metric.json')
   g1ab = cdblib.get('g_ab_1','metric.json')  # zero in RNC
   g2ab = cdblib.get('g_ab_2','metric.json')
   g3ab = cdblib.get('g_ab_3','metric.json')
   g4ab = cdblib.get('g_ab_4','metric.json')
   g5ab = cdblib.get('g_ab_5','metric.json')

   gab := @(g0ab) + @(g1ab) + @(g2ab) + @(g3ab) + @(g4ab) + @(g5ab).  # cdb (gab.001,gab)
   gxab := gx_{a b} -> @(gab).

   eps := d^{a}_{i} d^{b}_{j}.   # cdb(eps.001,eps)
   asym (eps,$^{a},^{b}$)        # cdb(eps.002,eps) # includes a factor of 1/2!

   # compute negative Ndetg rather than det g
   Ndetg := @(eps) gx_{p a} gx_{q b} g^{i p} g^{j q}.  # note 1/2! included in eps

   substitute       (Ndetg,gxab)
   distribute       (Ndetg)
   Ndetg = truncate (Ndetg,5)                                          # cdb (Ndetg.001,Ndetg)
   substitute       (Ndetg,$g^{a b} g_{b c} -> d^{a}_{c}$,repeat=True) # cdb (Ndetg.002,Ndetg)
   eliminate_kronecker (Ndetg)                                         # cdb (Ndetg.003,Ndetg)
   sort_product     (Ndetg)                                            # cdb (Ndetg.004,Ndetg)
   rename_dummies   (Ndetg)                                            # cdb (Ndetg.005,Ndetg)
   canonicalise     (Ndetg)                                            # cdb (Ndetg.006,Ndetg)

   # introduce the Ricci tensor

   substitute     (Ndetg,$R_{a b c d} g^{a c} -> R_{b d}$,repeat=True)                                  # cdb (Ndetg.101,Ndetg)
   substitute     (Ndetg,$\nabla_{a}{R_{b c d e}} g^{b d}  -> \nabla_{a}{R_{c e}}$,repeat=True)         # cdb (Ndetg.102,Ndetg)
   substitute     (Ndetg,$\nabla_{a b}{R_{c d e f}} g^{c e}  -> \nabla_{a b}{R_{d f}}$,repeat=True)     # cdb (Ndetg.103,Ndetg)
   substitute     (Ndetg,$\nabla_{a b c}{R_{d e f g}} g^{d f}  -> \nabla_{a b c}{R_{e g}}$,repeat=True) # cdb (Ndetg.104,Ndetg)

   # the following are based on sqrt-Ndetg.tex

   sqrtNdetg := 1/2 + (1/2) @(Ndetg)
               - (1/8) (1/9) R_{a b} R_{c d} x^{a} x^{b} x^{c} x^{d}
               - (1/4) (1/18) R_{a b} \nabla_{c}{R_{d e}} x^{a} x^{b} x^{c} x^{d} x^{e}.
               # cdb (sqrtNdetg.001,sqrtNdetg)

   sort_product   (sqrtNdetg)                                          # cdb (sqrtNdetg.002,sqrtNdetg)
   rename_dummies (sqrtNdetg)                                          # cdb (sqrtNdetg.003,sqrtNdetg)
   canonicalise   (sqrtNdetg)                                          # cdb (sqrtNdetg.004,sqrtNdetg)

   logNdetg := -1 + @(Ndetg)
               - (1/2) (1/9) R_{a b} R_{c d} x^{a} x^{b} x^{c} x^{d}
               - (1/18) R_{a b} \nabla_{c}{R_{d e}} x^{a} x^{b} x^{c} x^{d} x^{e}.
               # cdb (logNdetg.001,logNdetg)

   sort_product   (logNdetg)                                           # cdb (logNdetg.002,logNdetg)
   rename_dummies (logNdetg)                                           # cdb (logNdetg.003,logNdetg)
   canonicalise   (logNdetg)                                           # cdb (logNdetg.004,logNdetg)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{eps.001} \end{dmath*}
   \begin{dmath*} \cdb*{eps.002} \end{dmath*}
   \begin{dmath*} \cdb*{Ndetg.001} \end{dmath*}
   \begin{dmath*} \cdb*{Ndetg.002} \end{dmath*}
   \begin{dmath*} \cdb*{Ndetg.003} \end{dmath*}
   \begin{dmath*} \cdb*{Ndetg.004} \end{dmath*}
   \begin{dmath*} \cdb*{Ndetg.005} \end{dmath*}
   \begin{dmath*} \cdb*{Ndetg.006} \end{dmath*}
   \begin{dmath*} \cdb*{Ndetg.104} \end{dmath*}
   \begin{dmath*} \cdb*{sqrtNdetg.004} \end{dmath*}
   \begin{dmath*} \cdb*{logNdetg.004} \end{dmath*}
\end{dgroup*}

% =================================================================================================
% the remaining code is just for pretty printing

\clearpage

\begin{cadabra}
   # note: keeping numbering as is (out of order) to ensure R appears before \nabla R etc.
   def product_sort (obj):
       substitute (obj,$ x^{a}                            -> A000^{a}               $)
       substitute (obj,$ g^{a b}                          -> A001^{a b}             $)
       substitute (obj,$ \nabla_{c d e f}{R_{a b}}        -> A007_{a b c d e f}     $)
       substitute (obj,$ \nabla_{c d e}{R_{a b}}          -> A006_{a b c d e}       $)
       substitute (obj,$ \nabla_{c d}{R_{a b}}            -> A005_{a b c d}         $)
       substitute (obj,$ \nabla_{c}{R_{a b}}              -> A004_{a b c}           $)
       substitute (obj,$ \nabla_{e f g h}{R_{a b c d}}    -> A011_{a b c d e f g h} $)
       substitute (obj,$ \nabla_{e f g}{R_{a b c d}}      -> A010_{a b c d e f g}   $)
       substitute (obj,$ \nabla_{e f}{R_{a b c d}}        -> A009_{a b c d e f}     $)
       substitute (obj,$ \nabla_{e}{R_{a b c d}}          -> A008_{a b c d e}       $)
       substitute (obj,$ R_{a b}                          -> A002_{a b}             $)
       substitute (obj,$ R_{a b c d}                      -> A003_{a b c d}         $)
       sort_product   (obj)
       rename_dummies (obj)
       substitute (obj,$ A000^{a}                 -> x^{a}                          $)
       substitute (obj,$ A001^{a b}               -> g^{a b}                        $)
       substitute (obj,$ A002_{a b}               -> R_{a b}                        $)
       substitute (obj,$ A003_{a b c d}           -> R_{a b c d}                    $)
       substitute (obj,$ A004_{a b c}             -> \nabla_{c}{R_{a b}}            $)
       substitute (obj,$ A005_{a b c d}           -> \nabla_{c d}{R_{a b}}          $)
       substitute (obj,$ A006_{a b c d e}         -> \nabla_{c d e}{R_{a b}}        $)
       substitute (obj,$ A007_{a b c d e f}       -> \nabla_{c d e f}{R_{a b}}      $)
       substitute (obj,$ A008_{a b c d e}         -> \nabla_{e}{R_{a b c d}}        $)
       substitute (obj,$ A009_{a b c d e f}       -> \nabla_{e f}{R_{a b c d}}      $)
       substitute (obj,$ A010_{a b c d e f g}     -> \nabla_{e f g}{R_{a b c d}}    $)
       substitute (obj,$ A011_{a b c d e f g h}   -> \nabla_{e f g h}{R_{a b c d}}  $)

       return obj

   def get_term (obj,n):

       x^{a}::Weight(label=numx).

       foo := @(obj).
       bah  = Ex("numx = " + str(n))
       keep_weight (foo,bah)

       return foo

   def reformat (obj,scale):
       foo  = Ex(str(scale))
       bah := @(foo) @(obj).
       distribute     (bah)
       bah = product_sort (bah)
       rename_dummies (bah)
       canonicalise   (bah)
       sort_sum       (bah)
       factor_out     (bah,$x^{a?}$)
       ans := @(bah) / @(foo).
       return ans

   def rescale (obj,scale):
       foo  = Ex(str(scale))
       bah := @(foo) @(obj).
       distribute  (bah)
       factor_out  (bah,$x^{a?}$)
       return bah

   # ---------------------------------------------------------------
   # reformat Ndetg

   Rterm0 = get_term (Ndetg,0)       # cdb(Rterm0.701,Rterm0)
   Rterm1 = get_term (Ndetg,1)       # cdb(Rterm1.701,Rterm1)
   Rterm2 = get_term (Ndetg,2)       # cdb(Rterm2.701,Rterm2)
   Rterm3 = get_term (Ndetg,3)       # cdb(Rterm3.701,Rterm3)
   Rterm4 = get_term (Ndetg,4)       # cdb(Rterm4.701,Rterm4)
   Rterm5 = get_term (Ndetg,5)       # cdb(Rterm5.701,Rterm5)

   Rterm0 = reformat (Rterm0,  1)    # cdb(Rterm0.702,Rterm0)
   Rterm1 = reformat (Rterm1,  1)    # cdb(Rterm1.702,Rterm1)
   Rterm2 = reformat (Rterm2,  3)    # cdb(Rterm2.702,Rterm2)
   Rterm3 = reformat (Rterm3,  6)    # cdb(Rterm3.702,Rterm3)
   Rterm4 = reformat (Rterm4,180)    # cdb(Rterm4.702,Rterm4)
   Rterm5 = reformat (Rterm5, 90)    # cdb(Rterm5.702,Rterm5)

   Ndetg := @(Rterm0) + @(Rterm1) + @(Rterm2) + @(Rterm3) + @(Rterm4) + @(Rterm5).  # cdb (Ndetg.701,Ndetg)

   # ---------------------------------------------------------------
   # reformat sqrtNdetg

   Rterm0 = get_term (sqrtNdetg,0)   # cdb(Rterm0.801,Rterm0)
   Rterm1 = get_term (sqrtNdetg,1)   # cdb(Rterm1.801,Rterm1)
   Rterm2 = get_term (sqrtNdetg,2)   # cdb(Rterm2.801,Rterm2)
   Rterm3 = get_term (sqrtNdetg,3)   # cdb(Rterm3.801,Rterm3)
   Rterm4 = get_term (sqrtNdetg,4)   # cdb(Rterm4.801,Rterm4)
   Rterm5 = get_term (sqrtNdetg,5)   # cdb(Rterm5.801,Rterm5)

   Rterm0 = reformat (Rterm0,  1)    # cdb(Rterm0.802,Rterm0)
   Rterm1 = reformat (Rterm1,  1)    # cdb(Rterm1.802,Rterm1)
   Rterm2 = reformat (Rterm2,  6)    # cdb(Rterm2.802,Rterm2)
   Rterm3 = reformat (Rterm3, 12)    # cdb(Rterm3.802,Rterm3)
   Rterm4 = reformat (Rterm4,360)    # cdb(Rterm4.802,Rterm4)
   Rterm5 = reformat (Rterm5,360)    # cdb(Rterm5.802,Rterm5)

   sqrtNdetg := @(Rterm0) + @(Rterm1) + @(Rterm2) + @(Rterm3) + @(Rterm4) + @(Rterm5).  # cdb (sqrtNdetg.801,sqrtNdetg)

   # ---------------------------------------------------------------
   # reformat logNdetg

   Rterm0 = get_term (logNdetg,0)    # cdb(Rterm0.901,Rterm0)
   Rterm1 = get_term (logNdetg,1)    # cdb(Rterm1.901,Rterm1)
   Rterm2 = get_term (logNdetg,2)    # cdb(Rterm2.901,Rterm2)
   Rterm3 = get_term (logNdetg,3)    # cdb(Rterm3.901,Rterm3)
   Rterm4 = get_term (logNdetg,4)    # cdb(Rterm4.901,Rterm4)
   Rterm5 = get_term (logNdetg,5)    # cdb(Rterm5.901,Rterm5)

   Rterm0 = reformat (Rterm0,  1)    # cdb(Rterm0.902,Rterm0)
   Rterm1 = reformat (Rterm1,  1)    # cdb(Rterm1.902,Rterm1)
   Rterm2 = reformat (Rterm2,  3)    # cdb(Rterm2.902,Rterm2)
   Rterm3 = reformat (Rterm3,  6)    # cdb(Rterm3.902,Rterm3)
   Rterm4 = reformat (Rterm4,180)    # cdb(Rterm4.902,Rterm4)
   Rterm5 = reformat (Rterm5, 90)    # cdb(Rterm5.902,Rterm5)

   logNdetg := @(Rterm0) + @(Rterm1) + @(Rterm2) + @(Rterm3) + @(Rterm4) + @(Rterm5).  # cdb (logNdetg.901,logNdetg)

\end{cadabra}

\clearpage

% =================================================================================================
\section*{The metric determinant in Riemann normal coordinates}

\begin{dgroup*}
   \Dmath*{-\det g(x) = \cdb{Ndetg.701}+\BigO{\eps^6}}
\end{dgroup*}

% =================================================================================================
\section*{The volume element in Riemann normal coordinates}

If $-\det g(x)$ is non-negative then we also have
%
\begin{dgroup*}
   \Dmath*{\sqrt{-\det g(x)} = \cdb{sqrtNdetg.801}+\BigO{\eps^6}}
\end{dgroup*}

% =================================================================================================
\section*{The log of -detg in Riemann normal coordinates}

Apart from the signs, this matches exactly the expression given by Calzetta etal. (eq. A14)

\begin{dgroup*}
   \Dmath*{\log\left(-\det g(x)\right) = \cdb{logNdetg.901}+\BigO{\eps^6}}
\end{dgroup*}

\clearpage

% =================================================================================================
% export selected objects, these will later be imported into a library
% these are the objects that will appear in the paper

\begin{cadabra}
   cdblib.create ('detg2.export')

   cdblib.put ('Ndetg',    Ndetg,    'detg2.export')
   cdblib.put ('sqrtNdetg',sqrtNdetg,'detg2.export')
   cdblib.put ('logNdetg', logNdetg, 'detg2.export')

   checkpoint.append (Ndetg)
   checkpoint.append (sqrtNdetg)
   checkpoint.append (logNdetg)

\end{cadabra}

% =================================================================================================
% export checkpoints in json format

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   for i in range( len(checkpoint) ):
      cdblib.put ('check{:03d}'.format(i),checkpoint[i],checkpoint_file)
\end{cadabra}
\egroup

\end{document}

### 2019-09-24 source/cadabra/export.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   \nabla{#}::Derivative.
   \partial{#}::PartialDerivative.

   import cdblib

   # ------------------------------------------------------------------------------------
   # the metric

   # metric in terms of R and \partial R
   tmp = cdblib.get ('g_ab_3','metric.export')         # cdb (metric3,tmp)
   tmp = cdblib.get ('g_ab_4','metric.export')         # cdb (metric4,tmp)
   tmp = cdblib.get ('g_ab_5','metric.export')         # cdb (metric5,tmp)
   tmp = cdblib.get ('g_ab_6','metric.export')         # cdb (metric6,tmp)

   # metric in terms of R and \nabla R
   tmp = cdblib.get ('g_ab',  'metric.export')         # cdb (metric,tmp)

   tmp = cdblib.get ('g_ab_scaled0','metric.export')   # cdb (metric.scaled0,tmp)
   tmp = cdblib.get ('g_ab_scaled2','metric.export')   # cdb (metric.scaled2,tmp)
   tmp = cdblib.get ('g_ab_scaled3','metric.export')   # cdb (metric.scaled3,tmp)
   tmp = cdblib.get ('g_ab_scaled4','metric.export')   # cdb (metric.scaled4,tmp)
   tmp = cdblib.get ('g_ab_scaled5','metric.export')   # cdb (metric.scaled5,tmp)

   # ------------------------------------------------------------------------------------
   # the inverse metric

   # inverse metric in terms of R and \partial R
   tmp = cdblib.get ('g^ab_3','metric-inv.export')         # cdb (metric3.inv,tmp)
   tmp = cdblib.get ('g^ab_4','metric-inv.export')         # cdb (metric4.inv,tmp)
   tmp = cdblib.get ('g^ab_5','metric-inv.export')         # cdb (metric5.inv,tmp)
   tmp = cdblib.get ('g^ab_6','metric-inv.export')         # cdb (metric6.inv,tmp)

   # inverse metric in terms of R and \nabla R
   tmp = cdblib.get ('g^ab',  'metric-inv.export')         # cdb (metric.inv,tmp)

   tmp = cdblib.get ('g^ab_scaled0','metric-inv.export')   # cdb (metric.inv.scaled0,tmp)
   tmp = cdblib.get ('g^ab_scaled2','metric-inv.export')   # cdb (metric.inv.scaled2,tmp)
   tmp = cdblib.get ('g^ab_scaled3','metric-inv.export')   # cdb (metric.inv.scaled3,tmp)
   tmp = cdblib.get ('g^ab_scaled4','metric-inv.export')   # cdb (metric.inv.scaled4,tmp)
   tmp = cdblib.get ('g^ab_scaled5','metric-inv.export')   # cdb (metric.inv.scaled5,tmp)

   # ------------------------------------------------------------------------------------
   # the generalised connections

   # 4th order gen gamma
   tmp = cdblib.get ('gen_gamma_0_4th','genGamma.export')  # cdb (genGamma04,tmp)
   tmp = cdblib.get ('gen_gamma_1_4th','genGamma.export')  # cdb (genGamma14,tmp)

   # 6th order gen gamma
   tmp = cdblib.get ('gen_gamma_0','genGamma.export')  # cdb (genGamma0,tmp)
   tmp = cdblib.get ('gen_gamma_1','genGamma.export')  # cdb (genGamma1,tmp)
   tmp = cdblib.get ('gen_gamma_2','genGamma.export')  # cdb (genGamma2,tmp)
   tmp = cdblib.get ('gen_gamma_3','genGamma.export')  # cdb (genGamma3,tmp)

   # 6th order gen gamma scaled
   tmp = cdblib.get ('gen_gamma_0_scaled','genGamma.export')  # cdb (genGamma0scaled,tmp)
   tmp = cdblib.get ('gen_gamma_1_scaled','genGamma.export')  # cdb (genGamma1scaled,tmp)
   tmp = cdblib.get ('gen_gamma_2_scaled','genGamma.export')  # cdb (genGamma2scaled,tmp)
   tmp = cdblib.get ('gen_gamma_3_scaled','genGamma.export')  # cdb (genGamma3scaled,tmp)

   # ------------------------------------------------------------------------------------
   # the metric determinent and friends

   tmp = cdblib.get ('Ndetg','detg2.export')      # cdb (Ndetg,tmp)
   tmp = cdblib.get ('sqrtNdetg','detg2.export')  # cdb (sqrtNdetg,tmp)
   tmp = cdblib.get ('logNdetg','detg2.export')   # cdb (logNdetg,tmp)

   # ------------------------------------------------------------------------------------
   # the geodesic ivp

   # 4th order ivp terms, scaled
   tmp = cdblib.get ('ivp42','geodesic-ivp.export')  # cdb (ivp42,tmp)
   tmp = cdblib.get ('ivp43','geodesic-ivp.export')  # cdb (ivp43,tmp)

   # 6th order ivp terms, scaled
   tmp = cdblib.get ('ivp62','geodesic-ivp.export')  # cdb (ivp62,tmp)
   tmp = cdblib.get ('ivp63','geodesic-ivp.export')  # cdb (ivp63,tmp)
   tmp = cdblib.get ('ivp64','geodesic-ivp.export')  # cdb (ivp64,tmp)
   tmp = cdblib.get ('ivp65','geodesic-ivp.export')  # cdb (ivp65,tmp)

   # ------------------------------------------------------------------------------------
   # the geodesic bvp

   # 4th order ivp
   tmp = cdblib.get ('bvp4','geodesic-bvp.export')  # cdb (bvp4,tmp)

   # 6th order bvp terms, scaled
   tmp = cdblib.get ('bvp622','geodesic-bvp.export')  # cdb (bvp622,tmp)
   tmp = cdblib.get ('bvp623','geodesic-bvp.export')  # cdb (bvp623,tmp)
   tmp = cdblib.get ('bvp624','geodesic-bvp.export')  # cdb (bvp624,tmp)
   tmp = cdblib.get ('bvp625','geodesic-bvp.export')  # cdb (bvp625,tmp)

   tmp = cdblib.get ('bvp633','geodesic-bvp.export')  # cdb (bvp633,tmp)
   tmp = cdblib.get ('bvp634','geodesic-bvp.export')  # cdb (bvp634,tmp)
   tmp = cdblib.get ('bvp635','geodesic-bvp.export')  # cdb (bvp635,tmp)

   tmp = cdblib.get ('bvp644','geodesic-bvp.export')  # cdb (bvp644,tmp)
   tmp = cdblib.get ('bvp645','geodesic-bvp.export')  # cdb (bvp645,tmp)

   tmp = cdblib.get ('bvp655','geodesic-bvp.export')  # cdb (bvp655,tmp)

   # ------------------------------------------------------------------------------------
   # the geodesic lsq

   # 6th order lsq terms, scaled
   tmp = cdblib.get ('rnc61scaled','gen2rnc.export')  # cdb (rnc61,tmp)
   tmp = cdblib.get ('rnc62scaled','gen2rnc.export')  # cdb (rnc62,tmp)
   tmp = cdblib.get ('rnc63scaled','gen2rnc.export')  # cdb (rnc63,tmp)
   tmp = cdblib.get ('rnc64scaled','gen2rnc.export')  # cdb (rnc64,tmp)
   tmp = cdblib.get ('rnc65scaled','gen2rnc.export')  # cdb (rnc65,tmp)

   # ------------------------------------------------------------------------------------
   # the geodesic lsq

   # 3rd to 6th order lsq
   tmp = cdblib.get ('lsq3','geodesic-lsq.export')  # cdb (lsq3,tmp)
   tmp = cdblib.get ('lsq4','geodesic-lsq.export')  # cdb (lsq4,tmp)
   tmp = cdblib.get ('lsq5','geodesic-lsq.export')  # cdb (lsq5,tmp)
   tmp = cdblib.get ('lsq6','geodesic-lsq.export')  # cdb (lsq6,tmp)

   # 6th order lsq terms, scaled
   tmp = cdblib.get ('lsq60','geodesic-lsq.export')  # cdb (lsq60,tmp)
   tmp = cdblib.get ('lsq62','geodesic-lsq.export')  # cdb (lsq62,tmp)
   tmp = cdblib.get ('lsq63','geodesic-lsq.export')  # cdb (lsq63,tmp)
   tmp = cdblib.get ('lsq64','geodesic-lsq.export')  # cdb (lsq64,tmp)
   tmp = cdblib.get ('lsq65','geodesic-lsq.export')  # cdb (lsq65,tmp)

   # ------------------------------------------------------------------------------------
   # the dRabcd

   # 6th order dRabcd, scaled
   tmp = cdblib.get ('dRabcd61scaled','dRabcd.export')  # cdb (dRabcd61,tmp)
   tmp = cdblib.get ('dRabcd62scaled','dRabcd.export')  # cdb (dRabcd62,tmp)
   tmp = cdblib.get ('dRabcd63scaled','dRabcd.export')  # cdb (dRabcd63,tmp)
   tmp = cdblib.get ('dRabcd64scaled','dRabcd.export')  # cdb (dRabcd64,tmp)
   tmp = cdblib.get ('dRabcd65scaled','dRabcd.export')  # cdb (dRabcd65,tmp)

   # ------------------------------------------------------------------------------------
   # the dGamma

   # 6th order dGamma, scaled
   tmp = cdblib.get ('dGamma61scaled','dGamma.export')  # cdb (dGamma61,tmp)
   tmp = cdblib.get ('dGamma62scaled','dGamma.export')  # cdb (dGamma62,tmp)
   tmp = cdblib.get ('dGamma63scaled','dGamma.export')  # cdb (dGamma63,tmp)
   tmp = cdblib.get ('dGamma64scaled','dGamma.export')  # cdb (dGamma64,tmp)
   tmp = cdblib.get ('dGamma65scaled','dGamma.export')  # cdb (dGamma65,tmp)

\end{cadabra}

\end{document}

### 2019-09-24 source/cadabra/genGamma.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

% =================================================================================================
% create checkpoint file

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/genGamma.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{The generalised connections}

The generalised connections may be computed recursively using
\begin{align}
   \label{eq:GenGamma}
\Gamma^{a}{}_{b\uc d} = \Gamma^{a}{}_{(b\uc,d)}
                - (n+1) \Gamma^{a}{}_{p(\uc}
                        \Gamma^{p}{}_{bd)}
\end{align}
where $\uc$ contains $n>0$ indices. The sequence begins with the standard metric compatible connection
\begin{align}
   \Gamma^{d}_{ab} = \frac{1}{2} g^{dc}\left( g_{cb,a} + g_{ac,b} - g_{ab,c} \right)
\end{align}

Here we will use the results of {\tt metric.tex} and {\tt metric-inv.tex} to compute the metric connection
$\Gamma^{d}_{ab}$. But since the $g_{ab}$ and $g^{ab}$ provided by those codes are truncated at a
particular order in the curvatures (and thus are only approximations to the $g_{ab}$ and $g^{ab}$) similar
truncations will arise in the $\Gamma^{a}{}_{b\uc d}$.

Approximations will be denoted by the addition of an overbar to an object. In this notation the metric
$g$ can be written as
\begin{align}
   g = {\bar g} + \BigO{\eps^n}
\end{align}
in which ${\bar g}$ is the truncated polynomial approximation to $g$ and $\BigO{\eps^n}$ is the error term
(containing terms no smaller than $\eps^n$). The polynomial structure of ${\bar g}$ can be expressed as
\begin{align}
   \gabBar = \ngabBar{0}
           + \ngabBar{1}
           + \ngabBar{2}
           + \dots
           + \ngabBar{p}
\end{align}
in which each terms like $\overset{m}{\bar g}$ contains only terms of order $m$. This notation will be applied
to other quantities in particular the generalised connections.

The notation $\BigO{\eps^n}$ denotes terms in the curvatures that are of order $\eps^n$. What does this actually mean?
Each term in $R$ is of order $\eps^2$ while each derivative of $R$ carries an extra power of $\eps$.
Thus $R\cdot R = \BigO{\eps^4}$, $R\cdot R\cdot\nabla R = \BigO{\eps^7}$ and $R\cdot R\cdot\nabla^2R = \BigO{\eps^8}$.

We will also adopt the convention that an object is said to be an $\BigO{\eps^{m}}$ approximation when the corresponding error term is $\BigO{\eps^{m+1}}$.

Consider the $\BigO{\eps^{m}}$ approximation of the generalised connection, namely,
\begin{align}
   \GammaBar^{a}{}_{b\ucn d}
      = \nGammaBar{0}^{a}{}_{b\ucn d}
      + \nGammaBar{1}^{a}{}_{b\ucn d}
      + \nGammaBar{2}^{a}{}_{b\ucn d}
      + \dots
      + \nGammaBar{m}^{a}{}_{b\ucn d}
\end{align}
where $\ucn$ denotes a set of indices such as $c_1c_2c_3\dots c_n$.

The first thing to note is that
\begin{align}
   0 = \nGammaBar{1+n}^{a}{}_{(b\ucn,d)}
\end{align}

There are two proofs of this claim. For the first proof, note (by inspection) that the order $\BigO{\eps^p}$
approximation for $\GammaBar^{a}{}_{b\ucn d}$ is a polynomial in $x$ of degree $p-n-1$. Thus
$\nGammaBar{1+n}^{a}{}_{(b\ucn,d)}$ is a polynomial in $x$ of degree
zero, i.e., a constant. However, we know that all generalised connections vanish at the origin of the RNC frame.
Thus this constant must be zero. The second proof makes explicit use of the first (and second?) Bianchi identity,
that is $0=R_{a(bcd)}$. The term $\nGammaBar{1+n}^{a}{}_{(b\ucn,d)}$ will
itself consist of a sum of terms built from combinations of $x$, $R$, $\nabla R$ etc. The $x^{a}$ will always
appear in a contraction with one of the indices on $R_{abcd}$ or one of its derivatives. Consider any one of
these terms, denoted by $A$, and assume for the moment that $1+n$ is an even number, say $1+n=2p$. The indices
$(b\ucn,d)$ must somehow be assigned to the factors that comprise $A$. Our aim is to show that at least one $R$
factor in $A$ will receive 3 of these indices and thus by the Bianchi identities will be zero. If there are too
many $R$ factors then the Bianchi identities will not come into play. So how many $R$ factors can we expect?
Since $A$ is a term in an $\BigO{\eps^{(n+1)}}$ approximation there can be no more than $(n+1)/2=p$ Riemann
factors. There will be at least one $x$ term contracted with one of the $p$ Riemann factors. However, we have
$n+2=2p+1$ indices to distribute amongst the $x$ term and $p$ Riemann factors. One of the indices is a derivative
index and will have nett effect of transferring that index from $x$ to one of the Riemann factors. The remaining
$2p$ indices must be distributed amongst the $p$ Riemann factors. It is not possible to avoid assigning three
indices to at least one of the Riemann factors. Thus, by the Bianchi identity, this $A$ term must vanish. Similar
arguments can be applied to the other cases where the $A$ terms consists of products of $R$ and its derivatives
and in the case where $n+1$ is an odd number. The analysis always comes down to the distribution of the indices
$(b\ucn,d)$ amongst the factors of a typical $A$ term. In all cases the Bianchi identity will enter the play and
force $A$ to be zero.

A corollary of the second proof is that for all $m<n+2$
\begin{align}
   0 = \nGammaBar{m}^{a}{}_{b\ucn d}
\end{align}
The proof follows exactly that of the second proof given above.

We can use the above results to streamline the computation of the generalised connections.
We begin with the formal expression for the $\BigO{\eps^m}$ approximations
\begin{align}
   \Gamma^{a}{}_{bc}
      &= \nGammaBar{2}^{a}{}_{bc}
       + \nGammaBar{3}^{a}{}_{bc}
       + \nGammaBar{4}^{a}{}_{bc}
       + \dots
       + \nGammaBar{m}^{a}{}_{bc}\\
   \Gamma^{a}{}_{b\uc}
      &= \nGammaBar{n+1}^{a}{}_{b\uc}
       + \nGammaBar{n+2}^{a}{}_{b\uc}
       + \nGammaBar{n+3}^{a}{}_{b\uc}
       + \dots
       + \nGammaBar{m}^{a}{}_{b\uc}\\
   \Gamma^{a}{}_{b\uc d}
      &= \nGammaBar{n+2}^{a}{}_{b\uc d}
       + \nGammaBar{n+3}^{a}{}_{b\uc d}
       + \nGammaBar{n+4}^{a}{}_{b\uc d}
       + \dots
       + \nGammaBar{m}^{a}{}_{b\uc d}\label{eq:GenGammaA}
\end{align}
These can be substituted into equation (\ref{eq:GenGamma}) with the result
\def\m{\hskip 4pt}
\begin{align}
   \Gamma^{a}{}_{b\uc d}
      &= \nGammaBar{n+1}^{a}{}_{(b\uc,d)}
       + \nGammaBar{n+2}^{a}{}_{(b\uc,d)}
       + \nGammaBar{n+3}^{a}{}_{(b\uc,d)}
       + \dots
       + \nGammaBar{m}^{a}{}_{(b\uc,d)}
       -(n+1)\left(\m \nGammaBar{n+1}^{a}{}_{p\uc}
                    + \nGammaBar{n+2}^{a}{}_{p\uc}
                    + \nGammaBar{n+3}^{a}{}_{p\uc}
                    + \dots
                    + \nGammaBar{m}^{a}{}_{p\uc}\right)
             \left(   \nGammaBar{2}^{p}{}_{bd}
                    + \nGammaBar{3}^{p}{}_{bd}
                    + \nGammaBar{4}^{p}{}_{bd}
                    + \dots
                    + \nGammaBar{m}^{p}{}_{bd}\right)\label{eq:GenGammaB}
\end{align}
where it is understood that in expanding the pair of bracketed terms in the last result the terms should be
symmetrised over $b\uc d$ and also truncated to terms of order $\BigO{\eps^m}$. Note that the first term
on the right hand side of this equation vanishes by way of the results described above.

Comparing the order $m$ terms in equation (\ref{eq:GenGammaA}) and (\ref{eq:GenGammaB}) leads to the
following equation
\begin{align}
   \nGammaBar{m}^{a}{}_{b\uc d}
   = \nGammaBar{m}^{a}{}_{(b\uc,d)}
   - (n+1)\left(\m \nGammaBar{m-2}^{a}{}_{p(\uc}
                   \nGammaBar{  2}^{p}{}_{bd)}
                  +\nGammaBar{m-3}^{a}{}_{p(\uc}
                   \nGammaBar{  3}^{p}{}_{bd)}
                  +\nGammaBar{m-4}^{a}{}_{p(\uc}
                   \nGammaBar{  4}^{p}{}_{bd)}
                  + \dots
                  + \nGammaBar{  n+1}^{a}{}_{p(\uc}
                    \nGammaBar{m-n-1}^{p}{}_{bd)}
   \right)
   \label{eq:GenGammaC}
\end{align}
This one equation is all that is needed to compute all of the
$\nGammaBar{p}^{a}{}_{b\uc d}$ for $p=3,4,5,\dots m$ given just
the $\nGammaBar{p}^{a}{}_{bd}$ for $p=2,3,4,\dots m$. For example,
suppose $m=5$ and suppose that we are given
$\nGammaBar{p}^{a}{}_{bd}$ for $p=2,3,4,5$. Then with $n=1$ we can
use equation (\ref{eq:GenGammaC}) to compute in turn,
$\nGammaBar{p}^{a}{}_{bc_1d}$ for $p=3,4,5$. Then with $n=2$ we
compute
$\nGammaBar{p}^{a}{}_{bc_1c_2d}$ for $p=4,5$ and finally with $n=3$
we compute $\nGammaBar{p}^{a}{}_{bc_1c_2c_3d}$ for $p=5$. There
are no terms like
$\nGammaBar{p}^{a}{}_{bc_1c_2c_3c_4d}$ for $p\le5$ due to the
corollary given earlier.

\clearpage

The explicit computations for $m=5$ are as follows.

For $n=1$,
\begin{align}
   \nGammaBar{3}^{a}{}_{bc_1d}
   &=
   \nGammaBar{3}^{a}{}_{(bc_1,d)}\\
   %-----------------------------------------------------------------------
   \nGammaBar{4}^{a}{}_{bc_1d}
   &=
   \nGammaBar{4}^{a}{}_{(bc_1,d)}
   - 2 \nGammaBar{2}^{a}{}_{p(c_1}
       \nGammaBar{2}^{p}{}_{bd)}\\
   %-----------------------------------------------------------------------
   \nGammaBar{5}^{a}{}_{bc_1d}
   &=
   \nGammaBar{5}^{a}{}_{(bc_1,d)}
   - 2 \nGammaBar{3}^{a}{}_{p(c_1}
       \nGammaBar{2}^{p}{}_{bd)}
   - 2 \nGammaBar{2}^{a}{}_{p(c_1}
       \nGammaBar{3}^{p}{}_{bd)}
\end{align}

For $n=2$,
\begin{align}
   \nGammaBar{4}^{a}{}_{bc_1c_2d}
   &=
   \nGammaBar{4}^{a}{}_{(bc_1c_2,d)}\\
   %-----------------------------------------------------------------------
   \nGammaBar{5}^{a}{}_{bc_1c_2d}
   &=
   \nGammaBar{5}^{a}{}_{(bc_1c_2,d)}
   - 3 \nGammaBar{2}^{a}{}_{p(c_1c_2}
       \nGammaBar{2}^{p}{}_{bd)}
\end{align}

For $n=3$,
\begin{align}
   \nGammaBar{5}^{a}{}_{bc_1c_2c_3d}
   &=
   \nGammaBar{5}^{a}{}_{(bc_1c_2c_3,d)}
\end{align}

\clearpage

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,c1,c2,c3,c4,c5,w#}::Indices(position=independent).

   D{#}::Derivative.
   \nabla{#}::Derivative.
   \partial{#}::PartialDerivative.

   g_{a b}::Metric.
   g^{a b}::InverseMetric.
   g_{a}^{b}::KroneckerDelta.
   g^{a}_{b}::KroneckerDelta.
   \delta^{a}_{b}::KroneckerDelta.
   \delta_{a}^{b}::KroneckerDelta.

   R_{a b c d}::RiemannTensor.
   R^{a}_{b c d}::RiemannTensor.
   R_{a b c}^{d}::RiemannTensor.

   \Gamma^{a}_{b c}::TableauSymmetry(shape={2}, indices={1,2}).

   x^{a}::Depends(D{#}).

   g_{a b}::Depends(\partial{#}).
   R_{a b c d}::Depends(\partial{#}).
   R^{a}_{b c d}::Depends(\partial{#}).
   \Gamma^{a}_{b c}::Depends(\partial{#}).

   R_{a b c d}::Depends(\nabla{#}).
   R^{a}_{b c d}::Depends(\nabla{#}).

   import cdblib

   term0 = cdblib.get ('GammaRterm0','connection.json')
   term1 = cdblib.get ('GammaRterm2','connection.json')
   term2 = cdblib.get ('GammaRterm3','connection.json')
   term3 = cdblib.get ('GammaRterm4','connection.json')
   term4 = cdblib.get ('GammaRterm5','connection.json')

   # LCB: these terms were not computed in connection.tex so set them to zero
   #      maybe in the future I will compute down to term6.

   term5 := 0.
   term6 := 0.

   # genGmn : m = eps order of Rabcd terms
   #          n = number of c indices

   # --------------------------------------------------------------------------
   # rules for building the genGmn

   # note: after applying each rule, must symmetrise over (b c1 c2 ... cn d)

   # n = 0

   genG20 := genG2^{a}_{b d}.
   genG30 := genG3^{a}_{b d}.
   genG40 := genG4^{a}_{b d}.
   genG50 := genG5^{a}_{b d}.

   defG20 := genG2^{d}_{a b} -> @(term1).
   defG30 := genG3^{d}_{a b} -> @(term2).
   defG40 := genG4^{d}_{a b} -> @(term3).
   defG50 := genG5^{d}_{a b} -> @(term4).

   # LCB: rncGamma in connection.json limited to "term4" (ie. to 4th order in x)
   #      so can only compute genG3*, genG4* and genG5* (at this stage)
   #      but it doesn't hurt to provide the definitions for genG6*, genG7* etc. we just won't use them (at this atage)

   defG60 := genG6^{d}_{a b} -> @(term5).
   defG70 := genG7^{d}_{a b} -> @(term6).

   # n = 1

   defG31 := genG3^{a}_{b c1 d} -> D_{d}{genG3^{a}_{b c1}}.

   defG41 := genG4^{a}_{b c1 d} -> D_{d}{genG4^{a}_{b c1}}
                                   - 2 genG2^{a}_{p c1} genG2^{p}_{b d}.

   defG51 := genG5^{a}_{b c1 d} -> D_{d}{genG5^{a}_{b c1}}
                                   - 2 genG3^{a}_{p c1} genG2^{p}_{b d}
                                   - 2 genG2^{a}_{p c1} genG3^{p}_{b d}.

   defG61 := genG6^{a}_{b c1 d} -> D_{d}{genG6^{a}_{b c1}}
                                   - 2 genG4^{a}_{p c1} genG2^{p}_{b d}
                                   - 2 genG3^{a}_{p c1} genG3^{p}_{b d}
                                   - 2 genG3^{a}_{p c1} genG4^{p}_{b d}.

   defG71 := genG7^{a}_{b c1 d} -> D_{d}{genG7^{a}_{b c1}}
                                   - 2 genG5^{a}_{p c1} genG2^{p}_{b d}
                                   - 2 genG4^{a}_{p c1} genG3^{p}_{b d}
                                   - 2 genG3^{a}_{p c1} genG4^{p}_{b d}
                                   - 2 genG2^{a}_{p c1} genG5^{p}_{b d}.

   # n = 2

   defG42 := genG4^{a}_{b c1 c2 d} -> D_{d}{genG4^{a}_{b c1 c2}}.

   defG52 := genG5^{a}_{b c1 c2 d} -> D_{d}{genG5^{a}_{b c1 c2}}
                                      - 3 genG3^{a}_{p c1 c2} genG2^{p}_{b d}.

   defG62 := genG6^{a}_{b c1 c2 d} -> D_{d}{genG6^{a}_{b c1 c2}}
                                      - 3 genG4^{a}_{p c1 c2} genG2^{p}_{b d}
                                      - 3 genG3^{a}_{p c1 c2} genG3^{p}_{b d}.

   defG72 := genG7^{a}_{b c1 c2 d} -> D_{d}{genG7^{a}_{b c1 c2}}
                                      - 3 genG5^{a}_{p c1 c2} genG2^{p}_{b d}
                                      - 3 genG4^{a}_{p c1 c2} genG3^{p}_{b d}
                                      - 3 genG3^{a}_{p c1 c2} genG4^{p}_{b d}.

   # n = 3

   defG53 := genG5^{a}_{b c1 c2 c3 d} -> D_{d}{genG5^{a}_{b c1 c2 c3}}.

   defG63 := genG6^{a}_{b c1 c2 c3 d} -> D_{d}{genG6^{a}_{b c1 c2 c3}}
                                         - 4 genG3^{a}_{p c1 c2 c3} genG3^{p}_{b d}.

   defG73 := genG7^{a}_{b c1 c2 c3 d} -> D_{d}{genG7^{a}_{b c1 c2 c3}}
                                         - 4 genG4^{a}_{p c1 c2 c3} genG3^{p}_{b d}
                                         - 4 genG3^{a}_{p c1 c2 c3} genG4^{p}_{b d}.

   # n = 4

   defG64 := genG6^{a}_{b c1 c2 c3 c4 d} -> D_{d}{genG6^{a}_{b c1 c2 c3 c4}}.

   defG74 := genG7^{a}_{b c1 c2 c3 c4 d} -> D_{d}{genG7^{a}_{b c1 c2 c3 c4}}
                                            - 5 genG5^{a}_{p c1 c2 c3 c4} genG2^{p}_{b d}.

   # n = 5

   defG75 := genG7^{a}_{b c1 c2 c3 c4 c5 d} -> D_{d}{genG7^{a}_{b c1 c2 c3 c4 c5}}.

   # --------------------------------------------------------------------------
   # build the genGmn

   # ==========================================================================
   # n = 1

   genG31 := genG3^{a}_{b c1 d}.                              # cdb (genG31.000,genG31)
   genG41 := genG4^{a}_{b c1 d}.                              # cdb (genG41.000,genG41)
   genG51 := genG5^{a}_{b c1 d}.
   # genG61 := genG6^{a}_{b c1 d}.
   # genG71 := genG7^{a}_{b c1 d}.

   # --------------------------------------------------------------------------
   substitute     (genG20,defG20)                             # cdb (genG20.001,genG20)
   substitute     (genG30,defG30)                             # cdb (genG30.001,genG30)
   substitute     (genG40,defG40)                             # cdb (genG40.001,genG40)
   substitute     (genG50,defG50)                             # cdb (genG50.001,genG50)

   # --------------------------------------------------------------------------
   substitute     (genG31,defG31)                             # cdb (genG31.001,genG31)
   substitute     (genG31,defG30)                             # cdb (genG31.002,genG31)

   distribute     (genG31)                                    # cdb (genG31.002,genG31)
   unwrap         (genG31)                                    # cdb (genG31.003,genG31)
   product_rule   (genG31)                                    # cdb (genG31.004,genG31)
   distribute     (genG31)                                    # cdb (genG31.005,genG31)
   substitute     (genG31,$D_{a}{x^b}->\delta_{a}^{b}$)       # cdb (genG31.006,genG31)
   eliminate_kronecker (genG31)                               # cdb (genG31.007,genG31)
   sym            (genG31,$_{b}, _{c1}, _{d}$)
   sort_product   (genG31)                                    # cdb (genG31.008,genG31)
   rename_dummies (genG31)                                    # cdb (genG31.009,genG31)
   canonicalise   (genG31)                                    # cdb (genG31.010,genG31)

   # --------------------------------------------------------------------------
   substitute     (genG41,defG41)                             # cdb (genG41.001,genG41)
   substitute     (genG41,defG40)                             # cdb (genG41.002,genG41)
   substitute     (genG41,defG20,repeat=True)                 # cdb (genG41.003,genG41)

   distribute     (genG41)                                    # cdb (genG41.004,genG41)
   unwrap         (genG41)                                    # cdb (genG41.005,genG41)
   product_rule   (genG41)                                    # cdb (genG41.006,genG41)
   distribute     (genG41)                                    # cdb (genG41.007,genG41)
   substitute     (genG41,$D_{a}{x^b}->\delta_{a}^{b}$)       # cdb (genG41.008,genG41)
   eliminate_kronecker (genG41)                               # cdb (genG41.009,genG41)
   sym            (genG41,$_{b}, _{c1}, _{d}$)
   sort_product   (genG41)                                    # cdb (genG41.010,genG41)
   rename_dummies (genG41)                                    # cdb (genG41.011,genG41)
   canonicalise   (genG41)                                    # cdb (genG41.012,genG41)

   # --------------------------------------------------------------------------
   substitute     (genG51,defG51)
   substitute     (genG51,defG50)
   substitute     (genG51,defG30,repeat=True)
   substitute     (genG51,defG20,repeat=True)

   distribute     (genG51)
   unwrap         (genG51)
   product_rule   (genG51)
   distribute     (genG51)
   substitute     (genG51,$D_{a}{x^b}->\delta_{a}^{b}$)
   eliminate_kronecker (genG51)
   sym            (genG51,$_{b}, _{c1}, _{d}$)
   sort_product   (genG51)
   rename_dummies (genG51)
   canonicalise   (genG51)

   # update the rules

   defG31 := genG3^{a}_{b c1 d} -> @(genG31).
   defG41 := genG4^{a}_{b c1 d} -> @(genG41).
   defG51 := genG5^{a}_{b c1 d} -> @(genG51).

   # ==========================================================================
   # n = 2

   genG42 := genG4^{a}_{b c1 c2 d}.                           # cdb (genG42.000,genG42)
   genG52 := genG5^{a}_{b c1 c2 d}.
   # genG62 := genG6^{a}_{b c1 c2 d}.
   # genG72 := genG7^{a}_{b c1 c2 d}.

   # --------------------------------------------------------------------------
   substitute     (genG42,defG42)                             # cdb (genG42.001,genG42)
   substitute     (genG42,defG41)                             # cdb (genG42.002,genG42)

   distribute     (genG42)                                    # cdb (genG42.003,genG42)
   unwrap         (genG42)                                    # cdb (genG42.004,genG42)
   product_rule   (genG42)                                    # cdb (genG42.005,genG42)
   distribute     (genG42)                                    # cdb (genG42.006,genG42)
   substitute     (genG42,$D_{a}{x^b}->\delta_{a}^{b}$)       # cdb (genG42.007,genG42)
   eliminate_kronecker (genG42)                               # cdb (genG42.008,genG42)
   sym            (genG42,$_{b}, _{c1}, _{c2}, _{d}$)
   sort_product   (genG42)                                    # cdb (genG42.009,genG42)
   rename_dummies (genG42)                                    # cdb (genG42.010,genG42)
   canonicalise   (genG42)                                    # cdb (genG42.011,genG42)

   # --------------------------------------------------------------------------
   substitute     (genG52,defG52)
   substitute     (genG52,defG51)
   substitute     (genG52,defG31,repeat=True)
   substitute     (genG52,defG20,repeat=True)

   distribute     (genG52)
   unwrap         (genG52)
   product_rule   (genG52)
   distribute     (genG52)
   substitute     (genG52,$D_{a}{x^b}->\delta_{a}^{b}$)
   eliminate_kronecker (genG52)
   sym            (genG52,$_{b}, _{c1}, _{c2}, _{d}$)
   sort_product   (genG52)
   rename_dummies (genG52)
   canonicalise   (genG52)                                    # cdb (genG52.001,genG52)

   # update the rules

   defG42 := genG4^{a}_{b c1 c2 d} -> @(genG42).
   defG52 := genG5^{a}_{b c1 c2 d} -> @(genG52).

   # ==========================================================================
   # n = 3

   genG53 := genG5^{a}_{b c1 c2 c3 d}.
   # genG63 := genG6^{a}_{b c1 c2 c3 d}.
   # genG73 := genG7^{a}_{b c1 c2 c3 d}.

   # --------------------------------------------------------------------------
   substitute     (genG53,defG53)
   substitute     (genG53,defG52)

   distribute     (genG53)
   unwrap         (genG53)
   product_rule   (genG53)
   distribute     (genG53)
   substitute     (genG53,$D_{a}{x^b}->\delta_{a}^{b}$)
   eliminate_kronecker (genG53)
   sym            (genG53,$_{b}, _{c1}, _{c2}, _{c3}, _{d}$)
   sort_product   (genG53)
   rename_dummies (genG53)
   canonicalise   (genG53)                                    # cdb (genG53.001,genG53)

   # update the rules

   defG53 := genG5^{a}_{b c1 c2 c3 d} -> @(genG53).

\end{cadabra}

\clearpage

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{genG31.000} \end{dmath*}
   \begin{dmath*} \cdb*{genG31.001} \end{dmath*}
   \begin{dmath*} \cdb*{genG31.002} \end{dmath*}
   \begin{dmath*} \cdb*{genG31.003} \end{dmath*}
   \begin{dmath*} \cdb*{genG31.004} \end{dmath*}
   \begin{dmath*} \cdb*{genG31.005} \end{dmath*}
   \begin{dmath*} \cdb*{genG31.006} \end{dmath*}
   \begin{dmath*} \cdb*{genG31.007} \end{dmath*}
   \begin{dmath*} \cdb*{genG31.008} \end{dmath*}
   \begin{dmath*} \cdb*{genG31.009} \end{dmath*}
   \begin{dmath*} \cdb*{genG31.010} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{genG41.000} \end{dmath*}
   \begin{dmath*} \cdb*{genG41.001} \end{dmath*}
   \begin{dmath*} \cdb*{genG41.002} \end{dmath*}
   \begin{dmath*} \cdb*{genG41.003} \end{dmath*}
   \begin{dmath*} \cdb*{genG41.004} \end{dmath*}
   \begin{dmath*} \cdb*{genG41.005} \end{dmath*}
   \begin{dmath*} \cdb*{genG41.006} \end{dmath*}
   \begin{dmath*} \cdb*{genG41.007} \end{dmath*}
   \begin{dmath*} \cdb*{genG41.008} \end{dmath*}
   \begin{dmath*} \cdb*{genG41.009} \end{dmath*}
   % \begin{dmath*} \cdb*{genG41.010} \end{dmath*}
   % \begin{dmath*} \cdb*{genG41.011} \end{dmath*}
   \begin{dmath*} \cdb*{genG41.012} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{genG42.000} \end{dmath*}
   \begin{dmath*} \cdb*{genG42.001} \end{dmath*}
   \begin{dmath*} \cdb*{genG42.002} \end{dmath*}
   \begin{dmath*} \cdb*{genG42.003} \end{dmath*}
   \begin{dmath*} \cdb*{genG42.004} \end{dmath*}
   \begin{dmath*} \cdb*{genG42.005} \end{dmath*}
   % \begin{dmath*} \cdb*{genG42.006} \end{dmath*}
   % \begin{dmath*} \cdb*{genG42.007} \end{dmath*}
   % \begin{dmath*} \cdb*{genG42.008} \end{dmath*}
   % \begin{dmath*} \cdb*{genG42.009} \end{dmath*}
   % \begin{dmath*} \cdb*{genG42.010} \end{dmath*}
   \begin{dmath*} \cdb*{genG42.011} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{cadabra}
   # note: keeping numbering as is (out of order) to ensure R appears before \nabla R etc.
   def product_sort (obj):
       substitute (obj,$ A^{a}                            -> A001^{a}               $)
       substitute (obj,$ x^{a}                            -> A002^{a}               $)
       substitute (obj,$ g^{a b}                          -> A003^{a b}             $)
       substitute (obj,$ \nabla_{e f g h}{R_{a b c d}}    -> A008_{a b c d e f g h} $)
       substitute (obj,$ \nabla_{e f g}{R_{a b c d}}      -> A007_{a b c d e f g}   $)
       substitute (obj,$ \nabla_{e f}{R_{a b c d}}        -> A006_{a b c d e f}     $)
       substitute (obj,$ \nabla_{e}{R_{a b c d}}          -> A005_{a b c d e}       $)
       substitute (obj,$ R_{a b c d}                      -> A004_{a b c d}         $)
       sort_product   (obj)
       rename_dummies (obj)
       substitute (obj,$ A001^{a}                  -> A^{a}                         $)
       substitute (obj,$ A002^{a}                  -> x^{a}                         $)
       substitute (obj,$ A003^{a b}                -> g^{a b}                       $)
       substitute (obj,$ A004_{a b c d}            -> R_{a b c d}                   $)
       substitute (obj,$ A005_{a b c d e}          -> \nabla_{e}{R_{a b c d}}       $)
       substitute (obj,$ A006_{a b c d e f}        -> \nabla_{e f}{R_{a b c d}}     $)
       substitute (obj,$ A007_{a b c d e f g}      -> \nabla_{e f g}{R_{a b c d}}   $)
       substitute (obj,$ A008_{a b c d e f g h}    -> \nabla_{e f g h}{R_{a b c d}} $)

       return obj

   # --------------------------------------------------------------------------
   symG20 := @(genG20) A^{b} A^{d}.                           # cdb (symG20.100,symG20)

   distribute            (symG20)                             # cdb (symG20.101,symG20)
   symG20 = product_sort (symG20)                             # cdb (symG20.102,symG20)
   rename_dummies        (symG20)                             # cdb (symG20.103,symG20)
   canonicalise          (symG20)                             # cdb (symG20.104,symG20)

   # --------------------------------------------------------------------------
   symG30 := @(genG30) A^{b} A^{d}.                           # cdb (symG30.100,symG30)

   distribute            (symG30)                             # cdb (symG30.101,symG30)
   symG30 = product_sort (symG30)                             # cdb (symG30.102,symG30)
   rename_dummies        (symG30)                             # cdb (symG30.103,symG30)
   canonicalise          (symG30)                             # cdb (symG30.104,symG30)

   # --------------------------------------------------------------------------
   symG40 := @(genG40) A^{b} A^{d}.                           # cdb (symG40.100,symG40)

   distribute            (symG40)                             # cdb (symG40.101,symG40)
   symG40 = product_sort (symG40)                             # cdb (symG40.102,symG40)
   rename_dummies        (symG40)                             # cdb (symG40.103,symG40)
   canonicalise          (symG40)                             # cdb (symG40.104,symG40)

   # --------------------------------------------------------------------------
   symG50 := @(genG50) A^{b} A^{d}.                           # cdb (symG50.100,symG50)

   distribute            (symG50)                             # cdb (symG50.101,symG50)
   symG50 = product_sort (symG50)                             # cdb (symG50.102,symG50)
   rename_dummies        (symG50)                             # cdb (symG50.103,symG50)
   canonicalise          (symG50)                             # cdb (symG50.104,symG50)

   # --------------------------------------------------------------------------
   symG31 := @(genG31) A^{b} A^{c1} A^{d}.                    # cdb (symG31.100,symG31)

   distribute            (symG31)                             # cdb (symG31.101,symG31)
   symG31 = product_sort (symG31)                             # cdb (symG31.102,symG31)
   rename_dummies        (symG31)                             # cdb (symG31.103,symG31)
   canonicalise          (symG31)                             # cdb (symG31.104,symG31)

   # --------------------------------------------------------------------------
   symG41 := @(genG41) A^{b} A^{c1} A^{d}.                    # cdb (symG41.100,symG41)

   distribute            (symG41)                             # cdb (symG41.101,symG41)
   symG41 = product_sort (symG41)                             # cdb (symG41.102,symG41)
   rename_dummies        (symG41)                             # cdb (symG41.103,symG41)
   canonicalise          (symG41)                             # cdb (symG41.104,symG41)

   # --------------------------------------------------------------------------
   symG51 := @(genG51) A^{b} A^{c1} A^{d}.                    # cdb (symG51.100,symG51)

   distribute            (symG51)                             # cdb (symG51.101,symG51)
   symG51 = product_sort (symG51)                             # cdb (symG51.102,symG51)
   rename_dummies        (symG51)                             # cdb (symG51.103,symG51)
   canonicalise          (symG51)                             # cdb (symG51.104,symG51)

   # --------------------------------------------------------------------------
   symG42 := @(genG42) A^{b} A^{c1} A^{c2} A^{d}.             # cdb (symG42.100,symG42)

   distribute            (symG42)                             # cdb (symG42.101,symG42)
   symG42 = product_sort (symG42)                             # cdb (symG42.102,symG42)
   rename_dummies        (symG42)                             # cdb (symG42.103,symG42)
   canonicalise          (symG42)                             # cdb (symG42.104,symG42)

   # --------------------------------------------------------------------------
   symG52 := @(genG52) A^{b} A^{c1} A^{c2} A^{d}.             # cdb (symG52.100,symG52)

   distribute            (symG52)                             # cdb (symG52.101,symG52)
   symG52 = product_sort (symG52)                             # cdb (symG52.102,symG52)
   rename_dummies        (symG52)                             # cdb (symG52.103,symG52)
   canonicalise          (symG52)                             # cdb (symG52.104,symG52)

   # --------------------------------------------------------------------------
   symG53 := @(genG53) A^{b} A^{c1} A^{c2} A^{c3} A^{d}.      # cdb (symG53.100,symG53)

   distribute            (symG53)                             # cdb (symG53.101,symG53)
   symG53 = product_sort (symG53)                             # cdb (symG53.102,symG53)
   rename_dummies        (symG53)                             # cdb (symG53.103,symG53)
   canonicalise          (symG53)                             # cdb (symG53.104,symG53)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{symG31.100} \end{dmath*}
   \begin{dmath*} \cdb*{symG31.101} \end{dmath*}
   \begin{dmath*} \cdb*{symG31.102} \end{dmath*}
   \begin{dmath*} \cdb*{symG31.103} \end{dmath*}
   \begin{dmath*} \cdb*{symG31.104} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{symG41.100} \end{dmath*}
   \begin{dmath*} \cdb*{symG41.101} \end{dmath*}
   \begin{dmath*} \cdb*{symG41.102} \end{dmath*}
   \begin{dmath*} \cdb*{symG41.103} \end{dmath*}
   \begin{dmath*} \cdb*{symG41.104} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{symG51.104} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{symG42.104} \end{dmath*}
   \begin{dmath*} \cdb*{symG52.104} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{symG53.104} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{cadabra}
   def reformat (obj,scale):
       foo  = Ex(str(scale))
       bah := @(foo) @(obj).
       distribute (bah)
       factor_out (bah,$A^{a?},x^{b?}$)
       ans := @(bah) / @(foo).
       return ans

   fooG20 = reformat (symG20,3)
   fooG30 = reformat (symG30,12)
   fooG40 = reformat (symG40,360)
   fooG50 = reformat (symG50,180)

   fooG31 = reformat (symG31,2)
   fooG41 = reformat (symG41,120)
   fooG51 = reformat (symG51,180)

   fooG42 = reformat (symG42,15)
   fooG52 = reformat (symG52,90)

   fooG53 = reformat (symG53,3)

   genGamma0 := @(fooG20) + @(fooG30) + @(fooG40) + @(fooG50).  # cdb (genGamma0.000,genGamma0)
   genGamma1 := @(fooG31) + @(fooG41) + @(fooG51).              # cdb (genGamma1.000,genGamma1)
   genGamma2 := @(fooG42) + @(fooG52).                          # cdb (genGamma2.000,genGamma2)
   genGamma3 := @(fooG53).                                      # cdb (genGamma3.000,genGamma3)

   cdblib.create ('genGamma.json')

   cdblib.put ('genGamma0',genGamma0,'genGamma.json')
   cdblib.put ('genGamma1',genGamma1,'genGamma.json')
   cdblib.put ('genGamma2',genGamma2,'genGamma.json')
   cdblib.put ('genGamma3',genGamma3,'genGamma.json')

   cdblib.put ('genGamma01',fooG20,'genGamma.json')
   cdblib.put ('genGamma02',fooG30,'genGamma.json')
   cdblib.put ('genGamma03',fooG40,'genGamma.json')
   cdblib.put ('genGamma04',fooG50,'genGamma.json')

   cdblib.put ('genGamma11',fooG31,'genGamma.json')
   cdblib.put ('genGamma12',fooG41,'genGamma.json')
   cdblib.put ('genGamma13',fooG51,'genGamma.json')

   cdblib.put ('genGamma21',fooG42,'genGamma.json')
   cdblib.put ('genGamma22',fooG52,'genGamma.json')

   cdblib.put ('genGamma31',fooG53,'genGamma.json')

\end{cadabra}

\clearpage

% =================================================================================================
\section*{The generalised connection in Riemann normal coordinates}

\begin{dgroup*}
   \begin{dmath*} A^b A^c \Gamma^{a}_{b c}(x) = \cdb{genGamma0.000} \end{dmath*}
   \begin{dmath*} A^b A^c A^d \Gamma^{a}_{b c d}(x) = \cdb{genGamma1.000} \end{dmath*}
   \begin{dmath*} A^b A^c A^d A^e \Gamma^{a}_{b c d e}(x) = \cdb{genGamma2.000} \end{dmath*}
   \begin{dmath*} A^b A^c A^d A^e A^f \Gamma^{a}_{b c d e f}(x) = \cdb{genGamma3.000} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{cadabra}
   scaledGamma0 := 360 @(genGamma0).  # cdb (scaledGamma0.001,scaledGamma0)
   scaledGamma1 := 360 @(genGamma1).  # cdb (scaledGamma1.001,scaledGamma1)
   scaledGamma2 :=  90 @(genGamma2).  # cdb (scaledGamma2.001,scaledGamma2)
   scaledGamma3 :=   3 @(genGamma3).  # cdb (scaledGamma3.001,scaledGamma3)

\end{cadabra}

\clearpage

% =================================================================================================
\section*{The generalised connection in Riemann normal coordinates}

This is the same as the previous page but with a small change in the format to avoid fractions.

\begin{dgroup*}
   \begin{dmath*} 360 A^b A^c \Gamma^{a}_{b c}(x) = \cdb{scaledGamma0.001} \end{dmath*}
   \begin{dmath*} 360 A^b A^c A^d \Gamma^{a}_{b c d}(x) = \cdb{scaledGamma1.001} \end{dmath*}
   \begin{dmath*}  90 A^b A^c A^d A^e \Gamma^{a}_{b c d e}(x) = \cdb{scaledGamma2.001} \end{dmath*}
   \begin{dmath*}   3 A^b A^c A^d A^e A^f \Gamma^{a}_{b c d e f}(x) = \cdb{scaledGamma3.001} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
% export selected objects, these will later be imported into a library
% these are the objects that will appear in the paper

\begin{cadabra}
   tmp0 := @(fooG20) + @(fooG30).
   tmp1 := @(fooG31).

   alt0 := @(genGamma0).
   alt1 := @(genGamma1).
   alt2 := @(genGamma2).
   alt3 := @(genGamma3).

   alt0scaled := @(scaledGamma0).
   alt1scaled := @(scaledGamma1).
   alt2scaled := @(scaledGamma2).
   alt3scaled := @(scaledGamma3).

   substitute (tmp0, $A^{a}->1$)
   substitute (tmp1, $A^{a}->1$)

   substitute (alt0, $A^{a}->1$)
   substitute (alt1, $A^{a}->1$)
   substitute (alt2, $A^{a}->1$)
   substitute (alt3, $A^{a}->1$)

   substitute (alt0scaled, $A^{a}->1$)
   substitute (alt1scaled, $A^{a}->1$)
   substitute (alt2scaled, $A^{a}->1$)
   substitute (alt3scaled, $A^{a}->1$)

   cdblib.create ('genGamma.export')

   # 4th order gen gamma
   cdblib.put ('gen_gamma_0_4th',tmp0,'genGamma.export')
   cdblib.put ('gen_gamma_1_4th',tmp1,'genGamma.export')

   # 6th order gen gamma
   cdblib.put ('gen_gamma_0',alt0,'genGamma.export')
   cdblib.put ('gen_gamma_1',alt1,'genGamma.export')
   cdblib.put ('gen_gamma_2',alt2,'genGamma.export')
   cdblib.put ('gen_gamma_3',alt3,'genGamma.export')

   # 6th order gen gamma scaled
   cdblib.put ('gen_gamma_0_scaled',alt0scaled,'genGamma.export')
   cdblib.put ('gen_gamma_1_scaled',alt1scaled,'genGamma.export')
   cdblib.put ('gen_gamma_2_scaled',alt2scaled,'genGamma.export')
   cdblib.put ('gen_gamma_3_scaled',alt3scaled,'genGamma.export')

   checkpoint.append (tmp0)
   checkpoint.append (tmp1)

   checkpoint.append (alt0)
   checkpoint.append (alt1)
   checkpoint.append (alt2)
   checkpoint.append (alt3)

   checkpoint.append (alt0scaled)
   checkpoint.append (alt1scaled)
   checkpoint.append (alt2scaled)
   checkpoint.append (alt3scaled)

\end{cadabra}

% =================================================================================================
% export checkpoints in json format

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   for i in range( len(checkpoint) ):
      cdblib.put ('check{:03d}'.format(i),checkpoint[i],checkpoint_file)
\end{cadabra}
\egroup

\end{document}

### 2019-09-24 source/cadabra/geodesic-ivp.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\lstset{gobble=2}

\begin{document}

% =================================================================================================
% create checkpoint file

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/geodesic-ivp.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{Geodesic IVP}

Our game here is to find the solution of
\begin{align*}
   0 = \frac{d^2 x^{a}}{ds^2} + \Gamma^{a}_{bc}(x) \frac{dx^b}{ds} \frac{dx^c}{ds}
\end{align*}
subject to the initial conditions $x^{a}(s) = x^a$ and $dx^a(s)/ds={\Dot x}^{a}$ at $s=0$.

% =================================================================================================
\section*{Algorithm}

By successive differentiation of the above equation we can compute
\begin{align*}
   \frac{d^n x^{a}}{ds^n} = -\Gamma^{a}_{\udn}\frac{dx^{\udn}}{ds}
\end{align*}
at $s=0$ for $n=2,3,4,\dotsc$. The $\Gamma^{a}_{\udn}$ are the \emph{generalised connections}.

We can then construct the Taylor series solution for $x^{a}(s)$
\begin{align*}
   x^a(s) = x^a + s {\Dot x}^a - \sum_{k=2}^\infty\>\frac{s^{k}}{k!} \Gamma^{a}_{\udk}{\Dot x}^{\udk}
\end{align*}

\clearpage

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   \nabla{#}::Derivative.

   import cdblib

   sterm21 = cdblib.get ('genGamma01','genGamma.json')
   sterm22 = cdblib.get ('genGamma02','genGamma.json')
   sterm23 = cdblib.get ('genGamma03','genGamma.json')
   sterm24 = cdblib.get ('genGamma04','genGamma.json')

   sterm31 = cdblib.get ('genGamma11','genGamma.json')
   sterm32 = cdblib.get ('genGamma12','genGamma.json')
   sterm33 = cdblib.get ('genGamma13','genGamma.json')

   sterm41 = cdblib.get ('genGamma21','genGamma.json')
   sterm42 = cdblib.get ('genGamma22','genGamma.json')

   sterm51 = cdblib.get ('genGamma31','genGamma.json')

   sterm2 := @(sterm21) + @(sterm22) + @(sterm23) + @(sterm24).  # cdb (sterm2.000,sterm2)
   sterm3 := @(sterm31) + @(sterm32) + @(sterm33).               # cdb (sterm3.000,sterm3)
   sterm4 := @(sterm41) + @(sterm42).                            # cdb (sterm4.000,sterm4)
   sterm5 := @(sterm51).                                         # cdb (sterm5.000,sterm5)

   factor_out (sterm2,$A^{a?}$)                                  # cdb (sterm2.001,sterm2)
   factor_out (sterm3,$A^{a?}$)                                  # cdb (sterm3.001,sterm3)
   factor_out (sterm4,$A^{a?}$)                                  # cdb (sterm4.001,sterm4)
   factor_out (sterm5,$A^{a?}$)                                  # cdb (sterm5.001,sterm5)

   sterm2 := 360 @(sterm2).
   sterm3 := 360 @(sterm3).
   sterm4 :=  90 @(sterm4).
   sterm5 :=   3 @(sterm5).

   substitute (sterm2,$A^{a}->1$)                                # cdb (sterm2.002,sterm2)
   substitute (sterm3,$A^{a}->1$)                                # cdb (sterm3.002,sterm3)
   substitute (sterm4,$A^{a}->1$)                                # cdb (sterm4.002,sterm4)
   substitute (sterm5,$A^{a}->1$)                                # cdb (sterm5.002,sterm5)

\end{cadabra}

% =================================================================================================
% the remaining code is just for pretty printing

\clearpage

% =================================================================================================
\section*{The geodesic ivp}

\begin{align*}
   x^{d}(s) = x^{d}
            + s {\dot{x}^d}
            + \frac{s^2}{2!} {\dot{x}^a} {\dot{x}^b} A^{d}_{ab}
            + \frac{s^3}{3!} {\dot{x}^a} {\dot{x}^b} {\dot{x}^c} A^{d}_{abc}
            + \frac{s^4}{4!} {\dot{x}^a} {\dot{x}^b} {\dot{x}^c} {\dot{x}^d} A^{d}_{abce}
            + \frac{s^5}{5!} {\dot{x}^a} {\dot{x}^b} {\dot{x}^c} {\dot{x}^d} {\dot{x}^e} A^{d}_{abcef}
            + \dotsb
\end{align*}
\begin{dgroup*}
   \begin{dmath*} 360 A^{d}_{ab} = \cdb{sterm2.002} \end{dmath*}
   \begin{dmath*} 360 A^{d}_{abc} = \cdb{sterm3.002} \end{dmath*}
   \begin{dmath*}  90 A^{d}_{abce} = \cdb{sterm4.002} \end{dmath*}
   \begin{dmath*}   3 A^{d}_{abcef} = \cdb{sterm5.002} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
% export selected objects, these will later be imported into a library
% these are the objects that will appear in the paper

\begin{cadabra}
   sterm2short := @(sterm21) + @(sterm22).             # cdb (sterm2.short.001,sterm2short)
   sterm3short := @(sterm31).                          # cdb (sterm3.short.001,sterm3short)
   sterm2shortscaled := 12 @(sterm2short).             # cdb (sterm2.short.scaled.002,sterm2shortscaled)
   sterm3shortscaled :=  2 @(sterm3short).             # cdb (sterm3.short.scaled.002,sterm3shortscaled)

   substitute (sterm2shortscaled,$A^{a}->1$)           # cdb (sterm2.short.scaled.003,sterm2shortscaled)
   substitute (sterm3shortscaled,$A^{a}->1$)           # cdb (sterm3.short.scaled.003,sterm3shortscaled)

   cdblib.create ('geodesic-ivp.export')

   # 4th order ivp terms scaled
   cdblib.put ('ivp42',sterm2shortscaled,'geodesic-ivp.export')
   cdblib.put ('ivp43',sterm3shortscaled,'geodesic-ivp.export')

   # 6th order ivp terms scaled
   cdblib.put ('ivp62',sterm2,'geodesic-ivp.export')
   cdblib.put ('ivp63',sterm3,'geodesic-ivp.export')
   cdblib.put ('ivp64',sterm4,'geodesic-ivp.export')
   cdblib.put ('ivp65',sterm5,'geodesic-ivp.export')

   checkpoint.append (sterm2shortscaled)
   checkpoint.append (sterm3shortscaled)

   checkpoint.append (sterm2)
   checkpoint.append (sterm3)
   checkpoint.append (sterm4)
   checkpoint.append (sterm5)
\end{cadabra}

% just to check that we are exporting the correct 4th order terms

\begin{dgroup*}
   \begin{dmath*} \cdb*{sterm2.short.001} \end{dmath*}
   \begin{dmath*} \cdb*{sterm3.short.001} \end{dmath*}
   \begin{dmath*} \cdb*{sterm2.short.scaled.002} \end{dmath*}
   \begin{dmath*} \cdb*{sterm3.short.scaled.002} \end{dmath*}
   \begin{dmath*} \cdb*{sterm2.short.scaled.003} \end{dmath*}
   \begin{dmath*} \cdb*{sterm3.short.scaled.003} \end{dmath*}
\end{dgroup*}

% =================================================================================================
% export checkpoints in json format

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   for i in range( len(checkpoint) ):
      cdblib.put ('check{:03d}'.format(i),checkpoint[i],checkpoint_file)
\end{cadabra}
\egroup

\end{document}

### 2019-09-24 source/cadabra/geodesic-lsq-midpt.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

% =================================================================================================
% create checkpoint file

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/geodesic-lsq-midpt.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{Geodesic mid-point for arc-length}

\input{metric.cdbtex}
\input{geodesic-lsq.cdbtex}

This code uses the results of {\tts geodesic-lsq} and {\tts metric} to show that the 2nd and 3rd
order estimates for $L^2_{PQ}$ can be recovered using a mid-point estimate. For the 3rd order estimate
we have
\begin{align*}
   g_{ab}(x) &= \cdb{gab4.601} + \BigO{\eps^4}\\
   L^2_{PQ}  &= \cdb{lsq4.301} + \BigO{\eps^4}
\end{align*}
The code below verifies that
\begin{equation*}
   L^2_{PQ} = g_{ab}(\bar x) Dx^{a} Dx^{b} + \BigO{\eps^4}
\end{equation*}
where $\bar x$ is the \emph{coordinate} midpoint of the geodesic
\begin{equation*}
   {\bar x^a} = \frac{1}{2}\left(x^{a}_{P} + x^{a}_{Q}\right)
\end{equation*}
This result holds true only for the 2nd and 3rd order estimates. Note that the \emph{coordinate} midpoint
is not the \emph{geometric} midpoint of the geodesic.

It might be interesting to see if the higher order estimates could
be recovered by sampling the metric at points other than the mid point.

\clearpage

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   \nabla{#}::Derivative.

   g_{a b}::Metric.
   g^{a b}::InverseMetric.

   R_{a b c d}::RiemannTensor.

   import cdblib

   gab = cdblib.get('g_ab','metric.json')

   lsq2 = cdblib.get('lsq2','geodesic-lsq.json')
   lsq3 = cdblib.get('lsq3','geodesic-lsq.json')
   lsq4 = cdblib.get('lsq4','geodesic-lsq.json')
   lsq5 = cdblib.get('lsq5','geodesic-lsq.json')

   substitute (gab,$x^{a}->(p^{a}+q^{a})/2$)   # evaluate rnc gab at mid-point
   distribute (gab)

   defgab := g_{a b} -> @(gab).

   mid := g_{a b} (q^{a}-p^{a}) (q^{b}-p^{b}).

   substitute     (mid,defgab)
   distribute     (mid)
   sort_product   (mid)
   rename_dummies (mid)
   canonicalise   (mid)

   tst2 := @(lsq2) - @(mid).                             # cdb (tst2.201,tst2)
   tst3 := @(lsq3) - @(mid).                             # cdb (tst3.201,tst3)
   tst4 := @(lsq4) - @(mid).                             # cdb (tst4.201,tst4)
   tst5 := @(lsq5) - @(mid).                             # cdb (tst5.201,tst5)

   substitute     (tst2,$Dx^{a} -> q^{a}-p^{a}$)
   substitute     (tst2,$x^{a} -> p^{a}$)
   distribute     (tst2)
   sort_product   (tst2)
   rename_dummies (tst2)
   canonicalise   (tst2)                                 # cdb (tst2.202,tst2)

   substitute     (tst3,$Dx^{a} -> q^{a}-p^{a}$)
   substitute     (tst3,$x^{a} -> p^{a}$)
   distribute     (tst3)
   sort_product   (tst3)
   rename_dummies (tst3)
   canonicalise   (tst3)                                 # cdb (tst3.202,tst3)

   substitute     (tst4,$Dx^{a} -> q^{a}-p^{a}$)
   substitute     (tst4,$x^{a} -> p^{a}$)
   distribute     (tst4)
   sort_product   (tst4)
   rename_dummies (tst4)
   canonicalise   (tst4)                                 # cdb (tst4.202,tst4)

   substitute     (tst5,$Dx^{a} -> q^{a}-p^{a}$)
   substitute     (tst5,$x^{a} -> p^{a}$)
   distribute     (tst5)
   sort_product   (tst5)
   rename_dummies (tst5)
   canonicalise   (tst5)                                 # cdb (tst5.202,tst5)

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Reformatting}

\begin{cadabra}
   def truncateR (obj,n):

   # I would like to assign different weights to \nabla_{a}, \nabla_{a b}, \nabla_{a b c} etc. but no matter
   # what I do it appears that Cadabra assigns the same weight to all of these regardless of the number of subscripts.
   # It seems that the weight is assigned to the symbol \nabla alone. So I'm forced to use the following substitution trick.

       Q_{a b c d}::Weight(label=numR,value=2).
       Q_{a b c d e}::Weight(label=numR,value=3).
       Q_{a b c d e f}::Weight(label=numR,value=4).
       Q_{a b c d e f g}::Weight(label=numR,value=5).

       tmp := @(obj).

       substitute (tmp, $\nabla_{e f g}{R_{a b c d}} -> Q_{a b c d e f g}$)
       substitute (tmp, $\nabla_{e f}{R_{a b c d}} -> Q_{a b c d e f}$)
       substitute (tmp, $\nabla_{e}{R_{a b c d}} -> Q_{a b c d e}$)
       substitute (tmp, $R_{a b c d} -> Q_{a b c d}$)

       ans = Ex(0)

       for i in range (0,n+1):
          foo := @(tmp).
          bah = Ex("numR = " + str(i))
          keep_weight (foo, bah)
          ans = ans + foo

       substitute (ans, $Q_{a b c d e f g} -> \nabla_{e f g}{R_{a b c d}}$)
       substitute (ans, $Q_{a b c d e f} -> \nabla_{e f}{R_{a b c d}}$)
       substitute (ans, $Q_{a b c d e} -> \nabla_{e}{R_{a b c d}}$)
       substitute (ans, $Q_{a b c d} -> R_{a b c d}$)

       return ans

   tst2 = truncateR (tst2,2)  # cdb (tst2.301,tst2)
   tst3 = truncateR (tst3,3)  # cdb (tst3.301,tst3)
   tst4 = truncateR (tst4,4)  # cdb (tst4.301,tst4)
   tst5 = truncateR (tst5,5)  # cdb (tst5.301,tst5)
\end{cadabra}

\clearpage

% =================================================================================================
\section*{Errors is mid-point estimates for $L^2_{PQ}$}

\begin{dgroup*}
   \begin{dmath*} \left(L^2_{PQ} - g_{ab}(\bar x) Dx^a Dx^b\right)_2 = \cdb{tst2.301} \end{dmath*}
   \begin{dmath*} \left(L^2_{PQ} - g_{ab}(\bar x) Dx^a Dx^b\right)_3 = \cdb{tst3.301} \end{dmath*}
   \begin{dmath*} \left(L^2_{PQ} - g_{ab}(\bar x) Dx^a Dx^b\right)_4 = \cdb{tst4.301} \end{dmath*}
   \begin{dmath*} \left(L^2_{PQ} - g_{ab}(\bar x) Dx^a Dx^b\right)_5 = \cdb{tst5.301} \end{dmath*}
\end{dgroup*}

% =================================================================================================
% export checkpoints in json format

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   for i in range( len(checkpoint) ):
      cdblib.put ('check{:03d}'.format(i),checkpoint[i],checkpoint_file)
\end{cadabra}
\egroup

\end{document}

### 2019-09-24 source/cadabra/geodesic-lsq.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

% =================================================================================================
% create checkpoint file

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/geodesic-lsq.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{Geodesic arc-length}

Give a pair of points $P$ and $Q$ the geodesic arc-length can be computed using
\begin{align}
   L_{PQ} = \int_P^Q\>\left(g_{ab}(x)\frac{dx^a}{ds}\frac{dx^b}{ds}\right)^{1/2}\>ds
\end{align}
Since the path is a geodesic the integrand is constant and thus
\begin{align}
   L^2_{PQ} = \left.g_{ab}(x)\frac{dx^a}{ds}\frac{dx^b}{ds}\right\vert_{P}
\end{align}
where $s$ is a re-scaled parameter (0 at $P$ and 1 at $Q$). The point $P$ has
RNC coordinates $x^{a}$ while the point $Q$ has coordinates $x^{a} + Dx^{a}$.

The vector $dx^a/ds$ at $P$ is given by the solution of the geodesic boundary value
problem. This was found in the previous code ({\tts geodesic-bvp}). That is
\begin{align}
   \left.\frac{dx^b}{ds}\right\vert_{P} = y^{a}
\end{align}
and thus
\begin{align}
   \label{eq:lsq}
   L^2_{PQ} = g_{ab}(x) y^{a} y^{b}
\end{align}

It is possible to directly evaluate the right hand side of (\ref{eq:lsq}) using the results from
the {\tts geodesic-bvp} and {\tts metric} codes. The result would need to be truncated (to an
order consistent with the results form those codes). But doing so would be computationaly
expensive as at least half of the terms will be thrown away. A better approach is compute just
the terms that will survive the truncation. This is done by expanding $g_{ab}(x)$ and $y^{a}$ as
a truncated series in the curvatures and its derivatives.

The $g_{ab}(x)$ and $y^{a}$ are written in a (truncated) formal power series in the curvature and
its derivatives
\begin{align}
   y^{a} &= \ny{0}^a + \ny{2}^a + \ny{3}^a + \ny{4}^a + \ny{5}^a + \BigO{\eps^6}\\
   g_{a b}(x) &=   \ngab{0}_{a b}
                 + \ngab{2}_{a b}
                 + \ngab{3}_{a b}
                 + \ngab{4}_{a b}
                 + \ngab{5}_{a b}
                 + \BigO{\eps^6}
\end{align}
Note that this use of $\ny{i}$ differs from that used in {\tts geodesic-bvp}. Here the
index above $y^{a}$ denotes a particular term in the curvature expansion while in
{\tts geodesic-bvp} the index denoted the iteration number (in the fixed point scheme
used to solve the BVP for $y^{a}$).

% \clarepage

% =================================================================================================
\section*{Stage 1}
The formal curvature expansions are substituted into equation (\ref{eq:lsq}), expanded and
truncated to retain terms of order $\BigO{\eps^5}$ or less. The expansion to 4th order terms is
as follows.

\begin{dgroup*}
   \begin{dmath*} L^2_{PQ} = \cdb{lsq4.002} \end{dmath*}
\end{dgroup*}

\input{rnc2rnc.cdbtex}
\input{metric.cdbtex}

From {\tts geodesic-bvp} (actually from {\tts rnc2rnc which reformatted the results nicely}) we have
% \begin{align*}
%      y^{a} = \ny{0}^{a} + \ny{2}^{a} + \ny{3}^{a} + \ny{4}^{a} + \BigO{\eps^5}
% \end{align*}

\begin{dgroup*}
   \begin{dmath*} \ny{0}^{a} = \cdb{term0.102} \end{dmath*}
   \begin{dmath*} \ny{2}^{a} = \cdb{term2.102} \end{dmath*}
   \begin{dmath*} \ny{3}^{a} = \cdb{term3.102} \end{dmath*}
   \begin{dmath*} \ny{4}^{a} = \cdb{term4.102} \end{dmath*}
\end{dgroup*}

and from {\tts metric} we have

% \begin{align*}
%      g_{a b}(x) =
%      \ngab{0}_{a b}
%    + \ngab{2}_{a b}
%    + \ngab{3}_{a b}
%    + \ngab{4}_{a b}
%    + \BigO{\eps^6}
% \end{align*}
\begin{dgroup*}
   \begin{dmath*}     \ngab{0}_{a b} = \cdb{scaled0.601} \end{dmath*}
   \begin{dmath*}   3 \ngab{2}_{a b} = \cdb{scaled2.601} \end{dmath*}
   \begin{dmath*}   6 \ngab{3}_{a b} = \cdb{scaled3.601} \end{dmath*}
   \begin{dmath*} 180 \ngab{4}_{a b} = \cdb{scaled4.601} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 2}
The results from the {\tts geodesic-bvp} and {\tts metric} codes are read to provide
values for the $\ny{n}^{a}$ and $\ngab{m}_{ab}$. These are substituted into the result from
Stage 1, et volia, the final answer. To 4th-order terms the result is given by

\Dmath*{ L^2_{PQ} = \cdb{lsq5.301} + \BigO{\eps^5}}

\clearpage

% =================================================================================================
\section*{Shared properties}

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   D{#}::Derivative.
   \nabla{#}::Derivative.
   \partial{#}::PartialDerivative.

   g_{a b}::Metric.
   g^{a b}::InverseMetric.
   g_{a}^{b}::KroneckerDelta.
   g^{a}_{b}::KroneckerDelta.
   \delta^{a}_{b}::KroneckerDelta.
   \delta_{a}^{b}::KroneckerDelta.

   R_{a b c d}::RiemannTensor.
   R^{a}_{b c d}::RiemannTensor.
   R_{a b c}^{d}::RiemannTensor.

   \Gamma^{a}_{b c}::TableauSymmetry(shape={2}, indices={1,2}).
   \Gamma^{a}_{b c d}::TableauSymmetry(shape={3}, indices={1,2,3}).
   \Gamma^{a}_{b c d e}::TableauSymmetry(shape={4}, indices={1,2,3,4}).
   \Gamma^{a}_{b c d e f}::TableauSymmetry(shape={5}, indices={1,2,3,4,5}).

   x^{a}::Depends(D{#}).

   g_{a b}::Depends(\partial{#}).
   R_{a b c d}::Depends(\partial{#}).
   R^{a}_{b c d}::Depends(\partial{#}).
   \Gamma^{a}_{b c}::Depends(\partial{#}).

   R_{a b c d}::Depends(\nabla{#}).
   R^{a}_{b c d}::Depends(\nabla{#}).

   g0{#}::LaTeXForm ("\ngab{0}").
   g2{#}::LaTeXForm ("\ngab{2}").
   g3{#}::LaTeXForm ("\ngab{3}").
   g4{#}::LaTeXForm ("\ngab{4}").
   g5{#}::LaTeXForm ("\ngab{5}").

   y0{#}::LaTeXForm ("\ny{0}").
   y2{#}::LaTeXForm ("\ny{2}").
   y3{#}::LaTeXForm ("\ny{3}").
   y4{#}::LaTeXForm ("\ny{4}").
   y5{#}::LaTeXForm ("\ny{5}").

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Stage 1: The formal expansion}

\begin{cadabra}
   g0_{a b}::Symmetric.
   g2_{a b}::Symmetric.
   g3_{a b}::Symmetric.
   g4_{a b}::Symmetric.
   g5_{a b}::Symmetric.

   g0_{a b}::Weight(label=num,value=0).
   g2_{a b}::Weight(label=num,value=2).
   g3_{a b}::Weight(label=num,value=3).
   g4_{a b}::Weight(label=num,value=4).
   g5_{a b}::Weight(label=num,value=5).

   y0^{a}::Weight(label=num,value=0).
   y2^{a}::Weight(label=num,value=2).
   y3^{a}::Weight(label=num,value=3).
   y4^{a}::Weight(label=num,value=4).
   y5^{a}::Weight(label=num,value=5).

   # note: keeping numbering as is (out of order) to ensure R appears before \nabla R etc.
   def product_sort (obj):
       substitute (obj,$ A^{a}                            -> A001^{a}               $)
       substitute (obj,$ x^{a}                            -> A002^{a}               $)
       substitute (obj,$ Dx^{a}                           -> A003^{a}               $)
       substitute (obj,$ g_{a b}                          -> A004_{a b}             $)
       substitute (obj,$ g^{a b}                          -> A005^{a b}             $)
       substitute (obj,$ \nabla_{e f g h}{R_{a b c d}}    -> A010_{a b c d e f g h} $)
       substitute (obj,$ \nabla_{e f g}{R_{a b c d}}      -> A009_{a b c d e f g}   $)
       substitute (obj,$ \nabla_{e f}{R_{a b c d}}        -> A008_{a b c d e f}     $)
       substitute (obj,$ \nabla_{e}{R_{a b c d}}          -> A007_{a b c d e}       $)
       substitute (obj,$ R_{a b c d}                      -> A006_{a b c d}         $)
       sort_product   (obj)
       rename_dummies (obj)
       substitute (obj,$ A001^{a}                  -> A^{a}                         $)
       substitute (obj,$ A002^{a}                  -> x^{a}                         $)
       substitute (obj,$ A003^{a}                  -> Dx^{a}                        $)
       substitute (obj,$ A004_{a b}                -> g_{a b}                       $)
       substitute (obj,$ A005^{a b}                -> g^{a b}                       $)
       substitute (obj,$ A006_{a b c d}            -> R_{a b c d}                   $)
       substitute (obj,$ A007_{a b c d e}          -> \nabla_{e}{R_{a b c d}}       $)
       substitute (obj,$ A008_{a b c d e f}        -> \nabla_{e f}{R_{a b c d}}     $)
       substitute (obj,$ A009_{a b c d e f g}      -> \nabla_{e f g}{R_{a b c d}}   $)
       substitute (obj,$ A010_{a b c d e f g h}    -> \nabla_{e f g h}{R_{a b c d}} $)

       return obj

   def truncate (obj,n):
       ans = Ex(0)

       for i in range (0,n+1):
          foo := @(obj).
          bah = Ex("num = " + str(i))
          keep_weight (foo, bah)
          ans = ans + foo

       return ans

   # expansions wrt the curvature

   defgab := g_{a b} -> g0_{a b} + g2_{a b} + g3_{a b} + g4_{a b} + g5_{a b}.
   defy   := y^{a}   -> y0^{a} + y2^{a} + y3^{a} + y4^{a} + y5^{a}.

   lsq    := g_{a b} y^{a} y^{b}.

   substitute (lsq,defgab)
   substitute (lsq,defy)
   distribute (lsq)

   def tidy (obj):
       foo := @(obj).
       sort_product    (foo)
       rename_dummies  (foo)
       canonicalise    (foo)
       return foo

   lsq0 = tidy ( truncate (lsq,0) )  # cdb (lsq0.002,lsq0)
   lsq2 = tidy ( truncate (lsq,2) )  # cdb (lsq2.002,lsq2)
   lsq3 = tidy ( truncate (lsq,3) )  # cdb (lsq3.002,lsq3)
   lsq4 = tidy ( truncate (lsq,4) )  # cdb (lsq4.002,lsq4)
   lsq5 = tidy ( truncate (lsq,5) )  # cdb (lsq5.002,lsq5)

   d20 := @(lsq2) - @(lsq0).         # cdb (d20.001,d20)   # check, should contain only O(2) terms
   d32 := @(lsq3) - @(lsq2).         # cdb (d32.001,d32)   # check, should contain only O(3) terms
   d43 := @(lsq4) - @(lsq3).         # cdb (d43.001,d43)   # check, should contain only O(4) terms
   d54 := @(lsq5) - @(lsq4).         # cdb (d54.001,d54)   # check, should contain only O(5) terms

   d5 := @(lsq5) - @(lsq).           # cdb (d5.001,d5)
   d5  = tidy (d5)                   # cdb (d5.002,d5)  # all higher order terms, should see no O(5) terms

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{lsq0.002} \end{dmath*}
   \begin{dmath*} \cdb*{lsq2.002} \end{dmath*}
   \begin{dmath*} \cdb*{lsq3.002} \end{dmath*}
   \begin{dmath*} \cdb*{lsq4.002} \end{dmath*}
   \begin{dmath*} \cdb*{lsq5.002} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{d20.001} \end{dmath*}
   \begin{dmath*} \cdb*{d32.001} \end{dmath*}
   \begin{dmath*} \cdb*{d43.001} \end{dmath*}
   \begin{dmath*} \cdb*{d54.001} \end{dmath*}
   \begin{dmath*} \cdb*{d5.002} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 2: Substution of $\ny{n}^{a}$ and $\ngab{m}_{ab}$}

\begin{cadabra}
   import cdblib

   g0ab = cdblib.get('g_ab_0','metric.json')
   g2ab = cdblib.get('g_ab_2','metric.json')
   g3ab = cdblib.get('g_ab_3','metric.json')
   g4ab = cdblib.get('g_ab_4','metric.json')
   g5ab = cdblib.get('g_ab_5','metric.json')

   defg0ab := g0_{a b} -> @(g0ab).
   defg2ab := g2_{a b} -> @(g2ab).
   defg3ab := g3_{a b} -> @(g3ab).
   defg4ab := g4_{a b} -> @(g4ab).
   defg5ab := g5_{a b} -> @(g5ab).

   y0a = cdblib.get('y50','geodesic-bvp.json')
   y2a = cdblib.get('y52','geodesic-bvp.json')
   y3a = cdblib.get('y53','geodesic-bvp.json')
   y4a = cdblib.get('y54','geodesic-bvp.json')
   y5a = cdblib.get('y55','geodesic-bvp.json')

   defy0a := y0^{a} -> @(y0a).
   defy2a := y2^{a} -> @(y2a).
   defy3a := y3^{a} -> @(y3a).
   defy4a := y4^{a} -> @(y4a).
   defy5a := y5^{a} -> @(y5a).

   def substitute_gab_ya (obj):

      foo := @(obj).

      substitute (foo,defg0ab)
      substitute (foo,defg2ab)
      substitute (foo,defg3ab)
      substitute (foo,defg4ab)
      substitute (foo,defg5ab)

      substitute (foo,defy0a)
      substitute (foo,defy2a)
      substitute (foo,defy3a)
      substitute (foo,defy4a)
      substitute (foo,defy5a)

      distribute     (foo)
      sort_product   (foo)
      rename_dummies (foo)
      canonicalise   (foo)

      substitute     (foo,$g_{a b} g^{c b} -> \delta^{c}_{a}$)
      eliminate_kronecker (foo)
      foo = product_sort  (foo)
      rename_dummies      (foo)
      canonicalise        (foo)

      return foo

   def get_Rterm (obj,n):

   # I would like to assign different weights to \nabla_{a}, \nabla_{a b}, \nabla_{a b c} etc. but no matter
   # what I do it appears that Cadabra assigns the same weight to all of these regardless of the number of subscripts.
   # It seems that the weight is assigned to the symbol \nabla alone. So I'm forced to use the following substitution trick.

       Q_{a b c d}::Weight(label=numR,value=2).
       Q_{a b c d e}::Weight(label=numR,value=3).
       Q_{a b c d e f}::Weight(label=numR,value=4).
       Q_{a b c d e f g}::Weight(label=numR,value=5).

       tmp := @(obj).

       distribute (tmp)

       substitute (tmp, $\nabla_{e f g}{R_{a b c d}} -> Q_{a b c d e f g}$)
       substitute (tmp, $\nabla_{e f}{R_{a b c d}} -> Q_{a b c d e f}$)
       substitute (tmp, $\nabla_{e}{R_{a b c d}} -> Q_{a b c d e}$)
       substitute (tmp, $R_{a b c d} -> Q_{a b c d}$)

       foo := @(tmp).
       bah = Ex("numR = " + str(n))
       keep_weight (foo, bah)

       substitute (foo, $Q_{a b c d e f g} -> \nabla_{e f g}{R_{a b c d}}$)
       substitute (foo, $Q_{a b c d e f} -> \nabla_{e f}{R_{a b c d}}$)
       substitute (foo, $Q_{a b c d e} -> \nabla_{e}{R_{a b c d}}$)
       substitute (foo, $Q_{a b c d} -> R_{a b c d}$)

       return foo

   lsq2 = substitute_gab_ya (lsq2)  # cdb (lsq2.101,lsq2)
   lsq3 = substitute_gab_ya (lsq3)  # cdb (lsq3.101,lsq3)
   lsq4 = substitute_gab_ya (lsq4)  # cdb (lsq4.101,lsq4)
   lsq5 = substitute_gab_ya (lsq5)  # cdb (lsq5.101,lsq5)

   lsq50 = get_Rterm (lsq5,0)
   lsq52 = get_Rterm (lsq5,2)
   lsq53 = get_Rterm (lsq5,3)
   lsq54 = get_Rterm (lsq5,4)
   lsq55 = get_Rterm (lsq5,5)

   cdblib.create ('geodesic-lsq.json')

   cdblib.put ('lsq2',lsq2,'geodesic-lsq.json')
   cdblib.put ('lsq3',lsq3,'geodesic-lsq.json')
   cdblib.put ('lsq4',lsq4,'geodesic-lsq.json')
   cdblib.put ('lsq5',lsq5,'geodesic-lsq.json')

   cdblib.put ('lsq50',lsq50,'geodesic-lsq.json')
   cdblib.put ('lsq52',lsq52,'geodesic-lsq.json')
   cdblib.put ('lsq53',lsq53,'geodesic-lsq.json')
   cdblib.put ('lsq54',lsq54,'geodesic-lsq.json')
   cdblib.put ('lsq55',lsq55,'geodesic-lsq.json')

\end{cadabra}

\clearpage
\begin{dgroup*}
   \begin{dmath*} \cdb*{lsq2.101} \end{dmath*}
   \begin{dmath*} \cdb*{lsq3.101} \end{dmath*}
   \begin{dmath*} \cdb*{lsq4.101} \end{dmath*}
   \begin{dmath*} \cdb*{lsq5.101} \end{dmath*}
\end{dgroup*}

% =================================================================================================
% the remaining code is just for pretty printing

\clearpage

% =================================================================================================
\section*{Stage 3: Reformatting}

\begin{cadabra}
   def reformat (obj,scale):
      foo  = Ex(str(scale))
      bah := @(foo) @(obj).
      distribute     (bah)
      bah = product_sort (bah)
      rename_dummies (bah)
      canonicalise   (bah)
      substitute     (bah,$Dx^{b}->zzz^{b}$)
      factor_out     (bah,$x^{a?},zzz^{b?}$)
      substitute     (bah,$zzz^{b}->Dx^{b}$)
      ans := @(bah) / @(foo).
      return ans

   def rescale (obj,scale):
      foo  = Ex(str(scale))
      bah := @(foo) @(obj).
      distribute  (bah)
      substitute  (bah,$Dx^{b}->zzz^{b}$)
      factor_out  (bah,$x^{a?},zzz^{b?}$)
      substitute  (bah,$zzz^{b}->Dx^{b}$)
      return bah

   Rterm0 := @(lsq50).
   Rterm2 := @(lsq52).
   Rterm3 := @(lsq53).
   Rterm4 := @(lsq54).
   Rterm5 := @(lsq55).

   Rterm0 = reformat (Rterm0,   1)    # cdb(Rterm0.301,Rterm0) # LCB: returns Dx before g, not what I want
   Rterm2 = reformat (Rterm2,   3)    # cdb(Rterm2.301,Rterm2)
   Rterm3 = reformat (Rterm3,  12)    # cdb(Rterm3.301,Rterm3)
   Rterm4 = reformat (Rterm4, 360)    # cdb(Rterm4.301,Rterm4)
   Rterm5 = reformat (Rterm5,1080)    # cdb(Rterm5.301,Rterm5)

   Rterm0 := g_{a b} Dx^{a} Dx^{b}.   # LCB: fixes the order of terms, g before Dx,

   lsq3 := @(Rterm0) + @(Rterm2).                                      # cdb (lsq4.301,lsq3)
   lsq4 := @(Rterm0) + @(Rterm2) + @(Rterm3).                          # cdb (lsq4.301,lsq4)
   lsq5 := @(Rterm0) + @(Rterm2) + @(Rterm3) + @(Rterm4).              # cdb (lsq5.301,lsq5)
   lsq6 := @(Rterm0) + @(Rterm2) + @(Rterm3) + @(Rterm4) + @(Rterm5).  # cdb (lsq5.301,lsq6)

   lsq  := @(lsq6).                   # cdb (lsq.301,lsq)

   scaled0 = rescale (Rterm0,    1)   # cdb (scaled0.301,scaled0) # LCB: returns Dx before g, not what I want
   scaled2 = rescale (Rterm2,    3)   # cdb (scaled2.301,scaled2)
   scaled3 = rescale (Rterm3,   12)   # cdb (scaled3.301,scaled3)
   scaled4 = rescale (Rterm4,  360)   # cdb (scaled4.301,scaled4)
   scaled5 = rescale (Rterm5, 1080)   # cdb (scaled5.301,scaled5)

   scaled0 := g_{a b} Dx^{a} Dx^{b}.  # cdb (scaled0.301,scaled0) # LCB: fixes the order of terms, g before Dx, good

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Geodesic arc-length}

\begin{dgroup*}[spread=5pt]
   % LCB: which of these is correct?
   % \begin{dmath*} \left(\Delta s\right)^2 = \cdb{lsq.301} + \BigO{\eps^6,Dx^6} \end{dmath*}
   % \begin{dmath*} \left(\Delta s\right)^2 = \cdb{lsq.301} + \BigO{\eps^6,Dx^7} \end{dmath*}
   \begin{dmath*} \left(\Delta s\right)^2 = \cdb{lsq.301} + \BigO{\eps^6} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Geodesic arc-length curvature expansion}

\begin{align*}
   % LCB: which of these is correct?
   % \left(\Delta s\right)^2 = \nD{0} + \nD{2} + \nD{3} + \nD{4} + \nD{5} + \BigO{\eps^6,Dx^6}
   % \left(\Delta s\right)^2 = \nD{0} + \nD{2} + \nD{3} + \nD{4} + \nD{5} + \BigO{\eps^6,Dx^7}
   \left(\Delta s\right)^2 = \nD{0} + \nD{2} + \nD{3} + \nD{4} + \nD{5} + \BigO{\eps^6}
\end{align*}

\begin{dgroup*}[spread=5pt]
   \begin{dmath*}      \nD{0} = \cdb{scaled0.301} \end{dmath*}
   \begin{dmath*}    3 \nD{2} = \cdb{scaled2.301} \end{dmath*}
   \begin{dmath*}   12 \nD{3} = \cdb{scaled3.301} \end{dmath*}
   \begin{dmath*}  360 \nD{4} = \cdb{scaled4.301} \end{dmath*}
   \begin{dmath*} 1080 \nD{5} = \cdb{scaled5.301} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
% export selected objects, these will later be imported into a library
% these are the objects that will appear in the paper

\begin{cadabra}
   cdblib.create ('geodesic-lsq.export')

   # 3rd to 6th order lsq
   cdblib.put ('lsq3',lsq3,'geodesic-lsq.export')
   cdblib.put ('lsq4',lsq4,'geodesic-lsq.export')
   cdblib.put ('lsq5',lsq5,'geodesic-lsq.export')
   cdblib.put ('lsq6',lsq6,'geodesic-lsq.export')

   # 6th order lsq terms, scaled
   cdblib.put ('lsq60',scaled0,'geodesic-lsq.export')
   cdblib.put ('lsq62',scaled2,'geodesic-lsq.export')
   cdblib.put ('lsq63',scaled3,'geodesic-lsq.export')
   cdblib.put ('lsq64',scaled4,'geodesic-lsq.export')
   cdblib.put ('lsq65',scaled5,'geodesic-lsq.export')

   checkpoint.append (lsq4)

   checkpoint.append (scaled0)
   checkpoint.append (scaled2)
   checkpoint.append (scaled3)
   checkpoint.append (scaled4)
   checkpoint.append (scaled5)
\end{cadabra}

% =================================================================================================
% export checkpoints in json format

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   for i in range( len(checkpoint) ):
      cdblib.put ('check{:03d}'.format(i),checkpoint[i],checkpoint_file)
\end{cadabra}
\egroup

\end{document}

### 2019-09-24 source/cadabra/metric-inv.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

% =================================================================================================
% create checkpoint file

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/metric-inv.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{The inverse metric tensor in Riemann normal coordinates}

Here we calculate the Riemann normal expansion of the inverse metric, $g^{ab}$, by developing
the recursive sequences
\begin{align}
\label{eq:pdgab}
g^{ab}{}_{,d\ue} &= - \left(g^{cb}\Gamma^{a}{}_{c(d}\right){}_{,\ue)}
                    - \left(g^{ac}\Gamma^{b}{}_{c(d}\right){}_{,\ue)}\\[10pt]
\label{eq:pdGamma}
(n+3)\Gamma^a{}_{d(b,c\ue)} &= (n+1)\left(R^a{}_{(bc\Dot d,\ue)} - \left(\Gamma^a{}_{f(c}\Gamma^f{}_{b{\Dot d}}\right){}_{,\ue)}\right)
\end{align}
for $n=1,2,3,\cdots$. Note in these equations that the (extended) index $\ue$ contains $n$
normal indices.

We then construct a Taylor series for the metric using
\begin{dmath*}[spread=5pt]
g^{ab}(x) = g^{ab} + g^{ab}{}_{,c}x^c + \frac{1}{2!} g^{ab}{}_{,cd}x^cx^d + \frac{1}{3!} g^{ab}{}_{,cde}x^cx^dx^e + \cdots
          = g^{ab} + \sum_{n=1}^\infty\> \frac{1}{n!}\>g^{ab}{}_{,\uc}\>x^{.\uc}
\end{dmath*}

% =================================================================================================
\section*{Stage 1: Symmetrised partial derivatives of $g^{ab}$}

In this stage, equation (\ref{eq:pdgab}) is used to express the symmetrised partial derivatives
of the metric in terms of the symmetrised partial derivatives of the connection.

\begin{dgroup*}
   \begin{dmath*} g^{ab}{}_{,c} A^{c} = \cdb{term1.200} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cd} A^{c} A^{d} = \cdb{term2.200} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cde} A^{c} A^{d} A^{e} = \cdb{term3.200} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 2: Replace derivatives of $\Gamma$ with partial derivs of $R$}

Now we use the results from {\verb|dGamma|} to replace derivatives of $\Gamma$ with
partial derivatives of $R$. These were computed in {\verb|dGamma|} using equation
(\ref{eq:pdGamma}) above.

\begin{dgroup*}
   \begin{dmath*} g^{ab}{}_{,c} A^{c} = \cdb{term1.200} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cd} A^{c} A^{d} = \cdb{term2.303} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cde} A^{c} A^{d} A^{e} = \cdb{term3.305} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 3: Replace partial derivs of $R$ with covariant derivs of $R$}

Next we use the results from {\verb|dRabcd|} to replace the partial derivatives of $R$ with
covariant deriavtives.

\begin{dgroup*}
   \begin{dmath*} g^{ab}{}_{,c} A^{c} = \cdb{term1.404} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cd} A^{c} A^{d} = \cdb{term2.404} \end{dmath*}
   \begin{dmath*} g^{ab}{}_{,cde} A^{c} A^{d} A^{e} = \cdb{term3.403} \end{dmath*}
\end{dgroup*}

% =================================================================================================
\section*{Stage 4: Build the Taylor series for $g_{ab}$, reformatting and output}

Each of the above expressions constitutues one term in the Taylor series for the metric.
We also make the trivial change $A\rightarrow x$. Then we do some trivial reformatting.

\begin{align*}
   g_{ab}(x) &=   g^{ab}
                + g^{ab}{}_{,c} x^c
                + \frac{1}{2!} g^{ab}{}_{,cd} x^c x^d
                + \frac{1}{3!} g^{ab}{}_{,cde} x^c x^d x^e +  \BigO{\eps^4}\\
             &= \cdb{metric4.501} + \BigO{\eps^4}
\end{align*}

\clearpage

% =================================================================================================
\section*{Shared properties}

\begin{cadabra}
   import time

   def flatten_Rabcd (obj):
       substitute (obj,$R^{a}_{b c d}   -> g^{a e} R_{e b c d}$)
       substitute (obj,$R_{a}^{b}_{c d} -> g^{b e} R_{a e c d}$)
       substitute (obj,$R_{a b}^{c}_{b} -> g^{c e} R_{a b e d}$)
       substitute (obj,$R_{a b c}^{d}   -> g^{d e} R_{a b c e}$)
       unwrap     (obj)
       sort_product   (obj)
       rename_dummies (obj)
       return obj

   def impose_rnc (obj):
       # hide the derivatives of Gamma
       substitute (obj,$\partial_{d}{\Gamma^{a}_{b c}} -> zzz_{d}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e}{\Gamma^{a}_{b c}} -> zzz_{d e}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f}{\Gamma^{a}_{b c}} -> zzz_{d e f}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f g}{\Gamma^{a}_{b c}} -> zzz_{d e f g}^{a}_{b c}$,repeat=True)
       substitute (obj,$\partial_{d e f g h}{\Gamma^{a}_{b c}} -> zzz_{d e f g h}^{a}_{b c}$,repeat=True)
       # set Gamma to zero
       substitute (obj,$\Gamma^{a}_{b c} -> 0$,repeat=True)
       # recover the derivatives Gamma
       substitute (obj,$zzz_{d}^{a}_{b c} -> \partial_{d}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e}^{a}_{b c} -> \partial_{d e}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f}^{a}_{b c} -> \partial_{d e f}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f g}^{a}_{b c} -> \partial_{d e f g}{\Gamma^{a}_{b c}}$,repeat=True)
       substitute (obj,$zzz_{d e f g h}^{a}_{b c} -> \partial_{d e f g h}{\Gamma^{a}_{b c}}$,repeat=True)
       return obj

   def get_xterm (obj,n):

       x^{a}::Weight(label=numx).

       foo := @(obj).
       bah  = Ex("numx = " + str(n))
       keep_weight (foo,bah)

       return foo

   # note: keeping numbering as is (out of order) to ensure R appears before \nabla R etc.
   def product_sort (obj):
       substitute (obj,$ A^{a}                             -> A001^{a}                  $)
       substitute (obj,$ x^{a}                             -> A002^{a}                  $)
       substitute (obj,$ g_{a b}                           -> A003_{a b}                $)
       substitute (obj,$ g^{a b}                           -> A004^{a b}                $)
       substitute (obj,$ \nabla_{e f g h}{R_{a b c d}}     -> A010_{a b c d e f g h}    $)
       substitute (obj,$ \nabla_{e f g}{R_{a b c d}}       -> A009_{a b c d e f g}      $)
       substitute (obj,$ \nabla_{e f}{R_{a b c d}}         -> A008_{a b c d e f}        $)
       substitute (obj,$ \nabla_{e}{R_{a b c d}}           -> A007_{a b c d e}          $)
       substitute (obj,$ \partial_{e f g h}{R_{a b c d}}   -> A014_{a b c d e f g h}    $)
       substitute (obj,$ \partial_{e f g}{R_{a b c d}}     -> A013_{a b c d e f g}      $)
       substitute (obj,$ \partial_{e f}{R_{a b c d}}       -> A012_{a b c d e f}        $)
       substitute (obj,$ \partial_{e}{R_{a b c d}}         -> A011_{a b c d e}          $)
       substitute (obj,$ \partial_{e f g h}{R^{a}_{b c d}} -> A018^{a}_{b c d e f g h}  $)
       substitute (obj,$ \partial_{e f g}{R^{a}_{b c d}}   -> A017^{a}_{b c d e f g}    $)
       substitute (obj,$ \partial_{e f}{R^{a}_{b c d}}     -> A016^{a}_{b c d e f}      $)
       substitute (obj,$ \partial_{e}{R^{a}_{b c d}}       -> A015^{a}_{b c d e}        $)
       substitute (obj,$ R_{a b c d}                       -> A005_{a b c d}            $)
       substitute (obj,$ R^{a}_{b c d}                     -> A006^{a}_{b c d}          $)
       sort_product   (obj)
       rename_dummies (obj)
       substitute (obj,$ A001^{a}                  -> A^{a}                             $)
       substitute (obj,$ A002^{a}                  -> x^{a}                             $)
       substitute (obj,$ A003_{a b}                -> g_{a b}                           $)
       substitute (obj,$ A004^{a b}                -> g^{a b}                           $)
       substitute (obj,$ A005_{a b c d}            -> R_{a b c d}                       $)
       substitute (obj,$ A006^{a}_{b c d}          -> R^{a}_{b c d}                     $)
       substitute (obj,$ A007_{a b c d e}          -> \nabla_{e}{R_{a b c d}}           $)
       substitute (obj,$ A008_{a b c d e f}        -> \nabla_{e f}{R_{a b c d}}         $)
       substitute (obj,$ A009_{a b c d e f g}      -> \nabla_{e f g}{R_{a b c d}}       $)
       substitute (obj,$ A010_{a b c d e f g h}    -> \nabla_{e f g h}{R_{a b c d}}     $)
       substitute (obj,$ A011_{a b c d e}          -> \partial_{e}{R_{a b c d}}         $)
       substitute (obj,$ A012_{a b c d e f}        -> \partial_{e f}{R_{a b c d}}       $)
       substitute (obj,$ A013_{a b c d e f g}      -> \partial_{e f g}{R_{a b c d}}     $)
       substitute (obj,$ A014_{a b c d e f g h}    -> \partial_{e f g h}{R_{a b c d}}   $)
       substitute (obj,$ A015^{a}_{b c d e}        -> \partial_{e}{R^{a}_{b c d}}       $)
       substitute (obj,$ A016^{a}_{b c d e f}      -> \partial_{e f}{R^{a}_{b c d}}     $)
       substitute (obj,$ A017^{a}_{b c d e f g}    -> \partial_{e f g}{R^{a}_{b c d}}   $)
       substitute (obj,$ A018^{a}_{b c d e f g h}  -> \partial_{e f g h}{R^{a}_{b c d}} $)

       return obj

   def reformat_xterm (obj,scale):
      foo  = Ex(str(scale))
      bah := @(foo) @(obj).
      distribute     (bah)
      bah = product_sort (bah)
      rename_dummies (bah)
      canonicalise   (bah)
      factor_out     (bah,$x^{a?}$)
      ans := @(bah) / @(foo).
      return ans

   def rescale_xterm (obj,scale):
      foo  = Ex(str(scale))
      bah := @(foo) @(obj).
      distribute  (bah)
      factor_out  (bah,$x^{a?}$)
      return bah

   def add_tags (obj,tag):
      n = 0
      ans = Ex('0')
      for i in obj.top().terms():
         foo = obj[i]
         bah = Ex(tag+'_{'+str(n)+'}')
         ans := @(ans) + @(bah) @(foo).
         n = n + 1
      return ans

   def clear_tags (obj,tag):
      ans := @(obj).
      foo  = Ex(tag+'_{a?} -> 1')
      substitute (ans,foo)
      return ans

   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   \nabla{#}::Derivative.
   \partial{#}::PartialDerivative.

   g_{a b}::Metric.
   g^{a b}::InverseMetric.
   g_{a}^{b}::KroneckerDelta.
   g^{a}_{b}::KroneckerDelta.

   R_{a b c d}::RiemannTensor.
   R^{a}_{b c d}::RiemannTensor.
   R_{a b c}^{d}::RiemannTensor.

   \Gamma^{a}_{b c}::TableauSymmetry(shape={2}, indices={1,2}).

   g_{a b}::Depends(\partial{#}).
   R_{a b c d}::Depends(\partial{#}).
   R^{a}_{b c d}::Depends(\partial{#}).
   \Gamma^{a}_{b c}::Depends(\partial{#}).

   R_{a b c d}::Depends(\nabla{#}).
   R^{a}_{b c d}::Depends(\nabla{#}).

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Stage 1: Symmetrised partial derivatives of $g^{ab}$}

\begin{cadabra}
   beg_stage_1 = time.time()

   # symmetrised partial derivatives of g^{ab}

   gab00:=g^{a b}.                                              # cdb (gab00.101,gab00)

   gab01:= - g^{c b}\Gamma^{a}_{c d} - g^{a c}\Gamma^{b}_{c d}. # cdb (gab01.101,gab01)

   gab02:=\partial_{e}{ @(gab01) }.                             # cdb (gab02.101,gab02)
   distribute   (gab02)                                         # cdb (gab02.102,gab02)
   product_rule (gab02)                                         # cdb (gab02.103,gab02)
   substitute   (gab02, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab02.104,gab02)
   distribute   (gab02)                                         # cdb (gab02.105,gab02)

   gab03:=\partial_{f}{ @(gab02) }.                             # cdb (gab03.101,gab03)
   distribute   (gab03)                                         # cdb (gab03.102,gab03)
   product_rule (gab03)                                         # cdb (gab03.103,gab03)
   substitute   (gab03, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab03.104,gab03)
   distribute   (gab03)                                         # cdb (gab03.105,gab03)

   gab04:=\partial_{g}{ @(gab03) }.                             # cdb (gab04.101,gab04)
   distribute   (gab04)                                         # cdb (gab04.102,gab04)
   product_rule (gab04)                                         # cdb (gab04.103,gab04)
   substitute   (gab04, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab04.104,gab04)
   distribute   (gab04)                                         # cdb (gab04.105,gab04)

   gab05:=\partial_{h}{ @(gab04) }.                             # cdb (gab05.101,gab05)
   distribute   (gab05)                                         # cdb (gab05.102,gab05)
   product_rule (gab05)                                         # cdb (gab05.103,gab05)
   substitute   (gab05, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab05.104,gab05)
   distribute   (gab05)                                         # cdb (gab05.105,gab05)

   gab00 = impose_rnc (gab00)   # cdb (gab00.102,gab00)
   gab01 = impose_rnc (gab01)   # cdb (gab01.102,gab01)
   gab02 = impose_rnc (gab02)   # cdb (gab02.106,gab02)
   gab03 = impose_rnc (gab03)   # cdb (gab03.106,gab03)
   gab04 = impose_rnc (gab04)   # cdb (gab04.106,gab04)
   gab05 = impose_rnc (gab05)   # cdb (gab05.106,gab05)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{gab00.101} \end{dmath*}
   \begin{dmath*} \cdb*{gab00.102} \end{dmath*}
   \begin{dmath*} \cdb*{gab01.101} \end{dmath*}
   \begin{dmath*} \cdb*{gab01.102} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{gab02.101} \end{dmath*}
   \begin{dmath*} \cdb*{gab02.102} \end{dmath*}
   \begin{dmath*} \cdb*{gab02.103} \end{dmath*}
   \begin{dmath*} \cdb*{gab02.104} \end{dmath*}
   \begin{dmath*} \cdb*{gab02.105} \end{dmath*}
   \begin{dmath*} \cdb*{gab02.106} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{gab03.101} \end{dmath*}
   \begin{dmath*} \cdb*{gab03.102} \end{dmath*}
   \begin{dmath*} \cdb*{gab03.103} \end{dmath*}
   \begin{dmath*} \cdb*{gab03.104} \end{dmath*}
   \begin{dmath*} \cdb*{gab03.105} \end{dmath*}
   \begin{dmath*} \cdb*{gab03.106} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{gab04.101} \end{dmath*}
   \begin{dmath*} \cdb*{gab04.102} \end{dmath*}
   \begin{dmath*} \cdb*{gab04.103} \end{dmath*}
   \begin{dmath*} \cdb*{gab04.104} \end{dmath*}
   \begin{dmath*} \cdb*{gab04.105} \end{dmath*}
   \begin{dmath*} \cdb*{gab04.106} \end{dmath*}
\end{dgroup*}

\begin{cadabra}
   # prepare first six terms in the Taylor series expansion of g^{ab}(x)

   term0:= @(gab00).
   distribute (term0)                             # cdb(term0.200,term0)

   term1:= @(gab01) A^d.
   distribute (term1)                             # cdb(term1.200,term1)

   term2:= @(gab02) A^d A^e.
   distribute (term2)                             # cdb(term2.200,term2)

   term3:= @(gab03) A^d A^e A^f.
   distribute (term3)                             # cdb(term3.200,term3)

   term4:= @(gab04) A^d A^e A^f A^g.
   distribute (term4)                             # cdb(term4.200,term4)

   term5:= @(gab05) A^d A^e A^f A^g A^h.
   distribute (term5)                             # cdb(term5.200,term5)

   end_stage_1 = time.time()
\end{cadabra}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term0.200} \end{dmath*}
   \begin{dmath*} \cdb*{term1.200} \end{dmath*}
   \begin{dmath*} \cdb*{term2.200} \end{dmath*}
   \begin{dmath*} \cdb*{term3.200} \end{dmath*}
   % \begin{dmath*} \cdb*{term4.200} \end{dmath*}
   % \begin{dmath*} \cdb*{term5.200} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 2: Replace derivatives of $\Gamma$ with partial derivs of $R$}

\begin{cadabra}
   import cdblib

   beg_stage_2 = time.time()

   dGamma01 = cdblib.get ('dGamma01','dGamma.json')  # cdb(dGamma01.300,dGamma01)
   dGamma02 = cdblib.get ('dGamma02','dGamma.json')  # cdb(dGamma02.300,dGamma02)
   dGamma03 = cdblib.get ('dGamma03','dGamma.json')  # cdb(dGamma03.300,dGamma03)
   dGamma04 = cdblib.get ('dGamma04','dGamma.json')  # cdb(dGamma04.300,dGamma04)
   dGamma05 = cdblib.get ('dGamma05','dGamma.json')  # cdb(dGamma05.300,dGamma05)

   # replace partial derivs of \Gamma with products and derivs of Riemann tensor

   substitute (term2,$\partial_{c}{\Gamma^{a}_{b d}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term2.301,term2)
   substitute (term2,$\partial_{c}{\Gamma^{a}_{d b}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term2.302,term2)
   distribute (term2)                                                                                             # cdb(term2.303,term2)

   substitute (term3,$\partial_{c e}{\Gamma^{a}_{d b}}A^{c}A^{b}A^{e} -> @(dGamma02)$,repeat=True)                # cdb(term3.301,term3)
   substitute (term3,$\partial_{c e}{\Gamma^{a}_{b d}}A^{c}A^{b}A^{e} -> @(dGamma02)$,repeat=True)                # cdb(term3.302,term3)
   substitute (term3,$\partial_{c}{\Gamma^{a}_{b d}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term3.303,term3)
   substitute (term3,$\partial_{c}{\Gamma^{a}_{d b}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term3.304,term3)
   distribute (term3)                                                                                             # cdb(term3.305,term3)

   substitute (term4,$\partial_{c e f}{\Gamma^{a}_{d b}}A^{c}A^{b}A^{e}A^{f} -> @(dGamma03)$,repeat=True)         # cdb(term4.301,term4)
   substitute (term4,$\partial_{c e f}{\Gamma^{a}_{b d}}A^{c}A^{b}A^{e}A^{f} -> @(dGamma03)$,repeat=True)         # cdb(term4.302,term4)
   substitute (term4,$\partial_{c e}{\Gamma^{a}_{d b}}A^{c}A^{b}A^{e} -> @(dGamma02)$,repeat=True)                # cdb(term4.303,term4)
   substitute (term4,$\partial_{c e}{\Gamma^{a}_{b d}}A^{c}A^{b}A^{e} -> @(dGamma02)$,repeat=True)                # cdb(term4.304,term4)
   substitute (term4,$\partial_{c}{\Gamma^{a}_{b d}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term4.305,term4)
   substitute (term4,$\partial_{c}{\Gamma^{a}_{d b}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term4.306,term4)
   distribute (term4)                                                                                             # cdb(term4.307,term4)

   substitute (term5,$\partial_{c e f g}{\Gamma^{a}_{d b}}A^{c}A^{b}A^{e}A^{f}A^{g} -> @(dGamma04)$,repeat=True)  # cdb(term5.301,term5)
   substitute (term5,$\partial_{c e f g}{\Gamma^{a}_{b d}}A^{c}A^{b}A^{e}A^{f}A^{g} -> @(dGamma04)$,repeat=True)  # cdb(term5.302,term5)
   substitute (term5,$\partial_{c e f}{\Gamma^{a}_{d b}}A^{c}A^{b}A^{e}A^{f} -> @(dGamma03)$,repeat=True)         # cdb(term5.303,term5)
   substitute (term5,$\partial_{c e f}{\Gamma^{a}_{b d}}A^{c}A^{b}A^{e}A^{f} -> @(dGamma03)$,repeat=True)         # cdb(term5.304,term5)
   substitute (term5,$\partial_{c e}{\Gamma^{a}_{d b}}A^{c}A^{b}A^{e} -> @(dGamma02)$,repeat=True)                # cdb(term5.305,term5)
   substitute (term5,$\partial_{c e}{\Gamma^{a}_{b d}}A^{c}A^{b}A^{e} -> @(dGamma02)$,repeat=True)                # cdb(term5.306,term5)
   substitute (term5,$\partial_{c}{\Gamma^{a}_{b d}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term5.307,term5)
   substitute (term5,$\partial_{c}{\Gamma^{a}_{d b}}A^{c}A^{b} -> @(dGamma01)$,repeat=True)                       # cdb(term5.308,term5)
   distribute (term5)                                                                                             # cdb(term5.309,term5)

   # ------------------------------------------------------------------------------------
   # this block only produces formatted output, it is not part of the main computation
   # ------------------------------------------------------------------------------------

   # the metric in terms of partial derivatives of Rabcd

   metric:=@(term0)
         + (1/1) @(term1)
         + (1/2) @(term2)
         + (1/6) @(term3)
         + (1/24) @(term4)
         + (1/120) @(term5).  # cdb(metric.301,metric)

   substitute (metric,$A^{a} -> x^{a}$)  # cdb (metric.302,metric)

   # reformat and tidy up

   Xterm0 := @(term0).
   Xterm1 := (1/1) @(term1).     # zero
   Xterm2 := (1/2) @(term2).
   Xterm3 := (1/6) @(term3).
   Xterm4 := (1/24) @(term4).
   Xterm5 := (1/120) @(term5).

   substitute (Xterm0,$A^{a} -> x^{a}$)
   substitute (Xterm1,$A^{a} -> x^{a}$)
   substitute (Xterm2,$A^{a} -> x^{a}$)
   substitute (Xterm3,$A^{a} -> x^{a}$)
   substitute (Xterm4,$A^{a} -> x^{a}$)
   substitute (Xterm5,$A^{a} -> x^{a}$)

   # Manipulating these expressions is hampered by the presence of the partial derivative on Rabcd.
   # Thus we can't freely rasie/lower indices on the dRabcd terms. But we can do so on the first
   # derivatives (since these are evaluated at x=0 where the connection vanishes).

   substitute       (Xterm2,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm2.301,Xterm2)
   substitute       (Xterm3,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm3.301,Xterm3)
   substitute       (Xterm4,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm4.301,Xterm4)
   substitute       (Xterm5,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm5.301,Xterm5)

   substitute       (Xterm2,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm2.302,Xterm2)
   substitute       (Xterm3,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm3.302,Xterm3)
   substitute       (Xterm4,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm4.302,Xterm4)
   substitute       (Xterm5,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm5.302,Xterm5)

   substitute       (Xterm2,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm2.303,Xterm2)
   substitute       (Xterm3,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm3.303,Xterm3)
   substitute       (Xterm4,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm4.303,Xterm4)
   substitute       (Xterm5,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm5.303,Xterm5)

   substitute       (Xterm2,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm2.304,Xterm2)
   substitute       (Xterm3,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm3.304,Xterm3)
   substitute       (Xterm4,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm4.304,Xterm4)
   substitute       (Xterm5,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm5.304,Xterm5)

   sort_product     (Xterm2)  # cdb(Xterm2.305,Xterm2)
   sort_product     (Xterm3)  # cdb(Xterm3.305,Xterm3)
   sort_product     (Xterm4)  # cdb(Xterm4.305,Xterm4)
   sort_product     (Xterm5)  # cdb(Xterm5.305,Xterm5)

   rename_dummies   (Xterm2)  # cdb(Xterm2.306,Xterm2)
   rename_dummies   (Xterm3)  # cdb(Xterm3.306,Xterm3)
   rename_dummies   (Xterm4)  # cdb(Xterm4.306,Xterm4)
   rename_dummies   (Xterm5)  # cdb(Xterm5.306,Xterm5)

   canonicalise     (Xterm2)  # cdb(Xterm2.307,Xterm2)
   canonicalise     (Xterm3)  # cdb(Xterm3.307,Xterm3)
   canonicalise     (Xterm4)  # cdb(Xterm4.307,Xterm4)
   canonicalise     (Xterm5)  # cdb(Xterm5.307,Xterm5)

   # We can simplify Xterm2 and Xterm3 by careful juggling of the indices (swapping free indices on selected terms)

   tmp = add_tags (Xterm2,'\\mu')       # cdb (tmp.001,tmp)
   zoom (tmp, $\mu_{1} Q??$)            # cdb (tmp.002,tmp)
   substitute (tmp, $R^{b}_{c d}^{a} x^{c} x^{d} -> R^{a}_{c d}^{b} x^{c} x^{d}$)  # cdb (tmp.003,tmp)
   unzoom (tmp)
   Xterm2 = clear_tags (tmp,'\\mu')     # cdb (Xterm2.401,Xterm2)

   tmp = add_tags (Xterm3,'\\mu')       # cdb (tmp.011,tmp)
   zoom (tmp, $\mu_{1} Q??$)            # cdb (tmp.012,tmp)
   substitute (tmp, $\partial_{c}{R^{b}_{d e}^{a}} x^{c} x^{d} x^{e} -> \partial_{c}{R^{a}_{d e}^{b}} x^{c} x^{d} x^{e}$)  # cdb (tmp.013,tmp)
   unzoom (tmp)
   Xterm3 = clear_tags (tmp,'\\mu')     # cdb (Xterm3.401,Xterm3)

   Xterm0 = reformat_xterm (Xterm0,  1)    # cdb(Xterm0.308,Xterm0)
   Xterm2 = reformat_xterm (Xterm2,  3)    # cdb(Xterm2.308,Xterm2)
   Xterm3 = reformat_xterm (Xterm3,  6)    # cdb(Xterm3.308,Xterm3)
   Xterm4 = reformat_xterm (Xterm4,360)    # cdb(Xterm4.308,Xterm4)
   Xterm5 = reformat_xterm (Xterm5,360)    # cdb(Xterm5.308,Xterm5)

   # metric to 4th and 6th order terms in powers of x^a

   Metric3 := @(Xterm0) + @(Xterm2).                                      # cdb (Metric3.301,Metric3)
   Metric4 := @(Xterm0) + @(Xterm2) + @(Xterm3).                          # cdb (Metric4.301,Metric4)
   Metric5 := @(Xterm0) + @(Xterm2) + @(Xterm3) + @(Xterm4).              # cdb (Metric5.301,Metric5)
   Metric6 := @(Xterm0) + @(Xterm2) + @(Xterm3) + @(Xterm4) + @(Xterm5).  # cdb (Metric6.301,Metric6)

   # ------------------------------------------------------------------------------------
   # end of format block
   # ------------------------------------------------------------------------------------

   end_stage_2 = time.time()
\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{term2.301} \end{dmath*}
   \begin{dmath*} \cdb*{term2.302} \end{dmath*}
   \begin{dmath*} \cdb*{term2.303} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term3.301} \end{dmath*}
   \begin{dmath*} \cdb*{term3.302} \end{dmath*}
   \begin{dmath*} \cdb*{term3.303} \end{dmath*}
   \begin{dmath*} \cdb*{term3.304} \end{dmath*}
   \begin{dmath*} \cdb*{term3.305} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term4.301} \end{dmath*}
   \begin{dmath*} \cdb*{term4.302} \end{dmath*}
   \begin{dmath*} \cdb*{term4.303} \end{dmath*}
   \begin{dmath*} \cdb*{term4.304} \end{dmath*}
   \begin{dmath*} \cdb*{term4.305} \end{dmath*}
   \begin{dmath*} \cdb*{term4.306} \end{dmath*}
   \begin{dmath*} \cdb*{term4.307} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} g^{ab}(x) = \cdb{Metric3.301} \end{dmath*}
   \begin{dmath*} g^{ab}(x) = \cdb{Metric4.301} \end{dmath*}
   \begin{dmath*} g^{ab}(x) = \cdb{Metric5.301} \end{dmath*}
   \begin{dmath*} g^{ab}(x) = \cdb{Metric6.301} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 3: Replace partial derivs of $R$ with covariant derivs of $R$}

\begin{cadabra}
   beg_stage_3 = time.time()

   # now convert partial derivs of Rabcd to covariant derivs

   dRabcd01 = cdblib.get ('dRabcd01','dRabcd.json')  # cdb(dRabcd01.400,dRabcd01)
   dRabcd02 = cdblib.get ('dRabcd02','dRabcd.json')  # cdb(dRabcd02.400,dRabcd02)
   dRabcd03 = cdblib.get ('dRabcd03','dRabcd.json')  # cdb(dRabcd03.400,dRabcd03)

   # term1 & term2 need no special care, just a bit of tidying

   eliminate_metric (term1)   # cdb(term1.401,term1)
   sort_product     (term1)   # cdb(term1.402,term1)
   rename_dummies   (term1)   # cdb(term1.403,term1)
   canonicalise     (term1)   # cdb(term1.404,term1)

   eliminate_metric (term2)   # cdb(term2.401,term2)
   sort_product     (term2)   # cdb(term2.402,term2)
   rename_dummies   (term2)   # cdb(term2.403,term2)
   canonicalise     (term2)   # cdb(term2.404,term2)

   # replace partial derivatives of Riemann tensor in term3, term4 etc. with covariant derivatives of Rabcd

   tmp01 := @(dRabcd01).      # cdb(tmp01.403,tmp01)
   tmp02 := @(dRabcd02).      # cdb(tmp02.403,tmp02)
   tmp03 := @(dRabcd03).      # cdb(tmp03.403,tmp03)

   substitute (term3,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c d b}} ->   @(tmp01)$,repeat=True)         # cdb(term3.401,term3)
   substitute (term3,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c b d}} -> - @(tmp01)$,repeat=True)         # cdb(term3.402,term3)
   distribute (term3)                                                                                # cdb(term3.403,term3)

   substitute (term4,$A^{c}A^{d}A^{e}A^{f}\partial_{e f}{R^{a}_{c d b}} ->   @(tmp02)$,repeat=True)  # cdb(term4.401,term4)
   substitute (term4,$A^{c}A^{d}A^{e}A^{f}\partial_{e f}{R^{a}_{c b d}} -> - @(tmp02)$,repeat=True)  # cdb(term4.402,term4)
   substitute (term4,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c d b}} ->   @(tmp01)$,repeat=True)         # cdb(term4.403,term4)
   substitute (term4,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c b d}} -> - @(tmp01)$,repeat=True)         # cdb(term4.404,term4)
   distribute (term4)                                                                                # cdb(term4.405,term4)

   substitute (term5,$A^{c}A^{d}A^{e}A^{f}A^{g}\partial_{e f g}{R^{a}_{c d b}} ->   @(tmp03)$,repeat=True)
   substitute (term5,$A^{c}A^{d}A^{e}A^{f}A^{g}\partial_{e f g}{R^{a}_{c b d}} -> - @(tmp03)$,repeat=True)
   substitute (term5,$A^{c}A^{d}A^{e}A^{f}\partial_{e f}{R^{a}_{c d b}} ->   @(tmp02)$,repeat=True)
   substitute (term5,$A^{c}A^{d}A^{e}A^{f}\partial_{e f}{R^{a}_{c b d}} -> - @(tmp02)$,repeat=True)
   substitute (term5,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c d b}} ->   @(tmp01)$,repeat=True)
   substitute (term5,$A^{c}A^{d}A^{e}\partial_{e}{R^{a}_{c b d}} -> - @(tmp01)$,repeat=True)
   distribute (term5)

   end_stage_3 = time.time()
\end{cadabra}

\begin{dgroup*}
   \begin{dmath*} \cdb*{tmp01.403} \end{dmath*}
   \begin{dmath*} \cdb*{tmp02.403} \end{dmath*}
   \begin{dmath*} \cdb*{tmp03.403} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{term1.401} \end{dmath*}
   \begin{dmath*} \cdb*{term1.402} \end{dmath*}
   \begin{dmath*} \cdb*{term1.403} \end{dmath*}
   \begin{dmath*} \cdb*{term1.404} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term2.401} \end{dmath*}
   \begin{dmath*} \cdb*{term2.402} \end{dmath*}
   \begin{dmath*} \cdb*{term2.403} \end{dmath*}
   \begin{dmath*} \cdb*{term2.404} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term3.401} \end{dmath*}
   \begin{dmath*} \cdb*{term3.402} \end{dmath*}
   \begin{dmath*} \cdb*{term3.403} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{term4.401} \end{dmath*}
   \begin{dmath*} \cdb*{term4.402} \end{dmath*}
   \begin{dmath*} \cdb*{term4.403} \end{dmath*}
   \begin{dmath*} \cdb*{term4.404} \end{dmath*}
   \begin{dmath*} \cdb*{term4.405} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Stage 4: Build the Taylor series for $g_{ab}$, reformatting and output}

\begin{cadabra}
   beg_stage_4 = time.time()

   # final housekeeping

   # lower the ^{ab} indices to _{uv}

   tmp0 := g_{a u} g_{b v} @(term0).
   tmp1 := g_{a u} g_{b v} @(term1).
   tmp2 := g_{a u} g_{b v} @(term2).
   tmp3 := g_{a u} g_{b v} @(term3).
   tmp4 := g_{a u} g_{b v} @(term4).
   tmp5 := g_{a u} g_{b v} @(term5).

   distribute           (tmp1)  # cdb(tmp1.501,tmp1)
   eliminate_metric     (tmp1)  # cdb(tmp1.502,tmp1)
   eliminate_kronecker  (tmp1)  # cdb(tmp1.503,tmp1)
   tmp1 = flatten_Rabcd (tmp1)
   canonicalise         (tmp1)  # cdb(tmp1.506,tmp1)

   distribute           (tmp2)  # cdb(tmp2.501,tmp2)
   eliminate_metric     (tmp2)  # cdb(tmp2.502,tmp2)
   eliminate_kronecker  (tmp2)  # cdb(tmp2.503,tmp2)
   tmp2 = flatten_Rabcd (tmp2)
   canonicalise         (tmp2)  # cdb(tmp2.506,tmp2)

   distribute           (tmp3)  # cdb(tmp3.501,tmp3)
   eliminate_metric     (tmp3)  # cdb(tmp3.502,tmp3)
   eliminate_kronecker  (tmp3)  # cdb(tmp3.503,tmp3)
   tmp3 = flatten_Rabcd (tmp3)
   canonicalise         (tmp3)  # cdb(tmp3.506,tmp3)

   distribute           (tmp4)  # cdb(tmp4.501,tmp4)
   eliminate_metric     (tmp4)  # cdb(tmp4.502,tmp4)
   eliminate_kronecker  (tmp4)  # cdb(tmp4.503,tmp4)
   tmp4 = flatten_Rabcd (tmp4)
   canonicalise         (tmp4)  # cdb(tmp4.506,tmp4)

   distribute           (tmp5)  # cdb(tmp5.501,tmp5)
   eliminate_metric     (tmp5)  # cdb(tmp5.502,tmp5)
   eliminate_kronecker  (tmp5)  # cdb(tmp5.503,tmp5)
   tmp5 = flatten_Rabcd (tmp5)
   canonicalise         (tmp5)  # cdb(tmp5.506,tmp5)

   # this is out final answer

   # raise the _{uv} indices to ^{ab}

   metric:= g^{a u} g^{b v} (  @(tmp0)
                              + (1/1) @(tmp1)
                              + (1/2) @(tmp2)
                              + (1/6) @(tmp3)
                              + (1/24) @(tmp4)
                              + (1/120) @(tmp5) ).  # cdb(metric.500,metric)

   distribute             (metric)   # cdb(metric.501,metric)
   eliminate_metric       (metric)   # cdb(metric.502,metric)
   eliminate_kronecker    (metric)   # cdb(metric.503,metric)
   metric = flatten_Rabcd (metric)   # cdb(metric.504,metric)
   canonicalise           (metric)   # cdb(metric.505,metric)

   substitute          (metric,$g_{a b} g^{b c} -> g_{a}^{c}$)
   substitute          (metric,$g_{b a} g^{b c} -> g_{a}^{c}$)
   substitute          (metric,$g_{b a} g^{c b} -> g_{a}^{c}$)
   substitute          (metric,$g_{a b} g^{c b} -> g_{a}^{c}$)
   eliminate_kronecker (metric)   # cdb(metric.506,metric)
   canonicalise        (metric)   # cdb(metric.507,metric)

   substitute (metric,$A^{a} -> x^{a}$)  # cdb (metric.508,metric)

   cdblib.create ('metric-inv.json')

   cdblib.put ('g^ab',metric,'metric-inv.json')

   # extract the terms of the metric in powers of x

   term0 = get_xterm (metric,0)   # cdb(term0.501,term0)
   term1 = get_xterm (metric,1)   # cdb(term1.501,term1)
   term2 = get_xterm (metric,2)   # cdb(term2.501,term2)
   term3 = get_xterm (metric,3)   # cdb(term3.501,term3)
   term4 = get_xterm (metric,4)   # cdb(term4.501,term4)
   term5 = get_xterm (metric,5)   # cdb(term5.501,term5)

   cdblib.put ('g^ab_0',term0,'metric-inv.json')
   cdblib.put ('g^ab_1',term1,'metric-inv.json')
   cdblib.put ('g^ab_2',term2,'metric-inv.json')
   cdblib.put ('g^ab_3',term3,'metric-inv.json')
   cdblib.put ('g^ab_4',term4,'metric-inv.json')
   cdblib.put ('g^ab_5',term5,'metric-inv.json')

   # this version of "metric" is used only in the commentary at the start of this notebook

   metric4:=@(term0) + @(term1) + @(term2) + @(term3).  # cdb(metric4.501,metric4)

\end{cadabra}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{term0.501} \end{dmath*}
   \begin{dmath*} \cdb*{term1.501} \end{dmath*}
   \begin{dmath*} \cdb*{term2.501} \end{dmath*}
   \begin{dmath*} \cdb*{term3.501} \end{dmath*}
   \begin{dmath*} \cdb*{term4.501} \end{dmath*}
   \begin{dmath*} \cdb*{term5.501} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{tmp2.501} \end{dmath*}
   \begin{dmath*} \cdb*{tmp2.502} \end{dmath*}
   \begin{dmath*} \cdb*{tmp2.503} \end{dmath*}
   \begin{dmath*} \cdb*{tmp2.506} \end{dmath*}
\end{dgroup*}

\begin{dgroup*}
   \begin{dmath*} \cdb*{tmp3.501} \end{dmath*}
   \begin{dmath*} \cdb*{tmp3.502} \end{dmath*}
   \begin{dmath*} \cdb*{tmp3.503} \end{dmath*}
   \begin{dmath*} \cdb*{tmp3.506} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{tmp4.501} \end{dmath*}
   \begin{dmath*} \cdb*{tmp4.502} \end{dmath*}
   \begin{dmath*} \cdb*{tmp4.503} \end{dmath*}
   \begin{dmath*} \cdb*{tmp4.506} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{tmp5.506} \end{dmath*}
\end{dgroup*}

\clearpage

\begin{dgroup*}
   \begin{dmath*} \cdb*{metric.500} \end{dmath*}
   \begin{dmath*} \cdb*{metric.501} \end{dmath*}
   \begin{dmath*} \cdb*{metric.502} \end{dmath*}
   \begin{dmath*} \cdb*{metric.503} \end{dmath*}
   \begin{dmath*} \cdb*{metric.504} \end{dmath*}
   \begin{dmath*} \cdb*{metric.505} \end{dmath*}
   \begin{dmath*} \cdb*{metric.506} \end{dmath*}
   \begin{dmath*} \cdb*{metric.507} \end{dmath*}
   \begin{dmath*} \cdb*{metric.508} \end{dmath*}
\end{dgroup*}

% =================================================================================================
% the remaining code is just for pretty printing

\clearpage

\begin{cadabra}
   Rterm0 := @(term0).
   Rterm1 := @(term1).  # zero
   Rterm2 := @(term2).
   Rterm3 := @(term3).
   Rterm4 := @(term4).
   Rterm5 := @(term5).

   Rterm0 = reformat_xterm (Rterm0,  1)    # cdb(Rterm0.601,Rterm0)
   Rterm2 = reformat_xterm (Rterm2,  3)    # cdb(Rterm2.601,Rterm2)
   Rterm3 = reformat_xterm (Rterm3,  6)    # cdb(Rterm3.601,Rterm3)
   Rterm4 = reformat_xterm (Rterm4, 60)    # cdb(Rterm4.601,Rterm4)
   Rterm5 = reformat_xterm (Rterm5, 90)    # cdb(Rterm5.601,Rterm5)

   Metric := @(Rterm0) + @(Rterm2) + @(Rterm3) + @(Rterm4) + @(Rterm5).  # cdb (Metric.601,Metric)

   scaled0 = rescale_xterm (Rterm0,  1)    # cdb(scaled0.601,scaled0)
   scaled2 = rescale_xterm (Rterm2,  3)    # cdb(scaled2.601,scaled2)
   scaled3 = rescale_xterm (Rterm3,  6)    # cdb(scaled3.601,scaled3)
   scaled4 = rescale_xterm (Rterm4, 60)    # cdb(scaled4.601,scaled4)
   scaled5 = rescale_xterm (Rterm5, 90)    # cdb(scaled5.601,scaled5)

   end_stage_4 = time.time()
\end{cadabra}

\clearpage

% =================================================================================================
\section*{The inverse metric in Riemann normal coordinates}

\begin{dgroup*}
   \begin{dmath*} g^{a b}(x) = \cdb{Metric.601}+\BigO{\eps^6} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Curvature expansion of the inverse metric}
\begin{align*}
     g^{a b}(x) =
     \ngab{0}^{a b}
   + \ngab{2}^{a b}
   + \ngab{3}^{a b}
   + \ngab{4}^{a b}
   + \ngab{5}^{a b}+\BigO{\eps^6}
\end{align*}
\begin{dgroup*}
   \begin{dmath*}    \ngab{0}^{a b} = \cdb{scaled0.601} \end{dmath*}
   \begin{dmath*}  3 \ngab{2}^{a b} = \cdb{scaled2.601} \end{dmath*}
   \begin{dmath*}  6 \ngab{3}^{a b} = \cdb{scaled3.601} \end{dmath*}
   \begin{dmath*} 60 \ngab{4}^{a b} = \cdb{scaled4.601} \end{dmath*}
   \begin{dmath*} 90 \ngab{5}^{a b} = \cdb{scaled5.601} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
% export selected objects, these will later be imported into a library
% these are the objects that will appear in the paper

\begin{cadabra}
   cdblib.create ('metric-inv.export')

   cdblib.put ('g^ab_3',Metric3,'metric-inv.export')  # R and \partial R
   cdblib.put ('g^ab_4',Metric4,'metric-inv.export')
   cdblib.put ('g^ab_5',Metric5,'metric-inv.export')
   cdblib.put ('g^ab_6',Metric6,'metric-inv.export')

   cdblib.put ('g^ab',  Metric, 'metric-inv.export')  # R and \nabla R

   cdblib.put ('g^ab_scaled0',scaled0,'metric-inv.export')
   cdblib.put ('g^ab_scaled2',scaled2,'metric-inv.export')
   cdblib.put ('g^ab_scaled3',scaled3,'metric-inv.export')
   cdblib.put ('g^ab_scaled4',scaled4,'metric-inv.export')
   cdblib.put ('g^ab_scaled5',scaled5,'metric-inv.export')

   checkpoint.append (Metric4)
   checkpoint.append (Metric6)

   checkpoint.append (Metric)

   checkpoint.append (scaled0)
   checkpoint.append (scaled2)
   checkpoint.append (scaled3)
   checkpoint.append (scaled4)
   checkpoint.append (scaled5)

   # cdbBeg (timing)
   print ("Stage 1: {:7.1f} secs\\hfill\\break".format(end_stage_1-beg_stage_1))
   print ("Stage 2: {:7.1f} secs\\hfill\\break".format(end_stage_2-beg_stage_2))
   print ("Stage 3: {:7.1f} secs\\hfill\\break".format(end_stage_3-beg_stage_3))
   print ("Stage 4: {:7.1f} secs".format(end_stage_4-beg_stage_4))
   # cdbEnd (timing)

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Timing}

\cdb{timing}

% =================================================================================================
% export checkpoints in json format

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   for i in range( len(checkpoint) ):
      cdblib.put ('check{:03d}'.format(i),checkpoint[i],checkpoint_file)
\end{cadabra}
\egroup

\end{document}

### 2019-09-24 source/cadabra/metric.tex
Initial commit
This is where it all begins ...

--- 2019-09-24 source/cadabra/metric-inv.tex
@@
    import cdblib
-   checkpoint_file = 'tests/semantic/output/metric-inv.json'
+   checkpoint_file = 'tests/semantic/output/metric.json'
    cdblib.create (checkpoint_file)
@@
 % =================================================================================================
-\section*{The inverse metric tensor in Riemann normal coordinates}
+\section*{The metric tensor in Riemann normal coordinates}
 
-Here we calculate the Riemann normal expansion of the inverse metric, $g^{ab}$, by developing
-the recursive sequences
+In this notebook we compute the recursive sequences
 \begin{align}
 \label{eq:pdgab}
-g^{ab}{}_{,d\ue} &= - \left(g^{cb}\Gamma^{a}{}_{c(d}\right){}_{,\ue)}
-                    - \left(g^{ac}\Gamma^{b}{}_{c(d}\right){}_{,\ue)}\\[10pt]
+g_{ab,d\ue} &=  \left(g_{cb}\Gamma^{c}{}_{a(d}\right){}_{,\ue)}
+               + \left(g_{ac}\Gamma^{c}{}_{b(d}\right){}_{,\ue)}\\[10pt]
 \label{eq:pdGamma}
-(n+3)\Gamma^a{}_{d(b,c\ue)} &= (n+1)\left(R^a{}_{(bc\Dot d,\ue)} - \left(\Gamma^a{}_{f(c}\Gamma^f{}_{b{\Dot d}}\right){}_{,\ue)}\right)
+(n+3)\Gamma^a{}_{d(b,c\ue)} &= (n+1)\left(R^a{}_{(bc\Dot d,\ue)}
+                               - \left(\Gamma^a{}_{f(c}\Gamma^f{}_{b{\Dot d}}\right){}_{,\ue)}\right)
 \end{align}
@@
 \begin{dmath*}[spread=5pt]
-g^{ab}(x) = g^{ab} + g^{ab}{}_{,c}x^c + \frac{1}{2!} g^{ab}{}_{,cd}x^cx^d + \frac{1}{3!} g^{ab}{}_{,cde}x^cx^dx^e + \cdots
-          = g^{ab} + \sum_{n=1}^\infty\> \frac{1}{n!}\>g^{ab}{}_{,\uc}\>x^{.\uc}
+g_{ab}(x) = g_{ab} + g_{ab,c}x^c + \frac{1}{2!} g_{ab,cd}x^cx^d + \frac{1}{3!} g_{ab,cde}x^cx^dx^e + \cdots
+          = g_{ab} + \sum_{n=1}^\infty\> \frac{1}{n!}\>g_{ab,\uc}\>x^{.\uc}
 \end{dmath*}
@@
 % =================================================================================================
-\section*{Stage 1: Symmetrised partial derivatives of $g^{ab}$}
+\section*{Stage 1: Symmetrised partial derivatives of $g_{ab}$}
 
@@
 \begin{dgroup*}
-   \begin{dmath*} g^{ab}{}_{,c} A^{c} = \cdb{term1.200} \end{dmath*}
-   \begin{dmath*} g^{ab}{}_{,cd} A^{c} A^{d} = \cdb{term2.200} \end{dmath*}
-   \begin{dmath*} g^{ab}{}_{,cde} A^{c} A^{d} A^{e} = \cdb{term3.200} \end{dmath*}
+   \begin{dmath*} g_{ab,c} A^{c} = \cdb{term1.200} \end{dmath*}
+   \begin{dmath*} g_{ab,cd} A^{c} A^{d} = \cdb{term2.200} \end{dmath*}
+   \begin{dmath*} g_{ab,cde} A^{c} A^{d} A^{e} = \cdb{term3.200} \end{dmath*}
 \end{dgroup*}
@@
 \begin{dgroup*}
-   \begin{dmath*} g^{ab}{}_{,c} A^{c} = \cdb{term1.200} \end{dmath*}
-   \begin{dmath*} g^{ab}{}_{,cd} A^{c} A^{d} = \cdb{term2.303} \end{dmath*}
-   \begin{dmath*} g^{ab}{}_{,cde} A^{c} A^{d} A^{e} = \cdb{term3.305} \end{dmath*}
+   \begin{dmath*} g_{ab,c} A^{c} = \cdb{term1.200} \end{dmath*}
+   \begin{dmath*} g_{ab,cd} A^{c} A^{d} = \cdb{term2.303} \end{dmath*}
+   \begin{dmath*} g_{ab,cde} A^{c} A^{d} A^{e} = \cdb{term3.305} \end{dmath*}
 \end{dgroup*}
@@
 \begin{dgroup*}
-   \begin{dmath*} g^{ab}{}_{,c} A^{c} = \cdb{term1.404} \end{dmath*}
-   \begin{dmath*} g^{ab}{}_{,cd} A^{c} A^{d} = \cdb{term2.404} \end{dmath*}
-   \begin{dmath*} g^{ab}{}_{,cde} A^{c} A^{d} A^{e} = \cdb{term3.403} \end{dmath*}
+   \begin{dmath*} g_{ab,c} A^{c} = \cdb{term1.404} \end{dmath*}
+   \begin{dmath*} g_{ab,cd} A^{c} A^{d} = \cdb{term2.404} \end{dmath*}
+   \begin{dmath*} g_{ab,cde} A^{c} A^{d} A^{e} = \cdb{term3.403} \end{dmath*}
 \end{dgroup*}
@@
 \begin{align*}
-   g_{ab}(x) &=   g^{ab}
-                + g^{ab}{}_{,c} x^c
-                + \frac{1}{2!} g^{ab}{}_{,cd} x^c x^d
-                + \frac{1}{3!} g^{ab}{}_{,cde} x^c x^d x^e +  \BigO{\eps^4}\\
+   g_{ab}(x) &=   g_{ab}
+                + g_{ab,c} x^c
+                + \frac{1}{2!} g_{ab,cd} x^c x^d
+                + \frac{1}{3!} g_{ab,cde} x^c x^d x^e +  \BigO{\eps^4}\\
              &= \cdb{metric4.501} + \BigO{\eps^4}
@@
        unwrap     (obj)
-       sort_product   (obj)
-       rename_dummies (obj)
        return obj
@@
 
-   def add_tags (obj,tag):
-      n = 0
-      ans = Ex('0')
-      for i in obj.top().terms():
-         foo = obj[i]
-         bah = Ex(tag+'_{'+str(n)+'}')
-         ans := @(ans) + @(bah) @(foo).
-         n = n + 1
-      return ans
-
-   def clear_tags (obj,tag):
-      ans := @(obj).
-      foo  = Ex(tag+'_{a?} -> 1')
-      substitute (ans,foo)
-      return ans
-
    {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).
@@
 
-   # symmetrised partial derivatives of g^{ab}
+   # symmetrised partial derivatives of g_{ab}
 
-   gab00:=g^{a b}.                                              # cdb (gab00.101,gab00)
+   gab00:=g_{a b}.                                              # cdb (gab00.101,gab00)
 
-   gab01:= - g^{c b}\Gamma^{a}_{c d} - g^{a c}\Gamma^{b}_{c d}. # cdb (gab01.101,gab01)
+   gab01:=g_{c b}\Gamma^{c}_{a d} + g_{a c}\Gamma^{c}_{b d}.    # cdb (gab01.101,gab01)
 
@@
    product_rule (gab02)                                         # cdb (gab02.103,gab02)
-   substitute   (gab02, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab02.104,gab02)
+   substitute   (gab02, $\partial_{d}{g_{a b}} -> @(gab01)$)    # cdb (gab02.104,gab02)
    distribute   (gab02)                                         # cdb (gab02.105,gab02)
@@
    product_rule (gab03)                                         # cdb (gab03.103,gab03)
-   substitute   (gab03, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab03.104,gab03)
+   substitute   (gab03, $\partial_{d}{g_{a b}} -> @(gab01)$)    # cdb (gab03.104,gab03)
    distribute   (gab03)                                         # cdb (gab03.105,gab03)
@@
    product_rule (gab04)                                         # cdb (gab04.103,gab04)
-   substitute   (gab04, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab04.104,gab04)
+   substitute   (gab04, $\partial_{d}{g_{a b}} -> @(gab01)$)    # cdb (gab04.104,gab04)
    distribute   (gab04)                                         # cdb (gab04.105,gab04)
@@
    product_rule (gab05)                                         # cdb (gab05.103,gab05)
-   substitute   (gab05, $\partial_{d}{g^{a b}} -> @(gab01)$)    # cdb (gab05.104,gab05)
+   substitute   (gab05, $\partial_{d}{g_{a b}} -> @(gab01)$)    # cdb (gab05.104,gab05)
    distribute   (gab05)                                         # cdb (gab05.105,gab05)
@@
 \begin{cadabra}
-   # prepare first six terms in the Taylor series expansion of g^{ab}(x)
+   # prepare first six terms in the Taylor series expansion of g_{ab}(x)
 
@@
 
-   # ------------------------------------------------------------------------------------
-   # this block only produces formatted output, it is not part of the main computation
-   # ------------------------------------------------------------------------------------
+   end_stage_2 = time.time()
+
+   # -------------------------------------------------------------------------------------------
+   # this block of Xterms only produces formatted output, it's not part of the main computation
+   # -------------------------------------------------------------------------------------------
 
@@
    metric:=@(term0)
-         + (1/1) @(term1)
+         + (1/1) @(term1)  # zero
          + (1/2) @(term2)
@@
    Xterm0 := @(term0).
-   Xterm1 := (1/1) @(term1).     # zero
+   Xterm1 := (1/1) @(term1).
    Xterm2 := (1/2) @(term2).
@@
 
-   # Manipulating these expressions is hampered by the presence of the partial derivative on Rabcd.
-   # Thus we can't freely rasie/lower indices on the dRabcd terms. But we can do so on the first
-   # derivatives (since these are evaluated at x=0 where the connection vanishes).
+   substitute (Xterm2,$g_{a b} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm2.301,Xterm2)
+   substitute (Xterm3,$g_{a b} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm3.301,Xterm3)
+   substitute (Xterm4,$g_{a b} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm4.301,Xterm4)
+   substitute (Xterm5,$g_{a b} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm5.301,Xterm5)
 
-   substitute       (Xterm2,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm2.301,Xterm2)
-   substitute       (Xterm3,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm3.301,Xterm3)
-   substitute       (Xterm4,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm4.301,Xterm4)
-   substitute       (Xterm5,$g^{a b} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm5.301,Xterm5)
+   substitute (Xterm2,$g_{b a} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm2.301,Xterm2)
+   substitute (Xterm3,$g_{b a} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm3.301,Xterm3)
+   substitute (Xterm4,$g_{b a} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm4.301,Xterm4)
+   substitute (Xterm5,$g_{b a} \partial_{c}{R^{b}_{d e f}} -> \partial_{c}{R_{a d e f}}$)  # cdb(Xterm5.301,Xterm5)
 
-   substitute       (Xterm2,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm2.302,Xterm2)
-   substitute       (Xterm3,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm3.302,Xterm3)
-   substitute       (Xterm4,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm4.302,Xterm4)
-   substitute       (Xterm5,$g^{b a} R^{c}_{d e b} -> R^{c}_{d e}^{a}}$)  # cdb(Xterm5.302,Xterm5)
+   eliminate_metric (Xterm2)  # cdb(Xterm2.302,Xterm2)
+   eliminate_metric (Xterm3)  # cdb(Xterm3.302,Xterm3)
+   eliminate_metric (Xterm4)  # cdb(Xterm4.302,Xterm4)
+   eliminate_metric (Xterm5)  # cdb(Xterm5.302,Xterm5)
 
-   substitute       (Xterm2,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm2.303,Xterm2)
-   substitute       (Xterm3,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm3.303,Xterm3)
-   substitute       (Xterm4,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm4.303,Xterm4)
-   substitute       (Xterm5,$g^{a b} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm5.303,Xterm5)
+   sort_product     (Xterm2)  # cdb(Xterm2.303,Xterm2)
+   sort_product     (Xterm3)  # cdb(Xterm3.303,Xterm3)
+   sort_product     (Xterm4)  # cdb(Xterm4.303,Xterm4)
+   sort_product     (Xterm5)  # cdb(Xterm5.303,Xterm5)
 
-   substitute       (Xterm2,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm2.304,Xterm2)
-   substitute       (Xterm3,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm3.304,Xterm3)
-   substitute       (Xterm4,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm4.304,Xterm4)
-   substitute       (Xterm5,$g^{b a} \partial_{c}{R^{d}_{e f b}} -> \partial_{c}{R^{d}_{e f}^{a}$)  # cdb(Xterm5.304,Xterm5)
+   rename_dummies   (Xterm2)  # cdb(Xterm2.304,Xterm2)
+   rename_dummies   (Xterm3)  # cdb(Xterm3.304,Xterm3)
+   rename_dummies   (Xterm4)  # cdb(Xterm4.304,Xterm4)
+   rename_dummies   (Xterm5)  # cdb(Xterm5.304,Xterm5)
 
-   sort_product     (Xterm2)  # cdb(Xterm2.305,Xterm2)
-   sort_product     (Xterm3)  # cdb(Xterm3.305,Xterm3)
-   sort_product     (Xterm4)  # cdb(Xterm4.305,Xterm4)
-   sort_product     (Xterm5)  # cdb(Xterm5.305,Xterm5)
+   canonicalise     (Xterm2)  # cdb(Xterm2.305,Xterm2)
+   canonicalise     (Xterm3)  # cdb(Xterm3.305,Xterm3)
+   canonicalise     (Xterm4)  # cdb(Xterm4.305,Xterm4)
+   canonicalise     (Xterm5)  # cdb(Xterm5.305,Xterm5)
 
-   rename_dummies   (Xterm2)  # cdb(Xterm2.306,Xterm2)
-   rename_dummies   (Xterm3)  # cdb(Xterm3.306,Xterm3)
-   rename_dummies   (Xterm4)  # cdb(Xterm4.306,Xterm4)
-   rename_dummies   (Xterm5)  # cdb(Xterm5.306,Xterm5)
+   # push upper index to the left
+   def tidy_Rabcd (obj):
+       substitute (obj,$R_{a b c}^{d} -> - R^{d}_{c a b}$)
+       substitute (obj,$R_{a b}^{c}_{d} -> R^{c}_{d a b}$)
+       substitute (obj,$R_{a}^{b}_{c d} -> - R^{b}_{a c d}$)
+       return obj
 
-   canonicalise     (Xterm2)  # cdb(Xterm2.307,Xterm2)
-   canonicalise     (Xterm3)  # cdb(Xterm3.307,Xterm3)
-   canonicalise     (Xterm4)  # cdb(Xterm4.307,Xterm4)
-   canonicalise     (Xterm5)  # cdb(Xterm5.307,Xterm5)
+   Xterm0 = tidy_Rabcd (Xterm0)  # cdb(Xterm0.666,Xterm0)
+   Xterm2 = tidy_Rabcd (Xterm2)  # cdb(Xterm2.666,Xterm2)
+   Xterm3 = tidy_Rabcd (Xterm3)  # cdb(Xterm3.666,Xterm3)
+   Xterm4 = tidy_Rabcd (Xterm4)  # cdb(Xterm4.666,Xterm4)
+   Xterm5 = tidy_Rabcd (Xterm5)  # cdb(Xterm5.666,Xterm5)
 
-   # We can simplify Xterm2 and Xterm3 by careful juggling of the indices (swapping free indices on selected terms)
+   Xterm0 = reformat_xterm (Xterm0,  1)    # cdb(Xterm0.301,Xterm0)
+   Xterm2 = reformat_xterm (Xterm2,  3)    # cdb(Xterm2.301,Xterm2)
+   Xterm3 = reformat_xterm (Xterm3,  6)    # cdb(Xterm3.301,Xterm3)
+   Xterm4 = reformat_xterm (Xterm4,360)    # cdb(Xterm4.301,Xterm4)
+   Xterm5 = reformat_xterm (Xterm5,180)    # cdb(Xterm5.301,Xterm5)
 
-   tmp = add_tags (Xterm2,'\\mu')       # cdb (tmp.001,tmp)
-   zoom (tmp, $\mu_{1} Q??$)            # cdb (tmp.002,tmp)
-   substitute (tmp, $R^{b}_{c d}^{a} x^{c} x^{d} -> R^{a}_{c d}^{b} x^{c} x^{d}$)  # cdb (tmp.003,tmp)
-   unzoom (tmp)
-   Xterm2 = clear_tags (tmp,'\\mu')     # cdb (Xterm2.401,Xterm2)
+   # canonicalise from reformat_xterm will slide upper index from left hand side
+   # so now we slide the upper index back to the left
 
-   tmp = add_tags (Xterm3,'\\mu')       # cdb (tmp.011,tmp)
-   zoom (tmp, $\mu_{1} Q??$)            # cdb (tmp.012,tmp)
-   substitute (tmp, $\partial_{c}{R^{b}_{d e}^{a}} x^{c} x^{d} x^{e} -> \partial_{c}{R^{a}_{d e}^{b}} x^{c} x^{d} x^{e}$)  # cdb (tmp.013,tmp)
-   unzoom (tmp)
-   Xterm3 = clear_tags (tmp,'\\mu')     # cdb (Xterm3.401,Xterm3)
+   Xterm0 = tidy_Rabcd (Xterm0)  # cdb(Xterm0.667,Xterm0)
+   Xterm2 = tidy_Rabcd (Xterm2)  # cdb(Xterm2.667,Xterm2)
+   Xterm3 = tidy_Rabcd (Xterm3)  # cdb(Xterm3.667,Xterm3)
+   Xterm4 = tidy_Rabcd (Xterm4)  # cdb(Xterm4.667,Xterm4)
+   Xterm5 = tidy_Rabcd (Xterm5)  # cdb(Xterm5.667,Xterm5)
 
-   Xterm0 = reformat_xterm (Xterm0,  1)    # cdb(Xterm0.308,Xterm0)
-   Xterm2 = reformat_xterm (Xterm2,  3)    # cdb(Xterm2.308,Xterm2)
-   Xterm3 = reformat_xterm (Xterm3,  6)    # cdb(Xterm3.308,Xterm3)
-   Xterm4 = reformat_xterm (Xterm4,360)    # cdb(Xterm4.308,Xterm4)
-   Xterm5 = reformat_xterm (Xterm5,360)    # cdb(Xterm5.308,Xterm5)
-
-   # metric to 4th and 6th order terms in powers of x^a
+   # metric to 3rd, 4th, 5th and 6th order terms in powers of x^a
 
@@
 
-   end_stage_2 = time.time()
 \end{cadabra}
@@
 \begin{dgroup*}
-   \begin{dmath*} g^{ab}(x) = \cdb{Metric3.301} \end{dmath*}
-   \begin{dmath*} g^{ab}(x) = \cdb{Metric4.301} \end{dmath*}
-   \begin{dmath*} g^{ab}(x) = \cdb{Metric5.301} \end{dmath*}
-   \begin{dmath*} g^{ab}(x) = \cdb{Metric6.301} \end{dmath*}
+   \begin{dmath*} g_{ab}(x) = \cdb{Metric3.301} \end{dmath*}
+   \begin{dmath*} g_{ab}(x) = \cdb{Metric4.301} \end{dmath*}
+   \begin{dmath*} g_{ab}(x) = \cdb{Metric5.301} \end{dmath*}
+   \begin{dmath*} g_{ab}(x) = \cdb{Metric6.301} \end{dmath*}
 \end{dgroup*}
@@
    beg_stage_4 = time.time()
-
    # final housekeeping
 
-   # lower the ^{ab} indices to _{uv}
+   term1 = flatten_Rabcd (term1)         # cdb(term1.501,term1)
+   term2 = flatten_Rabcd (term2)         # cdb(term2.501,term2)
+   term3 = flatten_Rabcd (term3)         # cdb(term3.501,term3)
+   term4 = flatten_Rabcd (term4)         # cdb(term4.501,term4)
+   term5 = flatten_Rabcd (term5)         # cdb(term5.501,term5)
 
-   tmp0 := g_{a u} g_{b v} @(term0).
-   tmp1 := g_{a u} g_{b v} @(term1).
-   tmp2 := g_{a u} g_{b v} @(term2).
-   tmp3 := g_{a u} g_{b v} @(term3).
-   tmp4 := g_{a u} g_{b v} @(term4).
-   tmp5 := g_{a u} g_{b v} @(term5).
+   eliminate_metric (term1)
+   eliminate_metric (term2)
+   eliminate_metric (term3)
+   eliminate_metric (term4)
+   eliminate_metric (term5)
 
-   distribute           (tmp1)  # cdb(tmp1.501,tmp1)
-   eliminate_metric     (tmp1)  # cdb(tmp1.502,tmp1)
-   eliminate_kronecker  (tmp1)  # cdb(tmp1.503,tmp1)
-   tmp1 = flatten_Rabcd (tmp1)
-   canonicalise         (tmp1)  # cdb(tmp1.506,tmp1)
+   eliminate_kronecker (term1)
+   eliminate_kronecker (term2)
+   eliminate_kronecker (term3)
+   eliminate_kronecker (term4)
+   eliminate_kronecker (term5)
 
-   distribute           (tmp2)  # cdb(tmp2.501,tmp2)
-   eliminate_metric     (tmp2)  # cdb(tmp2.502,tmp2)
-   eliminate_kronecker  (tmp2)  # cdb(tmp2.503,tmp2)
-   tmp2 = flatten_Rabcd (tmp2)
-   canonicalise         (tmp2)  # cdb(tmp2.506,tmp2)
+   sort_product (term1)
+   sort_product (term2)
+   sort_product (term3)
+   sort_product (term4)
+   sort_product (term5)
 
-   distribute           (tmp3)  # cdb(tmp3.501,tmp3)
-   eliminate_metric     (tmp3)  # cdb(tmp3.502,tmp3)
-   eliminate_kronecker  (tmp3)  # cdb(tmp3.503,tmp3)
-   tmp3 = flatten_Rabcd (tmp3)
-   canonicalise         (tmp3)  # cdb(tmp3.506,tmp3)
+   rename_dummies (term1)
+   rename_dummies (term2)
+   rename_dummies (term3)
+   rename_dummies (term4)
+   rename_dummies (term5)
 
-   distribute           (tmp4)  # cdb(tmp4.501,tmp4)
-   eliminate_metric     (tmp4)  # cdb(tmp4.502,tmp4)
-   eliminate_kronecker  (tmp4)  # cdb(tmp4.503,tmp4)
-   tmp4 = flatten_Rabcd (tmp4)
-   canonicalise         (tmp4)  # cdb(tmp4.506,tmp4)
-
-   distribute           (tmp5)  # cdb(tmp5.501,tmp5)
-   eliminate_metric     (tmp5)  # cdb(tmp5.502,tmp5)
-   eliminate_kronecker  (tmp5)  # cdb(tmp5.503,tmp5)
-   tmp5 = flatten_Rabcd (tmp5)
-   canonicalise         (tmp5)  # cdb(tmp5.506,tmp5)
+   canonicalise (term1)                  # cdb(term1.502,term1)
+   canonicalise (term2)                  # cdb(term2.502,term2)
+   canonicalise (term3)                  # cdb(term3.502,term3)
+   canonicalise (term4)                  # cdb(term4.502,term4)
+   canonicalise (term5)                  # cdb(term5.502,term5)
 
@@
 
-   # raise the _{uv} indices to ^{ab}
+   metric:=@(term0)
+         + (1/1) @(term1)
+         + (1/2) @(term2)
+         + (1/6) @(term3)
+         + (1/24) @(term4)
+         + (1/120) @(term5).             # cdb(metric.501,metric)
 
-   metric:= g^{a u} g^{b v} (  @(tmp0)
-                              + (1/1) @(tmp1)
-                              + (1/2) @(tmp2)
-                              + (1/6) @(tmp3)
-                              + (1/24) @(tmp4)
-                              + (1/120) @(tmp5) ).  # cdb(metric.500,metric)
+   substitute (metric,$A^{a} -> x^{a}$)  # cdb (metric.502,metric)
 
-   distribute             (metric)   # cdb(metric.501,metric)
-   eliminate_metric       (metric)   # cdb(metric.502,metric)
-   eliminate_kronecker    (metric)   # cdb(metric.503,metric)
-   metric = flatten_Rabcd (metric)   # cdb(metric.504,metric)
-   canonicalise           (metric)   # cdb(metric.505,metric)
+   cdblib.create ('metric.json')
 
-   substitute          (metric,$g_{a b} g^{b c} -> g_{a}^{c}$)
-   substitute          (metric,$g_{b a} g^{b c} -> g_{a}^{c}$)
-   substitute          (metric,$g_{b a} g^{c b} -> g_{a}^{c}$)
-   substitute          (metric,$g_{a b} g^{c b} -> g_{a}^{c}$)
-   eliminate_kronecker (metric)   # cdb(metric.506,metric)
-   canonicalise        (metric)   # cdb(metric.507,metric)
-
-   substitute (metric,$A^{a} -> x^{a}$)  # cdb (metric.508,metric)
-
-   cdblib.create ('metric-inv.json')
-
-   cdblib.put ('g^ab',metric,'metric-inv.json')
+   cdblib.put ('g_ab',metric,'metric.json')
 
@@
 
-   term0 = get_xterm (metric,0)   # cdb(term0.501,term0)
-   term1 = get_xterm (metric,1)   # cdb(term1.501,term1)
-   term2 = get_xterm (metric,2)   # cdb(term2.501,term2)
-   term3 = get_xterm (metric,3)   # cdb(term3.501,term3)
-   term4 = get_xterm (metric,4)   # cdb(term4.501,term4)
-   term5 = get_xterm (metric,5)   # cdb(term5.501,term5)
+   term0 = get_xterm (metric,0)          # cdb(term0.503,term0)
+   term1 = get_xterm (metric,1)          # cdb(term1.503,term1)
+   term2 = get_xterm (metric,2)          # cdb(term2.503,term2)
+   term3 = get_xterm (metric,3)          # cdb(term3.503,term3)
+   term4 = get_xterm (metric,4)          # cdb(term4.503,term4)
+   term5 = get_xterm (metric,5)          # cdb(term5.503,term5)
 
-   cdblib.put ('g^ab_0',term0,'metric-inv.json')
-   cdblib.put ('g^ab_1',term1,'metric-inv.json')
-   cdblib.put ('g^ab_2',term2,'metric-inv.json')
-   cdblib.put ('g^ab_3',term3,'metric-inv.json')
-   cdblib.put ('g^ab_4',term4,'metric-inv.json')
-   cdblib.put ('g^ab_5',term5,'metric-inv.json')
+   cdblib.put ('g_ab_0',term0,'metric.json')
+   cdblib.put ('g_ab_1',term1,'metric.json')
+   cdblib.put ('g_ab_2',term2,'metric.json')
+   cdblib.put ('g_ab_3',term3,'metric.json')
+   cdblib.put ('g_ab_4',term4,'metric.json')
+   cdblib.put ('g_ab_5',term5,'metric.json')
 
@@
 \begin{dgroup*}
-   \begin{dmath*} \cdb*{term0.501} \end{dmath*}
-   \begin{dmath*} \cdb*{term1.501} \end{dmath*}
    \begin{dmath*} \cdb*{term2.501} \end{dmath*}
+   \begin{dmath*} \cdb*{term2.502} \end{dmath*}
+\end{dgroup*}
+
+\begin{dgroup*}
    \begin{dmath*} \cdb*{term3.501} \end{dmath*}
+   \begin{dmath*} \cdb*{term3.502} \end{dmath*}
+\end{dgroup*}
+
+\begin{dgroup*}
    \begin{dmath*} \cdb*{term4.501} \end{dmath*}
+   \begin{dmath*} \cdb*{term4.502} \end{dmath*}
+\end{dgroup*}
+
+\begin{dgroup*}
    \begin{dmath*} \cdb*{term5.501} \end{dmath*}
+   \begin{dmath*} \cdb*{term5.502} \end{dmath*}
 \end{dgroup*}
@@
 \begin{dgroup*}
-   \begin{dmath*} \cdb*{tmp2.501} \end{dmath*}
-   \begin{dmath*} \cdb*{tmp2.502} \end{dmath*}
-   \begin{dmath*} \cdb*{tmp2.503} \end{dmath*}
-   \begin{dmath*} \cdb*{tmp2.506} \end{dmath*}
-\end{dgroup*}
-
-\begin{dgroup*}
-   \begin{dmath*} \cdb*{tmp3.501} \end{dmath*}
-   \begin{dmath*} \cdb*{tmp3.502} \end{dmath*}
-   \begin{dmath*} \cdb*{tmp3.503} \end{dmath*}
-   \begin{dmath*} \cdb*{tmp3.506} \end{dmath*}
+   \begin{dmath*} \cdb*{metric.501} \end{dmath*}
+   \begin{dmath*} \cdb*{metric.502} \end{dmath*}
 \end{dgroup*}
@@
 \begin{dgroup*}
-   \begin{dmath*} \cdb*{tmp4.501} \end{dmath*}
-   \begin{dmath*} \cdb*{tmp4.502} \end{dmath*}
-   \begin{dmath*} \cdb*{tmp4.503} \end{dmath*}
-   \begin{dmath*} \cdb*{tmp4.506} \end{dmath*}
-\end{dgroup*}
-
-\clearpage
-
-\begin{dgroup*}
-   \begin{dmath*} \cdb*{tmp5.506} \end{dmath*}
-\end{dgroup*}
-
-\clearpage
-
-\begin{dgroup*}
-   \begin{dmath*} \cdb*{metric.500} \end{dmath*}
-   \begin{dmath*} \cdb*{metric.501} \end{dmath*}
-   \begin{dmath*} \cdb*{metric.502} \end{dmath*}
-   \begin{dmath*} \cdb*{metric.503} \end{dmath*}
-   \begin{dmath*} \cdb*{metric.504} \end{dmath*}
-   \begin{dmath*} \cdb*{metric.505} \end{dmath*}
-   \begin{dmath*} \cdb*{metric.506} \end{dmath*}
-   \begin{dmath*} \cdb*{metric.507} \end{dmath*}
-   \begin{dmath*} \cdb*{metric.508} \end{dmath*}
+   \begin{dmath*} \cdb*{term0.503} \end{dmath*}
+   \begin{dmath*} \cdb*{term1.503} \end{dmath*}
+   \begin{dmath*} \cdb*{term2.503} \end{dmath*}
+   \begin{dmath*} \cdb*{term3.503} \end{dmath*}
+   \begin{dmath*} \cdb*{term4.503} \end{dmath*}
+   \begin{dmath*} \cdb*{term5.503} \end{dmath*}
 \end{dgroup*}
@@
 \begin{cadabra}
-   Rterm0 := @(term0).
-   Rterm1 := @(term1).  # zero
-   Rterm2 := @(term2).
-   Rterm3 := @(term3).
-   Rterm4 := @(term4).
-   Rterm5 := @(term5).
+   Xterm0 := @(term0).
+   Xterm1 := @(term1).  # zero
+   Xterm2 := @(term2).
+   Xterm3 := @(term3).
+   Xterm4 := @(term4).
+   Xterm5 := @(term5).
 
-   Rterm0 = reformat_xterm (Rterm0,  1)    # cdb(Rterm0.601,Rterm0)
-   Rterm2 = reformat_xterm (Rterm2,  3)    # cdb(Rterm2.601,Rterm2)
-   Rterm3 = reformat_xterm (Rterm3,  6)    # cdb(Rterm3.601,Rterm3)
-   Rterm4 = reformat_xterm (Rterm4, 60)    # cdb(Rterm4.601,Rterm4)
-   Rterm5 = reformat_xterm (Rterm5, 90)    # cdb(Rterm5.601,Rterm5)
+   Xterm0 = reformat_xterm (Xterm0,  1)    # cdb(Xterm0.601,Xterm0)
+   Xterm2 = reformat_xterm (Xterm2,  3)    # cdb(Xterm2.601,Xterm2)
+   Xterm3 = reformat_xterm (Xterm3,  6)    # cdb(Xterm3.601,Xterm3)
+   Xterm4 = reformat_xterm (Xterm4,180)    # cdb(Xterm4.601,Xterm4)
+   Xterm5 = reformat_xterm (Xterm5, 90)    # cdb(Xterm5.601,Xterm5)
 
-   Metric := @(Rterm0) + @(Rterm2) + @(Rterm3) + @(Rterm4) + @(Rterm5).  # cdb (Metric.601,Metric)
+   gab3   := @(Xterm0) + @(Xterm2).                                      # cdb (gab3.601,gab3)
+   gab4   := @(Xterm0) + @(Xterm2) + @(Xterm3).                          # cdb (gab4.601,gab4)
+   gab5   := @(Xterm0) + @(Xterm2) + @(Xterm3) + @(Xterm4).              # cdb (gab5.601,gab5)
+   gab6   := @(Xterm0) + @(Xterm2) + @(Xterm3) + @(Xterm4) + @(Xterm5).  # cdb (gab6.601,gab6)
 
-   scaled0 = rescale_xterm (Rterm0,  1)    # cdb(scaled0.601,scaled0)
-   scaled2 = rescale_xterm (Rterm2,  3)    # cdb(scaled2.601,scaled2)
-   scaled3 = rescale_xterm (Rterm3,  6)    # cdb(scaled3.601,scaled3)
-   scaled4 = rescale_xterm (Rterm4, 60)    # cdb(scaled4.601,scaled4)
-   scaled5 = rescale_xterm (Rterm5, 90)    # cdb(scaled5.601,scaled5)
+   Metric := @(Xterm0) + @(Xterm2) + @(Xterm3) + @(Xterm4) + @(Xterm5).  # cdb (Metric.601,Metric)
+
+   scaled0 = rescale_xterm (Xterm0,  1)    # cdb(scaled0.601,scaled0)
+   scaled2 = rescale_xterm (Xterm2,  3)    # cdb(scaled2.601,scaled2)
+   scaled3 = rescale_xterm (Xterm3,  6)    # cdb(scaled3.601,scaled3)
+   scaled4 = rescale_xterm (Xterm4,180)    # cdb(scaled4.601,scaled4)
+   scaled5 = rescale_xterm (Xterm5, 90)    # cdb(scaled5.601,scaled5)
 
@@
 % =================================================================================================
-\section*{The inverse metric in Riemann normal coordinates}
+\section*{The metric in Riemann normal coordinates}
 
 \begin{dgroup*}
-   \begin{dmath*} g^{a b}(x) = \cdb{Metric.601}+\BigO{\eps^6} \end{dmath*}
+   \begin{dmath*} g_{a b}(x) = \cdb{Metric.601}+\BigO{\eps^6} \end{dmath*}
 \end{dgroup*}
@@
 % =================================================================================================
-\section*{Curvature expansion of the inverse metric}
+\section*{Curvature expansion of the metric}
 \begin{align*}
-     g^{a b}(x) =
-     \ngab{0}^{a b}
-   + \ngab{2}^{a b}
-   + \ngab{3}^{a b}
-   + \ngab{4}^{a b}
-   + \ngab{5}^{a b}+\BigO{\eps^6}
+     g_{a b}(x) =
+     \ngab{0}_{a b}
+   + \ngab{2}_{a b}
+   + \ngab{3}_{a b}
+   + \ngab{4}_{a b}
+   + \ngab{5}_{a b}+\BigO{\eps^6}
 \end{align*}
 \begin{dgroup*}
-   \begin{dmath*}    \ngab{0}^{a b} = \cdb{scaled0.601} \end{dmath*}
-   \begin{dmath*}  3 \ngab{2}^{a b} = \cdb{scaled2.601} \end{dmath*}
-   \begin{dmath*}  6 \ngab{3}^{a b} = \cdb{scaled3.601} \end{dmath*}
-   \begin{dmath*} 60 \ngab{4}^{a b} = \cdb{scaled4.601} \end{dmath*}
-   \begin{dmath*} 90 \ngab{5}^{a b} = \cdb{scaled5.601} \end{dmath*}
+   \begin{dmath*}     \ngab{0}_{a b} = \cdb{scaled0.601} \end{dmath*}
+   \begin{dmath*}   3 \ngab{2}_{a b} = \cdb{scaled2.601} \end{dmath*}
+   \begin{dmath*}   6 \ngab{3}_{a b} = \cdb{scaled3.601} \end{dmath*}
+   \begin{dmath*} 180 \ngab{4}_{a b} = \cdb{scaled4.601} \end{dmath*}
+   \begin{dmath*}  90 \ngab{5}_{a b} = \cdb{scaled5.601} \end{dmath*}
 \end{dgroup*}
@@
 \begin{cadabra}
-   cdblib.create ('metric-inv.export')
+   cdblib.create ('metric.export')
 
-   cdblib.put ('g^ab_3',Metric3,'metric-inv.export')  # R and \partial R
-   cdblib.put ('g^ab_4',Metric4,'metric-inv.export')
-   cdblib.put ('g^ab_5',Metric5,'metric-inv.export')
-   cdblib.put ('g^ab_6',Metric6,'metric-inv.export')
+   cdblib.put ('g_ab_3',Metric3,'metric.export')  # R and \partial R
+   cdblib.put ('g_ab_4',Metric4,'metric.export')
+   cdblib.put ('g_ab_5',Metric5,'metric.export')
+   cdblib.put ('g_ab_6',Metric6,'metric.export')
 
-   cdblib.put ('g^ab',  Metric, 'metric-inv.export')  # R and \nabla R
+   cdblib.put ('g_ab',  Metric, 'metric.export')  # R and \nabla R
 
-   cdblib.put ('g^ab_scaled0',scaled0,'metric-inv.export')
-   cdblib.put ('g^ab_scaled2',scaled2,'metric-inv.export')
-   cdblib.put ('g^ab_scaled3',scaled3,'metric-inv.export')
-   cdblib.put ('g^ab_scaled4',scaled4,'metric-inv.export')
-   cdblib.put ('g^ab_scaled5',scaled5,'metric-inv.export')
+   cdblib.put ('g_ab_scaled0',scaled0,'metric.export')
+   cdblib.put ('g_ab_scaled2',scaled2,'metric.export')
+   cdblib.put ('g_ab_scaled3',scaled3,'metric.export')
+   cdblib.put ('g_ab_scaled4',scaled4,'metric.export')
+   cdblib.put ('g_ab_scaled5',scaled5,'metric.export')
 

### 2019-09-24 source/cadabra/rnc2rnc.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

% =================================================================================================
% create checkpoint file

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   import cdblib
   checkpoint_file = 'tests/semantic/output/rnc2rnc.json'
   cdblib.create (checkpoint_file)
   checkpoint = []
\end{cadabra}
\egroup

% =================================================================================================
\section*{From one RNC to another}

Consider an RNC frame with RNC cooridnates $x^{a}$.

In the {\tts geodesic-bvp} code the two point boundary value problem (for the geodesic connecting
two points) was solved. There is a bonus in that calculation -- it can be trivaly adapted to the
case of transforming form one RNC into another.

The starting point is the basic equation for the geodesic connecting $P$ (with coordinaties
$x^{a}$) to Q (with coordinates $x^{a} + Dx^{a}$)
\begin{equation*}
   x^a(s) = x^a_i + s y^a - \sum_{k=2}^\infty\>\frac{1}{k!}\>\Gamma^{a}{}_{\ubk}y^{.\ubk} s^k
\end{equation*}
The affine parameter $s$ varies form 0 (at $P$) to 1 (at $Q$).

A new RNC frame, with origin at $P$, can be defined via the $y^{a}$ with the coordinates of $Q$ in
the new RNC frame defined by $y^{a}$ (since $s=1$ at $Q$). Recall that in an RNC all geodesics
through the origin are described by $y^{a}(s) = s y^{a}$. Thus the transformation from $x^a$ to
$y^a$ satisfies
\begin{equation*}
   x^a = x^a_i + y^a - \sum_{k=2}^\infty\>\frac{1}{k!}\>\Gamma^{a}{}_{\ubk}y^{.\ubk}
\end{equation*}
where the $\Gamma^{a}{}_{\ubk}$ are the generalised connections of the $x^a$ frame evaluated at
$x^a=0$. This equation can be inverted to express $y^a$ in terms of $x^a$. This computation is
done in the {\tts geodesic-bvp} code -- we only quote the results here (at the end).

The new $y^a$ frame has origin at $P$. Its coordinate axes are aligned with those (at $P$) of the
origianl RNC frame. To see this just note that $\partial x^a/\partial y^b = \delta^a_b$ at $P$.
Thus the metric at $P$ in the new frame has values $g_{ab}(x)$ (i.e., exactly those of the
original RNC frame). Note that this means that the coordinate axes of the new frame are not
ncessarily orthogonal.

The calculations in this code are trivial. It uses the $y^{a}$ found in {\tts geodesic-bvp} as
the basis of the transformation from $x^{a}$ to $y^{a}$. Most of the code involves reformatting
the $y^{a}$.

\clearpage

\begin{cadabra}
   {a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}::Indices(position=independent).

   \nabla{#}::Derivative.

   g_{a b}::Metric.
   g^{a b}::InverseMetric.

   R_{a b c d}::RiemannTensor.
   R^{a}_{b c d}::RiemannTensor.

   # Dx{#}::LaTeXForm{"{\Dx}"}.  # LCB: currently causes a bug, it kills ::KeepWeight for Dx

   import cdblib

   Y5 = cdblib.get ('y5','geodesic-bvp.json')

   Y50 = cdblib.get ('y50','geodesic-bvp.json')
   Y52 = cdblib.get ('y52','geodesic-bvp.json')
   Y53 = cdblib.get ('y53','geodesic-bvp.json')
   Y54 = cdblib.get ('y54','geodesic-bvp.json')
   Y55 = cdblib.get ('y55','geodesic-bvp.json')

   # this copies y5* from geodesic-bvp.json to rnc2rnc.json

   cdblib.create ('rnc2rnc.json')

   cdblib.put ('rnc2rnc',Y5,'rnc2rnc.json')

   cdblib.put ('rnc2rnc0',Y50,'rnc2rnc.json')
   cdblib.put ('rnc2rnc2',Y52,'rnc2rnc.json')
   cdblib.put ('rnc2rnc3',Y53,'rnc2rnc.json')
   cdblib.put ('rnc2rnc4',Y54,'rnc2rnc.json')
   cdblib.put ('rnc2rnc5',Y55,'rnc2rnc.json')

\end{cadabra}

% =================================================================================================
% the remaining code is just for pretty printing

\clearpage

\begin{cadabra}
   # note: keeping numbering as is (out of order) to ensure R appears before \nabla R etc.
   def product_sort (obj):
       substitute (obj,$ x^{a}                            -> A001^{a}               $)
       substitute (obj,$ Dx^{a}                           -> A002^{a}               $)
       substitute (obj,$ g^{a b}                          -> A003^{a b}             $)
       substitute (obj,$ \nabla_{e f g h}{R_{a b c d}}    -> A008_{a b c d e f g h} $)
       substitute (obj,$ \nabla_{e f g}{R_{a b c d}}      -> A007_{a b c d e f g}   $)
       substitute (obj,$ \nabla_{e f}{R_{a b c d}}        -> A006_{a b c d e f}     $)
       substitute (obj,$ \nabla_{e}{R_{a b c d}}          -> A005_{a b c d e}       $)
       substitute (obj,$ R_{a b c d}                      -> A004_{a b c d}         $)
       sort_product   (obj)
       rename_dummies (obj)
       substitute (obj,$ A001^{a}                  -> x^{a}                         $)
       substitute (obj,$ A002^{a}                  -> Dx^{a}                        $)
       substitute (obj,$ A003^{a b}                -> g^{a b}                       $)
       substitute (obj,$ A004_{a b c d}            -> R_{a b c d}                   $)
       substitute (obj,$ A005_{a b c d e}          -> \nabla_{e}{R_{a b c d}}       $)
       substitute (obj,$ A006_{a b c d e f}        -> \nabla_{e f}{R_{a b c d}}     $)
       substitute (obj,$ A007_{a b c d e f g}      -> \nabla_{e f g}{R_{a b c d}}   $)
       substitute (obj,$ A008_{a b c d e f g h}    -> \nabla_{e f g h}{R_{a b c d}} $)

       return obj

   def get_xDxterm (obj,n,m):

       x^{a}::Weight(label=numx,value=1).
       Dx^{a}::Weight(label=numDx,value=1).

       tmp := @(obj).
       distribute  (tmp)

       foo = Ex("numx = " + str(n))
       bah = Ex("numDx = " + str(m))
       keep_weight (tmp, foo)
       keep_weight (tmp, bah)

       return tmp

   def reformat (obj,scale):
       foo  = Ex(str(scale))
       bah := @(foo) @(obj).
       distribute     (bah)
       bah = product_sort (bah)
       rename_dummies (bah)
       canonicalise   (bah)
       substitute     (bah,$Dx^{b}->zzz^{b}$)
       factor_out     (bah,$x^{a?},zzz^{b?}$)
       substitute     (bah,$zzz^{b}->Dx^{b}$)
       ans := @(bah) / @(foo).
       return ans

   def rescale (obj,scale):
       foo  = Ex(str(scale))
       bah := @(foo) @(obj).
       distribute  (bah)
       substitute  (bah,$Dx^{b}->zzz^{b}$)
       factor_out  (bah,$x^{a?},zzz^{b?}$)
       substitute  (bah,$zzz^{b}->Dx^{b}$)
       return bah

   term0 := @(Y50).  # cdb (term0.101,term0)
   term2 := @(Y52).  # cdb (term2.101,term2)
   term3 := @(Y53).  # cdb (term3.101,term3)
   term4 := @(Y54).  # cdb (term4.101,term4)
   term5 := @(Y55).  # cdb (term5.101,term5)

   term0 = reformat (term0,1)  # cdb (term0.102,term0)
   term2 = reformat (term2,1)  # cdb (term2.102,term2)
   term3 = reformat (term3,1)  # cdb (term3.102,term3)
   term4 = reformat (term4,1)  # cdb (term4.102,term4)
   term5 = reformat (term5,1)  # cdb (term5.102,term5)

   xDxterm12 = get_xDxterm (term2,1,2)   # cdb(xDxterm12.101,xDxterm12)

   xDxterm13 = get_xDxterm (term3,1,3)   # cdb(xDxterm13.101,xDxterm13)
   xDxterm22 = get_xDxterm (term3,2,2)   # cdb(xDxterm22.101,xDxterm22)

   xDxterm14 = get_xDxterm (term4,1,4)   # cdb(xDxterm14.101,xDxterm14)
   xDxterm23 = get_xDxterm (term4,2,3)   # cdb(xDxterm23.101,xDxterm23)
   xDxterm32 = get_xDxterm (term4,3,2)   # cdb(xDxterm32.101,xDxterm32)

   xDxterm15 = get_xDxterm (term5,1,5)   # cdb(xDxterm15.101,xDxterm15)
   xDxterm24 = get_xDxterm (term5,2,4)   # cdb(xDxterm24.101,xDxterm24)
   xDxterm33 = get_xDxterm (term5,3,3)   # cdb(xDxterm33.101,xDxterm33)
   xDxterm42 = get_xDxterm (term5,4,2)   # cdb(xDxterm42.101,xDxterm42)

   xDxterm12 = rescale ( reformat (xDxterm12,    3),     3 )   # cdb(xDxterm12.102,xDxterm12)

   xDxterm13 = rescale ( reformat (xDxterm13,   12),   -12 )   # cdb(xDxterm13.102,xDxterm13)
   xDxterm22 = rescale ( reformat (xDxterm22,   24),   -24 )   # cdb(xDxterm22.102,xDxterm22)

   xDxterm14 = rescale ( reformat (xDxterm14,  180),  -180 )   # cdb(xDxterm14.102,xDxterm14)
   xDxterm23 = rescale ( reformat (xDxterm23,  720),  -720 )   # cdb(xDxterm23.102,xDxterm23)
   xDxterm32 = rescale ( reformat (xDxterm32,  720),  -720 )   # cdb(xDxterm32.102,xDxterm32)

   xDxterm15 = rescale ( reformat (xDxterm15,  360),  -360 )   # cdb(xDxterm15.102,xDxterm15)
   xDxterm24 = rescale ( reformat (xDxterm24, 2160), -2160 )   # cdb(xDxterm24.102,xDxterm24)
   xDxterm33 = rescale ( reformat (xDxterm33, 1080), -1080 )   # cdb(xDxterm33.102,xDxterm33)
   xDxterm42 = rescale ( reformat (xDxterm42,  360),  -360 )   # cdb(xDxterm42.102,xDxterm42)

   checkpoint.append (term0)
   checkpoint.append (term2)
   checkpoint.append (term3)
   checkpoint.append (term4)
   checkpoint.append (term5)

\end{cadabra}

\clearpage

% =================================================================================================
\section*{Tranformation between two RNC frames}

\begin{align*}
     y^{a} = \ny{0}^{a} + \ny{2}^{a} + \ny{3}^{a} + \ny{4}^{a} + \ny{5}^{a} + \BigO{\eps^6}
\end{align*}

\begin{dgroup*}
   \begin{dmath*} \ny{0}^{a} = \cdb{term0.102} \end{dmath*}
   \begin{dmath*} \ny{2}^{a} = \cdb{term2.102} \end{dmath*}
   \begin{dmath*} \ny{3}^{a} = \cdb{term3.102} \end{dmath*}
   \begin{dmath*} \ny{4}^{a} = \cdb{term4.102} \end{dmath*}
   \begin{dmath*} \ny{5}^{a} = \cdb{term5.102} \end{dmath*}
\end{dgroup*}

\clearpage

% =================================================================================================
\section*{Tranformation between two RNC frames}

Same as before but with an improved format (maybe) for the expressions.

\begin{align}
   y^{a} = \ny{0}^{a} + \ny{2}^{a} + \ny{3}^{a} + \ny{4}^{a} + \ny{5}^{a} + \BigO{\eps^6}
\end{align}

\begin{dgroup}
   \begin{dmath} \ny{0}^{a} = Dx^{a} \end{dmath}
\end{dgroup}

\begin{dgroup}
   \begin{dmath} \ny{2}^{a} = \ny{2}^{a}_1 \end{dmath}
   \begin{dmath}   3 \ny{2}^{a}_1 = \cdb{xDxterm12.102} \end{dmath}
\end{dgroup}

\begin{dgroup}
   \begin{dmath} \ny{3}^{a} = \ny{3}^{a}_1 + \ny{3}^{a}_2 \end{dmath}
   \begin{dmath} -12 \ny{3}^{a}_1 = \cdb{xDxterm13.102} \end{dmath}
   \begin{dmath} -24 \ny{3}^{a}_2 = \cdb{xDxterm22.102} \end{dmath}
\end{dgroup}

\begin{dgroup}
   \begin{dmath} \ny{4}^{a} = \ny{4}^{a}_1 + \ny{4}^{a}_2 + \ny{4}^{a}_3 \end{dmath}
   \begin{dmath} -180 \ny{4}^{a}_1 = \cdb{xDxterm14.102} \end{dmath}
   \begin{dmath} -720 \ny{4}^{a}_2 = \cdb{xDxterm23.102} \end{dmath}
   \begin{dmath} -720 \ny{4}^{a}_3 = \cdb{xDxterm32.102} \end{dmath}
\end{dgroup}

\begin{dgroup}
   \begin{dmath} \ny{5}^{a} = \ny{5}^{a}_1 + \ny{5}^{a}_2 + \ny{5}^{a}_3 + \ny{5}^{a}_4 \end{dmath}
   \begin{dmath}  -360 \ny{5}^{a}_1 = \cdb{xDxterm15.102} \end{dmath}
   \begin{dmath} -2160 \ny{5}^{a}_2 = \cdb{xDxterm24.102} \end{dmath}
   \begin{dmath} -1080 \ny{5}^{a}_3 = \cdb{xDxterm33.102} \end{dmath}
   \begin{dmath}  -360 \ny{5}^{a}_4 = \cdb{xDxterm42.102} \end{dmath}
\end{dgroup}

% =================================================================================================
% export checkpoints in json format

\bgroup
\CdbSetup{action=hide}
\begin{cadabra}
   for i in range( len(checkpoint) ):
      cdblib.put ('check{:03d}'.format(i),checkpoint[i],checkpoint_file)
\end{cadabra}
\egroup

\end{document}

### 2019-09-24 source/cadabra/sqrt-detg.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{pylatex}

\begin{document}

Expand $\sqrt{g}$ in powers of $R,\nabla R$ etc.

\begin{python}
   from sympy import *

   eps, a, b, c, d = symbols('\epsilon a b c d')

   ans = sqrt(1+a*eps**2+b*eps**3+c*eps**4+d*eps**5)   # py(ans.001,ans)
   ans = ans.series(eps, 0, 6)                         # py(ans.002,ans)
   ans = simplify(ans)                                 # py(ans.003,ans)

\end{python}

\begin{dgroup*}
   \begin{dmath*} \py*{ans.001} \end{dmath*}
   \begin{dmath*} \py*{ans.002} \end{dmath*}
   \begin{dmath*} \py*{ans.003} \end{dmath*}
\end{dgroup*}

And while we're here, let's also expand $\log(g)$ in powers of $R,\nabla R$ etc.

\begin{python}
   from sympy import *

   eps, a, b, c, d = symbols('\epsilon a b c d')

   ans = log(1+a*eps**2+b*eps**3+c*eps**4+d*eps**5)    # py(ans.101,ans)
   ans = ans.series(eps, 0, 6)                         # py(ans.102,ans)
   ans = simplify(ans)                                 # py(ans.103,ans)

\end{python}

\begin{dgroup*}
   \begin{dmath*} \py*{ans.101} \end{dmath*}
   \begin{dmath*} \py*{ans.102} \end{dmath*}
   \begin{dmath*} \py*{ans.103} \end{dmath*}
\end{dgroup*}

\end{document}

### 2019-09-24 source/cadabra/tests/semantic/connection.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

\section*{\jobname}

\CdbSetup{action=hide}

\begin{cadabra}
   import shared

   import cdblib

   term00A = cdblib.get ('check000','expected/connection.json')
   term01A = cdblib.get ('check001','expected/connection.json')
   term02A = cdblib.get ('check002','expected/connection.json')
   term03A = cdblib.get ('check003','expected/connection.json')
   term04A = cdblib.get ('check004','expected/connection.json')
   term05A = cdblib.get ('check005','expected/connection.json')

   term00B = cdblib.get ('check000','output/connection.json')
   term01B = cdblib.get ('check001','output/connection.json')
   term02B = cdblib.get ('check002','output/connection.json')
   term03B = cdblib.get ('check003','output/connection.json')
   term04B = cdblib.get ('check004','output/connection.json')
   term05B = cdblib.get ('check005','output/connection.json')

   diff000 = shared.check (term00A,term00B)   # cdb (diff000,diff000)
   diff001 = shared.check (term01A,term01B)   # cdb (diff001,diff001)
   diff002 = shared.check (term02A,term02B)   # cdb (diff002,diff002)
   diff003 = shared.check (term03A,term03B)   # cdb (diff003,diff003)
   diff004 = shared.check (term04A,term04B)   # cdb (diff004,diff004)
   diff005 = shared.check (term05A,term05B)   # cdb (diff005,diff005)

\end{cadabra}

\begin{dgroup*}
   \Dmath*{ \cdb*{diff000} }
   \Dmath*{ \cdb*{diff001} }
   \Dmath*{ \cdb*{diff002} }
   \Dmath*{ \cdb*{diff003} }
   \Dmath*{ \cdb*{diff004} }
   \Dmath*{ \cdb*{diff005} }
\end{dgroup*}

\end{document}

### 2019-09-24 source/cadabra/tests/semantic/dGamma.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

\section*{\jobname}

\CdbSetup{action=hide}

\begin{cadabra}
   import shared

   import cdblib

   term00A = cdblib.get ('check000','expected/dGamma.json')
   term01A = cdblib.get ('check001','expected/dGamma.json')
   term02A = cdblib.get ('check002','expected/dGamma.json')
   term03A = cdblib.get ('check003','expected/dGamma.json')
   term04A = cdblib.get ('check004','expected/dGamma.json')

   term00B = cdblib.get ('check000','output/dGamma.json')
   term01B = cdblib.get ('check001','output/dGamma.json')
   term02B = cdblib.get ('check002','output/dGamma.json')
   term03B = cdblib.get ('check003','output/dGamma.json')
   term04B = cdblib.get ('check004','output/dGamma.json')

   diff000 = shared.check (term00A,term00B)   # cdb (diff000,diff000)
   diff001 = shared.check (term01A,term01B)   # cdb (diff001,diff001)
   diff002 = shared.check (term02A,term02B)   # cdb (diff002,diff002)
   diff003 = shared.check (term03A,term03B)   # cdb (diff003,diff003)
   diff004 = shared.check (term04A,term04B)   # cdb (diff004,diff004)

\end{cadabra}

\begin{dgroup*}
   \Dmath*{ \cdb*{diff000} }
   \Dmath*{ \cdb*{diff001} }
   \Dmath*{ \cdb*{diff002} }
   \Dmath*{ \cdb*{diff003} }
   \Dmath*{ \cdb*{diff004} }
\end{dgroup*}

\end{document}

### 2019-09-24 source/cadabra/tests/semantic/dRabcd.tex
Initial commit
This is where it all begins ...

--- 2019-09-24 source/cadabra/tests/semantic/dGamma.tex
@@
 
-   term00A = cdblib.get ('check000','expected/dGamma.json')
-   term01A = cdblib.get ('check001','expected/dGamma.json')
-   term02A = cdblib.get ('check002','expected/dGamma.json')
-   term03A = cdblib.get ('check003','expected/dGamma.json')
-   term04A = cdblib.get ('check004','expected/dGamma.json')
+   term00A = cdblib.get ('check000','expected/dRabcd.json')
+   term01A = cdblib.get ('check001','expected/dRabcd.json')
+   term02A = cdblib.get ('check002','expected/dRabcd.json')
+   term03A = cdblib.get ('check003','expected/dRabcd.json')
+   term04A = cdblib.get ('check004','expected/dRabcd.json')
 
-   term00B = cdblib.get ('check000','output/dGamma.json')
-   term01B = cdblib.get ('check001','output/dGamma.json')
-   term02B = cdblib.get ('check002','output/dGamma.json')
-   term03B = cdblib.get ('check003','output/dGamma.json')
-   term04B = cdblib.get ('check004','output/dGamma.json')
+   term00B = cdblib.get ('check000','output/dRabcd.json')
+   term01B = cdblib.get ('check001','output/dRabcd.json')
+   term02B = cdblib.get ('check002','output/dRabcd.json')
+   term03B = cdblib.get ('check003','output/dRabcd.json')
+   term04B = cdblib.get ('check004','output/dRabcd.json')
 

### 2019-09-24 source/cadabra/tests/semantic/detg2.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

\section*{\jobname}

\CdbSetup{action=hide}

\begin{cadabra}
   import shared

   import cdblib

   term00A = cdblib.get ('check000','expected/detg2.json')
   term01A = cdblib.get ('check001','expected/detg2.json')
   term02A = cdblib.get ('check002','expected/detg2.json')

   term00B = cdblib.get ('check000','output/detg2.json')
   term01B = cdblib.get ('check001','output/detg2.json')
   term02B = cdblib.get ('check002','output/detg2.json')

   diff000 = shared.check (term00A,term00B)   # cdb (diff000,diff000)
   diff001 = shared.check (term01A,term01B)   # cdb (diff001,diff001)
   diff002 = shared.check (term02A,term02B)   # cdb (diff002,diff002)

\end{cadabra}

\begin{dgroup*}
   \Dmath*{ \cdb*{diff000} }
   \Dmath*{ \cdb*{diff001} }
   \Dmath*{ \cdb*{diff002} }
\end{dgroup*}

\end{document}

### 2019-09-24 source/cadabra/tests/semantic/gen2rnc.tex
Initial commit
This is where it all begins ...

--- 2019-09-24 source/cadabra/tests/semantic/dRabcd.tex
@@
 
-   term00A = cdblib.get ('check000','expected/dRabcd.json')
-   term01A = cdblib.get ('check001','expected/dRabcd.json')
-   term02A = cdblib.get ('check002','expected/dRabcd.json')
-   term03A = cdblib.get ('check003','expected/dRabcd.json')
-   term04A = cdblib.get ('check004','expected/dRabcd.json')
+   term00A = cdblib.get ('check000','expected/gen2rnc.json')
+   term01A = cdblib.get ('check001','expected/gen2rnc.json')
+   term02A = cdblib.get ('check002','expected/gen2rnc.json')
+   term03A = cdblib.get ('check003','expected/gen2rnc.json')
+   term04A = cdblib.get ('check004','expected/gen2rnc.json')
 
-   term00B = cdblib.get ('check000','output/dRabcd.json')
-   term01B = cdblib.get ('check001','output/dRabcd.json')
-   term02B = cdblib.get ('check002','output/dRabcd.json')
-   term03B = cdblib.get ('check003','output/dRabcd.json')
-   term04B = cdblib.get ('check004','output/dRabcd.json')
+   term00B = cdblib.get ('check000','output/gen2rnc.json')
+   term01B = cdblib.get ('check001','output/gen2rnc.json')
+   term02B = cdblib.get ('check002','output/gen2rnc.json')
+   term03B = cdblib.get ('check003','output/gen2rnc.json')
+   term04B = cdblib.get ('check004','output/gen2rnc.json')
 

### 2019-09-24 source/cadabra/tests/semantic/genGamma.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

\section*{\jobname}

\CdbSetup{action=hide}

\begin{cadabra}
   import shared

   import cdblib

   term00A = cdblib.get ('check000','expected/genGamma.json')
   term01A = cdblib.get ('check001','expected/genGamma.json')
   term02A = cdblib.get ('check002','expected/genGamma.json')
   term03A = cdblib.get ('check003','expected/genGamma.json')
   term04A = cdblib.get ('check004','expected/genGamma.json')
   term05A = cdblib.get ('check005','expected/genGamma.json')
   term06A = cdblib.get ('check005','expected/genGamma.json')
   term07A = cdblib.get ('check005','expected/genGamma.json')
   term08A = cdblib.get ('check005','expected/genGamma.json')
   term09A = cdblib.get ('check005','expected/genGamma.json')

   term00B = cdblib.get ('check000','output/genGamma.json')
   term01B = cdblib.get ('check001','output/genGamma.json')
   term02B = cdblib.get ('check002','output/genGamma.json')
   term03B = cdblib.get ('check003','output/genGamma.json')
   term04B = cdblib.get ('check004','output/genGamma.json')
   term05B = cdblib.get ('check005','output/genGamma.json')
   term06B = cdblib.get ('check005','output/genGamma.json')
   term07B = cdblib.get ('check005','output/genGamma.json')
   term08B = cdblib.get ('check005','output/genGamma.json')
   term09B = cdblib.get ('check005','output/genGamma.json')

   diff000 = shared.check (term00A,term00B)   # cdb (diff000,diff000)
   diff001 = shared.check (term01A,term01B)   # cdb (diff001,diff001)
   diff002 = shared.check (term02A,term02B)   # cdb (diff002,diff002)
   diff003 = shared.check (term03A,term03B)   # cdb (diff003,diff003)
   diff004 = shared.check (term04A,term04B)   # cdb (diff004,diff004)
   diff005 = shared.check (term05A,term05B)   # cdb (diff005,diff005)
   diff006 = shared.check (term06A,term06B)   # cdb (diff006,diff006)
   diff007 = shared.check (term07A,term07B)   # cdb (diff007,diff007)
   diff008 = shared.check (term08A,term08B)   # cdb (diff008,diff008)
   diff009 = shared.check (term09A,term09B)   # cdb (diff009,diff009)

\end{cadabra}

\begin{dgroup*}
   \Dmath*{ \cdb*{diff000} }
   \Dmath*{ \cdb*{diff001} }
   \Dmath*{ \cdb*{diff002} }
   \Dmath*{ \cdb*{diff003} }
   \Dmath*{ \cdb*{diff004} }
   \Dmath*{ \cdb*{diff005} }
   \Dmath*{ \cdb*{diff006} }
   \Dmath*{ \cdb*{diff007} }
   \Dmath*{ \cdb*{diff008} }
   \Dmath*{ \cdb*{diff009} }
\end{dgroup*}

\end{document}

### 2019-09-24 source/cadabra/tests/semantic/geodesic-bvp.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

\section*{\jobname}

\CdbSetup{action=hide}

\begin{cadabra}
   import shared

   import cdblib

   term00A = cdblib.get ('check000','expected/geodesic-bvp.json')
   term01A = cdblib.get ('check001','expected/geodesic-bvp.json')
   term02A = cdblib.get ('check002','expected/geodesic-bvp.json')
   term03A = cdblib.get ('check003','expected/geodesic-bvp.json')
   term04A = cdblib.get ('check004','expected/geodesic-bvp.json')
   term05A = cdblib.get ('check005','expected/geodesic-bvp.json')
   term06A = cdblib.get ('check005','expected/geodesic-bvp.json')
   term07A = cdblib.get ('check005','expected/geodesic-bvp.json')
   term08A = cdblib.get ('check005','expected/geodesic-bvp.json')
   term09A = cdblib.get ('check005','expected/geodesic-bvp.json')
   term10A = cdblib.get ('check005','expected/geodesic-bvp.json')
   term00B = cdblib.get ('check000','output/geodesic-bvp.json')
   term01B = cdblib.get ('check001','output/geodesic-bvp.json')
   term02B = cdblib.get ('check002','output/geodesic-bvp.json')
   term03B = cdblib.get ('check003','output/geodesic-bvp.json')
   term04B = cdblib.get ('check004','output/geodesic-bvp.json')
   term05B = cdblib.get ('check005','output/geodesic-bvp.json')
   term06B = cdblib.get ('check005','output/geodesic-bvp.json')
   term07B = cdblib.get ('check005','output/geodesic-bvp.json')
   term08B = cdblib.get ('check005','output/geodesic-bvp.json')
   term09B = cdblib.get ('check005','output/geodesic-bvp.json')
   term10B = cdblib.get ('check005','output/geodesic-bvp.json')

   diff000 = shared.check (term00A,term00B)   # cdb (diff000,diff000)
   diff001 = shared.check (term01A,term01B)   # cdb (diff001,diff001)
   diff002 = shared.check (term02A,term02B)   # cdb (diff002,diff002)
   diff003 = shared.check (term03A,term03B)   # cdb (diff003,diff003)
   diff004 = shared.check (term04A,term04B)   # cdb (diff004,diff004)
   diff005 = shared.check (term05A,term05B)   # cdb (diff005,diff005)
   diff006 = shared.check (term06A,term06B)   # cdb (diff006,diff006)
   diff007 = shared.check (term07A,term07B)   # cdb (diff007,diff007)
   diff008 = shared.check (term08A,term08B)   # cdb (diff008,diff008)
   diff009 = shared.check (term09A,term09B)   # cdb (diff009,diff009)
   diff010 = shared.check (term10A,term10B)   # cdb (diff010,diff010)

\end{cadabra}

\begin{dgroup*}
   \Dmath*{ \cdb*{diff000} }
   \Dmath*{ \cdb*{diff001} }
   \Dmath*{ \cdb*{diff002} }
   \Dmath*{ \cdb*{diff003} }
   \Dmath*{ \cdb*{diff004} }
   \Dmath*{ \cdb*{diff005} }
   \Dmath*{ \cdb*{diff006} }
   \Dmath*{ \cdb*{diff007} }
   \Dmath*{ \cdb*{diff008} }
   \Dmath*{ \cdb*{diff009} }
   \Dmath*{ \cdb*{diff010} }
\end{dgroup*}

\end{document}

### 2019-09-24 source/cadabra/tests/semantic/geodesic-ivp.tex
Initial commit
This is where it all begins ...

--- 2019-09-24 source/cadabra/tests/semantic/connection.tex
@@
 
-   term00A = cdblib.get ('check000','expected/connection.json')
-   term01A = cdblib.get ('check001','expected/connection.json')
-   term02A = cdblib.get ('check002','expected/connection.json')
-   term03A = cdblib.get ('check003','expected/connection.json')
-   term04A = cdblib.get ('check004','expected/connection.json')
-   term05A = cdblib.get ('check005','expected/connection.json')
+   term00A = cdblib.get ('check000','expected/geodesic-ivp.json')
+   term01A = cdblib.get ('check001','expected/geodesic-ivp.json')
+   term02A = cdblib.get ('check002','expected/geodesic-ivp.json')
+   term03A = cdblib.get ('check003','expected/geodesic-ivp.json')
+   term04A = cdblib.get ('check004','expected/geodesic-ivp.json')
+   term05A = cdblib.get ('check005','expected/geodesic-ivp.json')
 
-   term00B = cdblib.get ('check000','output/connection.json')
-   term01B = cdblib.get ('check001','output/connection.json')
-   term02B = cdblib.get ('check002','output/connection.json')
-   term03B = cdblib.get ('check003','output/connection.json')
-   term04B = cdblib.get ('check004','output/connection.json')
-   term05B = cdblib.get ('check005','output/connection.json')
+   term00B = cdblib.get ('check000','output/geodesic-ivp.json')
+   term01B = cdblib.get ('check001','output/geodesic-ivp.json')
+   term02B = cdblib.get ('check002','output/geodesic-ivp.json')
+   term03B = cdblib.get ('check003','output/geodesic-ivp.json')
+   term04B = cdblib.get ('check004','output/geodesic-ivp.json')
+   term05B = cdblib.get ('check005','output/geodesic-ivp.json')
 

### 2019-09-24 source/cadabra/tests/semantic/geodesic-lsq.tex
Initial commit
This is where it all begins ...

--- 2019-09-24 source/cadabra/tests/semantic/connection.tex
@@
 
-   term00A = cdblib.get ('check000','expected/connection.json')
-   term01A = cdblib.get ('check001','expected/connection.json')
-   term02A = cdblib.get ('check002','expected/connection.json')
-   term03A = cdblib.get ('check003','expected/connection.json')
-   term04A = cdblib.get ('check004','expected/connection.json')
-   term05A = cdblib.get ('check005','expected/connection.json')
+   term00A = cdblib.get ('check000','expected/geodesic-lsq.json')
+   term01A = cdblib.get ('check001','expected/geodesic-lsq.json')
+   term02A = cdblib.get ('check002','expected/geodesic-lsq.json')
+   term03A = cdblib.get ('check003','expected/geodesic-lsq.json')
+   term04A = cdblib.get ('check004','expected/geodesic-lsq.json')
+   term05A = cdblib.get ('check005','expected/geodesic-lsq.json')
 
-   term00B = cdblib.get ('check000','output/connection.json')
-   term01B = cdblib.get ('check001','output/connection.json')
-   term02B = cdblib.get ('check002','output/connection.json')
-   term03B = cdblib.get ('check003','output/connection.json')
-   term04B = cdblib.get ('check004','output/connection.json')
-   term05B = cdblib.get ('check005','output/connection.json')
+   term00B = cdblib.get ('check000','output/geodesic-lsq.json')
+   term01B = cdblib.get ('check001','output/geodesic-lsq.json')
+   term02B = cdblib.get ('check002','output/geodesic-lsq.json')
+   term03B = cdblib.get ('check003','output/geodesic-lsq.json')
+   term04B = cdblib.get ('check004','output/geodesic-lsq.json')
+   term05B = cdblib.get ('check005','output/geodesic-lsq.json')
 

### 2019-09-24 source/cadabra/tests/semantic/metric-inv.tex
Initial commit
This is where it all begins ...

\documentclass[12pt]{cdblatex}

\begin{document}

\section*{\jobname}

\CdbSetup{action=hide}

\begin{cadabra}
   import shared

   import cdblib

   term00A = cdblib.get ('check000','expected/metric-inv.json')
   term01A = cdblib.get ('check001','expected/metric-inv.json')
   term02A = cdblib.get ('check002','expected/metric-inv.json')
   term03A = cdblib.get ('check003','expected/metric-inv.json')
   term04A = cdblib.get ('check004','expected/metric-inv.json')
   term05A = cdblib.get ('check005','expected/metric-inv.json')
   term06A = cdblib.get ('check005','expected/metric-inv.json')
   term07A = cdblib.get ('check005','expected/metric-inv.json')

   term00B = cdblib.get ('check000','output/metric-inv.json')
   term01B = cdblib.get ('check001','output/metric-inv.json')
   term02B = cdblib.get ('check002','output/metric-inv.json')
   term03B = cdblib.get ('check003','output/metric-inv.json')
   term04B = cdblib.get ('check004','output/metric-inv.json')
   term05B = cdblib.get ('check005','output/metric-inv.json')
   term06B = cdblib.get ('check005','output/metric-inv.json')
   term07B = cdblib.get ('check005','output/metric-inv.json')

   diff000 = shared.check (term00A,term00B)   # cdb (diff000,diff000)
   diff001 = shared.check (term01A,term01B)   # cdb (diff001,diff001)
   diff002 = shared.check (term02A,term02B)   # cdb (diff002,diff002)
   diff003 = shared.check (term03A,term03B)   # cdb (diff003,diff003)
   diff004 = shared.check (term04A,term04B)   # cdb (diff004,diff004)
   diff005 = shared.check (term05A,term05B)   # cdb (diff005,diff005)
   diff006 = shared.check (term06A,term06B)   # cdb (diff006,diff006)
   diff007 = shared.check (term07A,term07B)   # cdb (diff007,diff007)

\end{cadabra}

\begin{dgroup*}
   \Dmath*{ \cdb*{diff000} }
   \Dmath*{ \cdb*{diff001} }
   \Dmath*{ \cdb*{diff002} }
   \Dmath*{ \cdb*{diff003} }
   \Dmath*{ \cdb*{diff004} }
   \Dmath*{ \cdb*{diff005} }
   \Dmath*{ \cdb*{diff006} }
   \Dmath*{ \cdb*{diff007} }
\end{dgroup*}

\end{document}

### 2019-09-24 source/cadabra/tests/semantic/metric.tex
Initial commit
This is where it all begins ...

--- 2019-09-24 source/cadabra/tests/semantic/metric-inv.tex
@@
 
-   term00A = cdblib.get ('check000','expected/metric-inv.json')
-   term01A = cdblib.get ('check001','expected/metric-inv.json')
-   term02A = cdblib.get ('check002','expected/metric-inv.json')
-   term03A = cdblib.get ('check003','expected/metric-inv.json')
-   term04A = cdblib.get ('check004','expected/metric-inv.json')
-   term05A = cdblib.get ('check005','expected/metric-inv.json')
-   term06A = cdblib.get ('check005','expected/metric-inv.json')
-   term07A = cdblib.get ('check005','expected/metric-inv.json')
+   term00A = cdblib.get ('check000','expected/metric.json')
+   term01A = cdblib.get ('check001','expected/metric.json')
+   term02A = cdblib.get ('check002','expected/metric.json')
+   term03A = cdblib.get ('check003','expected/metric.json')
+   term04A = cdblib.get ('check004','expected/metric.json')
+   term05A = cdblib.get ('check005','expected/metric.json')
+   term06A = cdblib.get ('check005','expected/metric.json')
+   term07A = cdblib.get ('check005','expected/metric.json')
 
-   term00B = cdblib.get ('check000','output/metric-inv.json')
-   term01B = cdblib.get ('check001','output/metric-inv.json')
-   term02B = cdblib.get ('check002','output/metric-inv.json')
-   term03B = cdblib.get ('check003','output/metric-inv.json')
-   term04B = cdblib.get ('check004','output/metric-inv.json')
-   term05B = cdblib.get ('check005','output/metric-inv.json')
-   term06B = cdblib.get ('check005','output/metric-inv.json')
-   term07B = cdblib.get ('check005','output/metric-inv.json')
+   term00B = cdblib.get ('check000','output/metric.json')
+   term01B = cdblib.get ('check001','output/metric.json')
+   term02B = cdblib.get ('check002','output/metric.json')
+   term03B = cdblib.get ('check003','output/metric.json')
+   term04B = cdblib.get ('check004','output/metric.json')
+   term05B = cdblib.get ('check005','output/metric.json')
+   term06B = cdblib.get ('check005','output/metric.json')
+   term07B = cdblib.get ('check005','output/metric.json')
+
+   # bug: can't push this function into shared.py
+   #      no synatx error, but cadabra doesn't cancel equal terms
+   # see ~/cadabra/bugs/bug02
+
+   def check (objA,objB):
+       tmp := @(objA) - @(objB).
+       distribute         (tmp)
+       tmp = shared.standard_indices (tmp)
+       tmp = shared.product_sort (tmp)
+       rename_dummies     (tmp)
+       canonicalise       (tmp)
+
+       return tmp
 

### 2019-09-24 source/cadabra/tests/semantic/rnc2gen.tex
Initial commit
This is where it all begins ...

--- 2019-09-24 source/cadabra/tests/semantic/dRabcd.tex
@@
 
-   term00A = cdblib.get ('check000','expected/dRabcd.json')
-   term01A = cdblib.get ('check001','expected/dRabcd.json')
-   term02A = cdblib.get ('check002','expected/dRabcd.json')
-   term03A = cdblib.get ('check003','expected/dRabcd.json')
-   term04A = cdblib.get ('check004','expected/dRabcd.json')
+   term00A = cdblib.get ('check000','expected/rnc2gen.json')
+   term01A = cdblib.get ('check001','expected/rnc2gen.json')
+   term02A = cdblib.get ('check002','expected/rnc2gen.json')
+   term03A = cdblib.get ('check003','expected/rnc2gen.json')
+   term04A = cdblib.get ('check004','expected/rnc2gen.json')
 
-   term00B = cdblib.get ('check000','output/dRabcd.json')
-   term01B = cdblib.get ('check001','output/dRabcd.json')
-   term02B = cdblib.get ('check002','output/dRabcd.json')
-   term03B = cdblib.get ('check003','output/dRabcd.json')
-   term04B = cdblib.get ('check004','output/dRabcd.json')
+   term00B = cdblib.get ('check000','output/rnc2gen.json')
+   term01B = cdblib.get ('check001','output/rnc2gen.json')
+   term02B = cdblib.get ('check002','output/rnc2gen.json')
+   term03B = cdblib.get ('check003','output/rnc2gen.json')
+   term04B = cdblib.get ('check004','output/rnc2gen.json')
 

### 2019-09-24 source/cadabra/tests/semantic/rnc2rnc.tex
Initial commit
This is where it all begins ...

--- 2019-09-24 source/cadabra/tests/semantic/dRabcd.tex
@@
 
-   term00A = cdblib.get ('check000','expected/dRabcd.json')
-   term01A = cdblib.get ('check001','expected/dRabcd.json')
-   term02A = cdblib.get ('check002','expected/dRabcd.json')
-   term03A = cdblib.get ('check003','expected/dRabcd.json')
-   term04A = cdblib.get ('check004','expected/dRabcd.json')
+   term00A = cdblib.get ('check000','expected/rnc2rnc.json')
+   term01A = cdblib.get ('check001','expected/rnc2rnc.json')
+   term02A = cdblib.get ('check002','expected/rnc2rnc.json')
+   term03A = cdblib.get ('check003','expected/rnc2rnc.json')
+   term04A = cdblib.get ('check004','expected/rnc2rnc.json')
 
-   term00B = cdblib.get ('check000','output/dRabcd.json')
-   term01B = cdblib.get ('check001','output/dRabcd.json')
-   term02B = cdblib.get ('check002','output/dRabcd.json')
-   term03B = cdblib.get ('check003','output/dRabcd.json')
-   term04B = cdblib.get ('check004','output/dRabcd.json')
+   term00B = cdblib.get ('check000','output/rnc2rnc.json')
+   term01B = cdblib.get ('check001','output/rnc2rnc.json')
+   term02B = cdblib.get ('check002','output/rnc2rnc.json')
+   term03B = cdblib.get ('check003','output/rnc2rnc.json')
+   term04B = cdblib.get ('check004','output/rnc2rnc.json')
 

### 2019-09-24 source/tex/paper.tex
Initial commit
This is where it all begins ...

\documentclass[a4paper,12pt]{article}
\usepackage[style=numeric,
            sorting=none,
            sortcites=true]{biblatex}
% \usepackage[style=numeric,
%             sorting=none,
%             sortcites=true,
%             bibencoding=ascii,
%             backend=biber]{biblatex}
\addbibresource{bibdata.bib}
\usepackage[dvipsnames]{xcolor}
\usepackage{listings}
\usepackage{amsmath}
\usepackage{breqn}
% \usepackage[colorlinks=false]{hyperref}% for journals
\usepackage[colorlinks=true,citecolor=red]{hyperref}% for me
% \usepackage{showlabels}
% \usepackage{showkeys}
\usepackage{cdblatex}
\usepackage{geometry}
\usepackage{paper}

\geometry{a4paper,portrait,margin=2cm}

\numberwithin{equation}{section}

\begin{document}

\title{\bfseries Riemann Normal Coordinate expansions using Cadabra}
\author{%
Leo Brewin\\[10pt]%
School of Mathematical Sciences\\%
Monash University, 3800\\%
Australia}
% \date{10-Nov-2008}% 1st version started
% \date{29-Nov-2008}% 1st version finished
% \date{10-Mar-2009}% 1st version to gr-qc
% \date{22-Mar-2009}% minor corrections to section 8.2
% \date{09-Jul-2009}% minor grammatical corrections in response to 1st referee
% \date{12-Aug-2009}% small corrections arising from proofs received from CQG
\date{12-Aug-2009 (v1)\\
      19-Apr-2019 (v2)}% major changes to correct errors in 4th and higher order terms,
                       % all code will be available on github
\reference{Preprint: arXiv:0903.2087\\[5pt]
           Journal: {\it Class.Quantum.Grav.} {\bf 26} 175017 (2009)}

\maketitle

\begin{abstract}
\noindent
%
Riemann normal coordinate expansions of the metric and other geometrical quantities,
including the geodesic arc-length, will be presented. All of the results are given to
fifth-order in the curvature and were obtained using the computer algebra package Cadabra.

\end{abstract}

\BegNote{\bf\large Version 2}
The first version of this paper was written in 2009. This second version (written in 2019)
differs from the first in a number of important ways. First, the discussion on Cadabra has
been completely rewritten to account for the revised syntax introduced in the latest version
of Cadabra (itself updated in 2016). Second, and more importantly, a significant
computational error in the first version has been corrected. Though the algorithms given in
the first version were correct their implementation in Cadabra (by this author) was
incorrect. This lead to errors first in the cubic and higher order terms in the generalised
connections and then, down the line, to computations built upon the generalised connections
(e.g., the initial and boundary value problem and the arc-length equations). The Cadabra
codes have been corrected and thoroughly cross-checked. A number of minor typographical
errors have also been corrected.
\EndNote

% ============================================================================================
\section{Introduction}

In a previous paper \cite{brewin:2009-02} a series of simple examples were used to
demonstrate how the computer algebra program Cadabra
(\cite{peeters:2007-02},
\cite{peeters:2018-01},
\cite{peeters:2017-01},
\cite{peeters:2017-02})
could be employed to do the kinds of tensor computations often encountered in General
Relativity. The examples were deliberately chosen to be sufficiently simple as to allow the
reader to appreciate the basic functionality of Cadabra. This left open the question of how
well Cadabra might perform on more challenging computations. The purpose of this paper is to
address that point by showing how Cadabra can be used to express various geometrical
objects, such as the metric, the connection and the geodesic arc-length, in terms of Riemann
normal coordinates (also known as geodesic coordinates). This is a standard computation in
differential geometry that leads to a series expansion for the various geometrical objects
in powers of the curvature and its derivatives. The first few terms are rather easy to
compute by hand but further progress, to higher order terms, quickly becomes prohibitively
difficult. Thus some computer tool, in this case Cadabra, is essential when computing these
higher order terms.

The first part of this paper will be concerned with purely mathematical and algorithmic
issues including a definition of Riemann normal coordinates, how they can be constructed
(from generic coordinates) and what properties they confer upon geometrical objects (such as
the metric, the connection etc.). In the second part the attention will shift to issues
specific to Cadabra. This will be followed by full details of the 5th-order expansion of the
metric and other objects.

It should be emphasised that the primary objective of this paper is to demonstrate that
Cadabra is a mature program that can easily handle an otherwise prohibitive computation.
Some of the Cadabra results for the metric do replicate results by other authors (e.g.,
equation \eqref{eqn:metric} is consistent with that given in
\cite{hatzinikitas:2000-01},
\cite{higashijima:2002-01},
\cite{yamashita:1984-01},
and
\cite{muller:1997-01}
while some results are original (in particular, the solution for the two-point boundary value
problem, \eqref{eqn:bvp6}).

Before turning the discussion to the definition and construction of Riemann normal
coordinates, a simple coordinate transformation will be introduced. This will help in making
clear what is meant by the statement that an expression is an expansion to a given order.

% ============================================================================================
\section{Conformal coordinates}\label{sec:Conformal}

Each algorithm given later in this paper yields polynomial approximations to particular
geometric quantities (e.g., the metric). Higher order approximations are obtained by
recursive application of the algorithms.

The primary goal of this section is to provide meaning to the statement that the polynomial
$S_{\eps}$ is an expansion of $S$ up to and including terms of order $\BigO{\eps^n}$. The
key to this definition is the use of a conformal transformation of the original metric.

Consider some neighbourhood of $O$ and let $\eps$ be a typical length scale for $O$ (for
example, $\eps$ might be the length of the longest geodesic that passes through $O$ and
confined by the neighbourhood). Construct any regular set of coordinates $x^a$ (i.e., such
that the metric components are non-singular) in the neighbourhood of $O$ and let the
coordinates of $O$ be $x^a_{\star}$. The word \emph{patch} will be used to denote the
neighbourhood of $O$ in which these coordinates are defined. Now define a new set of
coordinates $y^a$ by
\[
x^a = x^a_{\star} + \eps y^a
\]
and thus
\[
  ds^2 = g_{ab}(x)dx^a dx^b
       = \eps^2 g_{ab}(x_{\star} + \eps y) dy^a dy^b
\]
The conformal metric $d\tilde s$ can now be defined by $d\tilde s = ds/\eps^2$. This leads to
\[
d{\tilde s}^2 = g_{ab}(x_{\star} + \eps y) dy^a dy^b
              = {\tilde g}_{ab}(y,\eps) dy^a dy^b
\]
and
\[
{\tilde g}_{ab} = g_{ab}\>,\qquad
{\tilde g}_{ab,c} = \eps g_{ab,c}\>,\qquad
{\tilde g}_{ab,cd} = \eps^2 g_{ab,cd}\qquad \text{at }O
\]
where the partial derivatives on the left are with respect to $y$ and those on
the right are with respect to $x$. Since $g_{ab}(x_{\star})$ does not depend on $\eps$
it is easy to see that
\begin{equation*}
{\tilde g}_{ab,i_1i_2i_3\cdots i_n} = \BigO{\eps^n}\qquad \text{at }O
\end{equation*}
From this it follows, by simple inspection of the standard equations, that
\begin{align*}
{\tilde\Gamma}^{a}{}_{bc,i_1i_2i_3\cdots i_n} &= \BigO{\eps^{n+1}}\qquad \text{at }O\\[5pt]
{\tilde R}^{a}{}_{bcd,i_1i_2i_3\cdots i_n} &= \BigO{\eps^{n+2}}\qquad \text{at }O
\end{align*}

There are now two ways to look at the patch. It can be viewed as a patch of length scale
$\eps$ with a curvature independent of $\eps$. Or it can be viewed as a patch of fixed size
but with a curvature that depends on $\eps$ (and where the limit $\eps\rightarrow0$
corresponds to flat space). This later view is useful since it ensures that the series
expansions around flat space can be made convergent (by choosing a sufficiently small
$\eps$).

These conformal coordinates will be used for the remainder of this paper. As there is no
longer any reason to distinguish between $x^a$ and $y^a$, the $y^a$ will be replaced with
$x^a$. The $x^a$ will now be treated as generic coordinates (but keep in mind that there is
an underlying conformal transformation built into these coordinates).

Finally, the statement that $S_\eps$ is an order $\BigO{\eps^n}$ expansion of $S$ will be
taken to mean that $n$ is the smallest positive integer for which
\[
0 < \lim_{\eps\rightarrow0} \frac{\vert S-S_\eps\vert}{\eps^{n+1}} < M
\]
for some finite positive $M$ i.e., if $S$ were expanded as a Taylor series in $\eps$ around
$\eps=0$ then $S$ and $S_\eps$ would differ by terms proportional to $\eps^{n+1}$.

% --------------------------------------------------------------------------------------------

% The bookkeeping of the terms in the expansions goes as follows. Every expansion is in the
% form of an infinite series in the curvature tensor and its derivatives. The leading
% curvature term $R_{abcd}$ is given a weight of $\eps^2$. Each derivative incurs an extra
% factor of $\eps$. Thus, for example, $R_{abcd,ef}$ is of order $\BigO{\eps^4}$ while
% terms such as $R_{abcd,e}R^{abcd,e}$ and $R_{abcd}R^{abpq}R^{cd}{}_{pq}$ are both of
% order $\BigO{\eps^6}$.

% ============================================================================================
\section{Riemann Normal Coordinates}\label{sec:RNC}

The basic idea behind Riemann normal coordinates is to use the geodesics through a given
point to define the coordinates for nearby points. Let the given point be $O$ (this will be
the origin of the Riemann normal frame) and consider some nearby point $P$. If $P$ is
sufficiently close to $O$ then there exists a unique geodesic joining $O$ to $P$. Let $v^a$
be the components of the unit tangent vector to this geodesic at $O$ and let $s$ be the
geodesic arc length measured from $O$ to $P$. Then the Riemann normal coordinates $x^a$ of
$P$ relative to $O$ are defined to be $x^a = s v^a$. These coordinates are well defined
provided the geodesics do not cross (which can always be ensured by choosing the
neighbourhood of $O$ to be sufficiently small). For more mathematical details see
(\cite{chern-chen-lam:2000-01},
\cite{chavel:2006-01},
\cite{eisenhart:1926-01})
and in particular the elegant exposition by Gray
(\cite{gray:1973-01}, \cite{willmore:1996-01}).

One trivial consequence of this definition is that all geodesics through $O$ are of the form
$x^a(s) = s v^a$ and that the $v^a$ are constant along each geodesic. This implies, by
direct substitution into the geodesic equation, that $\Gamma^c{}_{ab}=0$ at $O$ which in
turn implies that $g_{ab,c}=0$ at $O$. Suppose now that the metric can be written as a
Taylor series in $x^a$ about $O$. In that series only the zero, second and higher
derivatives of the $g_{ab}$ would appear. Thus the leading terms of the metric can be
expressed as a sum of a constant part (the leading term) plus a curvature part (from the
second and higher derivative terms). Similar expansions arise for other geometrical
quantities (e.g., geodesics, arc length) in terms of a flat space part plus a curvature
contribution.

Suppose, as is almost always the case, that the coordinates $x^a$ are \emph{not} in Riemann
normal form. How might they be transformed to a local set of Riemann normal coordinates? A
first attempt might be to make direct appeal to the basic definition, namely, $y^a = s v^a$.
Unfortunately this does raise an immediate problem. The quantities $v^a$ are rarely known
explicitly but must instead be computed by solving a two-point boundary value problem. This
is non-trivial but it can be dealt with in a number of ways. The approach adopted here is
to iterate over a sequence of initial value problems to provide a sequence of approximations
to the solution of the original boundary value problem. Here is a brief outline of the
process -- this serves as an introduction to the next two sections where the full details
will be given.

First, consider any method for computing the coordinates $x^a(s)$ of a typical geodesic that
originates from $O$. This solution of the typical initial value problem will depend on two
integration constants, $x^a$ and ${\Dot x}^a$, being the respective values of $x^a(s)$ and
$dx^a/ds$ at $s=0$. Next, embed this first step in an iterative scheme (e.g., a fixed-point
scheme) to produce successive approximations to ${\Dot x}^a$ so that the geodesic passes
through not only $O$ but also $P$. The final step, assuming the sequence converges, is to set
$v^a$ to be ${\Dot x}^a$ and finally $y^a = s v^a$.

The next two sub-sections provide further details on how the initial and boundary value
problems were solved.

% --------------------------------------------------------------------------------------------
\subsection{The initial value problem}\label{sub:CoordsIVP}

The aim here is to obtain a Taylor series, about the point $O$, for the solution of the
geodesic equation
\[
0 = \frac{d^2x^a}{ds^2} + \Gamma^a{}_{bc}(x) \frac{dx^b}{ds}\frac{dx^c}{ds}
\]
subject to the initial conditions $x^a(s) = x^a$ and $dx^a/ds={\Dot x}^a$ at $s=0$.

First choose $s=0$ at $O$ and then write the Taylor series for $x^a(s)$ as
\[
x^a(s) = \left.x^a\right\vert_{s=0}
       + s \left.\frac{dx^a}{ds}\right\vert_{s=0}
       + \sum_{n=2}^\infty\> \frac{s^n}{n!}\left.\frac{d^nx^a}{ds^n}\right\vert_{s=0}
\]
The second and higher derivatives can be obtained by successive differentiation of the
geodesic equation leading to
\begin{equation}
x^a(s) = x^a + s {\Dot x}^a
       - \sum_{n=2}^\infty \frac{s^n}{n!}\Gamma^{a}{}_{i_1i_2i_3\cdots i_n}
                                         {\Dot x}^{i_1}
                                         {\Dot x}^{i_2}
                                         {\Dot x}^{i_3}\cdots
                                         {\Dot x}^{i_n}
\pglabel{eqn:SolIVP}
\end{equation}
where the $\Gamma^{a}{}_{i_1i_2i_3\cdots i_n}$, known as \emph{generalised connections},
are defined recursively by
\begin{equation}
\Gamma^{a}{}_{i_1i_2i_3\cdots i_n} = \Gamma^{a}{}_{(i_1i_2i_3\cdots i_{n-1},i_n)}
                                 - n \Gamma^{a}{}_{p(i_2i_3\cdots i_{n-1}}
                                     \Gamma^{p}{}_{i_1i_n)}
\pglabel{eqn:GenGam}
\end{equation}
Note that the use of round brackets $(...)$ denotes total symmetrisation over the included
indices (see Appendix A for more details).

A convenient shorthand for equation \eqref{eqn:GenGam} in terms of covariant derivatives can
be obtained by ignoring (in this context alone) the single upper index. This leads to the
compact notation
\begin{equation}
\Gamma^{a}{}_{i_1i_2i_3\cdots i_n} = \Gamma^{a}{}_{(i_1i_2;i_3i_4i_5\cdots i_n)}
\pglabel{eqn:gamCovDeriv}
\end{equation}

% --------------------------------------------------------------------------------------------
\subsection{The boundary value problem}\label{sub:CoordsBVP}

The question here is: How can ${\Dot x}^a$ be chosen so that the geodesic passes through not
only $O$ but also $P$? The geodesic, by construction, already passes through $O$
so that leaves $P$ as the main focus of attention. Suppose that the coordinates of $P$ are
$x^a + \Delta x^a$ where $x^a$ are the coordinates of $O$. Let $s_P$ be the geodesic
distance from $O$ to $P$. The challenge now is to solve
\[
\Delta x^a = s_P {\Dot x}^a
           - \sum_{n=2}^\infty \frac{s_P^n}{n!}\Gamma^{a}{}_{i_1i_2i_3\cdots i_n}
                                               {\Dot x}^{i_1}
                                               {\Dot x}^{i_2}
                                               {\Dot x}^{i_3}\cdots
                                               {\Dot x}^{i_n}
\]
for ${\Dot x}^a$ in terms of $\Delta x^a$ and the generalised connections.

Put $y^a = s_P {\Dot x}^a$ (this introduces the Riemann normal coordinates) and re-arrange
the equation into the form
\begin{equation}
y^a = \Delta x^a + \sum_{n=2}^\infty \frac{1}{n!}\Gamma^{a}{}_{i_1i_2i_3\cdots i_n}
                   y^{i_1} y^{i_2} y^{i_3}\cdots y^{i_n}
\pglabel{eqn:solveBVP}
\end{equation}
In a small variation to the plan given above, the plan will be modified slightly to solve
this equation for $y^a$ (rather than ${\Dot x}^a$) by constructing a sequence of
approximations $y^a_m$ to $y^a$. The $y^a$ will, through the generalised connections, depend
on the Riemann curvature and its derivatives (at $O$). So in principle the $y^a$ could be
expanded as power series in $\eps$ (i.e., as a power series in the curvatures). The $y^a_m$
will be chosen to be the Taylor polynomial of $y^a$ to order $\eps^m$. That is, $y^a_m$ is a
polynomial in the curvatures (and its derivatives) up to and including terms of order
$\BigO{\eps^m}$. The $y^a_m$ can computed by truncating both sides of \eqref{eqn:solveBVP}
to terms no higher than $\BigO{\eps^m}$. But note that the $\Gamma^{a}{}_{i_1i_2i_3\cdots
i_n}$ are of order $\BigO{\eps^{n-1}}$. The upshot is that the infinite series may be
truncated at $n=m$ while still retaining all terms up to and including $\eps^m$. This leads
to
\[
y^a_m = \Delta x^a
      + T^m_{\eps}\left(
             \sum_{n=2}^{n=m} \frac{1}{n!}\Gamma^{a}{}_{i_1i_2i_3\cdots i_n}
                  y^{i_1}_m y^{i_2}_m y^{i_3}_m\cdots y^{i_n}_m
                  \right)
\]
where $T^m_{\eps}$ is a simple truncation operator (it deletes all terms of order
$\BigO{\eps^{m+1}}$ or higher). This is a marginal improvement on \eqref{eqn:solveBVP} (at
least it is a finite series) but it is still a non-linear equation for $y^a_m$. But
fortunately improvements are readily made. Notice, once again, that
$\Gamma^{a}{}_{i_1i_2i_3\cdots i_n} = \BigO{\eps^{n-1}}$ and this allows the use of lower
order estimates for $y^a$ in the product terms on the right hand side. This leads to
\begin{equation}
y^a_m = \Delta x^a
      + T^m_{\eps}\left(
             \sum_{n=2}^{n=m} \frac{1}{n!}\Gamma^{a}{}_{i_1i_2i_3\cdots i_n}
                  y^{i_1}_{m-n+1} y^{i_2}_{m-n+1} y^{i_3}_{m-n+1}\cdots y^{i_n}_{m-n+1}
                  \right)
\pglabel{eqn:iterBVP}
\end{equation}
This shows that $y^a_m$ appears only on the left hand side and thus this equation can be
used to recursively compute $y^a_p$ for $p=2,3,4,\cdots$. Here are the first few $y^a_m$.
Starting with the lowest order approximation,
\[
y^a_0 = \Delta x^a
\]
and as there are no $\eps^1$ terms in \eqref{eqn:solveBVP} this can also be written as
\[
y^a_1 = y^a_0 = \Delta x^a
\]
Now set $m=2$ in equation \eqref{eqn:iterBVP} to obtain
\begin{align*}
y^a_2 &= \Delta x^a
       + T^2_\eps\left(\frac{1}{2!} \Gamma^a{}_{i_1i_2} y^{i_1}_1 y^{i_2}_1\right)\\[3pt]
      &= \Delta x^a + \frac{1}{2} \Gamma^a{}_{i_1i_2} \Delta x^{i_1}_1 \Delta x^{i_2}_1
\end{align*}
and once more, with $m=3$, with the result
\begin{align*}
y^a_3 &= \Delta x^a
       + T^3_\eps\left(
             \frac{1}{2!} \Gamma^a{}_{i_1i_2} y^{i_1}_2 y^{i_2}_2
            + \frac{1}{3!} \Gamma^a{}_{i_1i_2i_3} y^{i_1}_1 y^{i_2}_1 y^{i_3}_1\right)\\[3pt]
      &= \Delta x^a + \frac{1}{2} \Gamma^a{}_{i_1i_2} \Delta x^{i_1} \Delta x^{i_2}
                    + \frac{1}{6} \left(  \Gamma^a{}_{bi_1}\Gamma^b{}_{i_2i_3}
                                        + \Gamma^a{}_{i_1i_2,i_3}\right) \Delta x^{i_1}
                                                                         \Delta x^{i_2}
                                                                         \Delta x^{i_3}
\end{align*}
This process may seem simple but looks can be deceiving -- the higher order $y^a_m$ contain
a profusion of terms that, when computed by hand, are largely unmanageable beyond $m\approx
7$. At this point there is little option but, if higher order terms are needed, to defer the
computations to a suitable computer algebra package. This point will be discussed in more
detail in section (\ref{sec:Cadabra}).

This completes the first objective -- to find a way to transform from any non-singular set
of coordinates to a local set of Riemann normal coordinates. The question to ask now is --
What form does the metric (and other objects) take in these coordinates? This is the subject
of the second next section. But before doing so, here is a short digression to introduce
some new notation.

% ============================================================================================
\section{Notation}\label{sec:Notation}

It has already been noted that the recursive nature of equations like \eqref{eqn:GenGam}
imposes a significant computational cost with each iteration. A less obvious side effect is
that number of tensor indices on each term will also increase with each iteration. The
equation for the generalised connections, \eqref{eqn:GenGam}, is a case in point. This
contains an index list of the form $i_1i_2i_3\cdots i_n$. Lists such as these are tedious to
write and are prone to transcription error (when in the hands of humans). It thus makes sense
to use a notation that is easy to read and write while also not detracting from the meaning
in the expression. The proposal is that a sequence of indices such as $i_1i_2i_3\cdots i_n$
be replaced with a single index of the form $\ui$. In this notation the equation for the
generalised connections \eqref{eqn:GenGam} would be written as
\[
\Gamma^{a}{}_{b\uc d} = \Gamma^{a}{}_{(b\uc,d)}
                - (n+1) \Gamma^{a}{}_{p(\uc}
                        \Gamma^{p}{}_{bd)}
\]
where $\uc$ contains $n>0$ indices. There will be cases where the number of hidden indices
needs to be made clear. In such cases one of two options can be taken -- either state
explicitly the number of hidden indices in words (as in the above example) or by including
the number as a subscript as in this example
\[
\Gamma^{a}{}_{b\uc_n d} = \Gamma^{a}{}_{(b\uc_n,d)}
                  - (n+1) \Gamma^{a}{}_{p(\uc_n}
                          \Gamma^{p}{}_{bd)}
\]
This notation can be extended to handle expressions such as
\[
\Gamma^a{}_{bc,i_1i_2i_3\cdots i_n}A^{i_1}A^{i_2}A^{i_3}\cdots A^{i_n}
\]
which would be written in condensed form as
\[
\Gamma^a{}_{bc,\ud}A^{\Xdot\ud}
\]
Note carefully the dot to the left of the $\ud$ superscript. Its purpose is to avoid an
ambiguity that would arise in using the condensed notation for an expression such as
\[
\Gamma^a{}_{bc,i_1i_2i_3\cdots i_n}A^{i_1i_2i_3\cdots i_n}
\]
which, following the earlier discussion, would be written as
\[
\Gamma^a{}_{bc,\ud}A^{\ud}
\]
These simple changes bring some degree of normalcy to the printed form but those gains
rapidly pale into insignificance when displaying the fifth order expansions generated by
Cadabra (in section (\ref{sec:fifthOrderRNC})). Cadabra's output does not employ the
familiar use of brackets to denote symmetrisation over an index list. Instead, it uses the
fully expanded form which, on paper, can lead to dramatic explosion in otherwise similar
looking terms. There seems little point in printing terms that differ only by swapping pairs
of indices. Thus the convention used here will be that symmetrisation over an index list on
the right hand side of an equation will be inferred from any explicit use of brackets for
symmetrisation written on the left hand side. Thus an equation like
\[
A_{(ab)} = B_a C_b
\]
will be taken to mean
\[
A_{(ab)} = \frac{1}{2!}\left (B_a C_b + B_b C_a\right)
\]
This convention will only be applied to results generated by Cadabra. The left hand side
will only contain a single term with just one pair of brackets.

The final notational device concerns cases where an index needs to be excluded from
symmetrisation. The normal practise is to enclose the index in a pair of vertical lines.
This involves two characters whereas a single dot over the chosen index can serve the same
purpose. Thus $(ab{\Dot c}d{\Dot e}fg)$ will denote symmetrisation over only $a,b,d,f$ and
$g$. In the standard notation this would have been written as $(ab\vert c\vert d\vert e\vert
fg)$.

% ============================================================================================
\section{The metric in Riemann normal form}\label{sec:MetricRNC}

In the preceding section the generic and Riemann normal coordinates distinguished by using
the symbols $x^a$ and $y^a$ respectively. Now, for notational convenience and to accord with
convention, the Riemann normal coordinates will be denoted $x^a$ while $y^a$ will be
stripped of any special meaning.

The aim of this section is to express the metric in Riemann normal form. This will take the
form of an infinite series in powers of the curvature and its derivatives. Start by writing
out the Taylor series for the metric around $x^a=0$
\[
g_{ab}(x) = g_{ab} + \sum_{n=1}^\infty\> \frac{1}{n!}\>g_{ab,\uc}\>x^{\Xdot\uc}
\]
where $\uc$ contains $n$ indices and $g_{ab}$ are constants (e.g.
$g_{ab}=\diag(1,1,1,\cdots)$).

The present task is to express the partial derivatives of the metric in terms of the Riemann
tensor. From the standard definition of a metric compatible connection, a series of partial
derivatives leads to
\[
g_{ab,c\ud} = \left(g_{ae}\Gamma^{e}{}_{bc}+g_{eb}\Gamma^{e}{}_{ac}\right)_{,\ud}
\]
and since $g_{ab,c\ud}$ is totally symmetric in $c\ud$ it is easy to see that
\[
g_{ab,c\ud} = \left(g_{ae}\Gamma^{e}{}_{b(c}\right){}_{\!,\ud)}
            + \left(g_{eb}\Gamma^{e}{}_{a(c}\right){}_{\!,\ud)}
\]
Two points should be noted, first, the connection appears only in the form
$\Gamma^{a}{}_{b(c,\ud)}$, second, the left hand side contains derivatives one order higher
than in the corresponding terms on the right hand side. The upshot is that this equation can
be used to recursively compute all of the metric derivatives solely in terms of the
$\Gamma^{a}{}_{b(c,\ud)}$ and the constants $g_{ab}$. In this way the above Taylor series
for the metric can be expressed solely in terms of the connection and its derivatives. But
more can be done -- the derivatives of the connection must surely tie in with the
curvatures. Thus attention turns to the standard definition for the curvature, which after a
series of derivatives, can be written in the form
\begin{equation}
R^a{}_{(bc{\Dot d},\ue)} = \Gamma^a{}_{d(bc,\ue)} - \Gamma^a{}_{(bc,\ue)d}
                         + \left(\Gamma^a{}_{i(c}\Gamma^i{}_{b{\Dot d}}\right){}_{\!,\ue)}
                         - \left(\Gamma^a{}_{id}\Gamma^i{}_{(bc}\right){}_{\!,\ue)}
\pglabel{eqn:RabcdDerivs}
\end{equation}
(Note that the $\Gamma^a{}_{dbc}$ in $\Gamma^a{}_{d(bc,\ue)}$ is the first of the
generalised connections discussed earlier in section (\ref{sub:CoordsIVP})). Can this result
be used to eliminate the connection and its derivatives from the metric? Yes, but only after
specialising to the Riemann normal coordinates.

Recall that, in Riemann normal coordinates, all geodesics through $O$ are of the form
\[
x^a(s) = s v^a
\]
which upon substitution into the geodesic equations leads to
\[
0 = \Gamma^a{}_{(b\uc)}\qquad\text{at $O$}
\]
It follows, by recursion on equation \eqref{eqn:gamCovDeriv}, that
\begin{equation}
0 = \Gamma^a{}_{(bc,\ud)}\qquad\text{at $O$}
\pglabel{eqn:rncGamDeriv}
\end{equation}
Recall that $\Gamma^a{}_{bc,\ue d}$ is separately symmetric in its first pair of indices and
in the remaining $(n+1)$ lower indices (assuming $\ue$ contains $n$ indices). Thus using
equation \eqref{eqn:doubleSymm} it follows that
\[
0 = (n+3)\Gamma^a{}_{(bc,\ue d)} = 2\Gamma^{a}{}_{d(b,c\ue)}
                                 + (n+1)\Gamma^{a}{}_{(bc,\ue)d}
\]
This result can be used to eliminate the $\Gamma^a{}_{(bc,\ue)d}$ term in equation
\eqref{eqn:RabcdDerivs} for the derivatives of the curvature. The result, after a minor
shuffling of terms is
\[
(n+3)\Gamma^a{}_{d(b,c\ue)} =
        (n+1)\left( R^a{}_{(bc\Dot d,\ue)}
                   - \left(\Gamma^a{}_{i(c}\Gamma^i{}_{b{\Dot d}}\right){}_{\!,\ue)}
                   + \left(\Gamma^a{}_{id}\Gamma^i{}_{(bc}\right){}_{\!,\ue)}
             \right)
\]
(the reason for rearranging the terms will become clear in a moment). Note also that the
last term in the previous equation can be eliminated by equation \eqref{eqn:rncGamDeriv} and
the product rule.

In summary, the equations of interest are
\begin{gather}
\pglabel{eqn:metricTaylor}
g_{ab}(x) = g_{ab} + \sum_{n=1}^\infty\> \frac{1}{n!}\>g_{ab,\uc}\>x^{\Xdot\uc}\\[5pt]
\pglabel{eqn:metricGamma}
g_{ab,c\ud} = \left(g_{ae}\Gamma^{e}{}_{b(c}\right){}_{\!,\ud)}
             + \left(g_{eb}\Gamma^{e}{}_{a(c}\right){}_{\!,\ud)}\\[10pt]
\pglabel{eqn:GammaRiemann}
(n+3)\Gamma^a{}_{d(b,c\ue)} =
        (n+1)\left( R^a{}_{(bc\Dot d,\ue)}
                   - \left(\Gamma^a{}_{i(c}\Gamma^i{}_{b{\Dot d}}\right){}_{\!,\ue)}
              \right)
\end{gather}
These equations could be used as follows. First, equation \eqref{eqn:GammaRiemann} is used to
recursively compute the $\Gamma^a{}_{b(c,\ue d)}$ in terms of the Riemann tensor and its
partial derivatives (this was the reason behind the shuffling of terms noted above). Note
that $\ue$ in equation \eqref{eqn:GammaRiemann} contains $n$ hidden indices. The
$\Gamma^a{}_{b(c,\ud)}$ are then substituted into \eqref{eqn:metricGamma} which in turn is
used to recursively express all of the $g_{ab,\uc}$ in terms of the Riemann tensor and its
partial derivatives. When the dust settles the result is a finite series expansion for the
metric in terms of the Riemann tensor and its partial derivatives. The result, to 3rd-order
in the curvature, is
%
\begin{dmath*}g_{ab}(x) = \cdb{metric4} + \BigO{\eps^4}\end{dmath*}
%
Though this result meets the stated aim -- to express the metric in terms of the curvatures
and its derivatives -- there is a small problem. The partial derivatives make subsequent
raising and lowering of indices very tedious. A far better result would be one that employs
covariant derivatives. Fortunately the work required to achieve that is not too onerous. The
details are described in the remainder of this section.

It is not hard to see that a series of covariant derivatives of the Riemann tensor,
would lead to an equation of the form, in any coordinate frame,
\begin{equation*}
R^{a}{}_{(bc{\Dot d};\ue)} = R^{a}{}_{(bc{\Dot d},\ue)} + Q^{a}{}_{(bc{\Dot d}\ue)}
\end{equation*}
where $Q^{a}{}_{(bc{\Dot d}\ue)}$ is a function of the $\Gamma^{a}{}_{bc}$, the
$R^{a}{}_{bcd}$ and their partial derivatives. If this is going to sit nicely with the
algorithm given above then it will be necessary to show, in the Riemann normal frame, that
this equation only contains connection terms of the form $\Gamma^{p}{}_{q(r,\us)}$.
Fortunately this is rather easy to do.

Each term of the form $\Gamma^{p}{}_{qr,\us}$ in $Q$ arose during one round of covariant
differentiation. Thus \emph{at least one} of the indices $q,r$ and \emph{all} of the indices
in $\us$ must be drawn from the index list $\ue$. If both $q$ and $r$ are contained in $\ue$
then the term is of the form $\Gamma^{p}{}_{(qr,\us)}$ and thus will vanish when specialised
to the Riemann normal frame. This completes the proof. By re-arranging the above equation
into the following form
\begin{equation}
R^{a}{}_{(bc{\Dot d},\ue)} = R^{a}{}_{(bc{\Dot d};\ue)} - Q^{a}{}_{(bc{\Dot d}\ue)}
\pglabel{eqn:RabcdPartials}
\end{equation}
it can be used to recursively compute all of the partial derivatives of the curvatures in
terms of their covariant derivatives. The $Q^{a}{}_{(bc{\Dot d}\ue)}$ will contain lower
order derivatives of the curvatures and partial derivatives of the connection all of which
can be eliminated (in favour of covariant derivatives) using previously computed results.
For the first two derivatives it turns out that the partial and covariant derivatives are
equal (as expected) but differences do appear in higher order derivatives. These differences
will be apparent when the fifth order results are given later in section
(\ref{sub:partialRabcd}).

% ============================================================================================
\section{The inverse metric in Riemann normal form}\label{sec:InvMetricRNC}

Most of the hard work is done and it is now time to develop algorithms for Riemann normal
expansions for other interesting quantities, in this instance the inverse metric
$g^{ab}(x)$. The previous section used $0=g_{ab;\uc}$ as a starting point to express the
metric in terms of the curvatures. On this occasion, for the inverse metric, the starting
point will be $0=g^{ab}{}_{;\uc}$. Then, following a path similar to that used in the
previous section, leads to the following equations
\begin{gather}
\pglabel{eqn:InvMetricTaylor}
g^{ab}(x) = g^{ab} + \sum_{n=1}^\infty\> \frac{1}{n!}\>g^{ab}{}_{,\uc}\>x^{\Xdot\uc}\\[5pt]
\pglabel{eqn:InvMetricGamma}
g^{ab}{}_{,c\ud} = - \left(g^{ae}\Gamma^{b}{}_{e(c}\right){}_{\!,\ud)}
                   - \left(g^{eb}\Gamma^{a}{}_{e(c}\right){}_{\!,\ud)}\\[10pt]
(n+3)\Gamma^a{}_{d(b,c\ue)} =
        (n+1)\left( R^a{}_{(bc\Dot d,\ue)}
                   - \left(\Gamma^a{}_{i(c}\Gamma^i{}_{b{\Dot d}}\right){}_{\!,\ue)}
              \right)\tag{\ref{eqn:GammaRiemann}}
\end{gather}
These equations can be used to construct the series expansion for $g^{ab}(x)$, which to
3rd-order is
%
\begin{dmath*}g^{ab}(x) = \cdb{metric4.inv} + \BigO{\eps^4}\end{dmath*}
%

% ============================================================================================
\section{Generalised connections}\label{sec:GenGamma}

In section (\ref{sub:CoordsIVP}) the generalised connections $\Gamma^{a}{}_{bc\ud}$ were
shown to arise from successive differentiation of the geodesic equation and that they can be
computed recursively using
\[
\Gamma^{a}{}_{b\uc d} = \Gamma^{a}{}_{(b\uc,d)}
                - (n+1) \Gamma^{a}{}_{p(\uc}
                        \Gamma^{p}{}_{bd)}
\tag{\ref{eqn:GenGam}}
\]
where the index list $\uc$ contains $n>0$ indices.

Here are the first three generalised connections
%
%-------------------------------------------------------------------------------
\def\genGammaA{{\Gamma^{a}_{(bc)}(x)}}%
\def\genGammaB{{\Gamma^{a}_{(bcd)}(x)}}%
\def\genGammaC{{\Gamma^{a}_{(bcde)}(x)}}%
\def\genGammaD{{\Gamma^{a}_{(bcdef)}(x)}}%
%-------------------------------------------------------------------------------
%
\begin{dgroup*}[compact,spread={3pt}]
\begin{dmath*}\genGammaA = \cdb{genGamma04}+\BigO{\eps^4}\end{dmath*}%
\begin{dmath*}\genGammaB = \cdb{genGamma14}+\BigO{\eps^4}\end{dmath*}%
\begin{dmath*}\genGammaC = \BigO{\eps^4}\end{dmath*}
\end{dgroup*}
%

% ============================================================================================
\section{Geodesics}\label{sec:GeodesicsRNC}

The discussion so far has been framed mostly around the set of geodesics that pass though
the point $O$. There are of course many geodesics within the patch of $O$ that do not pass
through $O$. The challenge now is to construct any geodesic within the patch of $O$. The
approach followed here will be similar to that taken previously in which solutions of the
geodesic initial value problem are used to solve the geodesic boundary value problem.

% --------------------------------------------------------------------------------------------
\subsection{Geodesic initial value problem}\label{sec:GeodesicIVP}

Consider a point $P$ distinct from $O$. At $P$ it is reasonable to assume that the
generalised connections do not vanish (which is generally true, the exception being flat
space). Thus the coordinates $x^a$ in the neighbourhood of $P$ do not constitute a Riemann
normal frame relative at $P$. But as $P$ lies in the patch for $O$ it follows (by definition
of the patch) that the metric is non-singular at $P$ and thus a new set of Riemann normal
coordinates, $y^a$, with $P$ as the origin, can always be constructed.

This problem has been discussed once before, in section (\ref{sub:CoordsBVP}). Using equation
\eqref{eqn:SolIVP} and the generalised connections from section (\ref{sec:GenGamma}) leads to
%
\begin{dmath*}x^{a}(s)
   = x^{a}
   + s \dotx^{a}
   + \frac{1}{24} s^2 {\Dot x}^b {\Dot x}^c \Bigl(\cdb{ivp42}\Bigr)
   + \frac{1}{12} s^3 {\Dot x}^b {\Dot x}^c {\Dot x}^d \cdb{ivp43}
   + \BigO{s^4,\eps^4}\end{dmath*}
%
% --------------------------------------------------------------------------------------------
\subsection{Geodesic boundary value problem}\label{sec:GeodesicBVP}

Consider now the case of three distinct points $O$, $P$ and $Q$. The goal in this section is
to compute the geodesic that passes through $P$ and $Q$. Start by using equation
\eqref{eqn:iterBVP} for the generalised connections from section (\ref{sec:GenGamma}) to
obtain
%
\begin{dgroup*}[compact,spread={3pt}]
\begin{dmath*} x^{a}(\lambda) = \cdb{bvp4} + \BigO{\eps^4} \end{dmath*}
\end{dgroup*}
where $\lambda=s/L_{PQ}$ is the scaled geodesic distance from $P(\lambda=0)$ to
$Q(\lambda=1)$.

% ============================================================================================
\section{Geodesic arc-length}\label{sec:ArcLength}

Given the explicit expressions for the metric and the geodesic that joins the points $P$ and
$Q$, the length of that geodesic can be computed by way of the integral
\[
L_{PQ} = \int_P^Q\>\left(g_{ab}(x)\frac{dx^a}{ds}\frac{dx^b}{ds}\right)^{1/2}\>ds
\]
Up to this point the parameter $s$ has been taken to be the proper distance along the
geodesic. However, after careful inspection of the geodesic path \eqref{eqn:SolIVP} it is
clear that any uniform scaling of $s$ is allowed. Thus it is possible to re-scale $s$ so
that $s=0$ at $P$ and $s=1$ at $Q$ (of course, the parameter $s$ no longer measures proper
distance so it may be better to use a different symbol, say $\lambda$). Furthermore, a
standard result states that the integrand is constant along the geodesic and can thus be
evaluated at any point such as $P$. The integral is now trivial to evaluate with the result
\[
L^2_{PQ} = \left.g_{ab}(x)\frac{dx^a}{ds}\frac{dx^b}{ds}\right\vert_{P}
\]
which, using previous results, leads to
%
\begin{dmath*}L^2_{PQ} = \cdb{lsq4} + \BigO{\eps^4}\end{dmath*}
%
A simple calculation shows that this result can also be written in the slightly more
suggestive form
\[
L^2_{PQ} = g_{ab}(\bar x) Dx^a Dx^b + \BigO{\eps^4}
\]
where ${\bar x}^a = (x^a_P+x^a_Q)/2$ is the \emph{coordinate} mid-point of the geodesic.
This does raise the question -- Can higher order estimates for $L^2$ (see equation
\ref{eqn:geodesicLSQ}) be obtained by sampling the metric at suitablly chosen points on
(or near) the geodesic? That question will be left for another occassion.

% ============================================================================================
\section{Cadabra}\label{sec:Cadabra}

Though the above equations seem simple they do impose a significant computational cost --
their recursive structure quickly leads to expressions that are too hard to manage (for
humans) beyond the first few iterations. The computations are best left to computer programs
specifically designed for tenor computations. One such program is Cadabra -- a {\tt C++}
program that reads plain text files to perform the various tensor computations. It was
initially designed for computations in high energy physics but it also very well suited to
tensor computations in general.

Cadabra's syntax is a hybrid of LaTeX to express tensor expressions, Python to coordinate the
computations and some unique Cadabra syntax to describe properties of various objects (e.g.,
index sets, symmetries, commutation rules etc.). A number of simple examples, including a
detailed discussion of Cadabra's syntax with particular emphasis for use in General
Relativity, can be found on the GitHub repository \cite{brewin:2019-01}.
This includes full Cadabra sources for all the examples and exercises (with solutions)
covering elementary topics (e.g., verifying $\nabla g=0$ for a metric connection) through
to more advanced topics (e.g., deriving the BSSN equations from the ADM equations).
This material was based on an earlier paper \cite{brewin:2009-02} (written for an
ealier version of Cadabra).

For those readers seeking a flavour of what Cadabra code looks like (before rushing off to
learn Cadabra in detail) here are two short examples. The first shows how Cadabra can be
used to verify that $\nabla g=0$ when $\nabla$ is the metric compatible covariant
derivative. The second example shows how the truncation operator $T^m_{\eps}$ introduced in
section (\ref{sub:CoordsBVP}) can be implemented in Cadabra.

% --------------------------------------------------------------------------------------------
\subsection{The metric connection}\label{sub:MetricGamma}

This simple example (adapted from the earlier paper \cite{brewin:2009-02})
verifies that $\nabla g =0$ when $\nabla$ is the metric compatible covariant derivative.

\lstset{numbers=left,gobble=2}
\CdbSetup{action=verbatim}

\begin{cadabra}
   # Define some properties

   {a,b,c,d,e,f,h,i,j,k,l,m,n,o,p,q,r,s,t,u#}::Indices.

   g_{a b}::Metric.
   g_{a}^{b}::KroneckerDelta.

   \partial_{#}::PartialDerivative.

   # Define a rule for the Christoffel symbol

   Gamma := \Gamma^{a}_{b c} -> (1/2) g^{a d} (  \partial_{b}{g_{d c}}
                                               + \partial_{c}{g_{b d}}
                                               - \partial_{d}{g_{b c}} );

   # Define the covariant derivative of the metric

   cderiv := \partial_{c}{g_{a b}} - g_{a d}\Gamma^{d}_{b c}
                                   - g_{d b}\Gamma^{d}_{a c};

   # Do the computations

   substitute          (cderiv, Gamma);
   distribute          (cderiv);
   eliminate_metric    (cderiv);
   eliminate_kronecker (cderiv);
   canonicalise        (cderiv);
\end{cadabra}

The code is rather easy to follow. The first few lines defines a set of indices (line 3) and
some objects with particular properties (lines 5 to 8). This is followed by two rules for
constructing a connection (line 12) and the covariant derivative of $g_{ab}$ (line 18). The
body of the calculations can be seen in the last few lines (23 to 27) where various
operations (known as \emph{algorithms} in Cadabra's lexicon) are applied. As expected, the
above code returns a value of zero for the object \cdbverb{cderiv}.

A few points are worth noting. Cadabra uses a hybrid syntax of Python and LaTeX. Tensor
equations are defined and results are returned using a subset of LaTeX while computations
(such as substitutions, simplifications etc.) are expressed using a Python syntax. This use
of Python and LaTeX makes for an easy entry into Cadabra programming as most users would
already have familiarity with Python and LaTeX. Using LaTeX to define and record
tensors means that it is very easy to carry results from one Cadabra notebook to another or
even to other LaTeX documents -- all of the Cadabra output shown in this paper were imported
without change from the Cadabra output generated by other documents.

% --------------------------------------------------------------------------------------------
\subsection{Truncation of polynomials}\label{sub:Truncation}

In section (\ref{sub:CoordsBVP}) a truncation operator $T^m_{\eps}$ was introduced. The
question here is -- How might that operator be implemented in Cadabra?

Suppose you are asked to extract the leading terms from an expression such as
\[
P^a(x) =  c^a
        + c^a_{b} x^b
        + c^a_{bc} x^b x^c
        + c^a_{bcd} x^b x^c x^d
        + c^a_{bcde} x^b x^c x^d x^e
\]
One approach (there are others, e.g., emulating a truncated Taylor series) is to use
Cadabra's \verb!::Weight! property and the \verb!keep_weight! algorithm. The idea is to
assign weights to nominated objects (through the \verb!::Weight! property) and then extract
terms matching a chosen weight (using the \verb!keep_weight! algorithm).

Here is a small piece of Cadabra code that does the job.
\begin{cadabra}
   def truncate (obj,n):

       # define the weight and give it a label
       x^{a}::Weight(label=\epsilon).

       # start with an empty espression
       ans = Ex(0)

       # loop over selected terms in the source
       for i in range (0,n+1):

          foo := @(obj).
          bah  = Ex("\epsilon = " + str(i))

          # extract a single term
          keep_weight (foo, bah)

          # update the running sum
          ans = ans + foo

       # all done, return final answer
       return ans

   # the quartic polynomial
   quarticPoly :=   c^{a}
                  + c^{a}_{b} x^{b}
                  + c^{a}_{b c} x^{b} x^{c}
                  + c^{a}_{b c d} x^{b} x^{c} x^{d}
                  + c^{a}_{b c d e} x^{b} x^{c} x^{d} x^{e}.

   # truncate to cubic terms
   cubicPoly = truncate (quarticPoly,3)
\end{cadabra}
The first thing to note is the Python function \cdbverb{truncate}. This simply does as its
name suggests -- it truncates an object to a certain order. How does it do the job? The
first line to note is line 4. This identifies $x^a$ as the target to carry the weights (and
is given the label \verb!\epsilon! to distinguish it from other targets declared by other
instances of \verb!::Weight!). Cadabra now sees the polynomial $P^a(x)$ as if it had been
written as
\[
P^a(x) =   c^a
         + c^a_{b} x^b \eps
         + c^a_{bc} x^b x^c \eps^2
         + c^a_{bcd} x^b x^c x^d \eps^3
         + c^a_{bcde} x^b x^c x^d x^e \eps^4
\]
The Python \cdbverb{for loop} then extracts the requested terms from the source returning
the final truncated answer in \cdbverb{ans}. The final result is exactly as expected -- the
leading cubic part of the original quartic polynomial. This function and variations on it
(e.g., extracting a single term) are used extensively in all of the Cadabra codes used in
this paper.

% ============================================================================================
\section{Expansions to fifth order}\label{sec:fifthOrderRNC}

All of the $\BigO{\eps^6}$ Cadabra programs were not overly demanding on computational
resources, taking between 0.5 second and 12 minutes to run and requiring
between 40 and 360 Mbyte of memory (on a Mac Pro running macOS 10.14.4).

The Cadabra codes and several support scripts are available from the author's GitHub site
\ \url{https://github.com/leo-brewin/cadabra-riemann-coords}.

Some of the following expansions can be compared directly with results obtained by
traditional methods. In particular the expansions for the metric and its inverse, equations
\eqref{eqn:metric} and \eqref{eqn:metricinv} agree exactly with those given in equations
(17) and (18) respectively of \cite{hatzinikitas:2000-01}. Also, equation
\eqref{eqn:genGamma0} agrees with that given in equation (A12) of \cite{calzetta:1988-01}
(though showing that they are equivalent does require some work using the first and second
Bianchi identities).

Note that in the following output any symmetrisation of indices on the right hand side
should be inferred from the explicit symmetrisation (the round brackets) on the left hand
side (as per the discussion in section \ref{sec:Notation}).

\clearpage

% ============================================================================================
% And now for the Cadabra output...
% ============================================================================================

\def\mysubsection#1{\vskip1.5\parskip\leftline{\bf #1}\vskip-1.5\parskip\ignorespaces}
\def\dMath#1#2#3#4{\begin{dmath*}[number={#1}]#2\pglabel{#3}#4(#1)\end{dmath*}}

% --------------------------------------------------------------------------------------------
\mysubsection{The metric}

\begin{dgroup*}[compact,spread={3pt}]
   \dMath{11.1}{g_{(a b)}(x) = \cdb{metric}+\BigO{\eps^6}}{eqn:metric}{\hfill}
\end{dgroup*}

% --------------------------------------------------------------------------------------------
\mysubsection{The inverse metric}

\begin{dgroup*}[compact,spread={3pt}]
   \dMath{11.2}{g^{(a b)}(x) = \cdb{metric.inv}+\BigO{\eps^6}}{eqn:metricinv}{\hfill}
\end{dgroup*}

% --------------------------------------------------------------------------------------------
\mysubsection{The metric determinant}

\begin{dgroup*}[compact,spread={3pt}]
   \dMath{11.3a}{-g(x)        = \cdb{Ndetg}     + \BigO{\eps^6}}{eqn:detg}{\hskip 1cm\hfill}
   \dMath{11.3b}{\sqrt{-g(x)} = \cdb{sqrtNdetg} + \BigO{\eps^6}}{eqn:sqrtdetg}{\hskip 1cm\hfill}
   \dMath{11.3c}{\log{-g(x)}  = \cdb{logNdetg}  + \BigO{\eps^6}}{eqn:logdetg}{\hfill}
\end{dgroup*}

\clearpage

% --------------------------------------------------------------------------------------------
\mysubsection{Generalised connections}

\vskip 10pt

The following results were obtained by recursive application of equation \eqref{eqn:GenGam}.
%
\begin{dgroup*}[compact,spread={3pt}]
   \dMath{11.4}{\Gamma^{a}_{(b c)}(x) = \cdb{genGamma0} + \BigO{\eps^6}}{eqn:genGamma0}{\hfill}
   \dMath{11.5}{\Gamma^{a}_{(b c d)}(x) = \cdb{genGamma1} + \BigO{\eps^6}}{eqn:genGamma1}{\hfill}
   \dMath{11.6}{\Gamma^{a}_{(b c d e)}(x) = \cdb{genGamma2} + \BigO{\eps^6}}{eqn:genGamma2}{\hfill}
   \dMath{11.7}{\Gamma^{a}_{(b c d e f)}(x) = \cdb{genGamma3} + \BigO{\eps^6}}{eqn:genGamma3}{\hfill}
\end{dgroup*}

% --------------------------------------------------------------------------------------------
\mysubsection{Symmetrised partial derivatives of the connection.}\label{sub:partialGamma}

\vskip 10pt

The following results were obtained from equation \eqref{eqn:GammaRiemann}
%
\begin{dgroup*}[compact,spread={3pt}]
   \dMath{11.8}{   3 \Gamma^a{}_{d(b,c)} = \cdb{dGamma61}}{eqn:dGamma61}{\hfill}
   \dMath{11.9}{   6 \Gamma^a{}_{d(b,ce)} =  \cdb{dGamma62}}{eqn:dGamma62}{\hfill}
   \dMath{11.10}{ 15 \Gamma^a{}_{d(b,cef)} =  \cdb{dGamma63}}{eqn:dGamma63}{\hfill}
   \dMath{11.11}{  9 \Gamma^a{}_{d(b,cefg)} = \cdb{dGamma64}}{eqn:dGamma64}{\hfill}
   \dMath{11.12}{252 \Gamma^a{}_{d(b,cefgh)} = \cdb{dGamma65}}{eqn:dGamma65}{\hfill}
\end{dgroup*}

% --------------------------------------------------------------------------------------------
\mysubsection{Symmetrised partial derivatives of the%
              Riemann curvature tensor.}\label{sub:partialRabcd}

\vskip 5pt

\vskip 10pt

The following results were obtained from equation \eqref{eqn:RabcdPartials}. Note that by
inspection of these reulst and those above it is easy to verify the claim made in
section (\ref{sec:MetricRNC}) that the connection terms that arise in
$R^{a}{}_{(bc{\Dot d};\ue)}$ are always of the form $\Gamma^{p}{}_{q(r,\us)}$.
%
\begin{dgroup*}[compact,spread={3pt}]
   \dMath{11.13}{    R^a{}_{(cd{\Dot b},e)} = \cdb{dRabcd61}}{eqn:dRabcd61}{\hfill}
   \dMath{11.14}{    R^a{}_{(cd{\Dot b},ef)} = \cdb{dRabcd62}}{eqn:dRabcd62}{\hfill}
   \dMath{11.15}{ -2 R^a{}_{(cd{\Dot b},efg)} = \cdb{dRabcd63}}{eqn:dRabcd63}{\hfill}
   \dMath{11.16}{ -5 R^a{}_{(cd{\Dot b},efgh)} = \cdb{dRabcd64}}{eqn:dRabcd64}{\hfill}
   \dMath{11.17}{ -3 R^a{}_{(cd{\Dot b},efghi)} = \cdb{dRabcd65}}{eqn:dRabcd65}{\hfill}
\end{dgroup*}

% --------------------------------------------------------------------------------------------
\mysubsection{Riemann normal coordinates}

\begin{equation*}
   y^a = \ny{0}^{a} + \ny{1}^{a} + \ny{2}^{a} + \ny{3}^{a} + \ny{4}^{a}\tag{11.18}
\end{equation*}

\begin{dgroup*}
   \dMath{11.18a}{    \ny{0}^{a} = \cdb{rnc61}}{eqn:rnc61}{\hfill}
   \dMath{11.18b}{  2 \ny{1}^{a} = \cdb{rnc62}}{eqn:rnc62}{\hfill}
   \dMath{11.18c}{  6 \ny{2}^{a} = \cdb{rnc63}}{eqn:rnc63}{\hfill}
   \dMath{11.18d}{ 24 \ny{3}^{a} = \cdb{rnc64}}{eqn:rnc64}{\hfill}
   \dMath{11.18e}{360 \ny{4}^{a} = \cdb{rnc65}}{eqn:rnc65}{\hskip 1cm}
\end{dgroup*}

% --------------------------------------------------------------------------------------------
\mysubsection{Geodesic IVP}

\begin{align}
   x^{d}(s) ={}& x^{d}
            + s {{\Dot x}^d}
            + \frac{s^2}{2!} {{\Dot x}^a} {{\Dot x}^b} A^{d}_{ab}
            + \frac{s^3}{3!} {{\Dot x}^a} {{\Dot x}^b} {{\Dot x}^c} A^{d}_{abc}
            + \frac{s^4}{4!} {{\Dot x}^a} {{\Dot x}^b} {{\Dot x}^c}
                                                       {{\Dot x}^d} A^{d}_{abce}\notag\\
           &+ \frac{s^5}{5!} {{\Dot x}^a} {{\Dot x}^b} {{\Dot x}^c} {{\Dot x}^d}
                                                       {{\Dot x}^e} A^{d}_{abcef}
            + \BigO{s^6,\eps^6}\tag{11.19}
\end{align}
\begin{dgroup*}
   \dMath{11.19a}{ 360 A^{d}_{ab} = \cdb{ivp62}}{eqn:ivp62}{\hfill}
   \dMath{11.19b}{ 360 A^{d}_{abc} = \cdb{ivp63}}{eqn:ivp63}{\hskip 0.5cm}
   \dMath{11.19c}{  90 A^{d}_{abce} = \cdb{ivp64}}{eqn:ivp64}{\hfill}
   \dMath{11.19d}{   3 A^{d}_{abcef} = \cdb{ivp65}}{eqn:ivp65}{\hfill}
\end{dgroup*}

% --------------------------------------------------------------------------------------------
\mysubsection{Geodesic BVP}

\begin{align}
   x^{a}(s) &= x^{a} + s Dx^{a}
                     + (s-s^2) x^{a}_2
                     + (s-s^3) x^{a}_3
                     + (s-s^4) x^{a}_4
                     + (s-s^5) x^{a}_5
                     + \BigO{s^6,\eps^6}\tag{11.20}\pglabel{eqn:bvp6}\\
   x^{a}_2 &= \nx{2}^{a}_2 + \nx{3}^{a}_2 + \nx{4}^{a}_2 + \nx{5}^{a}_2
            + \BigO{\eps^6}\tag{11.20a}\\
   x^{a}_3 &= \nx{3}^{a}_3 + \nx{4}^{a}_3 + \nx{5}^{a}_3 + \BigO{\eps^6}\tag{11.20b}\\
   x^{a}_4 &= \nx{4}^{a}_4 + \nx{5}^{a}_4 + \BigO{\eps^6}\tag{11.20c}\\
   x^{a}_5 &= \nx{5}^{a}_5 + \BigO{\eps^6}\tag{11.20d}
\end{align}
%
\begin{dgroup*}[spread=5pt]
   \dMath{11.20a.1}{    -3 \nx{2}^{a}_2 = \cdb{bvp622}}{eqn:bvp622}{\hfill}
   \dMath{11.20a.2}{   -24 \nx{3}^{a}_2 = \cdb{bvp623}}{eqn:bvp623}{\hfill}
   \dMath{11.20a.3}{  -720 \nx{4}^{a}_2 = \cdb{bvp624}}{eqn:bvp624}{\hskip 1cm}
   \dMath{11.20a.4}{  -360 \nx{5}^{a}_2 = \cdb{bvp625}}{eqn:bvp625}{\hfill}
%
   \dMath{11.20b.1}{   -12 \nx{3}^{a}_3 = \cdb{bvp633}}{eqn:bvp633}{\hfill}
   \dMath{11.20b.2}{  -720 \nx{4}^{a}_3 = \cdb{bvp634}}{eqn:bvp634}{\hskip 1cm}
   \dMath{11.20b.3}{ -1080 \nx{5}^{a}_3 = \cdb{bvp635}}{eqn:bvp635}{\hskip 1cm}
%
   \dMath{11.20c.1}{  -180 \nx{4}^{a}_4 = \cdb{bvp644}}{eqn:bvp644}{\hfill}
   \dMath{11.20c.2}{ -2160 \nx{5}^{a}_4 = \cdb{bvp645}}{eqn:bvp645}{\hskip 1cm}
%
   \dMath{11.20d.1}{  -360 \nx{5}^{a}_5 = \cdb{bvp655}}{eqn:bvp655}{\hskip 1cm}
\end{dgroup*}

% --------------------------------------------------------------------------------------------
\mysubsection{Geodesic arc-length}

\begin{equation*}
   \pglabel{eqn:geodesicLSQ}
   % LCB: which of these is correct?
   % \left(\Delta s\right)^2 = \nD{0} + \nD{2} + \nD{3} + \nD{4} + \nD{5} + \BigO{\eps^6,Dx^6}
   % \left(\Delta s\right)^2 = \nD{0} + \nD{2} + \nD{3} + \nD{4} + \nD{5} + \BigO{\eps^6,Dx^7}
   \left(\Delta s\right)^2 = \nD{0} + \nD{2} + \nD{3} + \nD{4} + \nD{5}
                           + \BigO{\eps^6}\tag{11.21}
\end{equation*}

\begin{dgroup*}[spread=5pt]
   \dMath{11.21a}{      \nD{0} = \cdb{lsq60}}{eqn:lsq60}{\hfill}
   \dMath{11.21b}{    3 \nD{2} = \cdb{lsq62}}{eqn:lsq62}{\hfill}
   \dMath{11.21c}{   12 \nD{3} = \cdb{lsq63}}{eqn:lsq63}{\hfill}
   \dMath{11.21d}{  360 \nD{4} = \cdb{lsq64}}{eqn:lsq64}{\hfill}
   \dMath{11.21e}{ 1080 \nD{5} = \cdb{lsq65}}{eqn:lsq65}{\hfill}
\end{dgroup*}

% ============================================================================================
\section{Source}

All of the Cadabra files used in preparing this paper are available at the GitHub repository
\ \url{https://github.com/leo-brewin/cadabra-riemann-coords}

% ============================================================================================
\section{Acknowledgements}

I am very grateful to Kasper Peeters for his many helpful suggestions. Any errors,
omissions or inaccuracies in regard to Cadabra are entirely my fault.

\appendix

% ============================================================================================
\section{Symmetrisation of tensors}

The totally symmetric part of a tensor $A_{i_1i_2i_3\cdots i_n}$ is commonly
defined by
\[
A_{\left(i_1i_2i_3\cdots i_n\right)}
= \frac{1}{n!}\left(  A_{i_1i_2i_3\cdots i_n}
                    + A_{i_1i_2i_3\cdots i_n}
                    + A_{i_1i_2i_3\cdots i_n}
                    + \cdots \right)
\]
where the sum on the right hand side includes every permutation of the indices
of $i_1i_2i_3\cdots i_n$. If the tensor $A_{i_1i_2i_3\cdots i_n}$ happens to be
symmetric in every pair of indices then
\[
A_{\left(i_1i_2i_3\cdots i_n\right)} = A_{i_1i_2i_3\cdots i_n}
\]
From the above definition it is very easy to establish the following theorems
\begin{gather*}
A_{(i_1i_2i_3\cdots(j_1j_2j_3\cdots j_m)\cdots i_n)}
= A_{(i_1i_2i_3\cdots j_1j_2j_3\cdots j_m\cdots i_n)}\\[5pt]
A_{(i_1i_2i_3\cdots i_n}B_{j_1j_2j_3\cdots j_m)}
= A_{((i_1i_2i_3\cdots i_n)}B_{(j_1j_2j_3\cdots j_m))}\\[5pt]
nA_{(i_1i_2i_3i_4\cdots i_n)} = A_{i_1(i_2i_3i_4\cdots i_n)}
                        + A_{i_2(i_1i_3i_4\cdots i_n)}
                        + A_{i_3(i_1i_2i_4\cdots i_n)}
                        + \cdots\notag\\[5pt]
                        {\hskip 0.7\textwidth} + A_{i_n(i_1i_2i_3\cdots i_{n-1})}\\[5pt]
nA_{(i_1i_2i_3i_4\cdots i_n)} = A_{(i_2i_3i_4\cdots i_n)i_1}
                        + A_{(i_1i_3i_4\cdots i_n)i_2}
                        + A_{(i_1i_2i_4\cdots i_n)i_3}
                        + \cdots\notag\\[5pt]
                        {\hskip 0.7\textwidth} + A_{(i_1i_2i_3\cdots i_{n-1})i_n}
\end{gather*}

Suppose now that $A_{i_1i_2i_3\cdots i_n} = A_{(i_1i_2i_3\cdots i_n)}$, that is,
$A_{i_1i_2i_3\cdots i_n}$ is totally symmetric. Then for any $B_j$
\begin{gather*}
\begin{split}
(n+1) A_{(i_1i_2i_3\cdots i_n}B_{j)}
&= A_{ji_2i_3\cdots i_n} B_{i_1}
+ A_{i_1ji_3\cdots i_n} B_{i_2}
+ A_{i_1i_2j\cdots i_n} B_{i_3}\\[5pt]
&\quad+ \cdots + A_{i_1i_2i_3\cdots i_{n-1}j} B_{i_n}
\end{split}\\[5pt]
\intertext{and}
\begin{split}
(n+1) A_{(i_1i_2i_3\cdots i_n,j)}
&= A_{ji_2i_3\cdots i_n,i_1}
+ A_{i_1ji_3\cdots i_n,i_2}
+ A_{i_1i_2j\cdots i_n,i_3}\\[5pt]
&\quad+\cdots
+A_{i_1i_2i_3\cdots i_{n-1}j,i_n}
+A_{i_1i_2i_3\cdots i_n,j}
\end{split}
\end{gather*}

All of the above are very easy to prove but one result which requires just a little more
thought is the following.

Suppose $A_{i_1i_2j_3j_4j_5\cdots j_n}$ is symmetric in the pair $i_1i_2$ and symmetric in
all the indices $j_3j_4j_5\cdots j_n$. That is, it is symmetric under the interchange of any
pair of $i$'s and any pair of $j$'s but it is \emph{not} necessarily symmetric when any $i$
is swapped with any $j$. What can be said about $A_{(i_1i_2j_3j_4j_5\cdots j_n)}$? Here is
the result
\begin{gather}
n A_{(i_1i_2i_3\cdots i_n)}
= 2 A_{i_n(i_1i_2i_3\cdots i_{n-1})}
+ (n-2) A_{(i_1i_2i_3\cdots i_{n-1})i_n}
\pglabel{eqn:doubleSymm}
\end{gather}
The proof is very easy. Begin by writing out $n!A_{(i_1i_2i_3\cdots i_n)}$ in full. Then
partition the terms into two disjoint sets, one set in which $i_n$ appears in one of the
first two index slots, the other set in which $i_n$ appears in any of the remaining $n-2$
slots. The terms in the first set are exactly those that define $A_{i_n(i_1i_2i_3\cdots
i_{n-1})}$ while those in the second set define $A_{(i_1i_2i_3\cdots i_{n-1})i_n}$. The
above equation follows by simply counting the number of terms in each set ($2(n-1)!$ and
$(n-2)(n-1)!$ respectively) and the simple observation that $n!A_{(i_1i_2i_3\cdots i_n)}$
equals the sum of the terms from both sets.

Finally note that if $Q=A_{i_1i_2i_3\cdots i_n}x^{i_1}x^{i_2}x^{i_3}\cdots x^{i_n}$
then
\begin{gather*}
Q_{,i_1i_2i_3\cdots i_n} = n! A_{(i_1i_2i_3\cdots i_n)}\\[10pt]
% \intertext{and}
Q = A_{(i_1i_2i_3\cdots i_n)}x^{i_1}x^{i_2}x^{i_3}\cdots x^{i_n}
\end{gather*}

% ============================================================================================
\printbibliography  % create the bibliography using biblatex & biber

% ============================================================================================
% \bibliographystyle{bibdata}
% \bibliography{bibdata}

\end{document}

### 2019-09-24 source/build.sh

@@ -1,0 +1,10 @@
+#! /bin/bash
+
+(cd cadabra; make github)
+(cd tex; make github)
+
+mkdir -p ../pdf
+
+cp -rf tex/paper.pdf                           ../pdf/paper.pdf
+cp -rf cadabra/pdf/summary.pdf                 ../pdf/summary.pdf
+cp -rf cadabra/pdf/collection.pdf              ../pdf/collection.pdf

### 2019-09-24 source/cadabra/tests.sh

@@ -1,0 +1,9 @@
+#! /bin/bash
+
+echo "---------------------------------------------------------"
+echo "> diff text files"
+tests-text.sh
+
+echo "---------------------------------------------------------"
+echo "> checkpoint files, results in tests/semantic/summary.pdf"
+tests-semantic.sh

### 2019-09-24 source/cadabra/tests/semantic/shared.py

@@ -1,0 +1,88 @@
+from cadabra2_defaults import *
+__cdbkernel__ = cadabra2.__cdbkernel__
+
+__cdbtmp__ = Indices(Ex(r'''{a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}'''), Ex(r'''position=independent)''') )
+
+__cdbtmp__ = TableauSymmetry(Ex(r'''\Gamma^{a}_{b c}'''), Ex(r'''shape={2}, indices={1,2})''') )
+
+__cdbtmp__ = Derivative(Ex(r'''D{#}'''), Ex(r'''''') )
+__cdbtmp__ = Derivative(Ex(r'''\nabla{#}'''), Ex(r'''''') )
+__cdbtmp__ = PartialDerivative(Ex(r'''\partial{#}'''), Ex(r'''''') )
+
+__cdbtmp__ = Metric(Ex(r'''g_{a b}'''), Ex(r'''''') )
+__cdbtmp__ = InverseMetric(Ex(r'''g^{a b}'''), Ex(r'''''') )
+
+__cdbtmp__ = RiemannTensor(Ex(r'''R_{a b c d}'''), Ex(r'''''') )
+
+__cdbtmp__ = RiemannTensor(Ex(r'''R^{a}_{b c d}'''), Ex(r'''''') )
+# R_{a b c}^{d}::RiemannTensor.
+__cdbtmp__ = RiemannTensor(Ex(r'''R^{a}_{b c d}'''), Ex(r'''''') )
+__cdbtmp__ = RiemannTensor(Ex(r'''R^{a}_{b}^{d}_{c}'''), Ex(r'''''') )
+
+# note: this function only covers the cases that I have seen
+#       I want to keep this as short as possible
+def standard_indices (obj):
+    # single upstairs index on Rabcd
+    substitute (obj,Ex(r'''R_{a}^{b}_{c d}   -> - R^{b}_{a c d}''', False))
+    substitute (obj,Ex(r'''R_{a b}^{c}_{d}   ->   R^{c}_{d a b}''', False))
+    substitute (obj,Ex(r'''R_{a b c}^{d}     -> - R^{d}_{c a b}''', False))
+    # double upstairs index on Rabcd
+    substitute (obj,Ex(r'''R_{a}^{b}_{c}^{d} ->   R^{b}_{a}^{d}_{c}''', False))
[…]

### 2019-09-24 support/scripts/cdblatex.sh

@@ -1,0 +1,102 @@
+#!/bin/bash
+
+export CADABRA_NO_UNICODE=1
+export PYTHONWARNINGS="ignore"
+
+file="<none>"
+silent="no"
+keep="no"
+skiplatex="no"
+Timer=""
+CDB=/usr/local/bin/
+sty=""
+nowarn=""
+
+# -----------------------------------------------------------------------------------------
+# Parse the command-line options
+
+OPTIND=1
+
+while getopts 'i:I:P:sktTxhN' option
+do
+   case "$option" in
+   "i")  file="$OPTARG"           ;;
+   "I")  sty="-I$OPTARG"          ;;
+   "P")  CDB="$OPTARG"            ;;
+   "s")  silent="yes"             ;;
+   "k")  keep="yes"               ;;
+   "t")  Timer="/usr/bin/time"    ;;
+   "T")  Timer="/usr/bin/time -l" ;;
+   "x")  skiplatex="yes"          ;;
+   "N")  nowarn="-N"              ;;
+   "h")  echo "usage : cdblatex.sh -i file [-P<path to Cadabra bin dir>]"
+         echo "                            [-I<path to cdbmacros.sty>] [-s] [-k] [-x] [-N] [-h]"
+         echo "options :  -i file : source file (with or without .tex extension)"
+         echo "           -I file : full path to cdbmacros.sty file"
+         echo "           -P file : path to Cadabra bin directory"
+         echo "           -s : silent, don't open the pdf file"
+         echo "           -k : keep all temporary files"
+         echo "           -t : report brief cpu time"
+         echo "           -T : report detailed cpu time plus memory usage"
+         echo "           -x : don't call latex"
+         echo "           -N : don't warn if errors found in the output for some tags"
+         echo "           -h : this help messag"
+         echo "example : cdblatex.sh -i file"
+         exit                ;;
+   ?)    echo "cdblatex.sh : Unknown option specified."
+         exit                ;;
+   esac
+done
+
+# strip .tex if given
+file=$(basename -s ".tex" "$file")
+name=$file
+
+if [[ $file = "<none>" ]]; then
+   echo "> no source file given, use cdblatex.sh -i file"
+   exit 1;
+fi;
+
[…]

### 2019-09-24 support/scripts/cdbpreproc.py

@@ -1,0 +1,480 @@
+#!/usr/local/bin/python3
+
+import re
+import sys
+import os.path
+
+re_beg_cadabra_environ     = re.compile (r'^\s*\\begin{cadabra}')
+re_end_cadabra_environ     = re.compile (r'^\s*\\end{cadabra}')
+
+re_beg_cadabra_verbatim    = re.compile (r'^\s*\\CdbSetup{(.*?)action=verbatim')
+re_end_cadabra_verbatim    = re.compile (r'^\s*\\CdbSetup{(.*?)action=(show|hide)')
+
+re_beg_latex_document      = re.compile (r'^\s*\\begin{document}')
+re_end_latex_document      = re.compile (r'^\s*\\end{document}')
+
+re_indent          = re.compile (r'(^ *)')
+re_assignment      = re.compile (r'(^\s*)([a-zA-Z0-9]+\s*:?=)')
+re_empty_line      = re.compile (r'(^ *$)')
+re_latex_comment   = re.compile (r'(^ *%)')
+re_cadabra_comment = re.compile (r'(^ *#)')
+re_cadabra_markup  = re.compile (r'(# *(cdb\s*\(|cdbBeg\s*\(|cdbEnd))')
+re_hidden_markup   = re.compile (r'(^ *#.*#\s*(cdb\s*\(|cdbBeg\s*\(|cdbEnd))')
+re_pure_markup     = re.compile (r'(^\s*#\s*(cdb\s*\(|cdbBeg\s*\(|cdbEnd))')
+re_capture         = re.compile (r'(# *((cdbBeg|cdbEnd)\s*\(\s*([a-zA-Z0-9_.-]+)\)))')
+re_beg_capture     = re.compile (r'(# *(cdbBeg\s*\(\s*([a-zA-Z0-9_.-]+)\)))')
+re_end_capture     = re.compile (r'(# *(cdbEnd\s*(\(\s*([a-zA-Z0-9_.-]+)\))?))')
+re_cdb_tag         = re.compile (r'(# *(cdb\s*\(\s*([a-zA-Z0-9_.-]+)\s*(,\s*([a-zA-Z0-9]+)\s*)?\)))')
+                                # Allow two forms of tag, py(foo,bah) and py(bah).
+                                # In both cases bah must be a valid Cadabra expression.
+
+re_inline_comment  = re.compile (r'(.*?)( +#)')  # any text followed by " #"
+
+re_algorithm       = re.compile (r'(^\s*)('
[…]

### 2019-09-24 source/tex/paper.tex
Fixed one typo in paper.tex
What can I say?

@@
 The following results were obtained from equation \eqref{eqn:RabcdPartials}. Note that by
-inspection of these reulst and those above it is easy to verify the claim made in
+inspection of these results and those above it is easy to verify the claim made in
 section (\ref{sec:MetricRNC}) that the connection terms that arise in

### 2019-10-11 source/cadabra/dGamma.tex
Minor housekeeping. No major changes.
Fixed a few typos in some of the tex and sh files.

@@
 
-Here we impose the RNC conition by setting the $\Gamma^{a}{}_{bc}$ to zero (but not their derivatives).
+Here we impose the RNC condition by setting the $\Gamma^{a}{}_{bc}$ to zero (but not their derivatives).
 
@@
 % =================================================================================================
-\section*{Stage 3: Replace partial derivatives $\Gamma$ with partial derivatives $R$}
+\section*{Stage 3: Replace partial derivatives of $\Gamma$ with partial derivatives of $R$}
 

### 2019-10-18 source/cadabra/sqrt-detg.tex
Updated Makefile to *not* make sqrt-detg
The sqrt-detg code requires pylatex.sh and friends which are not included in this repo. (see https://github.com/leo-brewin/hybrid-latex). The sqrt-detg results are not required by any code here. It's output was hand coded in detg.tex.

Also, the cdbpreproc.py was updated (copied from hybrid-latex repo.).
See https://github.com/leo-brewin/hybrid-latex/commit/199d8e817b54ca515c1221aee749990829b69538

@@
 
-   ans = sqrt(1+a*eps**2+b*eps**3+c*eps**4+d*eps**5)   # py(ans.001,ans)
-   ans = ans.series(eps, 0, 6)                         # py(ans.002,ans)
-   ans = simplify(ans)                                 # py(ans.003,ans)
+   ans = sqrt(1+a*eps**2+b*eps**3+c*eps**4+d*eps**5)   # py (ans.001,ans)
+   ans = ans.series(eps, 0, 6)                         # py (ans.002,ans)
+   ans = simplify(ans)                                 # py (ans.003,ans)
 
@@
 
-   ans = log(1+a*eps**2+b*eps**3+c*eps**4+d*eps**5)    # py(ans.101,ans)
-   ans = ans.series(eps, 0, 6)                         # py(ans.102,ans)
-   ans = simplify(ans)                                 # py(ans.103,ans)
+   ans = log(1+a*eps**2+b*eps**3+c*eps**4+d*eps**5)    # py (ans.101,ans)
+   ans = ans.series(eps, 0, 6)                         # py (ans.102,ans)
+   ans = simplify(ans)                                 # py (ans.103,ans)
 

### 2019-12-20 source/tex/paper.tex
Updated to match version on arxiv
Updated to match version on arxiv.

@@
+\pdfoutput=1% forces arxiv to use pdflatex
+% --------------------------------------------------------------------------------------------
+\def\Title{Riemann Normal Coordinate expansions using Cadabra.}
+\def\Author{Leo Brewin}
+% --------------------------------------------------------------------------------------------
 \documentclass[a4paper,12pt]{article}
-\usepackage[style=numeric,
-            sorting=none,
-            sortcites=true]{biblatex}
-% \usepackage[style=numeric,
-%             sorting=none,
-%             sortcites=true,
-%             bibencoding=ascii,
-%             backend=biber]{biblatex}
+%
+% biber will replace LaTeX accented characters like \'{e} with utf-8 characters.
+% The utf-8 characters are written to the .bbl and this causes problems with pdflatex.
+%    First  solution use: biber tutorial and \usepackage[latin1]{inputenc}
+%    Second solution use: biber --decodecharsset=null tutorial
+\usepackage[style=numeric, sorting=none, sortcites=true]{biblatex}
 \addbibresource{bibdata.bib}
-\usepackage[dvipsnames]{xcolor}
-\usepackage{listings}
-\usepackage{amsmath}
-\usepackage{breqn}
-% \usepackage[colorlinks=false]{hyperref}% for journals
-\usepackage[colorlinks=true,citecolor=red]{hyperref}% for me
+%
 % \usepackage{showlabels}
@@
 \usepackage{cdblatex}
+\usepackage{hyperref}
 \usepackage{geometry}
@@
 
+% \usepackage[colorlinks=false]{hyperref}% for journals
+% \usepackage[colorlinks=true,citecolor=red]{hyperref}% for me
+\hypersetup{colorlinks=true,
+            citecolor=blue,
+            linkcolor=red,
+            pdfauthor= \Author,
+            pdftitle = \Title}
+
 \numberwithin{equation}{section}
 
+% ============================================================================================
 \begin{document}
 
-\title{\bfseries Riemann Normal Coordinate expansions using Cadabra}
-\author{%
-Leo Brewin\\[10pt]%
+\input{../cadabra/lib/export.cdbtex}% pulls in all the Cadabra results
+
+\CdbSetup{action=verbatim}% we won't be processing any Cadabra code,
+                          % just listing code from other sources
+
+\lstset{numbers=left,gobble=2,basicstyle=\ttfamily\small,escapechar=|}
+
+% ============================================================================================
+\title{\Title}
+\author{\Author\\[10pt]%
 School of Mathematical Sciences\\%
@@
 \date{12-Aug-2009 (v1)\\
-      19-Apr-2019 (v2)}% major changes to correct errors in 4th and higher order terms,
+      10-Dec-2019 (v2)}% major changes to correct errors in 4th and higher order terms,
                        % all code will be available on github
@@
 been completely rewritten to account for the revised syntax introduced in the latest version
-of Cadabra (itself updated in 2016). Second, and more importantly, a significant
-computational error in the first version has been corrected. Though the algorithms given in
-the first version were correct their implementation in Cadabra (by this author) was
-incorrect. This lead to errors first in the cubic and higher order terms in the generalised
-connections and then, down the line, to computations built upon the generalised connections
-(e.g., the initial and boundary value problem and the arc-length equations). The Cadabra
-codes have been corrected and thoroughly cross-checked. A number of minor typographical
-errors have also been corrected.
+of Cadabra (itself updated in 2016)\footnote{The Cadabra results in this paper are based on
+\CdbVersion}. Second, and more importantly, a significant computational error in the first
+version (of this paper) has been corrected. Though the algorithms given in the first version
+were correct their implementation in Cadabra (by this author) was incorrect. This lead to
+errors first in the quartic and higher order terms in the generalised connections and then,
+down the line, to computations built upon the generalised connections (e.g., the initial and
+boundary value problem and the arc-length equations). The Cadabra codes have been corrected
+and thoroughly cross-checked. A number of minor typographical errors have also been
+corrected.
 \EndNote
@@
 resources, taking between 0.5 second and 12 minutes to run and requiring
-between 40 and 360 Mbyte of memory (on a Mac Pro running macOS 10.14.4).
+between 40 and 360 Mbyte of memory (on a Mac Pro (2013) running macOS 10.14.4).
 
 The Cadabra codes and several support scripts are available from the author's GitHub site
-\ \url{https://github.com/leo-brewin/cadabra-riemann-coords}.
+\ \url{https://github.com/leo-brewin/riemann-normal-coords}.
 
@@
    \dMath{11.3b}{\sqrt{-g(x)}  = \cdb{sqrtNdetg} + \BigO{\eps^6}}{eqn:sqrtdetg}{\hskip 1cm\hfill}
-   \dMath{11.3c}{\log{-g(x)}  = \cdb{logNdetg}  + \BigO{\eps^6}}{eqn:logdetg}{\hfill}
+   \dMath{11.3c}{\log{(-g(x))} = \cdb{logNdetg}  + \BigO{\eps^6}}{eqn:logdetg}{\hfill}
 \end{dgroup*}
@@
 All of the Cadabra files used in preparing this paper are available at the GitHub repository
-\ \url{https://github.com/leo-brewin/cadabra-riemann-coords}
+\ \url{https://github.com/leo-brewin/riemann-normal-coords}
 

### 2019-12-20 source/build.sh

@@ -1,10 +1,6 @@
 #! /bin/bash
 
-(cd cadabra; make github)
-(cd tex; make github)
-
 mkdir -p ../pdf
 
-cp -rf tex/paper.pdf                           ../pdf/paper.pdf
-cp -rf cadabra/pdf/summary.pdf                 ../pdf/summary.pdf
-cp -rf cadabra/pdf/collection.pdf              ../pdf/collection.pdf
+(cd cadabra; make github)
+(cd tex; make github)

### 2020-02-20 source/cadabra/geodesic-ivp.tex
Corrected typos on free indices in eqtns. 11.19*
The LaTeX source for eqtns. 11.19* (geodesic-ivp results) had incorrect free indices on the left hand side. The indices were written by hand (not by Cadabra). The Cadabra code/results are unchanged.
Corrected +/- sign error on s^2, s^3 etc.  terms in eqs. 11.19a-d
I had forgotten to take account of the minus sign in eq. 3.1 when constructing the solution to the geodesic IVP. This correction has no effect on any other results.

@@
 
-   sterm21 = cdblib.get ('genGamma01','genGamma.json')
-   sterm22 = cdblib.get ('genGamma02','genGamma.json')
-   sterm23 = cdblib.get ('genGamma03','genGamma.json')
-   sterm24 = cdblib.get ('genGamma04','genGamma.json')
+   # change signs to account for - sign in front of the sum for x^a(s), see above preamble
 
-   sterm31 = cdblib.get ('genGamma11','genGamma.json')
-   sterm32 = cdblib.get ('genGamma12','genGamma.json')
-   sterm33 = cdblib.get ('genGamma13','genGamma.json')
+   def flip_sign (obj):
+       return Ex(0) - obj
 
-   sterm41 = cdblib.get ('genGamma21','genGamma.json')
-   sterm42 = cdblib.get ('genGamma22','genGamma.json')
+   sterm21 = flip_sign (cdblib.get ('genGamma01','genGamma.json'))
+   sterm22 = flip_sign (cdblib.get ('genGamma02','genGamma.json'))
+   sterm23 = flip_sign (cdblib.get ('genGamma03','genGamma.json'))
+   sterm24 = flip_sign (cdblib.get ('genGamma04','genGamma.json'))
 
-   sterm51 = cdblib.get ('genGamma31','genGamma.json')
+   sterm31 = flip_sign (cdblib.get ('genGamma11','genGamma.json'))
+   sterm32 = flip_sign (cdblib.get ('genGamma12','genGamma.json'))
+   sterm33 = flip_sign (cdblib.get ('genGamma13','genGamma.json'))
+
+   sterm41 = flip_sign (cdblib.get ('genGamma21','genGamma.json'))
+   sterm42 = flip_sign (cdblib.get ('genGamma22','genGamma.json'))
+
+   sterm51 = flip_sign (cdblib.get ('genGamma31','genGamma.json'))
 
@@
 \begin{align*}
-   x^{d}(s) = x^{d}
-            + s {\dot{x}^d}
-            + \frac{s^2}{2!} {\dot{x}^a} {\dot{x}^b} A^{d}_{ab}
-            + \frac{s^3}{3!} {\dot{x}^a} {\dot{x}^b} {\dot{x}^c} A^{d}_{abc}
-            + \frac{s^4}{4!} {\dot{x}^a} {\dot{x}^b} {\dot{x}^c} {\dot{x}^d} A^{d}_{abce}
-            + \frac{s^5}{5!} {\dot{x}^a} {\dot{x}^b} {\dot{x}^c} {\dot{x}^d} {\dot{x}^e} A^{d}_{abcef}
+   x^{a}(s) = x^{a}
+            + s {\dot{x}^a}
+            + \frac{s^2}{2!} {\dot{x}^b} {\dot{x}^c} A^{a}_{bc}
+            + \frac{s^3}{3!} {\dot{x}^b} {\dot{x}^c} {\dot{x}^d} A^{a}_{bcd}
+            + \frac{s^4}{4!} {\dot{x}^b} {\dot{x}^c} {\dot{x}^d} {\dot{x}^e} A^{a}_{bcde}
+            + \frac{s^5}{5!} {\dot{x}^b} {\dot{x}^c} {\dot{x}^d} {\dot{x}^e} {\dot{x}^f} A^{a}_{bcdef}
             + \dotsb
@@
 \begin{dgroup*}
-   \begin{dmath*} 360 A^{d}_{ab} = \cdb{sterm2.002} \end{dmath*}
-   \begin{dmath*} 360 A^{d}_{abc} = \cdb{sterm3.002} \end{dmath*}
-   \begin{dmath*}  90 A^{d}_{abce} = \cdb{sterm4.002} \end{dmath*}
-   \begin{dmath*}   3 A^{d}_{abcef} = \cdb{sterm5.002} \end{dmath*}
+   \begin{dmath*} 360 A^{a}_{bc} = \cdb{sterm2.002} \end{dmath*}
+   \begin{dmath*} 360 A^{a}_{bcd} = \cdb{sterm3.002} \end{dmath*}
+   \begin{dmath*}  90 A^{a}_{bcde} = \cdb{sterm4.002} \end{dmath*}
+   \begin{dmath*}   3 A^{a}_{bcdef} = \cdb{sterm5.002} \end{dmath*}
 \end{dgroup*}

### 2021-11-25 INSTALL.sh
Simplified installation, updated hybrid-latex tools
1. Non-trivial refactoring of the build/install scripts.
   This was both to tidy the file structure as well as to simplify the installation
   process. The default install is now as simple as "source SETPATHS; make install".
   Almost all install/build/test/clean actions are now controlled by Makefiles.

2. Updated cdbpreproc.py.
   The Cadabra preprocessor will strip any in-line comments (such as cdb tags) from
   what it thinks are short lines. This failed for some of the examples (some in line
   comments were not deleted). This was fixed

@@ -1,0 +1,34 @@
+#!/bin/bash
+
+# defaults, edit to suit
+
+if [[ $HERE = '' ]]; then
+   HERE=$HOME/local/rnc/
+fi;
+
+MyBin=$HERE/bin/
+MyLib=$HERE/lib/
+MyTex=$HERE/tex/
+
+# Parse the command-line options
+
+while getopts 'b:l:t:h' option
+do
+   case "$option" in
+   "b")  MyBin="$OPTARG"      ;;
+   "l")  MyLib="$OPTARG"      ;;
+   "t")  MyTex="$OPTARG"      ;;
+   "h")  echo "usage : INSTALL.sh [-bBIN] [-lLIB] [-tTEX]"
+         echo "options : -b : where to find the binaries and shell scripts"
+         echo "          -l : where to find the python libraries"
+         echo "          -t : where to find the LaTeX files "
+         echo "          -h : this help message"
+         exit                 ;;
+   ?)    echo "INSTALL.sh: Unknown option specified."
+         exit                 ;;
+   esac
+done
+
+mkdir -p $MyBin $MyLib $MyTex
+
+(cd hybrid-latex; INSTALL.sh -b $MyBin -l $MyLib -t $MyTex)

### 2021-11-25 hybrid-latex/INSTALL.sh

@@ -1,0 +1,36 @@
+#!/bin/bash
+
+# defaults, edit to suit
+
+if [[ $HERE = '' ]]; then
+   HERE=$HOME/local/rnc/
+fi;
+
+MyBin=$HERE/bin/
+MyLib=$HERE/lib/
+MyTex=$HERE/tex/
+
+# Parse the command-line options
+
+while getopts 'b:l:t:h' option
+do
+   case "$option" in
+   "b")  MyBin="$OPTARG"      ;;
+   "l")  MyLib="$OPTARG"      ;;
+   "t")  MyTex="$OPTARG"      ;;
+   "h")  echo "usage : INSTALL.sh [-bBIN] [-lLIB] [-tTEX]"
+         echo "options : -b : where to find the binaries and shell scripts"
+         echo "          -l : where to find the python libraries"
+         echo "          -t : where to find the LaTeX files "
+         echo "          -h : this help message"
+         exit                 ;;
+   ?)    echo "INSTALL.sh: Unknown option specified."
+         exit                 ;;
+   esac
+done
+
+mkdir -p $MyBin $MyLib $MyTex
+
+cp -rf scripts/* $MyBin
+cp -rf python/*  $MyLib
+cp -rf latex/*   $MyTex

### 2021-11-25 source/build.sh

@@ -2,5 +2,5 @@
 
 mkdir -p ../pdf
 
-(cd cadabra; make github)
-(cd tex; make github)
+(cd cadabra; make; make veryclean)
+(cd tex; make; make veryclean)

### 2021-11-25 source/cadabra/tests.sh

@@ -5,5 +5,5 @@ echo "> diff text files"
 tests-text.sh
 
 echo "---------------------------------------------------------"
-echo "> checkpoint files, results in tests/semantic/summary.pdf"
+echo "> checkpoint files, results in source/cadabra/tests/semantic/summary.pdf"
 tests-semantic.sh

### 2021-11-25 source/cadabra/tests/semantic/shared.py

@@ -1,88 +1,0 @@
-from cadabra2_defaults import *
-__cdbkernel__ = cadabra2.__cdbkernel__
-
-__cdbtmp__ = Indices(Ex(r'''{a,b,c,d,e,f,g,h,i,j,k,l,m,n,o,p,q,r,s,t,u,v,w#}'''), Ex(r'''position=independent)''') )
-
-__cdbtmp__ = TableauSymmetry(Ex(r'''\Gamma^{a}_{b c}'''), Ex(r'''shape={2}, indices={1,2})''') )
-
-__cdbtmp__ = Derivative(Ex(r'''D{#}'''), Ex(r'''''') )
-__cdbtmp__ = Derivative(Ex(r'''\nabla{#}'''), Ex(r'''''') )
-__cdbtmp__ = PartialDerivative(Ex(r'''\partial{#}'''), Ex(r'''''') )
-
-__cdbtmp__ = Metric(Ex(r'''g_{a b}'''), Ex(r'''''') )
-__cdbtmp__ = InverseMetric(Ex(r'''g^{a b}'''), Ex(r'''''') )
-
-__cdbtmp__ = RiemannTensor(Ex(r'''R_{a b c d}'''), Ex(r'''''') )
-
-__cdbtmp__ = RiemannTensor(Ex(r'''R^{a}_{b c d}'''), Ex(r'''''') )
-# R_{a b c}^{d}::RiemannTensor.
-__cdbtmp__ = RiemannTensor(Ex(r'''R^{a}_{b c d}'''), Ex(r'''''') )
-__cdbtmp__ = RiemannTensor(Ex(r'''R^{a}_{b}^{d}_{c}'''), Ex(r'''''') )
-
-# note: this function only covers the cases that I have seen
-#       I want to keep this as short as possible
-def standard_indices (obj):
-    # single upstairs index on Rabcd
-    substitute (obj,Ex(r'''R_{a}^{b}_{c d}   -> - R^{b}_{a c d}''', False))
-    substitute (obj,Ex(r'''R_{a b}^{c}_{d}   ->   R^{c}_{d a b}''', False))
-    substitute (obj,Ex(r'''R_{a b c}^{d}     -> - R^{d}_{c a b}''', False))
-    # double upstairs index on Rabcd
-    substitute (obj,Ex(r'''R_{a}^{b}_{c}^{d} ->   R^{b}_{a}^{d}_{c}''', False))
[…]

### 2023-05-29 source/tex/paper.tex
Correct free indices on eqtns 11.19* take 2.
In the previous commit I made the changes (on the free indices) in the sources but forgot to do so in the main paper.tex file. Now fixed. Finally. Sheesh.
Simplified installation, updated hybrid-latex tools
1. Non-trivial refactoring of the build/install scripts.
   This was both to tidy the file structure as well as to simplify the installation
   process. The default install is now as simple as "source SETPATHS; make install".
   Almost all install/build/test/clean actions are now controlled by Makefiles.

2. Updated cdbpreproc.py.
   The Cadabra

@@
 %    First  solution use: biber tutorial and \usepackage[latin1]{inputenc}
-%    Second solution use: biber --decodecharsset=null tutorial
+%    Second solution use: biber --decodecharsset=null file
 \usepackage[style=numeric, sorting=none, sortcites=true]{biblatex}
@@
 \date{12-Aug-2009 (v1)\\
-      10-Dec-2019 (v2)}% major changes to correct errors in 4th and higher order terms,
+      10-Dec-2019 (v2)\\  % major changes to correct errors in 4th and higher order terms,
                           % all code will be available on github
+       4-Nov-2022 (v3)\\  % corrected typos on lhs of eqtns 11.19a-d,
+       1-Dec-2022 (v4)}   % corrected typo on equation (3.2) in section (3.1)
 \reference{Preprint: arXiv:0903.2087\\[5pt]
@@
 
-\BegNote{\bf\large Version 2}
+\bgroup
+\BegNote{\bf Version 2}
+\small%
 The first version of this paper was written in 2009. This second version (written in 2019)
@@
 corrected.
+% \EndNote
+\vskip 2pt
+
+\BegNote{\bf Version 3}
+\small%
+This version corrects the indicies (and sign) of the left hand side of equations 11.19a-d.
+This brings the arXiv version in line with the version on GitHub
+\ \url{https://github.com/leo-brewin/riemann-normal-coords}.
+% \EndNote
+\vskip 2pt
+
+\BegNote{\bf Version 4}
+\small%
+Equation (3.2) in Section 3.1 had a small error -- the coefficient in front of the $\Gamma \Gamma$
+term should be $(n-1)$ rather than $n$. Other instances of equation (3.2) (in section 4 and 7)
+were correct. This was a simple typographical error. All of the Cadabra results are unchanged.
 \EndNote
+\egroup
 
@@
 \begin{equation}
-\Gamma^{a}{}_{i_1i_2i_3\cdots i_n} = \Gamma^{a}{}_{(i_1i_2i_3\cdots i_{n-1},i_n)}
-                                 - n \Gamma^{a}{}_{p(i_2i_3\cdots i_{n-1}}
+  \Gamma^{a}{}_{i_1i_2i_3\cdots i_n}
+= \Gamma^{a}{}_{(i_1i_2i_3\cdots i_{n-1},i_n)}
+      - (n-1) \Gamma^{a}{}_{p(i_2i_3\cdots i_{n-1}}
               \Gamma^{p}{}_{i_1i_n)}
@@
    + \frac{1}{24} s^2 {\Dot x}^b {\Dot x}^c \Bigl(\cdb{ivp42}\Bigr)
-   + \frac{1}{12} s^3 {\Dot x}^b {\Dot x}^c {\Dot x}^d \cdb{ivp43}
+   + \frac{1}{12} s^3 {\Dot x}^b {\Dot x}^c {\Dot x}^d \bigl(\cdb{ivp43}\bigr)
    + \BigO{s^4,\eps^4}\end{dmath*}
@@
 % --------------------------------------------------------------------------------------------
-\mysubsection{Symmetrised partial derivatives of the%
+\mysubsection{Symmetrised partial derivatives of the %
               Riemann curvature tensor.}\label{sub:partialRabcd}
@@
 \begin{align}
-   x^{d}(s) ={}& x^{d}
-            + s {{\Dot x}^d}
-            + \frac{s^2}{2!} {{\Dot x}^a} {{\Dot x}^b} A^{d}_{ab}
-            + \frac{s^3}{3!} {{\Dot x}^a} {{\Dot x}^b} {{\Dot x}^c} A^{d}_{abc}
-            + \frac{s^4}{4!} {{\Dot x}^a} {{\Dot x}^b} {{\Dot x}^c}
-                                                       {{\Dot x}^d} A^{d}_{abce}\notag\\
-           &+ \frac{s^5}{5!} {{\Dot x}^a} {{\Dot x}^b} {{\Dot x}^c} {{\Dot x}^d}
-                                                       {{\Dot x}^e} A^{d}_{abcef}
+   x^{a}(s) ={}& x^{a}
+            + s {{\Dot x}^a}
+            + \frac{s^2}{2!} {{\Dot x}^b} {{\Dot x}^c} A^{a}_{bc}
+            + \frac{s^3}{3!} {{\Dot x}^b} {{\Dot x}^c} {{\Dot x}^d} A^{a}_{bcd}
+            + \frac{s^4}{4!} {{\Dot x}^b} {{\Dot x}^c} {{\Dot x}^d}
+                                                       {{\Dot x}^e} A^{a}_{bcde}\notag\\
+           &+ \frac{s^5}{5!} {{\Dot x}^b} {{\Dot x}^c} {{\Dot x}^d}
+                                                       {{\Dot x}^e}
+                                                       {{\Dot x}^f} A^{a}_{bcdef}
             + \BigO{s^6,\eps^6}\tag{11.19}
@@
 \begin{dgroup*}
-   \dMath{11.19a}{ 360 A^{d}_{ab} = \cdb{ivp62}}{eqn:ivp62}{\hfill}
-   \dMath{11.19b}{ 360 A^{d}_{abc} = \cdb{ivp63}}{eqn:ivp63}{\hskip 0.5cm}
-   \dMath{11.19c}{  90 A^{d}_{abce} = \cdb{ivp64}}{eqn:ivp64}{\hfill}
-   \dMath{11.19d}{   3 A^{d}_{abcef} = \cdb{ivp65}}{eqn:ivp65}{\hfill}
+   \dMath{11.19a}{ 360 A^{a}_{bc} = \cdb{ivp62}}{eqn:ivp62}{\hfill}
+   \dMath{11.19b}{ 360 A^{a}_{bcd} = \cdb{ivp63}}{eqn:ivp63}{\hskip 0.5cm}
+   \dMath{11.19c}{  90 A^{a}_{bcde} = \cdb{ivp64}}{eqn:ivp64}{\hfill}
+   \dMath{11.19d}{   3 A^{a}_{bcdef} = \cdb{ivp65}}{eqn:ivp65}{\hfill}
 \end{dgroup*}

### 2023-11-23 source/tex/paper.tex
Replaced Dx with \Delta x
Previous versions used both Dx and Delta x for the same thing -- the change in coordinates. Now I use just Delta x

@@
 \author{\Author\\[10pt]%
-School of Mathematical Sciences\\%
+School of Mathematics\\%
 Monash University, 3800\\%
@@
        4-Nov-2022 (v3)\\  % corrected typos on lhs of eqtns 11.19a-d,
-       1-Dec-2022 (v4)}   % corrected typo on equation (3.2) in section (3.1)
+       1-Dec-2022 (v4)\\  % corrected typo on equation (3.2) in section (3.1)
+      21-Nov-2023 (v5)}   % reformatted Cadabra ouput to print \Delta x rather than Dx
 \reference{Preprint: arXiv:0903.2087\\[5pt]
@@
 were correct their implementation in Cadabra (by this author) was incorrect. This lead to
-errors first in the quartic and higher order terms in the generalised connections and then,
-down the line, to computations built upon the generalised connections (e.g., the initial and
+errors first in the quartic and higher order terms in the generalised connections and later to computations built upon the generalised connections (e.g., the initial and
 boundary value problem and the arc-length equations). The Cadabra codes have been corrected
@@
 were correct. This was a simple typographical error. All of the Cadabra results are unchanged.
+
+% \EndNote
+\vskip 2pt
+
+\BegNote{\bf Version 5}
+\small%
+Versions 2 to 4 used both $Dx^a$ and $\Delta x^a$ to record the change in cooridnates between
+two nearby points. This version replaces the $Dx^a$ with $\Delta x^a$.
 \EndNote
@@
 \[
-L^2_{PQ} = g_{ab}(\bar x) Dx^a Dx^b + \BigO{\eps^4}
+L^2_{PQ} = g_{ab}(\bar x) \Dx^a \Dx^b + \BigO{\eps^4}
 \]
@@
 resources, taking between 0.5 second and 12 minutes to run and requiring
-between 40 and 360 Mbyte of memory (on a Mac Pro (2013) running macOS 10.14.4).
-
-The Cadabra codes and several support scripts are available from the author's GitHub site
-\ \url{https://github.com/leo-brewin/riemann-normal-coords}.
+between 40 and 360 Mbyte of memory (on a Mac Studio M1 running macOS 14.1.1).
 
@@
 side (as per the discussion in section \ref{sec:Notation}).
+
+Note also that in the following equations, $x^a$ denotes the Riemann normal coordinates of
+a typical point, $P$, while those of a nearby point $Q$ are denoted by $x^a + \Delta x^a$.
+The single exception to this rule occurs in equations (11.18) that displays a coordinate
+transformation from generic coordinates, $x^a$, to (non-orthogonal) Riemann normal
+coordinates, $y^a$.
 
@@
 \begin{align}
-   x^{a}(s) &= x^{a} + s Dx^{a}
+   x^{a}(s) &= x^{a} + s \Dx^{a}
                      + (s-s^2) x^{a}_2
@@
    \pglabel{eqn:geodesicLSQ}
-   % LCB: which of these is correct?
-   % \left(\Delta s\right)^2 = \nD{0} + \nD{2} + \nD{3} + \nD{4} + \nD{5} + \BigO{\eps^6,Dx^6}
-   % \left(\Delta s\right)^2 = \nD{0} + \nD{2} + \nD{3} + \nD{4} + \nD{5} + \BigO{\eps^6,Dx^7}
-   \left(\Delta s\right)^2 = \nD{0} + \nD{2} + \nD{3} + \nD{4} + \nD{5}
+   L^2_{PQ} = \nD{0} + \nD{2} + \nD{3} + \nD{4} + \nD{5}
             + \BigO{\eps^6}\tag{11.21}
@@
 
-All of the Cadabra files used in preparing this paper are available at the GitHub repository
-\ \url{https://github.com/leo-brewin/riemann-normal-coords}
+This paper along with all of the Cadabra files and several support scripts are available on
+the GitHub repository \ \url{https://github.com/leo-brewin/riemann-normal-coords}
 

### 2024-05-28 source/cadabra/export.tex
Simplified installation, updated hybrid-latex tools
1. Non-trivial refactoring of the build/install scripts.
   This was both to tidy the file structure as well as to simplify the installation
   process. The default install is now as simple as "source SETPATHS; make install".
   Almost all install/build/test/clean actions are now controlled by Makefiles.

2. Updated cdbpreproc.py.
   The Cadabra preprocessor will strip any in-line comments (such as cdb tags) from
   what it thinks are short lines. This failed for some of the examples (some in line
   comments were not deleted). This was fixed

@@
+\def\Date{19 Jan 2024}
+% \def\FileID{file:}
+
 \documentclass[12pt]{cdblatex}
@@
 
+   Dx{#}::LaTeXForm{"{\Dx}"}.  # LCB: This used to cause a bug, it kills ::KeepWeight for Dx
+                               # LCB: But I'm not sure that this is still true.
+                               # LCB: So just to be safe I'll leave all Cadabra sources unchanged.
+                               # LCB: I can use this trick in this file because it doesn't use ::KeepWeight
    import cdblib
@@
    tmp = cdblib.get ('gen_gamma_3_scaled','genGamma.export')  # cdb (genGamma3scaled,tmp)
+
+   # gen gamma in terms of partial derivs of Gamma^{a}_{bc}
+   tmp = cdblib.get ('gen_gamma_pderiv0','genGamma.export')  # cdb (genGammaPderiv0,tmp)
+   tmp = cdblib.get ('gen_gamma_pderiv1','genGamma.export')  # cdb (genGammaPderiv1,tmp)
+   tmp = cdblib.get ('gen_gamma_pderiv2','genGamma.export')  # cdb (genGammaPderiv2,tmp)
 

### 2024-05-28 source/cadabra/genGamma.tex
Simplified installation, updated hybrid-latex tools
1. Non-trivial refactoring of the build/install scripts.
   This was both to tidy the file structure as well as to simplify the installation
   process. The default install is now as simple as "source SETPATHS; make install".
   Almost all install/build/test/clean actions are now controlled by Makefiles.

2. Updated cdbpreproc.py.
   The Cadabra preprocessor will strip any in-line comments (such as cdb tags) from
   what it thinks are short lines. This failed for some of the examples (some in line
   comments were not deleted). This was fixed

@@
 % =================================================================================================
+% compute the first three generalised Gamms (including \Gamma^{a}_{bc})
+% this is just provide the LaTeX code for the first three \Gamma's in section 7 of the paper
+% these results are never used anywhere else
+
+\begin{cadabra}
+   deriv01:=B^{a}:
+
+   deriv02:=-\Gamma^{a}_{b c} B^{b} B^{c}:                # cdb (deriv02.100,deriv02)
+
+   deriv03:=\nabla{@(deriv02)}.                           # cdb (deriv03.100,deriv03)
+   distribute     (deriv03)
+   product_rule   (deriv03)                               # cdb (deriv03.101,deriv03)
+   substitute     (deriv03,$\nabla{B^{a}}->@(deriv02)$)   # cdb (deriv03.102,deriv03)
+   substitute     (deriv03,$\nabla{\Gamma^{m}_{s t}}->B^{d}\partial_{d}{\Gamma^{m}_{s t}}$)   # cdb (deriv03.103,deriv03)
+   sort_product   (deriv03)                               # cdb (deriv03.104,deriv03)
+   rename_dummies (deriv03)                               # cdb (deriv03.105,deriv03)
+   canonicalise   (deriv03)                               # cdb (deriv03.106,deriv03)
+
+   deriv04:=\nabla{@(deriv03)}.                           # cdb (deriv04.100,deriv04)
+   distribute     (deriv04)
+   product_rule   (deriv04)                               # cdb (deriv04.101,deriv04)
+   substitute     (deriv04,$\nabla{B^{a}}->@(deriv02)$)   # cdb (deriv04.102,deriv04)
+   substitute     (deriv04,$\nabla{\Gamma^{m}_{s t}}->B^{d}\partial_{d}{\Gamma^{m}_{s t}}$)   # cdb (deriv04.103,deriv04)
+   substitute     (deriv04,$\nabla{\partial_{e}{\Gamma^{m}_{s t}}}->B^{d}\partial_{d e}{\Gamma^{m}_{s t}}$)   # cdb (deriv04.104,deriv04)
+   sort_product   (deriv04)                               # cdb (deriv04.105,deriv04)
+   rename_dummies (deriv04)                               # cdb (deriv04.106,deriv04)
+   canonicalise   (deriv04)                               # cdb (deriv04.107,deriv04)
+
+   pderiv02 := -@(deriv02).                # cdb (pderiv02.100,pderiv02)
+   factor_out (pderiv02, $B^{a?}$)         # cdb (pderiv02.101,pderiv02)
+   substitute (pderiv02, $B^{a} -> 1$)     # cdb (pderiv02.102,pderiv02)
+
+   pderiv03 := -@(deriv03).                # cdb (pderiv03.100,pderiv03)
+   factor_out (pderiv03, $B^{a?}$)         # cdb (pderiv03.101,pderiv03)
+   substitute (pderiv03, $B^{a} -> 1$)     # cdb (pderiv03.102,pderiv03)
+
+   pderiv04 := -@(deriv04).                # cdb (pderiv04.100,pderiv04)
+   factor_out (pderiv04, $B^{a?}$)         # cdb (pderiv04.101,pderiv04)
+   substitute (pderiv04, $B^{a} -> 1$)     # cdb (pderiv04.102,pderiv04)
+\end{cadabra}
+
+\clearpage
+
+% =================================================================================================
+\section*{The generalised connection in generic coordinates (for the paper section 7)}
+
+\begin{dgroup*}[spread=5pt]
+   \begin{dmath*} {\Gamma}^{a}_{(bc)}(x) = \Cdb*{pderiv02.102} \end{dmath*}
+   \begin{dmath*} {\Gamma}^{a}_{(bcd)}(x) = \Cdb*{pderiv03.102} \end{dmath*}
+   \begin{dmath*} {\Gamma}^{a}_{(bcde)}(x) = \Cdb*{pderiv04.102} \end{dmath*}
+\end{dgroup*}
+
+\clearpage
+
+% =================================================================================================
 % export selected objects, these will later be imported into a library
@@
 
+   # gen gamma in terms of partial derivs of Gamma^{a}_{bc}
+   cdblib.put ('gen_gamma_pderiv0',pderiv02,'genGamma.export')
+   cdblib.put ('gen_gamma_pderiv1',pderiv03,'genGamma.export')
+   cdblib.put ('gen_gamma_pderiv2',pderiv04,'genGamma.export')
+
    checkpoint.append (tmp0)
@@
 
+   checkpoint.append (pderiv02)
+   checkpoint.append (pderiv03)
+   checkpoint.append (pderiv04)
 \end{cadabra}

### 2024-05-28 source/tex/paper.tex
Minor style changes.
Added a date to each pdf. Reinstated, in Section 7, the equations for the generalised connections in terms of the partial derivatives of the metric connection.

@@
 \small%
-The first version of this paper was written in 2009. This second version (written in 2019)
-differs from the first in a number of important ways. First, the discussion on Cadabra has
-been completely rewritten to account for the revised syntax introduced in the latest version
-of Cadabra (itself updated in 2016)\footnote{The Cadabra results in this paper are based on
-\CdbVersion}. Second, and more importantly, a significant computational error in the first
-version (of this paper) has been corrected. Though the algorithms given in the first version
-were correct their implementation in Cadabra (by this author) was incorrect. This lead to
-errors first in the quartic and higher order terms in the generalised connections and later to computations built upon the generalised connections (e.g., the initial and
-boundary value problem and the arc-length equations). The Cadabra codes have been corrected
-and thoroughly cross-checked. A number of minor typographical errors have also been
-corrected.
+The first version of this paper was written in 2009. This second version (written in
+2019) differs from the first in a number of important ways. First, the discussion on
+Cadabra has been completely rewritten to account for the revised syntax introduced in the
+latest version of Cadabra (itself updated in 2016)\footnote{The Cadabra results in this
+paper are based on \CdbVersion}. Second, and more importantly, a significant
+computational error in the first version (of this paper) has been corrected. Though the
+algorithms given in the first version were correct their implementation in Cadabra (by
+this author) was incorrect. This lead to errors first in the quartic and higher order
+terms in the generalised connections and later to computations built upon the generalised
+connections (e.g., the initial and boundary value problem and the arc-length equations).
+The Cadabra codes have been corrected and thoroughly cross-checked. A number of minor
+typographical errors have also been corrected.
 % \EndNote
+
 \vskip 2pt
@@
 % \EndNote
+
 \vskip 2pt
@@
 \small%
-Equation (3.2) in Section 3.1 had a small error -- the coefficient in front of the $\Gamma \Gamma$
-term should be $(n-1)$ rather than $n$. Other instances of equation (3.2) (in section 4 and 7)
-were correct. This was a simple typographical error. All of the Cadabra results are unchanged.
+Equation (3.2) in Section 3.1 had a small error -- the coefficient in front of the
+$\Gamma \Gamma$ term should be $(n-1)$ rather than $n$. Other instances of equation (3.2)
+(in section 4 and 7) were correct. This was a simple typographical error. All of the
+Cadabra results are unchanged.
 
@@
 \small%
-Versions 2 to 4 used both $Dx^a$ and $\Delta x^a$ to record the change in cooridnates between
-two nearby points. This version replaces the $Dx^a$ with $\Delta x^a$.
-\EndNote
+Versions 2 to 4 used both $Dx^a$ and $\Delta x^a$ to record the change in cooridnates
+between two nearby points. This version replaces the $Dx^a$ with $\Delta x^a$.
+
+% \EndNote
+\vskip 2pt
+
+\BegNote{\bf Version 6}
+\small%
+Added back, in section 7, expressions for the generalised connections in terms of the
+partial derivatives of the metric connection. These are the same expressions that
+appeared in version 1 but were dropped in versions 2-4.
+% \EndNote
 \egroup
@@
 
-Here are the first three generalised connections
+Here are the first three generalised connections in generic coordinates
 %
@@
 %-------------------------------------------------------------------------------
+%
+\begin{dgroup*}[compact]
+\begin{dmath*}\genGammaA = \cdb{genGammaPderiv0}\end{dmath*}%
+\begin{dmath*}\genGammaB = \cdb{genGammaPderiv1}\end{dmath*}%
+\begin{dmath*}\genGammaC = \cdb{genGammaPderiv2}\end{dmath*}%
+\end{dgroup*}
+%-------------------------------------------------------------------------------
+%
+and when the Riemman normal conditions are applied
 %

### 2024-06-18 source/tex/paper.tex
Correct syntax errors on ::LaTeXForm and ::Indices for Cadabra 2.5.2
Prior to version 2.5.2 Cadabra would allow LaTeXForm{"foo"} but since 2.5.2 must use round brackets. Also, since 2.5.2, cannot have whitespace between ::Indices and the round brackets, must use ::Indices(foo). Ditto for other properties (e.g., LaTeXForm("foo")). These were errors on my part that earlier versions of Cadabra let slip through. But 2.5.2 closes these loopholes.

@@
        1-Dec-2022 (v4)\\  % corrected typo on equation (3.2) in section (3.1)
-      21-Nov-2023 (v5)}   % reformatted Cadabra ouput to print \Delta x rather than Dx
+      21-Nov-2023 (v5)\\  % reformatted Cadabra ouput to print \Delta x rather than Dx
+      24-May-2024 (v6)}   % reinstated equations for generalised connection in section 7,
+                          % this matches section 7 of version 1
 \reference{Preprint: arXiv:0903.2087\\[5pt]

### 2024-10-28 source/cadabra/geodesic-ivp.tex
Simplified installation, updated hybrid-latex tools
1. Non-trivial refactoring of the build/install scripts.
   This was both to tidy the file structure as well as to simplify the installation
   process. The default install is now as simple as "source SETPATHS; make install".
   Almost all install/build/test/clean actions are now controlled by Makefiles.

2. Updated cdbpreproc.py.
   The Cadabra preprocessor will strip any in-line comments (such as cdb tags) from
   what it thinks are short lines. This failed for some of the examples (some in line
   comments were not deleted). This was fixed

@@
+\def\Date{30 Jul 2024}
+% \def\FileID{file:}
+
 \documentclass[12pt]{cdblatex}
@@
    sterm51 = flip_sign (cdblib.get ('genGamma31','genGamma.json'))
+
+   # note: the various ivp21, ivp31  etc. are the pieces of the Taylor series
+   #       for the ivp but *without* the leading 1/n! of the Taylor series
+
+   ivp21 := @(sterm21).                                          # cdb (ivp21.000,ivp21)
+
+   ivp31 := @(sterm21) + @(sterm22).                             # cdb (ivp31.000,ivp31)
+   ivp32 := @(sterm31).                                          # cdb (ivp32.000,ivp32)
+
+   ivp41 := @(sterm21) + @(sterm22) + @(sterm23).                # cdb (ivp41.000,ivp41)
+   ivp42 := @(sterm31) + @(sterm32).                             # cdb (ivp42.000,ivp42)
+   ivp43 := @(sterm41).                                          # cdb (ivp43.000,ivp43)
+
+   ivp51 := @(sterm21) + @(sterm22) + @(sterm23) + @(sterm24).   # cdb (ivp51.000,ivp51)
+   ivp52 := @(sterm31) + @(sterm32) + @(sterm33).                # cdb (ivp52.000,ivp52)
+   ivp53 := @(sterm41) + @(sterm42).                             # cdb (ivp53.000,ivp53)
+   ivp54 := @(sterm51).                                          # cdb (ivp54.000,ivp54)
+
+   factor_out (ivp21,$A^{a?}$)                                   # cdb (ivp21.001,ivp21)
+
+   factor_out (ivp31,$A^{a?}$)                                   # cdb (ivp31.001,ivp31)
+   factor_out (ivp32,$A^{a?}$)                                   # cdb (ivp32.001,ivp32)
+
+   factor_out (ivp41,$A^{a?}$)                                   # cdb (ivp41.001,ivp41)
+   factor_out (ivp42,$A^{a?}$)                                   # cdb (ivp42.001,ivp42)
+   factor_out (ivp43,$A^{a?}$)                                   # cdb (ivp43.001,ivp43)
+
+   factor_out (ivp51,$A^{a?}$)                                   # cdb (ivp51.001,ivp51)
+   factor_out (ivp52,$A^{a?}$)                                   # cdb (ivp52.001,ivp52)
+   factor_out (ivp53,$A^{a?}$)                                   # cdb (ivp53.001,ivp53)
+   factor_out (ivp54,$A^{a?}$)                                   # cdb (ivp54.001,ivp54)
+
+   v{#}::LaTeXForm("\dot{x}").
+
+   substitute (ivp21, $A^{a} -> v^{a}$)                          # cdb (ivp21.002,ivp21)
+
+   substitute (ivp31, $A^{a} -> v^{a}$)                          # cdb (ivp31.002,ivp31)
+   substitute (ivp32, $A^{a} -> v^{a}$)                          # cdb (ivp32.002,ivp32)
+
+   substitute (ivp41, $A^{a} -> v^{a}$)                          # cdb (ivp41.002,ivp41)
+   substitute (ivp42, $A^{a} -> v^{a}$)                          # cdb (ivp42.002,ivp42)
+   substitute (ivp43, $A^{a} -> v^{a}$)                          # cdb (ivp43.002,ivp43)
+
+   substitute (ivp51, $A^{a} -> v^{a}$)                          # cdb (ivp51.002,ivp51)
+   substitute (ivp52, $A^{a} -> v^{a}$)                          # cdb (ivp52.002,ivp52)
+   substitute (ivp53, $A^{a} -> v^{a}$)                          # cdb (ivp53.002,ivp53)
+   substitute (ivp54, $A^{a} -> v^{a}$)                          # cdb (ivp54.002,ivp54)
+
+   # build the Taylor series
+   # note the inclusion of the 1/n! factors
+
+   ivp2 := x^{a} + s v^{a} + (1/2) (s**2) @(ivp21).
+   ivp3 := x^{a} + s v^{a} + (1/2) (s**2) @(ivp31) + (1/6) (s**3) @(ivp32).
+   ivp4 := x^{a} + s v^{a} + (1/2) (s**2) @(ivp41) + (1/6) (s**3) @(ivp42) + (1/24) (s**4) @(ivp43).
+   ivp5 := x^{a} + s v^{a} + (1/2) (s**2) @(ivp51) + (1/6) (s**3) @(ivp52) + (1/24) (s**4) @(ivp53) + (1/120) (s**5) @(ivp54).
+
+   # cdb (ivp2.000,ivp2)
+   # cdb (ivp3.000,ivp3)
+   # cdb (ivp4.000,ivp4)
+   # cdb (ivp5.000,ivp5)
+
+   # now construct the scaled terms for ivp5
 
@@
    checkpoint.append (sterm5)
+
+   cdblib.create ('geodesic-ivp.json')
+
+   cdblib.put ('ivp21',ivp21,'geodesic-ivp.json')
+
+   cdblib.put ('ivp31',ivp31,'geodesic-ivp.json')
+   cdblib.put ('ivp32',ivp32,'geodesic-ivp.json')
+
+   cdblib.put ('ivp41',ivp41,'geodesic-ivp.json')
+   cdblib.put ('ivp42',ivp42,'geodesic-ivp.json')
+   cdblib.put ('ivp43',ivp43,'geodesic-ivp.json')
+
+   cdblib.put ('ivp51',ivp51,'geodesic-ivp.json')
+   cdblib.put ('ivp52',ivp52,'geodesic-ivp.json')
+   cdblib.put ('ivp53',ivp53,'geodesic-ivp.json')
+   cdblib.put ('ivp54',ivp54,'geodesic-ivp.json')
+
+   cdblib.put ('ivp2',ivp2,'geodesic-ivp.json')
+   cdblib.put ('ivp3',ivp3,'geodesic-ivp.json')
+   cdblib.put ('ivp4',ivp4,'geodesic-ivp.json')
+   cdblib.put ('ivp5',ivp5,'geodesic-ivp.json')
 \end{cadabra}
@@
    \begin{dmath*} \cdb*{sterm3.short.scaled.003} \end{dmath*}
+\end{dgroup*}
+
+\begin{dgroup*}
+   \begin{dmath*} \cdb*{ivp21.002} \end{dmath*}
+   \begin{dmath*} \cdb*{ivp31.002} \end{dmath*}
+   \begin{dmath*} \cdb*{ivp32.002} \end{dmath*}
+   \begin{dmath*} \cdb*{ivp41.002} \end{dmath*}
+   \begin{dmath*} \cdb*{ivp42.002} \end{dmath*}
+   \begin{dmath*} \cdb*{ivp43.002} \end{dmath*}
+   \begin{dmath*} \cdb*{ivp51.002} \end{dmath*}
+   \begin{dmath*} \cdb*{ivp52.002} \end{dmath*}
+   \begin{dmath*} \cdb*{ivp53.002} \end{dmath*}
+   \begin{dmath*} \cdb*{ivp54.002} \end{dmath*}
+\end{dgroup*}
+
+\begin{dgroup*}
+   \begin{dmath*} \cdb*{ivp2.000} \end{dmath*}
+   \begin{dmath*} \cdb*{ivp3.000} \end{dmath*}
+   \begin{dmath*} \cdb*{ivp4.000} \end{dmath*}
+   % \begin{dmath*} \cdb*{ivp5.000} \end{dmath*}
 \end{dgroup*}

### 2024-10-28 source/cadabra/metric-inv.tex
Simplified installation, updated hybrid-latex tools
1. Non-trivial refactoring of the build/install scripts.
   This was both to tidy the file structure as well as to simplify the installation
   process. The default install is now as simple as "source SETPATHS; make install".
   Almost all install/build/test/clean actions are now controlled by Makefiles.

2. Updated cdbpreproc.py.
   The Cadabra preprocessor will strip any in-line comments (such as cdb tags) from
   what it thinks are short lines. This failed for some of the examples (some in line
   comments were not deleted). This was fixed

@@
 
+   # these versions of "metric" are created just to add to the metric.json library
+   # note: term1 = 0, I could have used this fact above but ...
+
+   metric2:=@(term0) + @(term2).
+   metric3:=@(term0) + @(term2) + @(term3).
+   metric4:=@(term0) + @(term2) + @(term3) + @(term4).
+   metric5:=@(term0) + @(term2) + @(term3) + @(term4) + @(term5).
+
+   cdblib.put ('g^ab2',metric2,'metric-inv.json')
+   cdblib.put ('g^ab3',metric3,'metric-inv.json')
+   cdblib.put ('g^ab4',metric4,'metric-inv.json')
+   cdblib.put ('g^ab5',metric5,'metric-inv.json')
 \end{cadabra}

### 2024-10-28 source/cadabra/metric.tex
Simplified installation, updated hybrid-latex tools
1. Non-trivial refactoring of the build/install scripts.
   This was both to tidy the file structure as well as to simplify the installation
   process. The default install is now as simple as "source SETPATHS; make install".
   Almost all install/build/test/clean actions are now controlled by Makefiles.

2. Updated cdbpreproc.py.
   The Cadabra preprocessor will strip any in-line comments (such as cdb tags) from
   what it thinks are short lines. This failed for some of the examples (some in line
   comments were not deleted). This was fixed

@@
 
+   # these versions of "metric" are created just to add to the metric.json library
+   # note: term1 = 0, I could have used this fact above but ...
+
+   metric2:=@(term0) + @(term2).
+   metric3:=@(term0) + @(term2) + @(term3).
+   metric4:=@(term0) + @(term2) + @(term3) + @(term4).
+   metric5:=@(term0) + @(term2) + @(term3) + @(term4) + @(term5).
+
+   cdblib.put ('g_ab2',metric2,'metric.json')
+   cdblib.put ('g_ab3',metric3,'metric.json')
+   cdblib.put ('g_ab4',metric4,'metric.json')
+   cdblib.put ('g_ab5',metric5,'metric.json')
 \end{cadabra}

### 2025-04-12 source/cadabra/checks/check-genGamma.tex
Simplified installation, updated hybrid-latex tools
1. Non-trivial refactoring of the build/install scripts.
   This was both to tidy the file structure as well as to simplify the installation
   process. The default install is now as simple as "source SETPATHS; make install".
   Almost all install/build/test/clean actions are now controlled by Makefiles.

2. Updated cdbpreproc.py.
   The Cadabra preprocessor will strip any in-line comments (such as cdb tags) from
   what it thinks are short lines. This failed for some of the examples (some in line
   comments were not deleted). This was fixed

@@
+\def\Date{18 May 2024}
+% \def\FileID{file:}
+
 \documentclass[12pt]{cdblatex}
@@
 This code was written as a check for the {\tt genGamma.tex} code. I had a discrepency between my
-newly created Cadabra v2.0 codes and my old Cadabra v1.0 codes (which were the basis of my lcb09-03 paper).
-I found that my new codes agreed with this code and thus my old codes were wrong. I found the errors
-in the old code (see the updated codes in v1.0/rnc-new/gen-gamma.cdbp, see also the file v1.0/rnc-new/NOTES.txt).
+newly created Cadabra v2.0 codes and my old Cadabra v1.0 codes (which were the basis of my
+lcb09-03 paper). I found that my new v2.0 codes agreed with this code and thus my old v1.0 codes
+were wrong. I found the errors in the old code but it seems that (at some time) I deletd the
+details of those errors. I have no idea why I did that -- clearly another mistake on my part.
+What I do recall is that there were two errors, 1) I had declared {\tt
+\nabla{#}::PartialDerivative} and 2) In some crucial part of the generalised gamma computation
+the Cadabra {\tt asym} algorithm did not correctly do the full anti-symmetrisation. The first
+mistake is really a stupid mistake on my part -- it forces $\nabla_{ab} = \nabla_{ba}$. So all
+second and higher order derivatives of $R_{abcd}$ are wrong. This means that the expressions (in
+my v1.0 results) that depend upon the generalised gammas (e.g., the geodesic bvp, ivp and Lsq's)
+are wrong (from the second oder terms and above, yikes). The seond problem, to do with incomplete
+and incorrect anti-symmetrisations, was a Cadabra problem. The new v2.0 codes fixes these problem.
 

### 2025-04-12 source/cadabra/detg2.tex
Minor housekeeping. No major changes.
Fixed a few typos in some of the tex and sh files.
Simplified installation, updated hybrid-latex tools
1. Non-trivial refactoring of the build/install scripts.
   This was both to tidy the file structure as well as to simplify the installation
   process. The default install is now as simple as "source SETPATHS; make install".
   Almost all install/build/test/clean actions are now controlled by Makefiles.

2. Updated cdbpreproc.py.
   The Cadabra preprocessor will strip any in-line comments (such as cdb tags) from
   what it thinks are short lines. This fa

@@
 
-{\bf Note that Calzetta uses an opposite sign for $R_{abcd}$ so when comparing my results
-against Calzetta do take note of this flipped sign in $R_{abcd}$.}
+{\bf Note} that Calzetta etal. use an opposite sign for $R_{abcd}$ so when comparing the
+following results against Calzetta do take note of this flipped sign in $R_{abcd}$.
 
@@
 
-   # the following are based on sqrt-Ndetg.tex
+   # the following are based on sqrt-detg.tex
 

### 2025-04-12 source/cadabra/sqrt-detg.tex
Removed dependancy on cadabra2python
I used cadabra2python as a backdoor way to import .cdb files.
This would create a .py file which would then be imported using
a standard Python import statement.

But Cadabra can now (since at least version 2.5.x) import .cdb files directly.
So I no longer need the backdoor hack.

@@
+\def\Date{19 Jan 2024}
+% \def\FileID{file:}
+
 \documentclass[12pt]{pylatex}
+\usepackage{config}
 
@@
 
-Expand $\sqrt{g}$ in powers of $R,\nabla R$ etc.
+% =================================================================================================
+\section*{Side calculations for detg2.tex}
+
+This is a short computation for the expansions of $\sqrt{-\det g(x)}$ and $\log\left(-\det g(x)\right)$
+in powers of $R$ and its derivatives. The results wil be used in {\tts detg2.tex}.
+
+The starting point is to write $\det g(x)$ in the form
+\begin{dgroup*}
+   \begin{dmath*} -g(x) = 1 + a \epsilon^2 + b \epsilon^3 + c \epsilon^4 + d \epsilon^5 + \BigO{\epsilon^6} \end{dmath*}
+\end{dgroup*}
+where $\eps$ is introduced as an expansion parameter and where $a,b,c$ and $d$
+are simple expressions in $R$ and its derivatives. These can be read off directly
+from the expansion for $\det g(x)$ given in {\tts detg2.tex}.
+
+\clearpage
 
@@
 
+   det = 1 + eps**2*a + eps**3*b + eps**4*c + eps**5*d   # py (det.001,det)
+   # this foo will be used in detg2.tex
+   foo = Rational(1,2) + det/2 - (eps**4*a**2)/8 - (eps**5*a*b)/4   # py (foo.001,foo)
+   err = simplify (ans-foo)   # py (err.001,err)
+
 \end{python}
@@
 \end{dgroup*}
+
+\begin{dgroup*}
+   \begin{dmath*} \py*{det.001} \end{dmath*}
+   \begin{dmath*} \py*{foo.001} \end{dmath*}
+   \begin{dmath*} \py*{err.001} \end{dmath*}
+\end{dgroup*}
+
+\clearpage
 
@@
 
+   det = 1 + eps**2*a + eps**3*b + eps**4*c + eps**5*d   # py (det.001,det)
+   # this foo will be used in detg2.tex
+   foo = -1 + det - (eps**4*a**2)/2 - (eps**5*a*b)       # py (foo.002,foo)
+   err = simplify (ans-foo)   # py (err.002,err)
 \end{python}
@@
 
+\begin{dgroup*}
+   \begin{dmath*} \py*{det.001} \end{dmath*}
+   \begin{dmath*} \py*{foo.002} \end{dmath*}
+   \begin{dmath*} \py*{err.002} \end{dmath*}
+\end{dgroup*}
+
 \end{document}

### 2025-04-12 INSTALL.sh

@@ -1,34 +1,0 @@
-#!/bin/bash
-
-# defaults, edit to suit
-
-if [[ $HERE = '' ]]; then
-   HERE=$HOME/local/rnc/
-fi;
-
-MyBin=$HERE/bin/
-MyLib=$HERE/lib/
-MyTex=$HERE/tex/
-
-# Parse the command-line options
-
-while getopts 'b:l:t:h' option
-do
-   case "$option" in
-   "b")  MyBin="$OPTARG"      ;;
-   "l")  MyLib="$OPTARG"      ;;
-   "t")  MyTex="$OPTARG"      ;;
-   "h")  echo "usage : INSTALL.sh [-bBIN] [-lLIB] [-tTEX]"
-         echo "options : -b : where to find the binaries and shell scripts"
-         echo "          -l : where to find the python libraries"
-         echo "          -t : where to find the LaTeX files "
-         echo "          -h : this help message"
-         exit                 ;;
-   ?)    echo "INSTALL.sh: Unknown option specified."
-         exit                 ;;
-   esac
-done
-
-mkdir -p $MyBin $MyLib $MyTex
-
-(cd hybrid-latex; INSTALL.sh -b $MyBin -l $MyLib -t $MyTex)

### 2025-04-12 hybrid-latex/INSTALL.sh

@@ -1,36 +1,0 @@
-#!/bin/bash
-
-# defaults, edit to suit
-
-if [[ $HERE = '' ]]; then
-   HERE=$HOME/local/rnc/
-fi;
-
-MyBin=$HERE/bin/
-MyLib=$HERE/lib/
-MyTex=$HERE/tex/
-
-# Parse the command-line options
-
-while getopts 'b:l:t:h' option
-do
-   case "$option" in
-   "b")  MyBin="$OPTARG"      ;;
-   "l")  MyLib="$OPTARG"      ;;
-   "t")  MyTex="$OPTARG"      ;;
-   "h")  echo "usage : INSTALL.sh [-bBIN] [-lLIB] [-tTEX]"
-         echo "options : -b : where to find the binaries and shell scripts"
-         echo "          -l : where to find the python libraries"
-         echo "          -t : where to find the LaTeX files "
-         echo "          -h : this help message"
-         exit                 ;;
-   ?)    echo "INSTALL.sh: Unknown option specified."
-         exit                 ;;
-   esac
-done
-
-mkdir -p $MyBin $MyLib $MyTex
-
-cp -rf scripts/* $MyBin
-cp -rf python/*  $MyLib
-cp -rf latex/*   $MyTex

### 2025-04-12 hybrid-latex/scripts/pylatex.sh

@@ -1,16 +1,14 @@
 #!/bin/bash
 
-export CADABRA_NO_UNICODE=1
-export PYTHONWARNINGS="ignore"
-
 file="<none>"
 silent="no"
 keep="no"
 skiplatex="no"
 Timer=""
-CDB=/usr/local/bin/
+Python="python"
 sty=""
 nowarn=""
+mixed=""
 reformat="yes"
 
 # -----------------------------------------------------------------------------------------
@@ -18,35 +16,37 @@ reformat="yes"
 
 OPTIND=1
 
-while getopts 'i:I:P:sktTxhNX' option
+while getopts 'i:I:P:sktTxhNM' option
 do
    case "$option" in
    "i")  file="$OPTARG"           ;;
    "I")  sty="-I$OPTARG"          ;;
-   "P")  CDB="$OPTARG"            ;;
+   "P")  Python="$OPTARG"         ;;
    "s")  silent="yes"             ;;
    "k")  keep="yes"               ;;
    "t")  Timer="/usr/bin/time"    ;;
    "T")  Timer="/usr/bin/time -l" ;;
    "x")  skiplatex="yes"          ;;
    "N")  nowarn="-N"              ;;
+   "M")  mixed="-M"               ;;
    "X")  reformat="no"            ;;
-   "h")  echo "usage : cdblatex.sh -i file [-P<path to Cadabra bin dir>]"
-         echo "                            [-I<path to cdbmacros.sty>] [-sktTxNXh]"
+   "h")  echo "usage : pylatex.sh -i file [-P<path to python>]"
+         echo "                           [-I<path to pymacros.sty>] [-sktTxNMh]"
          echo "options :  -i file : source file (with or without .tex extension)"
-         echo "           -I file : full path to cdbmacros.sty file"
-         echo "           -P file : path to Cadabra bin directory"
+         echo "           -I file : full path to pymacros.sty file"
+         echo "           -P path : full path to the Python binary"
          echo "           -s : silent, don't open the pdf file"
          echo "           -k : keep all temporary files"
          echo "           -t : report brief cpu time"
          echo "           -T : report detailed cpu time plus memory usage"
          echo "           -x : don't call latex"
          echo "           -N : don't warn if errors found in the output for some tags"
[…]
